---
title: "Scaling the Narrative Identification of Fiscal Changes using LLMs"
date: last-modified
date-format: long
abstract: |
  *Draft abstract (written last, once the body is settled).* We build and
  validate an LLM-assisted pipeline that transfers the Romer & Romer (2010)
  narrative approach to fiscal-shock identification from the United States to
  emerging markets. Country-agnostic codebooks are validated on 44 labelled US
  acts under the Halterman & Keith (2025) framework, then deployed to Malaysia
  (1980–2023), where an agentic component layer produces an expert-ready
  inventory of statutory tax, spending, and tax-incentive changes — each
  carrying a motivation and a preliminary exogeneity read. We demonstrate the
  dataset with a first procyclicality decomposition.
number-sections: true
section-numbering: "1.1.1"
format:
  html:
    toc: true
    comments:
      hypothesis: true
  docx: default
  jats: default
  academic-typst: default
execute:
  echo: false
  warning: false
  cache: true
  freeze: auto
---

```{r}
#| label: setup
#| cache: false
#| message: false
#| warning: false

pacman::p_load(targets, tidyverse, tinytable, here, scales, patchwork)

here::i_am("index.qmd")
tar_config_set(store = here("_targets"))

source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))     # pretty_motivation() + palette
source(here("R/tax_shock_report.R"))
source(here("R/spending_shock_report.R"))
source(here("R/incentive_shock_report.R"))
source(here("R/iteration_reporting.R"))   # H&K S1-S3 figure/table helpers
source(here("R/pipeline_diagram.R"))      # plot_pipeline_flow()
source(here("R/dataset_summary.R"))       # cross-component summaries
source(here("R/codebook_display.R"))      # verbatim codebook exhibits
source(here("R/recall_scorecard.R"))      # frozen recall audits
theme_set(theme_minimal())

# ---- Empty-safe reads (rendering is read-only: tar_read only, never tar_make) --
# Malaysia deliverables (API-gated enrichment tails — may be unbuilt)
tax_shocks       <- tryCatch(tar_read(tax_shocks),       error = \(e) tibble())
spending_shocks  <- tryCatch(tar_read(spending_shocks),  error = \(e) tibble())
incentive_shocks <- tryCatch(tar_read(incentive_shocks), error = \(e) tibble())
have_tax        <- nrow(tax_shocks) > 0
have_spending   <- nrow(spending_shocks) > 0
have_incentives <- nrow(incentive_shocks) > 0
have_any        <- have_tax || have_spending || have_incentives

# Deployment corpus (API-free)
country_body   <- tryCatch(bind_rows(tar_read(country_body)),   error = \(e) tibble())
country_chunks <- tryCatch(bind_rows(tar_read(country_chunks)), error = \(e) tibble())
have_corpus    <- nrow(country_body) > 0

# H&K validation logs (API-free; parsed from prompts/iterations/*.yml)
il      <- tryCatch(tar_read(iteration_logs), error = \(e) NULL)
have_il <- !is.null(il)

# Select the latest formal (Claude) iteration of a codebook within a parsed log.
latest <- function(tbl, cb) {
  tbl |>
    dplyr::filter(codebook == cb, grepl("claude", model)) |>
    dplyr::filter(iteration == max(iteration))
}

# Malaysia EN/BM consistency targets (precomputed; may be unbuilt)
malay_er_c1_measures <- tryCatch(tar_read(malay_er_c1_measures), error = \(e) tibble())
malay_metrics        <- tryCatch(tar_read(malay_er_consistency_metrics), error = \(e) NULL)
malay_inv            <- if (is.null(malay_metrics)) tibble() else malay_metrics$inventory
have_malay           <- nrow(malay_inv) > 0
```

::: {.callout-note}
**Draft status.** The body sections and their exhibits are in place. The
introduction, abstract, and conclusion are written last, from the populated
body. Figures whose upstream targets are API-gated and not yet built render as
a "pending" note rather than an error.

The Malaysia figures below are built from the deliverable as frozen in June
2026. An evidence-scoping defect found in August 2026 (§[-@sec-defect]) has been
fixed in the pipeline but the enrichment tails have not yet been re-run, so four
of the 51 motivation labels are known to be wrong and are corrected in the
narrative, not yet in the exhibits.
:::

# Introduction

*Intent: the "why" and the "what is new." For most emerging markets we lack
consistent, comparable measures of exogenous fiscal shocks — the input credible
fiscal analysis needs [@vegh_how_2015; @romer2010]. LLMs make the narrative
approach scalable for the first time. Our contribution is a **method plus a
demonstration dataset**: country-agnostic codebooks, validated on the US and
deployed to Malaysia, that hand an expert an audit-ready inventory of fiscal-change
events. We position against the three LLM-narrative competitors
[@das_ai_2026; @bhasin_identifying_2026; @fritsch_high-frequency_nodate] on four
wedges they each cede — primary national documents, the tax side, full-richness
rows (measure × exo/endo × sign), and ground-truth-free validation. Prose sources:
`docs/brainstorm.qmd`, `docs/lit_review.qmd` §1/§4/§7.*

# The narrative specification and the validation philosophy

*Intent: two ideas the rest of the paper depends on. (1) The **specification** —
Romer & Romer's narrative method reads documents to recover, per act, the fiscal
*measure*, its *motivation* (the 4-way taxonomy), the *exogenous* filter, the
*sign*, and — in full — *timing* and *magnitude* [@romer2010; @romer_presidential_2023].
(2) The **discipline** — Halterman & Keith's five-stage framework treats codebook
conceptualisation as a first-order concern and shows off-the-shelf LLMs follow real
codebooks imperfectly zero-shot, so every codebook must pass behavioural tests,
zero-shot evaluation, and error analysis before it is trusted [@halterman2025].
Prose sources: `docs/lit_review.qmd` §2/§5, `docs/strategy.md`,
`docs/codebook_sources.md`.*

# The pipeline and what it produces {#sec-pipeline}

The method reads primary-source government documents and returns an inventory of
fiscal-change events, each carrying a motivation, a sign, and a preliminary
exogeneity read. @fig-pipeline is the route from one to the other, and it is the
map for the rest of the paper: §[-@sec-build] walks the route for Malaysia,
§[-@sec-us-validation] reports what the US benchmark certifies about the blue
stages, and §[-@sec-coverage] states what the other stages rest on instead.

```{r}
#| label: fig-pipeline
#| fig-cap: "**From primary sources to an event inventory.** Two LLM codebook stages (blue) are validated against US ground truth; the act-identification step between them (orange) is an agentic narrative search whose output a human stamps before it enters the chain."
#| fig-width: 6.5
#| fig-height: 2.4
plot_pipeline_flow("none")
```

Two design commitments distinguish this route from a single end-to-end prompt.
The first is that **classification and discovery are separated**. Deciding
whether a passage describes a fiscal measure, and why the measure was enacted,
is a coding task with a codebook, a gold standard, and a validation record.
Deciding which acts exist in a corpus and where each begins and ends is a
retrieval-and-consolidation task with no gold standard, and it is handled
separately and audited differently. The second is that **every stage is
reversible to its source**: each event carries the documents, chunks, and
verbatim quotes it was built from, so a reviewer disagreeing with a label can go
to the passage rather than to the model.

## The two frameworks the pipeline implements

The pipeline implements one methodological specification and one validation
discipline. @tbl-rr-steps maps the specification: what Romer and Romer do by
hand, and which component does it here. @tbl-hk-stages maps the discipline: the
stages a codebook must clear before we trust its output.

```{r}
#| label: tbl-rr-steps
#| tbl-cap: "**The narrative specification.** Romer & Romer's procedure and the component that carries each step here. Steps RR5 and RR6 are where classification credibility lives; RR3 is deliberately not automated."
tibble::tribble(
  ~Step,  ~`Romer & Romer's step`,                          ~`Our component`,                     ~Status,
  "RR1",  "Read primary-source policy documents",           "Corpus acquisition and chunking",    "Built",
  "RR2",  "Identify legislated fiscal changes",             "C1 measure identification (LLM)",    "Validated on US",
  "RR3",  "Consolidate mentions into distinct acts",        "Agentic narrative identification",   "Human-stamped",
  "RR4",  "Record the size of each change",                 "Extracted with the act",             "Not separately validated",
  "RR5",  "Classify the motivation, filter to exogenous",   "C2a evidence, C2b classification (LLM)", "Validated on US",
  "RR6",  "Date the change and assemble the series",        "Announced and effective year",       "Annual only; C3 pending"
) |>
  tt(width = c(0.08, 0.34, 0.32, 0.26)) |>
  tt_theme_report() |>
  style_tt(j = 2:4, align = "l")
```

```{r}
#| label: tbl-hk-stages
#| tbl-cap: "**The validation discipline.** Halterman & Keith's stages and how far each codebook has been taken. S4 is a last resort we have not needed."
tibble::tribble(
  ~Stage, ~`What it establishes`,                                          ~C1,        ~C2,
  "S0",   "Machine-readable codebook, expert-approved",                    "Complete", "Complete",
  "S1",   "Behavioural tests: legal outputs, memorisation, order",         "Passed",   "Passed",
  "S2",   "Zero-shot evaluation against labelled data",                    "Passed",   "Passed",
  "S3",   "Error analysis, ablation, manual adjudication",                 "Passed",   "Passed",
  "S4",   "Fine-tuning, if S3 exhausts codebook improvements",             "Not needed", "Not needed"
) |>
  tt(width = c(0.09, 0.51, 0.20, 0.20)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

## What the pipeline produces {#sec-what-it-produces}

```{r}
#| label: pending-any
#| eval: !expr "!have_any"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe Malaysia deliverable targets are empty or",
    "unbuilt (their C2a/C2b enrichment is API-gated). The dataset exhibits",
    "below render once they are built.\n:::\n")
```

@fig-dataset-overview is the whole Malaysia dataset in one frame: 51
narratively-announced fiscal events between 1980 and 2023, split into the three
instrument families the pipeline covers. Three features of the data are visible
before any table is read. Statutory tax changes are dominated by rate *cuts*
spread evenly across the period and are mostly read as exogenous. Spending
events cluster tightly on 1998, 2009, and 2020–21 and are almost all read as
endogenous — which is what a crisis-response spending record should look like.
Tax incentives run the full period and are almost uniformly liability-reducing
and exogenous, which is what a long-run investment-promotion regime looks like.

```{r}
#| label: fig-dataset-overview
#| fig-cap: "**The Malaysia inventory at a glance.** One mark per announced fiscal event, positioned at its year of effect; marks stack where several events share a year. Colour is the direction of the change, fill is the preliminary exogeneity read."
#| fig-width: 6.5
#| fig-height: 4.2
#| eval: !expr have_any
p <- plot_dataset_overview(tax_shocks, spending_shocks, incentive_shocks)
if (is.null(p)) cat("No component data available.\n") else p
```

```{r}
#| label: tbl-dataset-summary
#| tbl-cap: "**Inventory summary.** Coverage and composition per component. Source documents and evidence chunks count the primary-source passages each event was built from; the exogenous share is the model's preliminary read, not an adjudicated finding."
#| eval: !expr have_any
dataset_summary_stats(tax_shocks, spending_shocks, incentive_shocks) |>
  tt() |>
  tt_theme_report()
```

The unit of observation is one narratively-announced act, per instrument.
@tbl-one-row is the schema; Exhibit 1 below is one real row printed in full,
which is the faster way to see what a row means.

```{r}
#| label: tbl-one-row
#| tbl-cap: "**Unit of observation.** One row per narratively-announced act × instrument, carrying a magnitude, a motivation label, and a preliminary exogeneity read pending expert adjudication."

tibble::tribble(
  ~Field,            ~Meaning,
  "act",             "One narratively-announced fiscal act (phased steps collapse to one row)",
  "instrument",      "CIT / PIT / CONSUMPTION (tax), Expenditure (spending), or incentive category",
  "year",            "Announced and effective year (quarterly timing is deferred)",
  "magnitude",       "Δpp statutory rate (tax or rate-bearing incentive) or ringgit amount (spending)",
  "sign",            "Direction of the effect on fiscal liabilities (increase / decrease / no change)",
  "motivation",      "Romer–Romer 4-way label from the validated classification codebook",
  "exogenous",       "Preliminary exogenous/endogenous read — an input for the expert, not a finding",
  "provenance",      "Source documents, member chunks, and the verbatim quotes behind each field"
) |>
  tt(width = c(0.22, 0.78)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

```{r}
#| label: exhibit-anatomy
#| echo: false
#| output: asis
#| eval: !expr have_tax
tax_exhibit(
  tax_shocks, "MY-CIT-01",
  exhibit_label = "Exhibit 1",
  note = "One row in full. Every field below is either read from a named primary source or produced by a validated codebook from the quoted evidence."
)
```

## The Malaysia inventory in detail

The figures below are the headline output of the pipeline, for Malaysia
(1980–2023). Downloads, data dictionaries, and the full worked exhibits live on
the reviewer-facing dataset page,
[malaysia_dataset.qmd](notebooks/malaysia_dataset.qmd).

### Statutory tax changes

```{r}
#| label: pending-tax
#| eval: !expr "!have_tax"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `tax_shocks` deliverable target is empty or",
    "unbuilt (its C2a/C2b enrichment is API-gated). The tax figures below render",
    "once it is built.\n:::\n")
```

```{r}
#| label: tbl-tax-inventory
#| tbl-cap: "**Statutory tax-change inventory.** One row per announced act × tax type; Δpp is the headline-rate change; motivation and exogeneity from the validated classification codebook."
#| eval: !expr have_tax
tax_inventory_table(tax_shocks)
```

```{r}
#| label: fig-tax-rate-paths
#| fig-cap: "**Statutory rate paths by tax type.** Step path of the headline statutory rate implied by the rate-bearing acts."
#| eval: !expr have_tax
plot_tax_rate_paths(tax_shocks)
```

```{r}
#| label: fig-tax-timeline
#| fig-cap: "**Tax-change timeline.** Acts per year × tax type as a diverging stacked bar (hikes up, cuts down); colour denotes the motivation label."
#| eval: !expr have_tax
plot_tax_timeline(tax_shocks)
```

### Government spending shocks

```{r}
#| label: pending-spending
#| eval: !expr "!have_spending"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `spending_shocks` deliverable target is empty or",
    "unbuilt. The spending figures below render once it is built.\n:::\n")
```

```{r}
#| label: tbl-spending-inventory
#| tbl-cap: "**Government spending-shock inventory.** One row per announced spending act; exogeneity from the validated classification codebook, alongside the @das_ai_2026 two-condition preliminary screen."
#| eval: !expr have_spending
spending_inventory_table(spending_shocks)
```

```{r}
#| label: fig-spending-timeline
#| fig-cap: "**Spending-shock timeline.** Acts per year as a diverging stacked bar (increases up, decreases down); colour denotes the exogenous/endogenous read."
#| eval: !expr have_spending
plot_spending_timeline(spending_shocks)
```

### Tax incentives and holidays

```{r}
#| label: pending-incentives
#| eval: !expr "!have_incentives"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `incentive_shocks` deliverable target is empty or",
    "unbuilt. The incentive figures below render once it is built.\n:::\n")
```

```{r}
#| label: tbl-incentive-inventory
#| tbl-cap: "**Tax-incentive shock inventory.** One row per announced incentive change; exogeneity from the validated classification codebook."
#| eval: !expr have_incentives
incentive_inventory_table(incentive_shocks)
```

```{r}
#| label: fig-incentive-timeline
#| fig-cap: "**Incentive-shock timeline.** Acts per year as a diverging stacked bar (new or expanded down, repeals up); colour denotes the exogenous/endogenous read."
#| eval: !expr have_incentives
plot_incentive_timeline(incentive_shocks)
```

# How the Malaysia dataset was built {#sec-build}

This section walks @fig-pipeline for Malaysia, stage by stage, so the dataset's
provenance is reconstructible. The short version: a 168-document primary-source
corpus was scanned exhaustively by a validated codebook, the scan was used as a
seed rather than as the event list, acts were consolidated by an agentic
narrative search that was audited for recall and stamped by a human, and only
then were the validated motivation codebooks run against each act's own
evidence.

## The corpus {#sec-corpus}

```{r}
#| label: pending-corpus
#| eval: !expr "!have_corpus"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe deployment corpus targets are unbuilt, so the",
    "corpus table cannot render.\n:::\n")
```

Malaysia's fiscal record is published across four document series, none of which
alone covers the period. Budget Speeches announce measures but do not always
report the rates they replace; Economic Reports carry the retrospective rate
tables; Bank Negara annual reports carry the macro framing that motivation
classification depends on; and the Five-Year Plans carry the structural and
development-expenditure commitments. The corpus takes all four, plus the
standalone crisis packages, and it is bilingual: ten Economic Reports are
available only in Bahasa Malaysia, which is what makes the cross-language test
in §[-@sec-my-validation] possible.

```{r}
#| label: tbl-corpus
#| tbl-cap: "**The deployment corpus.** Primary-source document series behind the Malaysia dataset. Pages are extracted pages, before chunking."
#| eval: !expr have_corpus
corpus_scope_stats(country_body) |>
  tt(width = c(0.36, 0.20, 0.14, 0.12, 0.18)) |>
  tt_theme_report() |>
  style_tt(j = 1, align = "l")
```

```{r}
#| label: corpus-chunks
#| echo: false
#| output: asis
#| eval: !expr "nrow(country_chunks) > 0"
cat(sprintf(
  "\nAfter extraction and cleaning, the corpus was segmented into %s chunks across %d documents, which is the unit the measure-identification codebook reads.\n\n",
  scales::comma(nrow(country_chunks)),
  dplyr::n_distinct(country_chunks$doc_id)
))
```

## The codebook scan and what it is for {#sec-scan}

The measure-identification codebook reads every chunk and returns the fiscal
measures discussed in it. On the Malaysia corpus it surfaced **7,676
measure-mentions** across 1,193 chunks, at a median of five measures per chunk
and a maximum of 42 — a density that matters twice below, once here and once in
§[-@sec-defect].

That scan is a *seed*, not the event list, for two reasons. The first is
mechanical: the handoff to the motivation stage keeps only the highest-ranked
measure per chunk, which discards **84.5%** of the mentions and leaves 1,036
measure rows. The second is conceptual: a scan returns mentions, and one act
generates many mentions across many documents and both languages, while some
acts generate mentions the scan phrases differently each time. Turning mentions
into acts is a separate problem, and it is the subject of the next two
subsections.

## Agentic narrative identification {#sec-agentic}

Acts are identified per instrument by an agentic procedure that reads the corpus
directly, in seven steps.

1. **Instrument configuration.** The concept is defined (statutory corporate
   income tax; broad consumption tax; discretionary spending; investment
   incentives), together with the surface terms it takes in the corpus, in both
   English and Bahasa Malaysia, and the document series that anchor it.
2. **Keyword scan.** The configured terms are matched against the codebook's
   measure pool. This is the recall floor and it is logged as a count, with
   false positives named.
3. **Semantic sweep.** The matched measures and their chunk text are read, and
   the search broadens past the keyword list to catch events phrased without the
   vocabulary — a rate change written as "reduce the rate to 24 per cent" and
   never as "corporate income tax".
4. **Recall recovery.** For rate instruments, the implied statutory rate path is
   reconstructed and checked for missing steps; for spending and incentives, the
   candidate set is checked against pre-declared known-act checkpoints. Where a
   step is missing, the anchor document's full text is read directly, behind the
   chunked view. This step is what recovered Malaysia's 2016 corporate rate cut
   and its 2022 Cukai Makmur surcharge, neither of which the scan surfaced.
5. **Consolidation.** Surface forms and recovered events are grouped into
   announced acts. A phased multi-step cut is one row carrying a schedule, not
   four rows; an omnibus act touching two taxes is two rows, one per instrument.
   A preliminary exogeneity read is recorded, anchored in a verbatim quote.
6. **Human review.** The consolidated events, the recall audit, and every open
   adjudication question are presented for review. Nothing is written until the
   researcher confirms the events, the consolidation grain, and the preliminary
   reads.
7. **Freeze.** The stamped dataset is written with a provenance record naming
   the reviewer, the commit, and the date, and a notebook that renders its
   contents back from the frozen file so the write-up cannot drift from the
   artifact.

The step that carries the method's honesty is the recall audit
(@tbl-recall-cit). Precision can be checked by anyone holding the dataset:
each row names its sources and quotes. Recall cannot — a reader cannot see what
was never found. The scorecard is therefore the only record of what was
searched, what the scan missed, and how each miss was recovered, and it is
frozen alongside the data.

The corporate income tax audit is reproduced below (@tbl-recall-cit); the other
four instruments carry the same structure and are in @sec-appendix-recall.

```{r}
#| label: tbl-recall-cit
#| tbl-cap: "**Recall audit, corporate income tax.** Frozen with the component dataset. Rows naming a recall miss record an event the codebook scan did not surface and the direct-reading pass recovered; outcome text is truncated for display."
#| eval: !expr have_tax
tb <- recall_scorecard_table(tax_shocks, instrument = "Corporate income tax")
if (is.null(tb)) cat("No recall scorecard available.\n") else tb
```

Two results in that audit are worth stating plainly because they cut against the
design's own expectations. First, the keyword floor was **not** near-empty for
spending and incentives, as a tax-scoped scan would predict: 189 spending
measures and 151 incentive measures were surfaced, because Malaysian budget and
central-bank documents discuss stimulus, subsidy, and incentive policy alongside
tax. The scan is a genuinely rich first pass across all three instrument
families. Second, and against that, the tax passes still recorded six recall
misses on the single track the scan should have been best at — the statutory
rate path — each recovered only by reading the anchor documents directly.

## Act consolidation without automated record linkage {#sec-no-c0}

We built an automated act-aggregation step for this purpose and did not use it.
The step clusters the scan's surface forms into canonical acts and was scored
against the 49 Romer–Romer reference acts, where it recovers most of them. On
the Malaysia pool of 877 distinct surface forms it fails in the way that matters
for an event layer: it fragments the instrument tracks that need to stay whole —
the corporate-rate track shattered into near-singletons, the consumption-tax
track split across nine clusters — and it never bridged the English and Bahasa
Malaysia names of the same act. Both failures are recorded in the frozen recall
audits above.

The finding is a negative result about clustering, not about LLMs: consolidating
mentions into acts requires knowing which mentions describe the same *statutory
instrument*, and surface-form similarity does not carry that. The agentic
procedure of §[-@sec-agentic] succeeds where clustering fails because it works
from the instrument down — reconstruct the rate path, then ask which mentions
account for each step — rather than from the mentions up. That is also why it
needs a human stamp: the reconstruction is a judgement about what counts as one
act, and the researcher owns it.

## A defect the evidence chain surfaced {#sec-defect}

The motivation stage is two codebooks: an extraction step that pulls the
motivation evidence for an act out of a passage, and a classification step that
assigns the four-way label from that evidence. During reviewer inspection of two
corporate-tax rows in August 2026, both codebooks turned out to be reading the
wrong act's evidence, through two independent mechanisms.

The first was a join. Evidence had been extracted once per chunk, under whichever
measure the scan ranked first — the 84.5% discard of §[-@sec-scan] — and events
were matched to that store on chunk identity alone. But the identification
procedure cites a chunk because the act's sentence is in it, not because the act
was ranked first. 95 of 100 tax member-chunk rows, and every spending and
incentive row, were structurally exposed. The second was the extraction codebook
itself, which had been instructed to extract *all* motivation evidence in a
chunk. That is harmless on the US development chunks, which discuss one act
each, and wrong on Malaysian budget speeches carrying a median of five measures
and a maximum of 42.

An audit of the 14 events where the model and the identification pass disagreed
about exogeneity found four classifications built on another act's evidence: two
corporate-tax rows, two personal-income-tax rows. Both mechanisms are fixed —
extraction is now scoped to the named act, and evidence is extracted per event
rather than per chunk — and the enrichment has not yet been re-run, so the four
rows remain wrong in the exhibits above.

We report this as a result rather than as a maintenance note, because it is the
kind of failure the validation framework exists to catch and did not. Both
codebooks passed every US test. They failed on a distributional difference
between the validation corpus and the deployment corpus — single-act chunks
versus multi-measure budget speeches — that no US test could have exposed. The
cost is quantified because the provenance chain made it quantifiable: four wrong
labels out of 51, found by following each label back to the passage it came
from. A pipeline without that chain would have shipped them.

# Validation I — the US benchmark {#sec-us-validation}

The codebooks are validated on 44 labelled US Romer–Romer acts, under the
Halterman & Keith stage sequence: behavioural tests, then zero-shot evaluation,
then error analysis. @fig-pipeline-validated marks what this covers, and the
gap is as important as the coverage: the benchmark certifies the two LLM
classification stages and says nothing about corpus construction or act
identification, which are validated in other ways or not at all
(§[-@sec-coverage]).

```{r}
#| label: fig-pipeline-validated
#| fig-cap: "**What the US benchmark covers.** Stages carrying a formal validation record against US ground truth are lit; the rest are attested by other means."
#| fig-width: 6.5
#| fig-height: 2.4
plot_pipeline_flow("validated")
```

One caveat frames everything that follows. The published Romer–Romer shock list
and the US source documents are almost certainly in the models' pretraining
data, so strong US agreement is **consistent with memorisation**
[@carlini_quantifying_2022]. This is why the behavioural battery includes an
explicit memorisation probe, and why the burden of proof sits on the
in-language Malaysia evidence in §[-@sec-my-validation], where the documents are
far less likely to have been seen.

```{r}
#| label: pending-il
#| eval: !expr "!have_il"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `iteration_logs` target is unbuilt, so the H&K",
    "validation battery cannot render. Build it (API-free) and re-render.\n:::\n")
```

```{r}
#| label: tbl-hk-tests
#| tbl-cap: "**The behavioural battery.** What each test probes and what passing it licenses. Tests I–IV run before any evaluation data is touched; Tests V–VII run alongside the ablation and the manual error analysis."
tt_hk_tests_table()
```

## Measure identification {#sec-c1}

The measure-identification codebook decides whether a chunk substantively
discusses a specific fiscal measure, and names the measures it finds. The
instruction block is reproduced verbatim below; the full codebook, with all
inclusion and exclusion clarifications, is in @sec-appendix-codebooks.

```{r}
#| label: c1-instructions
#| echo: false
#| output: asis
codebook_instructions_md(here("prompts/c1_measure_id.yml"))
```

```{r}
#| label: tbl-c1-classes
#| tbl-cap: "**Measure-identification classes.** The two labels and their definitions, with the count of inclusion and exclusion clarifications the codebook attaches to each."
codebook_classes_table(here("prompts/c1_measure_id.yml")) |>
  tt(width = c(0.20, 0.60, 0.10, 0.10)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

**This codebook is tuned for recall, deliberately.** It sits at the front of the
pipeline, and a measure it misses is invisible to every stage after it, while a
measure it wrongly surfaces is cheap: the act-identification pass discards it,
and the motivation stage never sees it. So the operating point is set to
maximise the fiscal policy available downstream, and precision is treated as
recoverable rather than binding. The diagnostic benchmarks reflect that ordering
— Tier 1 recall ≥95%, combined recall ≥90%, precision ≥70% — and none of them is
a hard gate. The gate is the manual error analysis below.

### Behavioural tests

```{r}
#| label: fig-c1-s1
#| fig-cap: "**Measure identification, behavioural tests.** Tests I–III as percent correct; Test IV as Fleiss' kappa across label orderings. Test II is the memorisation probe."
#| fig-height: 5
#| eval: !expr have_il
plot_s1_behavioral(latest(il$s1, "c1"))
```

The battery passes cleanly, and that is the expected result rather than an
achievement: every frontier commercial model we tried clears these tests on this
codebook. Their purpose is not to discriminate between models but to prove the
harness is sound — that the schema, the parsing, and the token budget let the
model express what the codebook asks for. They earned that keep twice. An
apparent Test I failure turned out to be an output-token cap truncating long
measure lists rather than a codebook defect, and it was the test, not the
evaluation, that isolated it.

```{r}
#| label: fig-c1-s2
#| fig-cap: "**Measure identification, zero-shot evaluation.** Point estimates with 95% bootstrap intervals; crosses mark the diagnostic benchmark where one is defined."
#| fig-height: 3.6
#| eval: !expr have_il
p <- plot_s2_metrics(latest(il$s2, "c1"))
if (is.null(p)) cat("No S2 metrics available.\n") else p
```

```{r}
#| label: fig-c1-s3
#| fig-cap: "**Measure identification, codebook-adherence tests.** Tests V–VII against the full-codebook baseline. Test VI compares accuracy under informative and neutral label names; Test VII puts label names and definitions in conflict."
#| fig-height: 5
#| eval: !expr have_il
plot_s3_behavioral(latest(il$s3, "c1"), latest(il$s3_ablation, "c1"))
```

### Manual error analysis and the label-noise correction {#sec-c1-manual}

The gate is a human reading of every disagreement between model and gold label,
sorted into the six-category framework of @halterman2025: the model was right
(A), the gold label was wrong (B), the document was corrupted (C), the model
ignored the instructions (D), the model reasoned incorrectly (E), or something
else (F).

```{r}
#| label: tbl-c1-manual
#| tbl-cap: "**Manual error analysis.** Distribution over the six-category framework for the measure-identification codebook."
#| eval: !expr have_il
tt_manual_analysis_table(latest(il$s3_manual, "c1"))
```

The Category B count is the finding, not a footnote. Our US gold labels are not
hand-coded chunk by chunk; they are attached by matching Romer–Romer act names
against chunk text. Verbatim passage matches are reliable, but the second tier
matches act-name *fragments*, and act-name fragments appear in book
indexes, tables of contents, glossaries, and administrative appendices. Those
chunks arrive carrying a positive gold label and no fiscal content whatsoever.
When the model correctly declines to call a book index a fiscal measure, the
evaluation scores it a false negative.

**The label-generating process is noisier than the thing it is measuring**, and
that is a general hazard for LLM content analysis at this scale: automating gold
label construction to make evaluation feasible imports exactly the errors the
evaluation is meant to detect. The response is not to discard the noisy labels
but to quantify them. The manual analysis identifies which disagreements are
label noise, and the gate is recomputed with those chunks excluded from the
denominator, as @tbl-c1-confusion and @tbl-c1-gates report. Ambiguous cases are
kept in the denominator rather than excluded, so the correction is conservative.

```{r}
#| label: tbl-c1-confusion
#| tbl-cap: "**Measure identification, corrected confusion matrix.** Counts after excluding the chunks the manual analysis judged mislabelled in the gold standard."
#| eval: !expr have_il
tb <- tt_confusion_matrix(latest(il$s3_manual_bias_corrected, "c1"))
if (is.null(tb)) cat("No corrected confusion matrix available.\n") else tb
```

```{r}
#| label: tbl-c1-gates
#| tbl-cap: "**Measure identification, corrected performance.** Label-noise-adjusted rates against the diagnostic benchmarks. Recall is the binding quantity here; precision is recoverable downstream."
#| eval: !expr have_il
tt_gate_metrics(
  latest(il$s3_manual_bias_corrected, "c1"),
  targets = c(recall = 0.90, tier1_recall = 0.95, precision = 0.70)
)
```

## Motivation and exogeneity {#sec-c2}

### Why this is two codebooks {#sec-c2-split}

Assigning a Romer–Romer motivation label to a passage is two tasks wearing one
name. The first is evidentiary: find, in a document that discusses many things,
the sentences stating why *this* measure was enacted. The second is
interpretive: decide which of the four motivation categories that stated reason
belongs to, and whether it makes the act exogenous to current output.

Fusing them into one prompt makes failures uninterpretable — a wrong label could
be a reading failure or a reasoning failure, and the transcript cannot tell you
which. So the stage is split. An **extraction** codebook returns the motivation
evidence for a named act as verbatim quotes. A **classification** codebook sees
only those quotes and returns the label, the sign, and the exogeneity call. The
split buys three things: errors localise to a stage, the extracted evidence
becomes the audit trail a reviewer reads instead of the label, and the
classification step can be evaluated on clean evidence independently of
retrieval quality.

The split does not, however, let us validate the two halves separately and
declare the chain sound. What the pipeline deploys is the composition, so the
gate is applied to the composition: extraction runs on real chunks, and the
classifier is scored on whatever evidence extraction actually produced, errors
included. The extraction codebook carries its own behavioural and zero-shot
record; there is no separate error analysis for it, because the analysis that
matters is of the bundle. §[-@sec-defect] is what that choice bought — a defect
in the *interface* between the two stages, which neither stage's own metrics
would have surfaced.

```{r}
#| label: c2b-instructions
#| echo: false
#| output: asis
codebook_instructions_md(here("prompts/c2b_classification.yml"))
```

```{r}
#| label: tbl-c2-classes
#| tbl-cap: "**Motivation classes.** The Romer–Romer four-way taxonomy as implemented, with the count of inclusion and exclusion clarifications attached to each label. The first two are endogenous to current output; the last two are exogenous."
codebook_classes_table(here("prompts/c2b_classification.yml")) |>
  tt(width = c(0.22, 0.58, 0.10, 0.10)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

### Behavioural tests and zero-shot evaluation

```{r}
#| label: fig-c2-s1
#| fig-cap: "**Motivation classification, behavioural tests.** Tests I–III as percent correct; Test IV as Fleiss' kappa across label orderings. Test II is the memorisation probe for the classification codebook."
#| fig-height: 5
#| eval: !expr have_il
plot_s1_behavioral(latest(il$s1, "c2b"))
```

```{r}
#| label: fig-c2-s2
#| fig-cap: "**Motivation classification, zero-shot evaluation.** Point estimates with 95% bootstrap intervals; crosses mark the gate where one is defined. Exogenous precision and sign accuracy on true-exogenous acts are the two gates that matter downstream."
#| fig-height: 3.6
#| eval: !expr have_il
p <- plot_s2_metrics(latest(il$s2, "c2b"))
if (is.null(p)) cat("No S2 metrics available.\n") else p
```

```{r}
#| label: fig-c2-s3
#| fig-cap: "**Motivation classification, codebook-adherence tests.** Tests V–VII against the full-codebook baseline."
#| fig-height: 5
#| eval: !expr have_il
plot_s3_behavioral(latest(il$s3, "c2b"), latest(il$s3_ablation, "c2b"))
```

The adherence tests carry an uncomfortable result that we report rather than
bury. When label names and label definitions are put in conflict (Test VII), the
model follows the *names* almost always and the definitions almost never. Read
together with a small ablation penalty — stripping the codebook's definitions
and clarifications costs only a few points of accuracy, and even the fully
stripped condition retains most of the four-way task — the natural reading is
that the model arrives already knowing what "countercyclical" and "long-run"
mean in fiscal policy, from pretraining, and that our codebook is calibrating a
prior rather than teaching a construct.

That is good news for transfer and bad news for control. It is good news because
a prior learned from the general fiscal-policy literature is not tied to US
institutions, which is the property cross-country deployment needs. It is bad
news because a researcher who redefines a category away from its conventional
meaning should not assume the model will follow, and because agreement on US
data is weaker evidence than it appears. It is the sharpest reason the Malaysia
evidence, not the US benchmark, has to carry the argument.

### Manual error analysis and the label-noise correction

```{r}
#| label: tbl-c2-manual
#| tbl-cap: "**Manual error analysis.** Distribution over the six-category framework for the motivation codebooks, adjudicated on the composed extraction-and-classification chain."
#| eval: !expr have_il
tt_manual_analysis_table(latest(il$s3_manual, "c2b"))
```

The label noise here has a different source than in §[-@sec-c1-manual] and is
harder to correct. Motivation gold labels come from Romer and Romer's own
classifications, which are act-level judgements published with their reasoning.
Where a chunk discusses an act only glancingly, or where the act's motivation is
genuinely mixed, the act-level label is not a defensible chunk-level target —
and unlike an index page, there is no mechanical signature that marks such a
case. Every Category B act here is of that kind: the act carries two motivations
in its own stated reasons, and the gold label follows a tie-breaking convention
rather than the codebook. Adjudicating those is itself an interpretive act, so
the Category B count here is less mechanical and more contestable than the
measure-side one. We again exclude only what the manual pass judged mislabelled
and keep the ambiguous cases in the denominator.

```{r}
#| label: tbl-c2-confusion
#| tbl-cap: "**Motivation classification, corrected confusion matrix.** Exogenous-versus-endogenous counts after excluding the acts the manual analysis judged mislabelled or undecidable in the gold standard."
#| eval: !expr have_il
tb <- tt_confusion_matrix(latest(il$s3_manual_bias_corrected, "c2b"))
if (is.null(tb)) cat("No corrected confusion matrix available.\n") else tb
```

```{r}
#| label: tbl-c2-gates
#| tbl-cap: "**Motivation classification, corrected performance.** Label-noise-adjusted rates against the two gates that bind: exogenous precision, and sign accuracy on acts that are truly exogenous."
#| eval: !expr have_il
tt_gate_metrics(
  latest(il$s3_manual_bias_corrected, "c2b"),
  targets = c(precision = 0.85, sign_accuracy = 0.90),
  labels  = c(precision = "Exogenous precision", recall = "Exogenous recall")
)
```

Exogenous precision is the quantity a downstream user depends on: it is the
share of the acts we hand over as exogenous that really are. It lands just below
its gate, on an effective sample small enough that the interval comfortably
contains the target — a result we report as borderline rather than as a pass.
Sign accuracy on truly-exogenous acts clears its gate. These two corrected
numbers, and the correction procedure that produces them, are the central
methodological output of the US benchmark.

# Validation II — Malaysia {#sec-my-validation}

*Intent: the credibility pillar in the deployment country, where no ground truth
exists. Three legs: (1) an **EN/BM cross-language consistency test** — the pipeline
should reason equivalently regardless of input language
[@ahuja_mega_2023; @rathje_gpt_2024]; (2) **worked disagreement exhibits** —
Romer–Romer-style act write-ups chosen where the model's read and the preliminary
narrative read diverge, the cases an expert must settle; and (3) the
**ground-truth-free expert-agreement protocol**, framed as human-parity
(model-vs-consensus against leave-one-out human-vs-consensus) rather than agreement
with absolute truth [@asirvatham_gpt_2026]. Prose sources:
`notebooks/malay_consistency.qmd`, `docs/phase_1/expert_review_protocol.md`.*

## Cross-language (EN/BM) consistency

The test exploits an accident of the corpus: ten Economic Reports exist in both
English and Bahasa Malaysia. The same fiscal year, the same underlying acts, two
languages. A pipeline that reasons about fiscal substance should return
comparable inventories from both; a pipeline anchored on English surface
patterns should not. Because no ground truth exists on either side, the
comparison is distributional — measure counts, act counts, and the motivation
label mix — rather than a matched act-by-act agreement rate.

```{r}
#| label: pending-malay
#| eval: !expr "!have_malay"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe Malaysia EN/BM consistency targets are unbuilt, so",
    "the cross-language figures cannot render. See malay_consistency.qmd.\n:::\n")
```

```{r}
#| label: fig-malay-c1
#| fig-cap: "**Measure identification across languages.** Distinct fiscal measures surfaced per Economic Report year, by language. Comparable counts indicate the extraction step is language-robust."
#| fig-height: 4
#| eval: !expr "nrow(malay_er_c1_measures) > 0"
p <- plot_malay_c1_comparability(malay_er_c1_measures)
if (is.null(p)) cat("No C1 measure data available.\n") else p
```

```{r}
#| label: fig-malay-timeline
#| fig-cap: "**Act inventory across languages.** Acts by year and motivation, faceted by language. Comparable structure across English and Bahasa Malaysia is the signal that the codebooks reason consistently across languages."
#| fig-height: 5
#| eval: !expr have_malay
p <- plot_malay_act_timeline(malay_inv, timing = "act_name")
if (is.null(p)) cat("No act inventory available.\n") else p
```

## Worked acts (disagreement cases)

We reproduce three acts in full — one per component — each chosen where the
model's classification and the preliminary narrative read **diverge**, in the
style of @romer2010 (their Exhibits 1–2). These are the methodologically
consequential cases an expert must adjudicate.

```{r}
#| label: exhibit-cit
#| echo: false
#| output: asis
#| eval: !expr have_tax
tax_exhibit(
  tax_shocks, "MY-CIT-03",
  exhibit_label = "Exhibit 2",
  note = "The narrative pass read this rate cut as exogenous; the classifier read the surrounding crisis framing and called it countercyclical. Both readings are defensible from the corpus, which is what makes it an adjudication case."
)
```

```{r}
#| label: exhibit-spending
#| echo: false
#| output: asis
#| eval: !expr have_spending
spending_exhibit(
  spending_shocks, "MY-SPEND-04",
  exhibit_label = "Exhibit 3",
  note = "The narrative pass applied the two-condition spending screen and read the subsidy restructuring as endogenous; the classifier read it as deficit-driven, hence exogenous. The disagreement is about whether fiscal-pressure motives count as cyclical."
)
```

```{r}
#| label: exhibit-incentive
#| echo: false
#| output: asis
#| eval: !expr have_incentives
incentive_exhibit(
  incentive_shocks, "MY-INCENT-12",
  exhibit_label = "Exhibit 4",
  note = "An investment incentive announced inside a pandemic recovery package. The narrative pass weighted the long-run investment-promotion objective, the classifier weighted the crisis vehicle it arrived in."
)
```

## The expert-agreement protocol

*Intent: prose describing the ground-truth-free validation design — sampled acts
for expert coding, the ≥80% (measure identification) and ≥70% (motivation)
agreement targets read as parity with an individual expert rather than with
truth, and expert-vs-expert agreement reported as the ceiling. Source:
`docs/phase_1/expert_review_protocol.md`.*

One gap in the protocol as currently written should be closed before it is run,
because §[-@sec-usecase] depends on it. The protocol selects acts for review by
purposive triage — every act predicted exogenous in a recession year, every act
carrying one of the two confusable motivation labels, and a convenience
spot-check of the rest. That is the right design for *finding* errors and the
wrong design for *correcting* an estimate: debiasing a downstream regression
against a validation set requires that set to be a probability sample, with
known inclusion probabilities [@angelopoulos_prediction-powered_2023;
@egami_using_2023]. The two purposes need either two samples or one design that
serves both.

# What is validated, and what is not {#sec-coverage}

The pipeline's stages rest on very different evidence, and the difference is
easy to lose once the output is a single tidy table. @fig-pipeline-unvalidated
inverts @fig-pipeline-validated: it lights the stages the US benchmark does
*not* cover, and @tbl-coverage states what each of them rests on instead.

```{r}
#| label: fig-pipeline-unvalidated
#| fig-cap: "**What the US benchmark does not cover.** The complement of @fig-pipeline-validated. These stages are attested by construction, by audit, or not yet at all."
#| fig-width: 6.5
#| fig-height: 2.4
plot_pipeline_flow("unvalidated")
```

```{r}
#| label: tbl-coverage
#| tbl-cap: "**Validation coverage.** What each stage of the pipeline rests on, and what is still missing. Nothing in the deliverable's exogeneity column is an adjudicated finding."
tibble::tribble(
  ~Stage,                    ~`Evidence we have`,                                                         ~`Evidence we lack`,
  "Corpus construction",     "Extraction QA; page and chunk counts reproducible from the pipeline",        "No recall audit over the universe of documents that exist",
  "Measure identification",  "Full behavioural, zero-shot and error-analysis record on 44 US acts",        "Deployment-corpus performance; the top-ranked-measure filter discards 84.5% of mentions",
  "Act identification",      "Per-instrument recall scorecards, frozen; human stamp on every event",       "No gold standard, so no recall or precision rate — only an audit of what was searched",
  "Act consolidation",       "Automated clustering built and scored against 49 US reference acts",         "Rejected for deployment, not validated for it",
  "Motivation and sign",     "Full record on US acts; label-noise-corrected gates",                        "Validated on single-act chunks, deployed on multi-measure speeches",
  "Magnitude",               "Extracted with each act and traceable to a quoted rate or amount",           "No codebook, no behavioural tests, no external rate anchor",
  "Timing",                  "Announced and effective year, from the document",                            "No quarterly timing; the annual grain limits the downstream use case",
  "Exogeneity of the output", "Two independent preliminary reads per act, and their disagreements flagged", "Expert adjudication not yet run; the column is a model read"
) |>
  tt(width = c(0.20, 0.42, 0.38)) |>
  tt_theme_report() |>
  style_tt(j = 1:3, align = "l")
```

Three entries deserve emphasis because they are the ones most likely to be
over-read.

**The exogeneity column is an input, not a finding.** Every act carries two
preliminary reads — one from the narrative identification pass, one from the
validated classification codebook — and where they disagree the act is flagged
rather than resolved. Those disagreements are the worked exhibits in
§[-@sec-my-validation]. Adjudicating them is the expert's job and it has not
happened yet.

**Magnitude and timing have no validation record at all.** They are extracted
agentically alongside each act and are traceable to a quoted rate or amount, but
the codebooks that would validate them are not built. Any use of the magnitude
column inherits that.

**The deployment distribution was never validated.** This is the general form of
§[-@sec-defect] and the most transferable lesson here. The codebooks were
validated on US chunks that discuss one act each and deployed on Malaysian
budget speeches that discuss many. That gap produced a real defect that no US
test could have caught, and it is a gap any cross-country transfer of this kind
will have in some form.

# A first use case: procyclicality decomposition {#sec-usecase}

::: {.callout-note}
**Section stub.** The figure for this section is deferred: it needs an annual
Malaysia real-GDP-growth (business-cycle) series added via a `{targets}` target
plus `R/procyclicality.R`, a data decision not yet taken. The framing and
caveats below stand; the estimate does not yet.
:::

*Intent: a demonstration, not an inferential claim. Using the exogenous/endogenous
split against the annual business cycle, decompose how much of Malaysia's
tax-policy procyclicality was exogenous all along — the decomposition
@vegh_how_2015 name as the missing input and could not perform with a statutory-rate
series. The value added of the paper is the dataset and the method; this is one use
case.*

*Generated-regressor caveat (load-bearing here): once an estimate is built on the
labels, valid inference is a measurement-error problem, not a classification-accuracy
one [@ludwig_large_2025; @battaglia_inference_2024]. The bias equals the correlation
of label error with the regressor, and the dangerous regressor is the business cycle
— the exact axis exogeneity turns on. So the decomposition must either show
mislabelling is uncorrelated with the cycle or scope the claim accordingly; a proper
downstream estimate would debias against a probability-sampled expert validation
set [@angelopoulos_prediction-powered_2023; @egami_using_2023].*

<!-- FIGURE PLACEHOLDER — procyclicality decomposition.
     Needs: annual Malaysia real-GDP-growth series (WDI) as a target;
            R/procyclicality.R for the decomposition + figure;
            overlay signed exogenous vs endogenous fiscal actions on the cycle. -->

# Limitations and roadmap

@tbl-coverage states what is and is not evidenced. Three further limitations are
about the design rather than its current state.

**Transfer is a claim to defend, not to assume.** Fixed text measures notoriously
fail to transfer across domains [@gentzkow_text_2019], and cross-codebook
fine-tuning overfits [@halterman2025], so frozen country-agnostic codebooks have
to be shown to survive contact with a new country rather than assumed to. The
cross-language test and the regional extension are what put that claim at risk.
The finding in §[-@sec-c2] that the classifier appears to be running on a
pretrained prior cuts both ways here: it is why transfer is plausible, and it is
why demonstrating transfer on US-adjacent material proves little.

**One country is one observation.** Malaysia was chosen for a stable political
window, a bilingual archive, and a documented fiscal record. Indonesia, Thailand,
the Philippines, and Vietnam follow the same pipeline, and the parts of the
method that turn out to be Malaysia-specific will only be visible then.

**The human stamp is load-bearing and does not scale linearly.** Act
identification and consolidation require a researcher to confirm the events and
the grain, per instrument, per country. The pipeline reduces months of work to
weeks; it does not remove the expert, and the claim in this paper is explicitly
about assistance, not automation.

# Appendix: recall audits {#sec-appendix-recall}

The remaining per-instrument recall audits, frozen with their datasets.

```{r}
#| label: tbl-recall-pit
#| tbl-cap: "**Recall audit, personal income tax.**"
#| eval: !expr have_tax
recall_scorecard_table(tax_shocks, instrument = "Personal income tax")
```

```{r}
#| label: tbl-recall-consumption
#| tbl-cap: "**Recall audit, consumption tax (GST/SST).**"
#| eval: !expr have_tax
recall_scorecard_table(tax_shocks, instrument = "Consumption tax (GST/SST)")
```

```{r}
#| label: tbl-recall-spending
#| tbl-cap: "**Recall audit, spending.**"
#| eval: !expr have_spending
recall_scorecard_table(spending = spending_shocks, instrument = "Spending")
```

```{r}
#| label: tbl-recall-incentive
#| tbl-cap: "**Recall audit, tax incentives.**"
#| eval: !expr have_incentives
recall_scorecard_table(incentive = incentive_shocks, instrument = "Tax incentives")
```

# Appendix: codebooks {#sec-appendix-codebooks}

The codebooks are reproduced verbatim, as the pipeline reads them.

## Measure identification

```{r}
#| label: appendix-c1
#| echo: false
#| output: asis
codebook_verbatim_md(here("prompts/c1_measure_id.yml"))
```

## Motivation evidence extraction

```{r}
#| label: appendix-c2a
#| echo: false
#| output: asis
codebook_verbatim_md(here("prompts/c2a_extraction.yml"))
```

## Motivation classification

```{r}
#| label: appendix-c2b
#| echo: false
#| output: asis
codebook_verbatim_md(here("prompts/c2b_classification.yml"))
```

## References {.unnumbered}

::: {#refs}
:::
