---
title: "Scaling the Narrative Identification of Fiscal Changes using LLMs"
date: last-modified
date-format: long
abstract: |
  *Draft abstract (second-pass prose pending).* We build and validate an
  LLM-assisted pipeline that transfers the Romer & Romer (2010) narrative approach
  to fiscal-shock identification from the United States to emerging markets. Four
  country-agnostic codebooks are validated on 44 labelled US acts under the
  Halterman & Keith (2025) framework, then deployed to Malaysia (1990–2023), where
  an agentic component layer produces an expert-ready inventory of statutory tax,
  spending, and tax-incentive changes — each carrying a motivation and a
  preliminary exogeneity read. We demonstrate the dataset with a first
  procyclicality decomposition.
number-sections: true
section-numbering: "1.1.1"
format:
  html:
    toc: true
    comments:
      hypothesis: true
  docx: default
  jats: default
  academic-typst: default
execute:
  echo: false
  warning: false
  cache: true
  freeze: auto
---

```{r}
#| label: setup
#| cache: false
#| message: false
#| warning: false

pacman::p_load(targets, tidyverse, tinytable, here, scales, patchwork)

here::i_am("index.qmd")
tar_config_set(store = here("_targets"))

source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))     # pretty_motivation() + palette
source(here("R/tax_shock_report.R"))
source(here("R/spending_shock_report.R"))
source(here("R/incentive_shock_report.R"))
source(here("R/iteration_reporting.R"))   # H&K S1–S3 figure/table helpers
theme_set(theme_minimal())

# ---- Empty-safe reads (rendering is read-only: tar_read only, never tar_make) --
# Malaysia deliverables (API-gated enrichment tails — may be unbuilt)
tax_shocks       <- tryCatch(tar_read(tax_shocks),       error = \(e) tibble())
spending_shocks  <- tryCatch(tar_read(spending_shocks),  error = \(e) tibble())
incentive_shocks <- tryCatch(tar_read(incentive_shocks), error = \(e) tibble())
have_tax        <- nrow(tax_shocks) > 0
have_spending   <- nrow(spending_shocks) > 0
have_incentives <- nrow(incentive_shocks) > 0

# H&K validation logs (API-free; parsed from prompts/iterations/*.yml)
il      <- tryCatch(tar_read(iteration_logs), error = \(e) NULL)
have_il <- !is.null(il)

# Select the latest formal (Claude) iteration of a codebook within a parsed log.
latest <- function(tbl, cb) {
  tbl |>
    dplyr::filter(codebook == cb, grepl("claude", model)) |>
    dplyr::filter(iteration == max(iteration))
}

# Malaysia EN/BM consistency targets (precomputed; may be unbuilt)
malay_er_c1_measures <- tryCatch(tar_read(malay_er_c1_measures), error = \(e) tibble())
malay_metrics        <- tryCatch(tar_read(malay_er_consistency_metrics), error = \(e) NULL)
malay_inv            <- if (is.null(malay_metrics)) tibble() else malay_metrics$inventory
have_malay           <- nrow(malay_inv) > 0
```

::: {.callout-note}
**Draft status.** This is the *structure-and-figures* pass of the manuscript: the
seven sections, their figures, and one-paragraph statements of intent are in place;
full section prose is a deliberate second pass. Figures whose upstream targets are
API-gated and not yet built render as a "pending" note rather than an error.
:::

# Introduction

*Intent: the "why" and the "what is new." For most emerging markets we lack
consistent, comparable measures of exogenous fiscal shocks — the input credible
fiscal analysis needs [@vegh_how_2015; @romer2010]. LLMs make the narrative
approach scalable for the first time. Our contribution is a **method plus a
demonstration dataset**: country-agnostic codebooks, validated on the US and
deployed to Malaysia, that hand an expert an audit-ready inventory of fiscal-change
events. We position against the three LLM-narrative competitors
[@das_ai_2026; @bhasin_identifying_2026; @fritsch_high-frequency_nodate] on four
wedges they each cede — primary national documents, the tax side, full-richness
rows (measure × exo/endo × sign), and ground-truth-free validation. Prose sources:
`docs/brainstorm.qmd`, `docs/lit_review.qmd` §1/§4/§7.*

[Let's leave the Intro, Conclusion and Abstract to the end; they should write themselves with what we populate the body with.]{.aside}

# The narrative specification and the validation philosophy

*Intent: two ideas the rest of the paper depends on. (1) The **specification** —
Romer & Romer's narrative method reads documents to recover, per act, the fiscal
*measure*, its *motivation* (the 4-way taxonomy), the *exogenous* filter, the
*sign*, and — in full — *timing* and *magnitude* [@romer2010; @romer_presidential_2023].
(2) The **discipline** — Halterman & Keith's five-stage framework treats codebook
conceptualisation as a first-order concern and shows off-the-shelf LLMs follow real
codebooks imperfectly zero-shot, so every codebook must pass behavioural tests,
zero-shot evaluation, and error analysis before it is trusted [@halterman2025].
Prose sources: `docs/lit_review.qmd` §2/§5, `docs/strategy.md`,
`docs/codebook_sources.md`.*

# The pipeline and what it produces {#sec-pipeline}

*Intent: the architecture, "what is one row," and the reconciliation of the two
production paths — then the headline Malaysia figures so the reader sees the output
of the method. **Path A** is the H&K-validated codebook chain (C1 measure ID → C0
act aggregation → C2a/C2b motivation, sign, and exogeneity), validated on 44 US
Romer–Romer acts; this is where classification credibility lives. **Path B** is the
agentic `identify-*` component layer that froze the Malaysia rows by reading the
corpus directly per instrument, then reused the validated C2a/C2b for the
motivation/exogeneity enrichment. Framing: the validated codebooks* certify the
classification; *the agentic component skills do measure discovery and structured
extraction under a shared output contract
(`docs/phase_1/{tax,spending,incentive}_shock_schema.md`).*

## Pipeline stages

```{r}
#| label: tbl-pipeline
#| tbl-cap: "**Pipeline stages.** Path A is the H&K-validated codebook chain (validated on US ground truth); Path B is the agentic component layer that freezes the country deliverable and reuses Path A's C2a/C2b for enrichment."

tibble::tribble(
  ~Stage,                    ~Component,                         ~Path, ~Output,
  "Corpus (RR1)",            "Primary-source extraction + QA",   "—",   "Document corpus",
  "Measure ID (RR2)",        "C1 codebook (LLM)",                "A",   "Candidate fiscal measures",
  "Act aggregation (RR6)",   "C0 canonical clustering (LLM)",    "A",   "Canonical acts",
  "Motivation (RR5)",        "C2a evidence → C2b class (LLM)",   "A",   "4-way label + exo/endo + sign",
  "Component identification", "identify-{cit,pit,vat,spending,incentives}", "B", "Frozen per-instrument acts",
  "Bind + enrich",           "reuse C2a/C2b on frozen acts",     "A+B", "Deliverable inventory"
) |>
  tt() |>
  tt_theme_report()
```

[@tbl-pipeline is illegible, even to me. Separate into two tables, one for the RR procedure and other for the HK procedure. ]{.aside}

## What is one row

```{r}
#| label: tbl-one-row
#| tbl-cap: "**Unit of observation.** One row per narratively-announced act × instrument, carrying a magnitude (statutory rate for tax; ringgit for spending), a motivation label, and a preliminary exogeneity read pending expert adjudication."

tibble::tribble(
  ~Field,            ~Meaning,
  "act",             "One narratively-announced fiscal act (phased steps collapse to one row)",
  "instrument",      "CIT / PIT / CONSUMPTION (tax), Expenditure (spending), or incentive category",
  "year",            "Announced / effective year (quarterly timing is deferred to C3)",
  "magnitude",       "Δpp statutory rate (tax/rate-bearing incentive) or ringgit amount (spending)",
  "sign",            "Direction of the effect on fiscal liabilities (increase / decrease / no change)",
  "motivation",      "R&R 4-way label from the validated C2b codebook",
  "exogenous",       "Preliminary exogenous/endogenous read — an input for the expert, not a finding"
) |>
  tt() |>
  tt_theme_report()
```

[@tbl-one-row is lazy. We should show what the data is through either (i) beautiful plots, or (ii) beautiful tables that convey the meaning and magnitudes of the data, i.e. summary stats of some sorts. This table sucks because the reader has not seen the data and is being asked to hold in memory some concepts she doesn't care about. Though we actually need a table like this; keep it and see if you can add some other visual to explain what the data is.]{.aside}

## What the pipeline produces: the Malaysia inventory

The figures below are the headline output of the pipeline, for Malaysia
(1990–2023). Downloads, data dictionaries, and the full worked exhibits live on the
reviewer-facing dataset page, [malaysia_dataset.qmd](notebooks/malaysia_dataset.qmd).

```{r}
#| label: pending-tax
#| eval: !expr "!have_tax"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `tax_shocks` deliverable target is empty or",
    "unbuilt (its C2a/C2b enrichment is API-gated). The tax figures below render",
    "once it is built.\n:::\n")
```

### Statutory tax changes

```{r}
#| label: tbl-tax-inventory
#| tbl-cap: "**Statutory tax-change inventory.** One row per announced act × tax type; Δpp is the headline-rate change; motivation and exogeneity from the validated C2b codebook (v0.9.1)."
#| eval: !expr have_tax
tax_inventory_table(tax_shocks)
```

```{r}
#| label: fig-tax-rate-paths
#| fig-cap: "**Statutory rate paths by tax type.** Step path of the headline statutory rate implied by the rate-bearing acts."
#| eval: !expr have_tax
plot_tax_rate_paths(tax_shocks)
```

```{r}
#| label: fig-tax-timeline
#| fig-cap: "**Tax-change timeline.** Acts per year × tax type as a diverging stacked bar (hikes up, cuts down); colour denotes the C2b motivation."
#| eval: !expr have_tax
plot_tax_timeline(tax_shocks)
```

### Government spending shocks

```{r}
#| label: pending-spending
#| eval: !expr "!have_spending"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `spending_shocks` deliverable target is empty or",
    "unbuilt. The spending figures below render once it is built.\n:::\n")
```

```{r}
#| label: tbl-spending-inventory
#| tbl-cap: "**Government spending-shock inventory.** One row per announced spending act; exogeneity from the validated C2b codebook, alongside the @das_ai_2026 two-condition preliminary screen."
#| eval: !expr have_spending
spending_inventory_table(spending_shocks)
```

```{r}
#| label: fig-spending-timeline
#| fig-cap: "**Spending-shock timeline.** Acts per year as a diverging stacked bar (increases up, decreases down); colour denotes the C2b exogenous/endogenous read."
#| eval: !expr have_spending
plot_spending_timeline(spending_shocks)
```

### Tax incentives and holidays

```{r}
#| label: pending-incentives
#| eval: !expr "!have_incentives"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `incentive_shocks` deliverable target is empty or",
    "unbuilt. The incentive figures below render once it is built.\n:::\n")
```

```{r}
#| label: tbl-incentive-inventory
#| tbl-cap: "**Tax-incentive shock inventory.** One row per announced incentive change; exogeneity from the validated C2b codebook (v0.9.1)."
#| eval: !expr have_incentives
incentive_inventory_table(incentive_shocks)
```

```{r}
#| label: fig-incentive-timeline
#| fig-cap: "**Incentive-shock timeline.** Acts per year as a diverging stacked bar (new/expanded down, repeals up); colour denotes the C2b exogenous/endogenous read."
#| eval: !expr have_incentives
plot_incentive_timeline(incentive_shocks)
```

# Validation I — the US benchmark {#sec-us-validation}

*Intent: the codebook credibility pillar. We validate each codebook on the 44
labelled US Romer–Romer acts under H&K S1–S3 (behavioural tests → zero-shot
evaluation → error analysis). Because the published R&R shock list and the US
source documents are almost certainly in pretraining, strong US agreement is
**consistent with memorisation** [@carlini_quantifying_2022]: the S1 memorisation
test (Test II below) and the contamination argument put the burden of proof on the
in-language Malaysia results in @sec-my-validation, where the documents are far
less likely to have been seen. Prose sources: `notebooks/c1_measure_id.qmd`,
`notebooks/c0_aggregator.qmd`, C2 iteration logs.*

[Include here a beutiful visual, a flow chart, that explains how we took primary source government documents and came up with the dataset. Thjis figure should guide the reader thorugh this section.]{.aside}

```{r}
#| label: pending-il
#| eval: !expr "!have_il"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `iteration_logs` target is unbuilt, so the H&K",
    "validation battery cannot render. Build it (API-free) and re-render.\n:::\n")
```

[Include here a table describing HK's tests: what do they test for?]{.aside}

## C1 — Measure identification

[Include the verbatim codebook prompt.]{.aside}



```{r}
#| label: fig-c1-s1
#| fig-cap: "C1 v0.7.0 S1 behavioural tests. Tests I/II as percent correct (Test II is the memorisation/contamination probe); Test IV as Fleiss' kappa."
#| fig-height: 5
#| eval: !expr have_il
plot_s1_behavioral(latest(il$s1, "c1"))
```

[It should say here that all the frontier commercial models perform equally good at S1, and that we used the test to see that there were no problems with the code/instructions. I'm unsure why I cannot see a stat Test IV in @fig-c1-s1, please check it out.]{.aside}

[Include here an explanation for how we chose to evaluate C1: we wanted to maximize recall so that we can evaluate all of the fiscal policy that was available.]{.aside}

```{r}
#| label: tbl-c1-s2
#| tbl-cap: "C1 v0.7.0 S2 zero-shot metrics with 95% bootstrap CIs."
#| eval: !expr have_il
tt_s2_metrics_table(latest(il$s2, "c1"))
```

[@tbl-c1-s2 is super tough in the eye. Let's try it with a plot instead. Also, fyi, this caption is totally illegible.]{.aside}

```{r}
#| label: fig-c1-s3
#| fig-cap: "C1 v0.7.0 S3 behavioural tests (Tests V–VII) against the full-codebook baseline F1."
#| fig-height: 5
#| eval: !expr have_il
plot_s3_behavioral(latest(il$s3, "c1"), latest(il$s3_ablation, "c1"))
```

[@fig-c1-s3 's aesthetics are inconsistent with @fig-c1-s1 's: stick to the point and line aesthetics.]{.aside}


[Include here a description of how we carried through the manual analysis; how the IKA process was noisier than the codebook's performance]{.aside}

```{r}
#| label: tbl-c1-manual
#| tbl-cap: "C1 v0.7.0 manual error-analysis distribution (H&K six-category framework)."
#| eval: !expr have_il
tt_manual_analysis_table(latest(il$s3_manual, "c1"))
```



```{r}
#| label: tbl-c1-biascorrected
#| tbl-cap: "C1 v0.7.0 bias-corrected gate metrics (Category B, incorrect gold, excluded)."
#| eval: !expr have_il
tt_bias_corrected_table(latest(il$s3_manual_bias_corrected, "c1"))
```


[Drop mentions of codebook versions everywhere and especially in captions. This methodological paper should discuss the current state of the project; not its development process.]{.aside}

[@tbl-c1-biascorrected is important but aesthetically akward. Let's try moving the absolute figures to a confusion matrix, and keep only the performance stats in a separate table.]{.aside}

## C2 — Motivation and exogeneity

[Describe here how we had to separate C2 into the extraction and classification parts, and how we tested it as a boundle]{.aside}

[Also discuss the IKA process for C2, how it was imperfect. Include]{.aside}

```{r}
#| label: fig-c2-s1
#| fig-cap: "C2b S1 behavioural tests. Test II is the memorisation/contamination probe for the classification codebook."
#| fig-height: 5
#| eval: !expr have_il
plot_s1_behavioral(latest(il$s1, "c2b"))
```

```{r}
#| label: tbl-c2-s2
#| tbl-cap: "C2b S2 zero-shot metrics with 95% bootstrap CIs."
#| eval: !expr have_il
tt_s2_metrics_table(latest(il$s2, "c2b"))
```

```{r}
#| label: fig-c2-s3
#| fig-cap: "C2b S3 behavioural tests (Tests V–VII) against the full-codebook baseline."
#| fig-height: 5
#| eval: !expr have_il
plot_s3_behavioral(latest(il$s3, "c2b"), latest(il$s3_ablation, "c2b"))
```

```{r}
#| label: tbl-c2-manual
#| tbl-cap: "C2b manual error-analysis distribution (H&K six-category framework)."
#| eval: !expr have_il
tt_manual_analysis_table(latest(il$s3_manual, "c2b"))
```


[Most of my notes on the C1 sections are applicable here. Also, we're missing the bias-corrected performance table(s), which are the most important methodological things we produced in this project.]{.aside}

## C0 — Act aggregation

*Intent: C0 is a record-linkage step, scored not by H&K S0–S3 but by recovery of
the 49 Romer–Romer reference acts (coverage, spurious rate, fragmentation, year
alignment). The full five-method comparison lives in
[c0_aggregator.qmd](notebooks/c0_aggregator.qmd); the headline of the leading
method (M5, LLM canonical clustering) is reproduced below.*

[I think we should drop this section entirely, as the final datasets we built did not use C0 at all (please verify this claim). Rather discuss how we used the agentic search to locate relevant information from every act that was a distinct act.]{.aside}

```{r}
#| label: tbl-c0-m5
#| tbl-cap: "C0 Method 5 (LLM canonical clustering) RR-recovery metrics under the keyword name gate. Full method comparison in c0_aggregator.qmd."
#| echo: false
res <- tryCatch({
  suppressWarnings(suppressMessages(source(here("R/c0_aggregator.R"))))
  m5 <- tar_read(c0_m5_rr_metrics)
  format_rr_method_ladder(m5, gate = "keyword") |> tt() |> tt_theme_report()
}, error = function(e) NULL)
if (is.null(res)) {
  cat("::: {.callout-warning}\nC0 RR-metrics unavailable (target unbuilt or helper",
      "not loaded); see c0_aggregator.qmd.\n:::\n")
} else {
  res
}
```

# Validation II — Malaysia {#sec-my-validation}

*Intent: the credibility pillar in the deployment country, where no ground truth
exists. Three legs: (1) an **EN/BM cross-language consistency test** — the pipeline
should reason equivalently regardless of input language
[@ahuja_mega_2023; @rathje_gpt_2024]; (2) **worked disagreement exhibits** —
Romer–Romer-style act write-ups chosen where the model's read and the preliminary
narrative read diverge, the cases an expert must settle; and (3) the
**ground-truth-free expert-agreement protocol**, framed as human-parity
(model-vs-consensus against leave-one-out human-vs-consensus) rather than agreement
with absolute truth [@asirvatham_gpt_2026]. Prose sources:
`notebooks/malay_consistency.qmd`, `docs/phase_1/expert_review_protocol.md`.*

## Cross-language (EN/BM) consistency

```{r}
#| label: pending-malay
#| eval: !expr "!have_malay"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe Malaysia EN/BM consistency targets are unbuilt, so",
    "the cross-language figures cannot render. See malay_consistency.qmd.\n:::\n")
```

```{r}
#| label: fig-malay-c1
#| fig-cap: "Distinct C1 fiscal measures per Economic Report year, by language (EN vs BM). Comparable counts indicate the extraction step is language-robust."
#| fig-height: 4
#| eval: !expr "nrow(malay_er_c1_measures) > 0"
p <- plot_malay_c1_comparability(malay_er_c1_measures)
if (is.null(p)) cat("No C1 measure data available.\n") else p
```

```{r}
#| label: fig-malay-timeline
#| fig-cap: "Act inventory by year and motivation, faceted by language. Comparable structure across EN and BM is the signal that the codebooks reason consistently across languages."
#| fig-height: 5
#| eval: !expr have_malay
p <- plot_malay_act_timeline(malay_inv, timing = "act_name")
if (is.null(p)) cat("No act inventory available.\n") else p
```

## Worked acts (disagreement cases)

We reproduce four acts in full — one per component — each chosen where the model's
classification and the preliminary narrative read **diverge**, in the style of
@romer2010 (their Exhibits 1–2). These are the methodologically consequential cases
an expert must adjudicate.

```{r}
#| label: exhibit-cit
#| echo: false
#| output: asis
#| eval: !expr have_tax
tax_exhibit(tax_shocks, "MY-CIT-03", exhibit_label = "Exhibit 1")
```

```{r}
#| label: exhibit-pit
#| echo: false
#| output: asis
#| eval: !expr have_tax
tax_exhibit(tax_shocks, "MY-PIT-01", exhibit_label = "Exhibit 2")
```

```{r}
#| label: exhibit-spending
#| echo: false
#| output: asis
#| eval: !expr have_spending
spending_exhibit(spending_shocks, "MY-SPEND-04", exhibit_label = "Exhibit 3")
```

```{r}
#| label: exhibit-incentive
#| echo: false
#| output: asis
#| eval: !expr have_incentives
incentive_exhibit(incentive_shocks, "MY-INCENT-01", exhibit_label = "Exhibit 4")
```

## The expert-agreement protocol

*Intent: prose describing the ground-truth-free validation design — probability-
sampled acts for expert coding, the ≥80% (C1) / ≥70% (C2) agreement targets read as
parity with an individual expert (not with truth), and expert-vs-expert agreement
reported as the ceiling. Source: `docs/phase_1/expert_review_protocol.md`.*

# A first use case: procyclicality decomposition {#sec-usecase}

::: {.callout-note}
**Section stub.** The figure for this section is deferred: it needs an annual
Malaysia real-GDP-growth (business-cycle) series added via a `{targets}` target +
`R/procyclicality.R`, a data decision not yet taken. The framing and caveats below
stand; the estimate does not yet.
:::

*Intent: a demonstration, not an inferential claim. Using the exogenous/endogenous
split against the annual business cycle, decompose how much of Malaysia's
tax-policy procyclicality was exogenous all along — the decomposition
@vegh_how_2015 name as the missing input and could not perform with a statutory-rate
series. The value added of the paper is the dataset and the method; this is one use
case.*

*Generated-regressor caveat (load-bearing here): once an estimate is built on the
labels, valid inference is a measurement-error problem, not a classification-accuracy
one [@ludwig_large_2025; @battaglia_inference_2024]. The bias equals the correlation
of label error with the regressor, and the dangerous regressor is the business cycle
— the exact axis exogeneity turns on. So the decomposition must either show
mislabelling is uncorrelated with the cycle or scope the claim accordingly; a proper
downstream estimate would debias against a probability-sampled expert validation
set [@angelopoulos_prediction-powered_2023; @egami_using_2023].*

<!-- FIGURE PLACEHOLDER — procyclicality decomposition.
     Needs: annual Malaysia real-GDP-growth series (WDI) as a target;
            R/procyclicality.R for the decomposition + figure;
            overlay signed exogenous vs endogenous fiscal actions on the cycle. -->

# Limitations and roadmap

*Intent: honest scoping. (i) **Timing and magnitude are not yet built** — C3
(quarterly timing) and C4 (%GDP magnitude) are planned, so the current rows carry
announced year and statutory Δpp / ringgit magnitude, not quarterly %GDP; the
procyclicality use case is therefore annual. (ii) **Expert validation is pending** —
the exogeneity column is a preliminary model read, not a finding; the Phase 2 expert
pass is the intended next input. (iii) **Transfer is a claim to defend, not assume**
— fixed text measures notoriously fail to transfer across domains
[@gentzkow_text_2019] and cross-codebook fine-tuning overfits [@halterman2025], so
the frozen country-agnostic codebooks must be shown to survive, which the EN/BM and
regional legs test. (iv) **Regional scaling** — Indonesia, Thailand, the
Philippines, and Vietnam follow the same pipeline. Prose sources: `CLAUDE.md`
status, `docs/lit_review.qmd` §6.*

## References {.unnumbered}

::: {#refs}
:::
