---
format:
  academic-typst: 
    brand: false
    # citeproc: true
    csl: chicago-author-date
    header: Please click [HERE](https://www.estebandegetau.com/Fiscal-shocks/) for the latest version.
    number-sections: true
    section-numbering: "1.1.1"
    number-depth: 3
    colorlinks: true
  jats: default
execute:
  echo: false
  warning: false
  cache: true
  freeze: auto
---

```{r}
#| label: setup
#| cache: false
#| message: false
#| warning: false

pacman::p_load(targets, tidyverse, tinytable, here, scales, patchwork)

here::i_am("index.qmd")
tar_config_set(store = here("_targets"))

source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))     # pretty_motivation() + palette
source(here("R/tax_shock_report.R"))
source(here("R/spending_shock_report.R"))
source(here("R/incentive_shock_report.R"))
source(here("R/iteration_reporting.R"))   # H&K S1-S3 figure/table helpers
source(here("R/pipeline_diagram.R"))      # plot_pipeline_flow()
source(here("R/dataset_summary.R"))       # cross-component summaries
source(here("R/codebook_display.R"))      # verbatim codebook exhibits
source(here("R/recall_scorecard.R"))      # frozen recall audits
source(here("R/procyclicality.R"))        # cycle + procyclicality decomposition
theme_set(theme_minimal())

# ---- Empty-safe reads (rendering is read-only: tar_read only, never tar_make) --
# Malaysia deliverables (API-gated enrichment tails — may be unbuilt)
tax_shocks       <- tryCatch(tar_read(tax_shocks),       error = \(e) tibble())
spending_shocks  <- tryCatch(tar_read(spending_shocks),  error = \(e) tibble())
incentive_shocks <- tryCatch(tar_read(incentive_shocks), error = \(e) tibble())
have_tax        <- nrow(tax_shocks) > 0
have_spending   <- nrow(spending_shocks) > 0
have_incentives <- nrow(incentive_shocks) > 0
have_any        <- have_tax || have_spending || have_incentives

# Deployment corpus (API-free)
country_body   <- tryCatch(bind_rows(tar_read(country_body)),   error = \(e) tibble())
country_chunks <- tryCatch(bind_rows(tar_read(country_chunks)), error = \(e) tibble())
have_corpus    <- nrow(country_body) > 0

# Corpus headline figures, so the prose can cite them inline without repeating
# the summary call. The last row of corpus_scope_stats() is the "All series"
# total; `Pending` counts catalogued-but-unacquired documents (zero pages).
corpus_stats  <- if (have_corpus) corpus_scope_stats(country_body) else tibble()
corpus_total  <- if (nrow(corpus_stats) > 0) dplyr::slice_tail(corpus_stats, n = 1) else NULL
n_docs_read   <- if (is.null(corpus_total)) NA_integer_ else corpus_total$Documents
n_docs_catlg  <- if (is.null(corpus_total)) NA_integer_ else corpus_total$Documents + corpus_total$Pending
n_pages_read  <- if (is.null(corpus_total)) NA_integer_ else corpus_total$Pages

# H&K validation logs (API-free; parsed from prompts/iterations/*.yml)
il      <- tryCatch(tar_read(iteration_logs), error = \(e) NULL)
have_il <- !is.null(il)

# Select the latest formal (Claude) iteration of a codebook within a parsed log.
latest <- function(tbl, cb) {
  tbl |>
    dplyr::filter(codebook == cb, grepl("claude", model)) |>
    dplyr::filter(iteration == max(iteration))
}

# Business cycle, decomposition and the external rate benchmark (API-free)
my_cycle       <- tryCatch(tar_read(my_cycle),                     error = \(e) tibble())
procyc_events  <- tryCatch(tar_read(procyclicality_events),        error = \(e) tibble())
procyc_decomp  <- tryCatch(tar_read(procyclicality_decomposition), error = \(e) NULL)
vv_cross       <- tryCatch(tar_read(cross_country_cyclicality),    error = \(e) tibble())
rate_concord   <- tryCatch(tar_read(rate_path_concordance),        error = \(e) NULL)
have_cycle     <- nrow(my_cycle) > 0
have_procyc    <- have_cycle && nrow(procyc_events) > 0 && !is.null(procyc_decomp)
have_concord   <- !is.null(rate_concord) && nrow(rate_concord$events) > 0

# Headline decomposition figures, so the prose can cite them inline rather than
# restating what the exhibit already computed. `pc_*` are all NA when unbuilt.
pc_tab   <- if (have_procyc) procyc_decomp$table else NULL
pc_proc  <- if (is.null(pc_tab)) NA_integer_ else sum(pc_tab["Procyclical", ])
pc_exo   <- if (is.null(pc_tab)) NA_integer_ else pc_tab["Procyclical", "Exogenous"]
pc_n     <- if (is.null(pc_tab)) NA_integer_ else sum(pc_tab)
pc_p     <- if (is.null(pc_tab)) NA_real_    else procyc_decomp$fisher$p_value
pc_frag  <- if (is.null(pc_tab)) NA_integer_ else procyc_decomp$fragility
pc_flag  <- if (is.null(pc_tab)) NA_integer_ else procyc_decomp$disagree_n
# The excluding-spending robustness row, cited inline where the caveat is made.
pc_nospend <- if (is.null(pc_tab)) NULL else
  dplyr::filter(procyc_decomp$robustness, spec == "Excluding spending")

# Malaysia EN/BM consistency targets (precomputed; may be unbuilt)
malay_er_c1_measures <- tryCatch(tar_read(malay_er_c1_measures), error = \(e) tibble())
malay_metrics        <- tryCatch(tar_read(malay_er_consistency_metrics), error = \(e) NULL)
malay_inv            <- if (is.null(malay_metrics)) tibble() else malay_metrics$inventory
have_malay           <- nrow(malay_inv) > 0
```

::: {.callout-note}
**Draft status.** The body, the introduction and the exhibits are in place,
including the procyclicality use case in @sec-usecase. Figures whose upstream
targets are API-gated and not yet built render as a "pending" note rather than
an error.

The Malaysia figures below are built from the deliverable as frozen in June
2026. An evidence-scoping defect found in August 2026 (@sec-defect) has been
fixed in the pipeline but the enrichment tails have not yet been re-run, so four
of the 51 motivation labels are known to be wrong and are corrected in the
narrative, not yet in the exhibits. This now has a quantitative consequence
rather than only a cosmetic one: @sec-usecase-belief reports that four
relabellings would overturn the decomposition, so the exhibits there should be
read as provisional until the tails are re-run.
:::

# Introduction

Credible fiscal policy analysis needs to distinguish which changes in taxes and
spending were responses to the economy and which were not.  @romer2010 read decades of
presidential messages, budget documents and congressional records to recover,
act by act, *why* each tax change was legislated, and kept only the acts whose
stated motivation was unrelated to current output--calling them exogenous fiscal shocks. The resulting series has been used to show that personal
and corporate income tax changes have sharply different macroeconomic effects
[@mertens2013], to discipline the anticipation problem that makes fiscal shocks
hard even to date [@ramey_identifying_2011], and to reconcile narrative and
time-series estimates of the tax multiplier that had disagreed for a decade
[@mertens_reconciliation_2014]. The method itself has been shown to travel, at
least to another rich country with a comparable archive
[@cloyne_discretionary_2013]. 

Fiscal policy in developing countries is procyclical: it amplifies booms and
busts rather than smoothing them. The stylized fact was named by
@kaminsky_when_2004, and @vegh_how_2015 sharpen it in the way that matters here,
showing that procyclicality holds for statutory tax *policy* and not merely for
realized tax revenue, which moves with the cycle mechanically whatever
policymakers do. They are also explicit about what their evidence cannot settle.
The correlation they document could reflect policy responding to the cycle or the
cycle responding to policy, and separating the two requires knowing why each
measure was enacted. They name the @romer2010 narrative classification as the
missing input. That is the gap this paper attacks.

What hangs on it is the credibility of essentially every fiscal policy
recommendation made outside the OECD. Fiscal multipliers are not a constant to be
borrowed from the United States: they vary with the exchange-rate regime, with
openness, and with the level of debt [@ilzetzki_how_2013], and whether a country
escapes procyclicality at all appears to turn on the quality of its institutions
rather than on its income [@frankel_graduation_2013]. For developing Asia
specifically, @jha_effectiveness_2014 find that tax cuts do more countercyclical
work than spending increases, while conceding that roughly 70% of the revenue
variation in their data is non-discretionary. 

**The method.** Narrative identification reads the documentary record of a policy
decision to recover the decision's stated motive, and then uses the motive to
decide whether the resulting change can be treated as a shock--calling it exogenous. @romer2010 and
their companion methodological paper [@romer] apply this to every significant
federal tax action in the United States between 1945 and 2007, sorting each into
one of four motivation categories and keeping only those whose stated purpose was
unrelated to current output. @romer_presidential_2023 later set out the
methodological doctrine explicitly, and @sec-frameworks takes those requirements
as the specification this pipeline has to satisfy. Two properties make the
approach attractive to automate using Large Language Models (LLMs): the evidence for every judgement is a passage
another researcher can re-read, and the judgement itself is a reading task rather
than an estimation task.

**Who has used it.** The approach has been replicated and extended into a small
industry. @cloyne_discretionary_2013 built the United Kingdom equivalent from
Budget documents and found magnitudes remarkably close to the American ones,
which is the single most important precedent for the transfer claim in this
paper. @mertens2013 exploited the act-level detail to separate personal from
corporate income tax changes. A parallel action-based tradition grew up at the IMF, where
@pescatori_new_2011 coded fiscal consolidation episodes from national policy
documents rather than from budget outturns, and @guajardo_expansionary_2014 used
that dataset to show the identification choice is not innocuous: action-based
measures overturn the "expansionary austerity" reading that conventional
cyclically-adjusted measures produce. @dabla-norris_macroeconomic_2023 extend the
same logic to more than 2,000 consolidation-era tax measures, coding each one's
yield, motivation, and date, which is close to the record this paper's pipeline
is built to produce.

**Where it has been attacked.** As one might expect from a subjective classification, the narrative approach is contested, and the
contest bears directly on how the results below should be read.
@mertens_reconciliation_2014 show that structural VARs cannot recover the tax
multiplier without assuming the output-elasticity of revenue, that narrative
series used as external instruments deliver substantially larger multipliers,
and, most relevant here, that measurement error in the narrative series is itself
consequential rather than a rounding concern. @hebous_revisiting_2018 press
harder: the widely-used narrative tax measures correlate only weakly with
cyclically-adjusted revenues in both the United States and the United Kingdom,
and under weak-instrument-robust inference the implied multipliers are frequently
insignificant, so previous literature may understate how uncertain narrative-based
estimates are.


The benchmark this paper scores itself against is a hand-built series with acknowledged measurement
error, produced by two economists reading documents and exercising judgement. The
useful question is therefore not whether an LLM reproduces it exactly, which
would be a suspicious result, but whether the disagreements are of the same
character as the disagreements two careful human coders would have.
@sec-us-validation is organised around that question.

Large language models make the reading step cheap, and a small but fast-moving
literature has already used them for narrative fiscal identification. Its three
leading entries share a design choice.

Each reads a **secondary, harmonised, English-language source**, precisely to
avoid the cost of primary national documents. @das_ai_2026 code Economist
Intelligence Unit country reports for 64 countries, recovering the direction of
spending actions but explicitly not their magnitude.
@bhasin_identifying_2026 read IMF Article IV reports for 17 OECD countries,
restricted to austerity episodes. @fritsch_high-frequency_nodate read news wires
at daily frequency and do recover magnitudes in percent of GDP, for the United
States alone, and for an expectations series rather than a realised one. All
three validate the same way, by concordance against a pre-existing narrative or
action-based series, which is a strategy that only works where such a series
already exists and therefore cannot travel to the countries the method is meant
to serve. @tbl-related-work sets the three designs against our own, side by side.

```{r}
#| label: tbl-related-work
#| tbl-cap: "**The three closest papers, and this one.** LLM-based narrative fiscal identification designs compared on the dimensions this paper inverts [@das_ai_2026; @bhasin_identifying_2026; @fritsch_high-frequency_nodate]."
tibble::tribble(
  ~Dimension,     ~`Das et al.`,               ~`Bhasin et al.`,              ~`Fritsch et al.`,            ~`This paper`,
  "Source",       "EIU country reports",       "IMF Article IV reports",      "News wires",                 "National budget speeches, economic reports, plans",
  "Language",     "English",                   "English",                     "English",                    "English and Bahasa Malaysia",
  "Countries",    "64",                        "17, all OECD",                "1, the United States",       "1 pilot, 4 planned",
  "Instruments",  "Spending",                  "Austerity episodes",          "Tax and spending",           "Statutory tax, spending, incentives",
  "Output unit",  "Direction of action",       "Episode",                     "Daily surprise",             "One row per act and instrument",
  "Magnitude",    "Not recovered",             "Percent of GDP",              "Percent of GDP",             "Extracted, not validated",
  "Validation",   "Concordance with benchmark", "Concordance with benchmark", "Correlation with benchmark",  "US benchmark, then two no-ground-truth legs"
) |>
  tt(width = c(0.15, 0.20, 0.20, 0.20, 0.25)) |>
  tt_theme_report() |>
  style_tt(j = 1:5, align = "l")
```

The alternative to concordance validation is visible just outside fiscal policy.
@aruoba_identifying_nodate identify monetary policy shocks from Federal Reserve
narrative records and validate not against an existing shock list but against
forecast-error predictability and theory-consistent impulse responses. That is
the right instinct for a country with no labelled data, and @sec-my-validation
explains why we cannot yet follow it.

The prior question, whether an LLM can stand in for a human coder at all, has
been answered fairly encouragingly. @gilardi_chatgpt_2023 find that zero-shot
models beat crowd workers on text annotation by a wide margin and with higher
intercoder agreement, and @tornberg_large_2025 find they beat expert coders and
supervised classifiers at multi-country political labelling. @ziems_can_2024
bound the claim, documenting taxonomic tasks where fine-tuned models still win
and concluding that LLMs augment the research pipeline rather than replace it.
That is the position this paper occupies. It is also a position @sec-c2 gives a
mechanism for: the models arrive already knowing the conventional constructs,
which is a strength on conventional tasks and a liability on bespoke ones.

Two further literatures constrain the design and are introduced where they bind
rather than here. @halterman2025 supply the validation framework the paper adopts
wholesale (@sec-frameworks). @gentzkow_text_2019 supply the standing objection
that fixed text measures notoriously fail to transfer across domains, which is
precisely the property our frozen codebooks claim and @sec-my-validation has to
defend.


We take the opposite side of each trade in @tbl-related-work. This paper builds
and validates a pipeline that reads **primary national documents in the local
language**, covers the **tax side** as well as spending, returns a **full row
per act** (the measure, its direction, its motivation, and a preliminary
exogeneity read, each traceable to a quoted passage), and is validated
**without assuming ground truth exists**. It is applied end to end to Malaysia,
1980–2023, producing an inventory of 51 announced fiscal events across
statutory tax, spending, and tax incentives.

The contribution is a method and a demonstration dataset, not a claim of
automation. The pipeline hands a domain expert an audit-ready inventory; the
expert still decides what is exogenous. Where our design differs most from a
single end-to-end prompt is that it separates each task to provide the most rigurous validation available.


The project this paper reports on is organised around four questions. This
paper answers the first two in part and delivers the fourth for one country.

:::{#cnj-1}

Can an LLM reproduce the narrative identification of fiscal changes for
the United States, given the same documents and judged against
Romer and Romer's own labels? 

:::

We answer this for three of the five
quantities the narrative method recovers: which passages describe a fiscal
measure, the motivation behind each act and whether it is exogenous, and the
direction of the change. Magnitude in percent of GDP and quarterly timing
are not yet built.

:::{#cnj-2}

Do the same codebooks, frozen, transfer to a country with no labelled
data?

:::

Malaysia is the test. Because no motivation ground truth exists there, the
question is answered by consistency evidence, by an external check on the
statutory rate paths, and by expert agreement rather than by an accuracy rate.

:::{#cnj-3}

Can the resulting series support comparable cross-country fiscal
multipliers?
   
:::

Out of scope here: it needs magnitude and quarterly timing,
and more than one country.

:::{#cnj-4}

Can the pipeline produce a fiscal-change dataset local researchers can
actually use?
   
:::

The Malaysia inventory is the demonstration, and its
provenance chain is the reason it is usable.

[Include a brief (1-2 paragraphs) summary of conclusions]{.aside}

The remainder of the paper is organized as follows. @sec-frameworks sets out the
two frameworks the pipeline implements: the narrative specification we are
automating, and the validation discipline we hold it to. @sec-pipeline describes
the pipeline and the dataset it produces. @sec-build reconstructs how the
Malaysia dataset was built, including a defect the provenance chain caught after
deployment. @sec-validation reports the two validation legs, the US benchmark
and Malaysia. @sec-coverage states plainly which stages carry evidence and which
do not. @sec-usecase reports the procyclicality decomposition the dataset was
built to make possible.

# The narrative specification and the validation philosophy {#sec-frameworks}

Two ideas carry the rest of the paper. The first is the **specification**: what
the narrative method actually asks a reader to extract from a document. The
second is the **discipline**: what has to be true before an LLM's output of that
extraction can be treated as a measurement. They come from different literatures
and they constrain each other, so it is worth stating both before any machinery
appears.

## The specification: what narrative identification asks for

This is the
specification it hands a coder. @romer_presidential_2023 set out four
requirements for narrative analysis to count as evidence rather than anecdote: a
reliable source, a clear idea of the information being sought, a dispassionate
and consistent approach to the source, and careful documentation of the
narrative evidence behind each judgement. Those
four requirements are what this pipeline has to satisfy, and they map onto it
directly. The reliable source is a corpus of primary government documents. The
clear idea of the information sought is a written codebook. The dispassionate
and consistent approach is a frozen prompt read at temperature zero, which is
consistency of a stronger kind than a human coder can offer. The documentation
requirement is met by carrying the verbatim evidence forward with every label,
so that each row of the output points back at the sentence that produced it.

Applied to a fiscal act, the method recovers five things: the **measure** (what
was changed), the **motivation** (why), the **sign** (whether liabilities rose
or fell), the **timing** (when liabilities actually changed, which is not when
the law passed), and the **magnitude** (how large the change was). This paper
builds the first three and defers the last two, a boundary @sec-coverage returns
to.

Motivation is where the identification lives. @romer2010 sort each act into one
of four categories according to the reason its own proponents gave for it. Two
are responses to the economy: **spending-driven** changes, which finance a
decision already taken on the expenditure side, and **countercyclical** changes,
which aim to return output growth to normal. Two are not: **deficit-driven**
changes, which address a budget deficit inherited from past decisions, and
**long-run** changes, which pursue growth, fairness, efficiency, or a different
size of government. Only the second pair is exogenous to current output, and
only that pair survives into the shock series.

The two boundaries that do the most work are the ones a coder gets wrong.
Between countercyclical and long-run, the test is directional: an act meant to
return growth to *normal* is countercyclical, while an act meant to raise growth
*above* normal is long-run, and an act proposed in normal times for long-run
reasons stays long-run even if the economy has weakened by the time it passes.
Between spending-driven and deficit-driven, the test is temporal: a tax increase
paying for a spending increase enacted within the past year is spending-driven,
and the same increase more than a year later is deficit-driven. @das_ai_2026
add a clarifier we adopt verbatim, because it corrects a failure mode LLMs fall
into readily: acknowledging current conditions does not by itself make an act
endogenous if the stated motive is explicitly non-cyclical.

## The discipline: what makes an LLM output a measurement

The obvious way to scale this is to describe the task to a capable model and
read off the answers. @halterman2025 show why that is not enough. Their premise
is that codebook conceptualisation remains a first-order concern even when the
coder is a machine, and their finding is that off-the-shelf models follow real
codebooks only imperfectly zero-shot: they attend to label *names* more than to
the definitions a researcher wrote under them, they lean on lexical overlap, and
they import background concepts a codebook explicitly excludes. An accuracy
number computed on top of that behaviour does not establish that the model
measured the construct the codebook defines. It may have measured something
adjacent that happens to correlate.

Their response is a staged framework, and we adopt it as the paper's validation
spine: write the codebook so both humans and models can read it, run behavioural
tests that probe whether the model can follow it at all, then evaluate zero-shot
against labelled data, then analyse the errors, and only then consider changing
the model itself. Each codebook here goes through that sequence, and none of
them is trusted before it does.

This project needs the discipline more than most, for a specific reason. The
published Romer and Romer act list, and the US documents it was built from, are
almost certainly in the pretraining data of any frontier model. Strong agreement
on US data is therefore consistent with the model recognising an act by name
rather than reasoning from the passage in front of it, and extractable
memorisation is known to rise with model size, with duplication in the training
corpus, and with the length of the prompt context [@carlini_quantifying_2022].
This is why the behavioural battery includes an explicit memorisation probe, and
it is the deeper reason the deployment country carries as much of the argument
as the benchmark does: Malaysian budget speeches in Bahasa Malaysia are the part
of the evidence a model is least likely to have already seen.

# The pipeline and what it produces {#sec-pipeline}

The method reads primary-source government documents and returns an inventory of
fiscal-change events, each carrying a motivation, a sign, and a preliminary
exogeneity read. @fig-pipeline represents the route we adopted to go from one to
the other, and it is the map for the rest of the paper: @sec-build walks the
route for Malaysia, @sec-us-validation reports what the US benchmark certifies
about the blue stages, and @sec-coverage states what the other stages rest on
instead.

```{r}
#| label: fig-pipeline
#| fig-cap: "**From primary sources to an event inventory.** Two LLM codebook stages (blue) are validated against US ground truth; the act-identification step between them (orange) is an agentic narrative search whose output a human stamps before it enters the chain."
#| fig-width: 6.5
#| fig-height: 4.2
plot_pipeline_flow("none", scale = 1.4)
```

Two design commitments distinguish this route from a single end-to-end prompt.
The first is that **classification and discovery are separated**. Deciding
whether a passage describes a fiscal measure, and why the measure was enacted,
is a coding task with a codebook, a gold standard, and a validation record.
Deciding which acts exist in a corpus and where each begins and ends is a
retrieval-and-consolidation task with no gold standard, and it is handled
separately and audited differently. The second is that **every stage is
reversible to its source**: each event carries the documents, chunks, and
verbatim quotes it was built from, so a reviewer disagreeing with a label can go
to the passage rather than to the model.

## The two frameworks the pipeline implements

@sec-frameworks stated the specification and the discipline as ideas; the two
tables here state them as work. @tbl-rr-steps maps the specification: what
@romer2010 do by hand, and which component does it here. One row carries no step
number, because the hand method never has to name it. A human reading a budget
speech does not need a rule for deciding that two mentions of a rate cut are the
same act; a pipeline does, and @sec-no-c0 is about how badly that goes when it
is automated. @tbl-hk-stages maps the discipline: the stages a codebook must
clear before its output is trusted.

```{r}
#| label: tbl-rr-steps
#| tbl-cap: "**The narrative specification.** Romer and Romer's -@romer2010 procedure and the component that carries each step here. Step RR5 is where classification credibility lives; act consolidation, which their hand method never has to name as a separate step, is deliberately not automated."
tibble::tribble(
  ~Step,  ~`Romer & Romer's step`,                          ~`Our component`,                     ~Status,
  "RR1",  "Compile primary-source policy documents",        "Corpus acquisition and chunking",    "Built",
  "RR2",  "Identify significant legislated fiscal changes", "C1 measure identification (LLM)",    "Validated on US",
  "—",    "Consolidate mentions into distinct acts",        "Agentic narrative identification",   "Human-stamped",
  "RR3",  "Record the sign and size of each change",        "Sign from C2b; size extracted with the act", "Sign validated; size not",
  "RR4",  "Date the change to the quarter of effect",       "Announced and effective year",       "Annual only; C3 not built",
  "RR5",  "Classify the motivation, filter to exogenous",   "C2a evidence, C2b classification (LLM)", "Validated on US",
  "RR6",  "Aggregate exogenous acts into a shock series",   "Signed annual event inventory",      "Built"
) |>
  tt(width = c(0.07, 0.31, 0.34, 0.28)) |>
  tt_theme_report() |>
  style_tt(j = 2:4, align = "l")
```

```{r}
#| label: tbl-hk-stages
#| tbl-cap: "**The validation discipline.** Halterman and Keith's -@halterman2025 stages and how far each codebook has been taken. S4 is a last resort we have not needed."
tibble::tribble(
  ~Stage, ~`What it establishes`,                                          ~C1,        ~C2,
  "S0",   "Machine-readable codebook, expert-approved",                    "Complete", "Complete",
  "S1",   "Behavioural tests: legal outputs, memorisation, order",         "Passed",   "Passed",
  "S2",   "Zero-shot evaluation against labelled data",                    "Passed",   "Passed",
  "S3",   "Error analysis, ablation, manual adjudication",                 "Passed",   "Passed",
  "S4",   "Fine-tuning, if S3 exhausts codebook improvements",             "Not needed", "Not needed"
) |>
  tt(width = c(0.09, 0.51, 0.20, 0.20)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

The last row of @tbl-hk-stages deserves more than a "not needed". Fine-tuning is
the obvious lever when a zero-shot codebook underperforms, and we ruled it out
before running anything, for reasons that are about the research design rather
than about the results.

The first is that fine-tuning and country-agnosticism are in direct conflict.
The only labelled data in existence is American, so a fine-tuned model would
have been fitted to US legislative vocabulary, US institutions, and US fiscal
history. The property the deployment depends on is that the same frozen codebook
can be pointed at a Malaysian budget speech and behave the same way, and
training on the benchmark is the most reliable way to destroy it. @halterman2025
find directly that fine-tuning across codebooks overfits, and @gentzkow_text_2019
make the more general point that fixed text measures notoriously fail to
transfer across domains. The second is arithmetic: 44 labelled acts is two
orders of magnitude below the 4,000 to 20,000 examples per dataset that the
fine-tuning results in @halterman2025 rest on. The third is replication. A
frozen prompt is a document another researcher can read, criticise, and re-run;
a set of fitted weights is not, and @das_ai_2026 make the same choice for the
same reason. The practical constraint, that our access is API-only so
fine-tuning would mean moving to a weaker open-weight model, is real but comes
fourth.

The consequence is that every improvement in this project had to come from
rewriting codebooks rather than from training, which is slower and leaves a
visible audit trail of what was changed and why. @sec-us-validation reports what
that bought, and where it fell short.

## What the pipeline produces {#sec-what-it-produces}

```{r}
#| label: pending-any
#| eval: !expr "!have_any"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe Malaysia deliverable targets are empty or",
    "unbuilt (their C2a/C2b enrichment is API-gated). The dataset exhibits",
    "below render once they are built.\n:::\n")
```

The pipeline covers three instrument families, and the boundaries between them
are definitional choices worth stating before any number is read, because they
determine what counts as one event.

**Statutory tax** changes are changes to a standing headline rate, and they
split three ways. Corporate income tax tracks the general statutory rate on
company profits, with SME, sectoral, and petroleum rates recorded separately.
Personal income tax is a progressive schedule rather than a single rate, so the
headline series is the top marginal rate and bracket or threshold changes are
recorded alongside it. Consumption tax is the broad statutory tax on
value added or on sales of goods and services, in whatever local form it takes;
introducing such a regime counts as an increase and abolishing it as a cut,
which is what makes Malaysia's 2015 GST introduction and its 2018 repeal two
events rather than an administrative footnote.

**Spending** covers discretionary changes: deliberate policy choices such as
stimulus and relief packages, subsidy-policy changes, large allocations,
Five-Year Plan launches, and consolidation or restraint. It excludes automatic
stabilisers and the routine year-on-year drift of the baseline, and it is typed
by the component families of @das_ai_2026.

**Tax incentives** are changes to selective relief: holidays, investment and
reinvestment allowances, concessionary investor rates, zones, sectoral and R&D
incentives, and personal reliefs.

Incentives and statutory rates both operate on tax liabilities, so they need a
rule that keeps them disjoint. Following @klemm_empirical_2012, the test is
selectivity, temporariness, and mechanism: a measure that is targeted on an
investor, sector, region, or activity, or is time-bounded, or works through the
base rather than the standing rate, is an incentive, and a universal, permanent
change to a standing statutory rate is not. Every event lives in exactly one
component, so the three inventories can be read together without
double-counting.

@fig-dataset-overview represents the whole Malaysia dataset in one frame: 51
narratively-announced fiscal events between 1980 and 2023, split into those
three families. Three features of the data are visible before any table is read.
Statutory tax changes are dominated by rate *cuts* spread evenly across the
period and are mostly read as exogenous. Spending events cluster tightly on
1998, 2009, and 2020–21 and are almost all read as endogenous, which is what a
crisis-response spending record should look like. Tax incentives run the full
period and are almost uniformly liability-reducing and exogenous, which is what
a long-run investment-promotion regime looks like.

```{r}
#| label: fig-dataset-overview
#| fig-cap: "**The Malaysia inventory at a glance.** One mark per announced fiscal event, positioned at its year of effect; marks stack where several events share a year. Colour is the direction of the change, fill is the preliminary exogeneity read."
#| fig-width: 6.5
#| fig-height: 4.2
#| eval: !expr have_any
p <- plot_dataset_overview(tax_shocks, spending_shocks, incentive_shocks)
if (is.null(p)) cat("No component data available.\n") else p
```

@tbl-dataset-summary puts numbers on that picture. Two columns in it are about
provenance rather than about fiscal policy: the source documents and evidence
chunks count the primary-source passages behind each component, and they are
what makes the exogenous share in the last column checkable rather than merely
reported. That share is the model's preliminary read, and the spread across
components (77% of tax events, 14% of spending events) is doing real work later
in the paper.

```{r}
#| label: tbl-dataset-summary
#| tbl-cap: "**Inventory summary.** Coverage and composition per component. Source documents and evidence chunks count the primary-source passages each event was built from; the exogenous share is the model's preliminary read, not an adjudicated finding."
#| eval: !expr have_any
dataset_summary_stats(tax_shocks, spending_shocks, incentive_shocks) |>
  tt() |>
  tt_theme_report()
```

The unit of observation is one narratively-announced act, per instrument.
@tbl-one-row sets out the schema. Exhibit 1 below then reproduces one real row
in full, which is the faster way to see what a row means.

```{r}
#| label: tbl-one-row
#| tbl-cap: "**Unit of observation.** One row per narratively-announced act × instrument, carrying a magnitude, a motivation label, and a preliminary exogeneity read pending expert adjudication."

tibble::tribble(
  ~Field,            ~Meaning,
  "act",             "One narratively-announced fiscal act (phased steps collapse to one row)",
  "instrument",      "CIT / PIT / CONSUMPTION (tax), Expenditure (spending), or incentive category",
  "year",            "Announced and effective year (quarterly timing is deferred)",
  "magnitude",       "Δpp statutory rate (tax or rate-bearing incentive) or ringgit amount (spending)",
  "sign",            "Direction of the effect on fiscal liabilities (increase / decrease / no change)",
  "motivation",      "Romer–Romer 4-way label from the validated classification codebook",
  "exogenous",       "Preliminary exogenous/endogenous read — an input for the expert, not a finding",
  "provenance",      "Source documents, member chunks, and the verbatim quotes behind each field"
) |>
  tt(width = c(0.22, 0.78)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

Exhibit 1 reproduces `MY-CIT-01`, the 1993 announcement cutting Malaysia's
corporate rate from 34% to 32% with effect in 1994, and it is worth reading
slowly, because every block in it comes from somewhere different and the
differences are the point.

The header block is read from the documents: the tax, the announced and
effective years, the rate before and after, and the resulting signed effect on
liabilities. The **identification** note is the human's, and it records a
judgement rather than a fact. This cut was one step of a longer reduction, and
the researcher decided to keep it as its own row because the steps were
announced in separate budgets rather than legislated once as a schedule. A
different, defensible decision would have produced one row instead of several,
which is why the note names the decision and the reasoning rather than hiding
them. The **motivation** paragraph and the exogeneity verdict come from the
classification codebook, reading only the evidence extracted for this act, and
the quoted passage below it is that evidence. The **preliminary narrative read**
is a second, independent opinion formed during act identification, and printing
both is deliberate: where they disagree, the act is flagged rather than
resolved. Here the narrative pass could not settle the question and recorded
`ambiguous`, so the row goes to the expert with the disagreement visible.
Finally, the **sources** are live links to the documents the act was built from.

The general property this illustrates is that no field in the dataset is
free-floating. Each is either read from a named primary source, produced by a
validated codebook from quoted evidence, or stamped by a human whose reasoning
is recorded. A reviewer who disagrees with the label has somewhere to go that is
not the model.

```{r}
#| label: exm-anatomy
#| echo: false
#| output: asis
#| eval: !expr have_tax
tax_exhibit(
  tax_shocks, "MY-CIT-01",
  exhibit_label = "Exhibit 1",
  note = "One row in full. Every field below is either read from a named primary source or produced by a validated codebook from the quoted evidence."
)
```



## The Malaysia inventory in detail

The exhibits below represent the headline output of the pipeline for Malaysia,
1980–2023, component by component. The three frozen datasets, their data
dictionaries, and the full set of worked act write-ups ship with the paper's
replication material.

### Statutory tax changes

```{r}
#| label: pending-tax
#| eval: !expr "!have_tax"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `tax_shocks` deliverable target is empty or",
    "unbuilt (its C2a/C2b enrichment is API-gated). The tax figures below render",
    "once it is built.\n:::\n")
```

The statutory tax component holds 22 acts: 9 corporate, 9 personal, and 4 on the
consumption base. The three tables that follow report them one tax at a time,
each row naming the act, the year the change took effect, the change in the
headline rate in percentage points, and the motivation and exogeneity the
classification codebook assigned.

The corporate record in @tbl-tax-inventory-cit is the cleanest downward path in
the dataset: seven cuts between 1994 and 2016 against a single increase, the
2022 Cukai Makmur one-off surcharge. Motivations are almost uniformly long-run,
with the model reading the sequence as investment-climate policy rather than
demand management. The two departures are informative. The 1998 cut is the one
corporate act read as countercyclical, and it lands in the middle of the Asian
Financial Crisis. The 2016 cut and the 2022 surcharge are both read as
deficit-driven, which is a different exogenous route into the series and one
that only appears once fiscal consolidation is on the agenda.

```{r}
#| label: tbl-tax-inventory-cit
#| tbl-cap: "**Corporate income tax inventory.** One row per announced act; Δpp is the headline-rate change; motivation and exogeneity from the validated classification codebook."
#| eval: !expr have_tax
tax_inventory_table(tax_shocks, tax_type = "CIT")
```

The personal record in @tbl-tax-inventory-pit is less monotone, because a
progressive schedule can move in two directions at once. Six of the nine acts
cut the top marginal rate, but three raise it, and the recent entries are
restructurings rather than simple cuts: the 2016 and 2020 acts add high-earner
bands above RM1m and RM2m, and the 2023 act cuts the middle of the schedule
while raising its top. Reading the top marginal rate alone would record that
last act as roughly neutral, which is why the inventory carries the structural
note alongside the rate change.

```{r}
#| label: tbl-tax-inventory-pit
#| tbl-cap: "**Personal income tax inventory.** One row per announced act; Δpp is the change in the top marginal rate, which for restructuring acts understates the change to the schedule."
#| eval: !expr have_tax
tax_inventory_table(tax_shocks, tax_type = "PIT")
```

The consumption record in @tbl-tax-inventory-consumption is the shortest and the
most eventful, because Malaysia changed regime twice in seven years:
the 2015 introduction of a 6% Goods and Services Tax, its 2018 repeal, and the
immediate reinstatement of the Sales and Services Tax in its place. The
introduction is read as deficit-driven and the repeal as countercyclical, which
tracks the documented politics of each.

```{r}
#| label: tbl-tax-inventory-consumption
#| tbl-cap: "**Consumption tax inventory.** One row per announced act. The 2018 repeal and the SST reinstatement that replaced it are two rows, since they are two changes to statutory liabilities."
#| eval: !expr have_tax
tax_inventory_table(tax_shocks, tax_type = "CONSUMPTION")
```

@fig-tax-rate-paths turns the rate-bearing acts back into the statutory rate
paths they imply, which is the form a user of the dataset will want: not a list
of changes but the level of each headline rate in each year. It also functions
as a check, since a path with an implausible jump or a missing step is a sign
that an act was missed.

```{r}
#| label: fig-tax-rate-paths
#| fig-cap: "**Statutory rate paths by tax type.** Step path of the headline statutory rate implied by the rate-bearing acts."
#| eval: !expr have_tax
plot_tax_rate_paths(tax_shocks)
```

Those paths admit an external check. @vegh_how_2015 assemble a hand-built
statutory-rate panel covering a large set of countries, Malaysia among them,
over part of this period. Comparing our reconstructed paths against theirs is
the most direct external validation this component admits.
@sec-rate-benchmark runs it: the paths agree exactly wherever the corpus is
dense, and every disagreement is either a year the corpus does not cover or an
act a top-rate series cannot represent.

The act-level record also carries a distinction an aggregate rate series cannot,
and the personal-income acts above are the case in point: several of them move a
bracket structure or a base rather than a headline rate. That separation is
consequential rather than bookkeeping. @dabla-norris_macroeconomic_2023, coding
more than 2,000 tax measures, find base-broadening markedly less contractionary
than rate increases, and @crispolti_cross-country_2022 show revenue yields differ
systematically by instrument and by whether a reform moves the rate or the base.
A dataset that recorded only the headline rate would collapse that distinction.

@fig-tax-timeline shows when the acts happened and what the model made of them.
The visible pattern is that tax policy activity is spread fairly evenly across
the period rather than bunched in crises, and that the bars are dominated by the
long-run label, with the countercyclical readings concentrated in 1998, 2002,
and 2018.

```{r}
#| label: fig-tax-timeline
#| fig-cap: "**Tax-change timeline.** Acts per year × tax type as a diverging stacked bar (hikes up, cuts down); colour denotes the motivation label."
#| eval: !expr have_tax
plot_tax_timeline(tax_shocks)
```

### Government spending shocks

```{r}
#| label: pending-spending
#| eval: !expr "!have_spending"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `spending_shocks` deliverable target is empty or",
    "unbuilt. The spending figures below render once it is built.\n:::\n")
```

The spending component in @tbl-spending-inventory holds 14 acts, and it is the
component that looks least like the tax side. Twelve of the fourteen are
increases, and the named packages are immediately recognisable as crisis
responses: the 1998 and 2001 stimulus packages, the two 2009 packages around the
global financial crisis, and the six COVID-era packages from PRIHATIN in 2020
through PEMULIH in 2021. The classification codebook reads twelve of the
fourteen as countercyclical and therefore endogenous. Only the two subsidy
rationalisations, in 2008 and 2010, come out exogenous, and they come out
deficit-driven, which is the reading their documents support.

That 14% exogenous share is not a weakness of the pipeline. A spending record
composed largely of endogenous crisis response is what the fiscal history of an
emerging market ought to produce, and the fact that the codebook recovers it
without being told to is one of the more reassuring results in the paper.

It is worth being explicit that there is nothing to check this component against.
The action-based consolidation datasets [@pescatori_new_2011;
@guajardo_expansionary_2014; @adler_updated_2024] are the benchmark class for
narrative spending and consolidation measures, and they are also the series the
LLM-based papers validate their own output against. Malaysia is not in them. That
absence is simultaneously the reason this pipeline exists and the reason this
component reports no concordance statistic. The contrast with @das_ai_2026 is the
sharpest available: they reach 64 countries by reading a harmonised secondary
source and recover the direction of each spending action without its size, while
the ringgit amounts in @tbl-spending-inventory are the trade made in the other
direction. Ours are extracted from the announcing document and traceable to it;
they are not validated.

```{r}
#| label: tbl-spending-inventory
#| tbl-cap: "**Government spending-shock inventory.** One row per announced spending act; exogeneity from the validated classification codebook, alongside the @das_ai_2026 two-condition preliminary screen."
#| eval: !expr have_spending
spending_inventory_table(spending_shocks)
```

@fig-spending-timeline makes the clustering visible. Three episodes carry almost
the whole component, and the years between them are close to empty, which is the
signature of a discretionary spending record rather than a budget series.

```{r}
#| label: fig-spending-timeline
#| fig-cap: "**Spending-shock timeline.** Acts per year as a diverging stacked bar (increases up, decreases down); colour denotes the exogenous/endogenous read."
#| eval: !expr have_spending
plot_spending_timeline(spending_shocks)
```

### Tax incentives and holidays

The incentive component in @tbl-incentive-inventory holds 15 acts and is the
only one spanning the full period, from the 1980 New Economic Policy equity
exemption to the 2023 electric-vehicle package. Every one of the fifteen reduces
liabilities, which is what an investment-promotion regime does: incentives are
announced and extended far more readily than they are repealed. Fourteen are
read as long-run, the exception being the 2020 relocation incentive announced
inside the PENJANA recovery package, which the classifier read as
countercyclical on the strength of the vehicle it arrived in. That single
disagreement is worked through in @sec-my-validation.

The `Base` column records the tax the incentive operates on. Thirteen sit on
corporate income, one on personal income, and one on the consumption base: the
2023 electric-vehicle package works through import-duty, excise, and sales-tax
exemptions rather than through a profits rate. That row is a reminder that the
component is defined by mechanism, not by base.

The nearest existing comparator is @klemm_empirical_2012, who assemble tax
holidays, allowances, credits, and corporate rates for more than 40 countries in
Latin America, the Caribbean, and Africa between 1985 and 2004. That is an
instrument inventory rather than a narrative one: it records what each incentive
was, not why it was granted, and it covers neither Malaysia nor this period. The
motivation column in @tbl-incentive-inventory is what this pipeline adds to that
class of dataset, and it is also the column carrying the most interpretive risk,
because an incentive's stated rationale is almost always developmental whatever
the timing of its announcement. The 2020 relocation incentive is the visible
instance. That these measures move real behaviour is not itself in doubt:
@agarwal_consumption_2017 find temporary consumption-tax holidays produce
substantial increases in spending on covered goods that are not offset before or
after.

```{r}
#| label: pending-incentives
#| eval: !expr "!have_incentives"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `incentive_shocks` deliverable target is empty or",
    "unbuilt. The incentive figures below render once it is built.\n:::\n")
```

```{r}
#| label: tbl-incentive-inventory
#| tbl-cap: "**Tax-incentive shock inventory.** One row per announced incentive change; exogeneity from the validated classification codebook."
#| eval: !expr have_incentives
incentive_inventory_table(incentive_shocks)
```

@fig-incentive-timeline shows the resulting timeline: a steady drip of
liability-reducing measures across four decades with no repeals, which is a
different fiscal object from either of the other two components and the reason
they are kept apart.

```{r}
#| label: fig-incentive-timeline
#| fig-cap: "**Incentive-shock timeline.** Acts per year as a diverging stacked bar (new or expanded down, repeals up); colour denotes the exogenous/endogenous read."
#| eval: !expr have_incentives
plot_incentive_timeline(incentive_shocks)
```

# How the Malaysia dataset was built {#sec-build}

This section walks @fig-pipeline for Malaysia, stage by stage, so the dataset's
provenance is reconstructible. The short version: a 168-document primary-source
corpus was scanned exhaustively by a validated codebook, the scan was used as a
seed rather than as the event list, acts were consolidated by an agentic
narrative search that was audited for recall and stamped by a human, and only
then were the validated motivation codebooks run against each act's own
evidence.


## The corpus {#sec-corpus}

```{r}
#| label: pending-corpus
#| eval: !expr "!have_corpus"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe deployment corpus targets are unbuilt, so the",
    "corpus table cannot render.\n:::\n")
```

Malaysia's fiscal record is published across four document series, none of which
alone covers the period. Budget Speeches announce measures but do not always
report the rates they replace; Economic Reports carry the retrospective rate
tables; Bank Negara annual reports carry the macro framing that motivation
classification depends on; and the Five-Year Plans carry the structural and
development-expenditure commitments. The corpus takes all four, plus the
standalone crisis packages. It is also bilingual: ten Economic Reports were
published in both English and Bahasa Malaysia, and that parallel decade is what
makes the cross-language test in @sec-my-validation possible.

@tbl-corpus reports what was actually read. The distinction the `Pending` column
draws is important and easy to lose: the acquisition pass catalogued
`r n_docs_catlg` documents, but `r n_docs_catlg - n_docs_read` of them sit in
archives we have not yet obtained, so the corpus the codebooks actually read is
`r n_docs_read` documents and `r scales::comma(n_pages_read)` extracted pages.
Counting the catalogue rather than the corpus would overstate coverage by a
third.

```{r}
#| label: tbl-corpus
#| tbl-cap: "**The deployment corpus.** Primary-source document series behind the Malaysia dataset. Documents, pages and year span describe the extracted corpus; `Pending` counts documents catalogued by the acquisition pass whose PDFs have not yet been obtained and which therefore contributed nothing."
#| eval: !expr have_corpus
corpus_scope_stats(country_body) |>
  tt(width = c(0.30, 0.15, 0.14, 0.12, 0.12, 0.17)) |>
  format_tt(j = "Pages", digits = 0, num_mark_big = ",") |>
  tt_theme_report() |>
  style_tt(j = 1, align = "l")
```

@fig-corpus-coverage shows how those pages are distributed over time, and it
carries a caveat the totals hide. The deep archive is recent. Before the
mid-1990s only the Five-Year Plans are available to us, and the annual series
begin in 1995 (Economic Reports), 1997 (Bank Negara) and 1999 (Budget
Speeches). The dataset nonetheless contains events back to 1980, because the
Economic Reports and Plan documents carry retrospective rate tables and
historical narrative. That is a legitimate route to an early event, but it is a
weaker one: an act recovered from a later document's summary table has less
motivation evidence attached than an act read from the speech that announced it.
Early-period rows should be read with that in mind.

```{r}
#| label: fig-corpus-coverage
#| fig-cap: "**Corpus coverage by year and series.** Extracted pages available in each year, stacked by document series. The bilingual Economic Reports (BM) mark the decade the cross-language consistency test uses."
#| fig-width: 6.5
#| fig-height: 3.6
#| eval: !expr have_corpus
p <- plot_corpus_coverage(country_body)
if (is.null(p)) cat("No corpus coverage data available.\n") else p
```

Extraction and cleaning produce a page-level text stream, which is then
segmented into the unit the measure-identification codebook actually reads:
`r scales::comma(nrow(country_chunks))` chunks across
`r dplyr::n_distinct(country_chunks$doc_id)` documents. The segmentation is a
sliding window over pages rather than a token
splitter: ten pages per chunk, advancing seven pages at a time, so consecutive
chunks overlap by three pages. Pages are joined with an explicit page-break
marker, chunks shorter than 100 characters are dropped as extraction artifacts,
and a 40,000-token ceiling raises a warning rather than splitting a chunk.

The choice of a page window over a semantic or token-based splitter is
deliberate. A budget speech states a measure in one sentence and its
justification several paragraphs later, sometimes on the next page; a splitter
that cuts on token count would routinely separate the two, and the motivation
stage would then be asked to classify a measure whose stated reason is in a
different chunk. The overlap exists for the same reason, so that a measure near
a page boundary appears whole in at least one chunk. The cost is redundancy,
since the same passage is read more than once, and that cost is paid at the
identification stage rather than passed to the reader.

## The codebook scan and what it is for {#sec-scan}

The measure-identification codebook reads every chunk and returns the fiscal
measures discussed in it. On the Malaysia corpus it surfaced **7,676
measure-mentions** across 1,193 chunks, at a median of five measures per chunk
and a maximum of 42 — a density that matters twice below, once here and once in
@sec-defect.

That scan is a *seed*, not the event list, for two reasons. The first is
mechanical: the handoff to the motivation stage keeps only the highest-ranked
measure per chunk, which discards **84.5%** of the mentions and leaves 1,036
measure rows. The second is conceptual: a scan returns mentions, and one act
generates many mentions across many documents and both languages, while some
acts generate mentions the scan phrases differently each time. Turning mentions
into acts is a separate problem, and it is the subject of the next two
subsections.

## Agentic narrative identification {#sec-agentic}

Acts are identified per instrument by an agentic procedure that reads the corpus
directly, in seven steps.

1. **Instrument configuration.** The concept is defined (statutory corporate
   income tax; broad consumption tax; discretionary spending; investment
   incentives), together with the surface terms it takes in the corpus, in both
   English and Bahasa Malaysia, and the document series that anchor it.
2. **Keyword scan.** The configured terms are matched against the codebook's
   measure pool. This is the recall floor and it is logged as a count, with
   false positives named.
3. **Semantic sweep.** The matched measures and their chunk text are read, and
   the search broadens past the keyword list to catch events phrased without the
   vocabulary — a rate change written as "reduce the rate to 24 per cent" and
   never as "corporate income tax".
4. **Recall recovery.** For rate instruments, the implied statutory rate path is
   reconstructed and checked for missing steps; for spending and incentives, the
   candidate set is checked against pre-declared known-act checkpoints. Where a
   step is missing, the anchor document's full text is read directly, behind the
   chunked view. This step is what recovered Malaysia's 2016 corporate rate cut
   and its 2022 Cukai Makmur surcharge, neither of which the scan surfaced.
5. **Consolidation.** Surface forms and recovered events are grouped into
   announced acts. A phased multi-step cut is one row carrying a schedule, not
   four rows; an omnibus act touching two taxes is two rows, one per instrument.
   A preliminary exogeneity read is recorded, anchored in a verbatim quote.
6. **Human review.** The consolidated events, the recall audit, and every open
   adjudication question are presented for review. Nothing is written until the
   researcher confirms the events, the consolidation grain, and the preliminary
   reads.
7. **Freeze.** The stamped dataset is written with a provenance record naming
   the reviewer, the commit, and the date, and a notebook that renders its
   contents back from the frozen file so the write-up cannot drift from the
   artifact.

**The agentic step earns its place at step 4.** The codebook scan is a strong
first pass, and better than the design expected. Although it is scoped to tax
liabilities, it surfaced 189 spending measures and 151 incentive measures,
because Malaysian budget and central-bank documents discuss stimulus, subsidy,
and incentive policy in the same breath as tax. On the track it should have been
best at, though, the statutory rate path, it was not sufficient on its own.
Reconstructing the corporate rate path and asking which mentions account for
each step exposed gaps that no amount of reading the scan's output would have
revealed, and reading the anchor documents directly, behind the chunked view,
recovered them. Malaysia's 2016 corporate rate cut and its 2022 Cukai Makmur
surcharge both entered the dataset that way.

That is a specific claim about division of labour, not a claim that the codebook
underperformed. A chunk-level classifier answers "does this passage discuss a
fiscal measure", which is the question it was validated on. It was never asked
"is the rate path complete", because a chunk cannot answer that: completeness is
a property of the corpus, not of any passage in it. The agentic pass is what
turns the second question into something answerable, by holding a model of the
instrument and checking the corpus against it.

Recall is nonetheless the property this dataset can least defend on its own.
Precision is checkable by anyone holding it, since each row names its sources
and quotes. Recall is not: no reader can see what was never found. The recall
scorecards frozen with each component (reproduced in @sec-appendix-recall)
record what was searched for and what the search returned, and they are useful
as a transparency record, but they are not a recall measurement. They are the
search describing itself, and a search cannot certify its own coverage.

The honest assessment has to come from outside, and it is the assessment
currently being collected: domain experts on Malaysian fiscal history are being
asked which significant acts are missing from the inventory, which is a question
only someone holding an independent mental list can answer. That procedure is
subjective, and we would rather say so than substitute a number that looks
objective and measures the wrong thing. @sec-my-validation sets out the design;
its results are not yet in.

## Act consolidation without automated record linkage {#sec-no-c0}

We built an automated act-aggregation step for this purpose and did not use it.
The step clusters the scan's surface forms into canonical acts and was scored
against the 49 Romer–Romer reference acts, where it recovers most of them. On
the Malaysia pool of 877 distinct surface forms it fails in the way that matters
for an event layer: it fragments the instrument tracks that need to stay whole —
the corporate-rate track shattered into near-singletons, the consumption-tax
track split across nine clusters, and it never bridged the English and Bahasa
Malaysia names of the same act. Both failures are recorded in the frozen recall
audits in @sec-appendix-recall.

What we ran into is the oldest problem in record linkage, and it helps to name it
as such. Deciding whether two mentions refer to the same underlying entity is the
decision problem @fellegi_theory_1969 formalised, and the modern taxonomy of
approaches to it is surveyed by @binette_almost_2022. Every method in that
literature trades off the same two errors we hit: merging records that belong
apart, and splitting records that belong together. Economics has shown the
trade-off can be managed well at scale, with @abramitzky_automated_2021
documenting automated historical linkage whose false-positive and match rates
stand comparison with hand linkage. Two features of our problem sit badly with
those results. Their linkage keys are person-level attributes with stable
distributions, whereas an act's surface form is a phrase a speechwriter chose. And
their algorithms are tuned and audited against hand-linked training pairs, which
for Malaysian acts do not exist. The conventional finding, that record linkage
automates well, is therefore compatible with ours, that this instance of it does
not.

The finding is a negative result about clustering, not about LLMs: consolidating
mentions into acts requires knowing which mentions describe the same *statutory
instrument*, and surface-form similarity does not carry that. The agentic
procedure of @sec-agentic succeeds where clustering fails because it works
from the instrument down, reconstructing the rate path and then asking which
mentions account for each step, rather than from the mentions up. That is also
why it needs a human stamp: the reconstruction is a judgement about what counts
as one act, and the researcher owns it. Keeping a human in that loop is the
conventional position in the linkage literature rather than a concession, since
the algorithm-versus-human comparisons that would license full automation are
exactly what a country with no hand-linked acts cannot run.

## A defect the evidence chain surfaced {#sec-defect}

The motivation stage is two codebooks: an extraction step that pulls the
motivation evidence for an act out of a passage, and a classification step that
assigns the four-way label from that evidence. During reviewer inspection of two
corporate-tax rows in August 2026, both codebooks turned out to be reading the
wrong act's evidence, through two independent mechanisms.

The first was a join. Evidence had been extracted once per chunk, under whichever
measure the scan ranked first — the 84.5% discard of @sec-scan — and events
were matched to that store on chunk identity alone. But the identification
procedure cites a chunk because the act's sentence is in it, not because the act
was ranked first. 95 of 100 tax member-chunk rows, and every spending and
incentive row, were structurally exposed. The second was the extraction codebook
itself, which had been instructed to extract *all* motivation evidence in a
chunk. That is harmless on the US development chunks, which discuss one act
each, and wrong on Malaysian budget speeches carrying a median of five measures
and a maximum of 42.

An audit of the 14 events where the model and the identification pass disagreed
about exogeneity found four classifications built on another act's evidence: two
corporate-tax rows, two personal-income-tax rows. Both mechanisms are fixed —
extraction is now scoped to the named act, and evidence is extracted per event
rather than per chunk — and the enrichment has not yet been re-run, so the four
rows remain wrong in the exhibits above.

We report this as a result rather than as a maintenance note, because it is the
kind of failure the validation framework exists to catch and did not. Both
codebooks passed every US test. They failed on a distributional difference
between the validation corpus and the deployment corpus — single-act chunks
versus multi-measure budget speeches — that no US test could have exposed. The
cost is quantified because the provenance chain made it quantifiable: four wrong
labels out of 51, found by following each label back to the passage it came
from. A pipeline without that chain would have shipped them.

# Validation {#sec-validation}

The pipeline described so far produces an inventory. This section asks whether
that inventory should be believed.

The **US benchmark** (@sec-us-validation) answers the first question: given the
same documents Romer and Romer read, and judged against the labels they
published, can an LLM reproduce their identification? That question has a
correct answer to score against, so it is answered with accuracy statistics.

**Malaysia** (@sec-my-validation) answers the second: do the same codebooks,
frozen, still work in a country with no labelled data? That question has no
correct answer to score against, so it is answered by consistency evidence and
expert agreement instead. The second is the harder question and the more
important one.

## The US benchmark {#sec-us-validation}

**Can an LLM, given the same primary documents and judged against Romer and
Romer's own labels, reproduce their identification of US fiscal changes?**

The ground truth we score against needs stating plainly, because everything in
this section depends on what it is and is not. @romer2010 published, alongside
their results, an act-level record, since archived as replication data
[@romer_replication_2019]: for each of the significant federal tax actions
between 1945 and 2007, the act's name, its motivation category, and the narrative
reasoning behind that classification, set out at length in their companion
narrative analysis [@romer]. After matching that record to
the US source documents our pipeline reads, 44 acts carry both a published label
and locatable passages in the corpus. Those 44 acts are the only ground truth
that exists for this task anywhere in the world, which is both why the US is the
benchmark and why the benchmark is small.

The labels are act-level, but the codebooks work on chunks of documents, so the
gold standard has to be projected from one grain to the other. That projection
is itself a procedure with error in it, and the error turns out to matter more
than any modelling choice in this section, so @sec-appendix-ika documents how
it was done and @sec-c1-manual quantifies what it cost.

This subsection proceeds in three parts. It first states what the benchmark
covers and the one caveat that qualifies all of it. It then takes the two
codebooks in pipeline order: measure identification (@sec-c1), then motivation
and exogeneity (@sec-c2). Each runs the same sequence, behavioural tests, then
zero-shot evaluation, then a manual error analysis that corrects for noise in
the gold labels, and in each case the manual analysis is the gate rather than
the automated metrics.

@fig-pipeline-validated marks what the benchmark covers. The gap is as important
as the coverage: it certifies the two LLM classification stages and says nothing
about corpus construction or act identification, which are attested in other
ways or not at all (@sec-coverage).

```{r}
#| label: fig-pipeline-validated
#| fig-cap: "**What the US benchmark covers.** Stages carrying a formal validation record against US ground truth are lit; the rest are attested by other means."
#| fig-width: 6.5
#| fig-height: 4.2
plot_pipeline_flow("validated", scale = 1.4)
```

One caveat frames everything that follows, and @sec-frameworks has already named
it. The published Romer–Romer shock list and the US source documents are almost
certainly in the models' pretraining data, so strong US agreement is
**consistent with memorisation** rather than proof against it
[@carlini_quantifying_2022]. This is why the behavioural battery includes an
explicit memorisation probe, and why the burden of proof sits on the in-language
Malaysia evidence in @sec-my-validation, where the documents are far less likely
to have been seen during model training.

```{r}
#| label: pending-il
#| eval: !expr "!have_il"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe `iteration_logs` target is unbuilt, so the H&K",
    "validation battery cannot render. Build it (API-free) and re-render.\n:::\n")
```

Both codebooks are put through the same battery of behavioural tests, and
@tbl-hk-tests serves as the reader's key to it. The column that matters is the last
one: each test licenses a specific claim, and a codebook that fails one has lost
the right to that claim regardless of how well it scores later. The tests divide
into two groups by when they run. Tests I to IV run before any evaluation data
is touched and ask whether the model can follow the codebook at all. Tests V to
VII run alongside the error analysis and ask the harder question of whether it
is following the codebook or something it already believed.

```{r}
#| label: tbl-hk-tests
#| tbl-cap: "**The behavioural battery.** What each test probes and what passing it licenses. Tests I–IV run before any evaluation data is touched; Tests V–VII run alongside the ablation and the manual error analysis [@halterman2025]."
tt_hk_tests_table()
```

### Measure identification {#sec-c1}

The measure-identification codebook decides whether a chunk substantively
discusses a specific fiscal measure, and names the measures it finds.
@tbl-c1-classes gives the two labels it chooses between and the definitions it
chooses by. Those definitions are the short form; each carries a set of
inclusion and exclusion clarifications that do most of the discriminating work
in practice, and the codebook is reproduced in full, exactly as the model reads
it, in @sec-appendix-codebooks. The same holds for every codebook in this
paper: the body gives the labels and definitions, the appendix gives the
artifact.

```{r}
#| label: tbl-c1-classes
#| tbl-cap: "**Measure-identification classes.** The two labels the codebook chooses between, and the definitions it chooses by."
codebook_classes_table(here("prompts/c1_measure_id.yml")) |>
  tt(width = c(0.24, 0.76)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

**This codebook is tuned for recall, deliberately.** It sits at the front of the
pipeline, and a measure it misses is invisible to every stage after it, while a
measure it wrongly surfaces is cheap: the act-identification pass discards it,
and the motivation stage never sees it. So the operating point is set to
maximise the fiscal policy available downstream, and precision is treated as
recoverable rather than binding. The diagnostic benchmarks reflect that ordering
— Tier 1 recall ≥95%, combined recall ≥90%, precision ≥70% — and none of them is
a hard gate. The gate is the manual error analysis below.

#### Behavioural tests

@fig-c1-s1 shows the battery passing cleanly, and that is the expected result
rather than an achievement: every frontier commercial model we tried clears
these tests on this codebook. Their purpose is not to discriminate between
models but to prove the harness is sound, that the schema, the parsing, and the
token budget let the model express what the codebook asks for. They earned their
keep anyway. An apparent Test I failure turned out to be an output-token cap
truncating long measure lists rather than a codebook defect, and it was the
test, not the evaluation, that isolated it.

```{r}
#| label: fig-c1-s1
#| fig-cap: "**Measure identification, behavioural tests.** Tests I–III as percent correct; Test IV as Fleiss' kappa across label orderings. See @tbl-hk-tests for more information on the tests."
#| fig-height: 5
#| eval: !expr have_il
plot_s1_behavioral(latest(il$s1, "c1"))
```

#### Zero-shot evaluation

@fig-c1-s2 represents the zero-shot evaluation against the projected gold
labels, and it reports recall at two tiers, which needs a word of explanation.
Because Romer and Romer's labels are attached to *acts* and the codebook reads
*chunks*, positives have to be located in the corpus, and they are located two
different ways with two different degrees of confidence. **Tier 1** chunks
contain an act's narrative passage verbatim, so they are as close to a certain
positive as this construction gets. **Tier 2** chunks merely *name* a known act,
matched through whitespace-normalised substring search, decomposition of
compound act names, and a small set of co-occurrence rules. Tier 2 is
deliberately loose, because an act name appears in the corpus in many more
places than its passage does, and @sec-appendix-ika documents the construction
in full.

The headline is that the codebook we tuned for recall is in fact stronger on
precision: 94.7% precision against 80.0% combined recall, with Tier 1 recall at
88.2%. Both recall figures sit below their diagnostic benchmarks. Read at face
value, that is a codebook that misses one positive chunk in five, which would be
bad news given where it sits in the pipeline. The manual analysis below is what
establishes that this is not what is happening.

```{r}
#| label: fig-c1-s2
#| fig-cap: "**Measure identification, zero-shot evaluation.** Point estimates with 95% bootstrap intervals; crosses mark the diagnostic benchmark where one is defined."
#| fig-height: 3.6
#| eval: !expr have_il
p <- plot_s2_metrics(latest(il$s2, "c1"))
if (is.null(p)) cat("No S2 metrics available.\n") else p
```

#### Codebook adherence

Accuracy alone cannot distinguish a model that read the codebook from one that
already agreed with it. @fig-c1-s3 reports the three adherence tests that can:
they perturb the codebook rather than the data, and ask whether the model's
output moves with it. This is asking whether the model is right for the right
reasons.

```{r}
#| label: fig-c1-s3
#| fig-cap: "**Measure identification, codebook-adherence tests.** Tests V–VII against the full-codebook baseline. Test VI compares accuracy under informative and neutral label names; Test VII puts label names and definitions in conflict."
#| fig-height: 5
#| eval: !expr have_il
plot_s3_behavioral(latest(il$s3, "c1"), latest(il$s3_ablation, "c1"))
```

#### Manual error analysis and the label-noise correction {#sec-c1-manual}

The gate is a human reading of every disagreement between model and gold label,
sorted into the six-category framework of @halterman2025: the model was right
(A), the gold label was wrong (B), the document was corrupted (C), the model
ignored the instructions (D), the model reasoned incorrectly (E), or something
else (F).

```{r}
#| label: tbl-c1-manual
#| tbl-cap: "**Manual error analysis.** Distribution over the six-category framework for the measure-identification codebook."
#| eval: !expr have_il
tt_manual_analysis_table(latest(il$s3_manual, "c1"))
```

The Category B count is the finding, not a footnote. Our US gold labels are not
hand-coded chunk by chunk; they are attached by matching Romer–Romer act names
against chunk text. Verbatim passage matches are reliable, but the second tier
matches act-name *fragments*, and act-name fragments appear in book
indexes, tables of contents, glossaries, and administrative appendices. Those
chunks arrive carrying a positive gold label and no fiscal content whatsoever.
When the model correctly declines to call a book index a fiscal measure, the
evaluation scores it a false negative.

**The label-generating process is noisier than the thing it is measuring**, and
that is a general hazard for LLM content analysis at this scale: automating gold
label construction to make evaluation feasible imports exactly the errors the
evaluation is meant to detect. It is the specific form, at scale, of the warning
that runs through the economics text-as-data literature: measurement from text
must be validated rather than assumed valid, and a validation is only ever as
good as its own labels [@ash_text_2023]. The response is not to discard the noisy
labels
but to quantify them. The manual analysis identifies which disagreements are
label noise, and the gate is recomputed with those chunks excluded from the
denominator, as @tbl-c1-confusion and @tbl-c1-gates report. Ambiguous cases are
kept in the denominator rather than excluded, so the correction is conservative.

```{r}
#| label: tbl-c1-confusion
#| tbl-cap: "**Measure identification, corrected confusion matrix.** Counts after excluding the chunks the manual analysis judged mislabelled in the gold standard."
#| eval: !expr have_il
tb <- tt_confusion_matrix(latest(il$s3_manual_bias_corrected, "c1"))
if (is.null(tb)) cat("No corrected confusion matrix available.\n") else tb
```


@tbl-c1-gates reports the corrected rates against the diagnostic benchmarks. The
correction moves the picture substantially: on the 33 chunks that survive it,
recall rises to 87.5% and Tier 1 recall to 80.0%, against raw figures of 80.0%
and 88.2%. Neither reaches its benchmark. We nonetheless took the codebook
forward, and the reasoning should be visible rather than buried in a "passed".
The benchmarks were set as diagnostics, not gates, precisely because we
anticipated that the projected gold labels would be noisy, and the manual
analysis confirmed that anticipation in a specific and correctable form. The
effective sample after correction is small enough that these rates carry wide
intervals. And the quantity the pipeline actually depends on is whether real
fiscal content reaches the identification stage, which the Malaysia deployment
then tests directly: the agentic pass found the corpus richly populated, and the
misses it did record were rate-path steps rather than whole measures. That is a
judgement call, made once, recorded here, and open to disagreement.

```{r}
#| label: tbl-c1-gates
#| tbl-cap: "**Measure identification, corrected performance.** Label-noise-adjusted rates against the diagnostic benchmarks. Recall is the binding quantity here; precision is recoverable downstream."
#| eval: !expr have_il
tt_gate_metrics(
  latest(il$s3_manual_bias_corrected, "c1"),
  targets = c(recall = 0.90, tier1_recall = 0.95, precision = 0.70)
)
```

### Motivation and exogeneity {#sec-c2}

#### Why this is two codebooks {#sec-c2-split}

Assigning a Romer–Romer motivation label to a passage is two tasks wearing one
name. The first is evidentiary: find, in a document that discusses many things,
the sentences stating why *this* measure was enacted. The second is
interpretive: decide which of the four motivation categories that stated reason
belongs to, and whether it makes the act exogenous to current output.

Fusing them into one prompt makes failures uninterpretable — a wrong label could
be a reading failure or a reasoning failure, and the transcript cannot tell you
which. So the stage is split. An **extraction** codebook returns the motivation
evidence for a named act as verbatim quotes. A **classification** codebook sees
only those quotes and returns the label, the sign, and the exogeneity call. The
split buys three things: errors localise to a stage, the extracted evidence
becomes the audit trail a reviewer reads instead of the label, and the
classification step can be evaluated on clean evidence independently of
retrieval quality.

The split creates a problem for validation, though, and the way we resolve it is
worth stating because it is the reason @sec-defect happened. The Halterman and
Keith framework is built to evaluate **classification**: its behavioural tests
perturb label names and definitions, its zero-shot stage compares a predicted
label against a gold label, and its error taxonomy sorts disagreements between
two labels. None of that has a natural analogue for a free-form extraction step
whose output is a set of quotes. There is no gold set of correct quotes to score
against, and no meaningful confusion matrix over spans of text.

So the framework is applied where it fits, to the classifier, and the extraction
step is evaluated only through its effect on the classifier's output. In
practice that means the gate is applied to the composition: extraction runs on
real chunks, and the classifier is scored on whatever evidence extraction
actually produced, errors included, rather than on clean evidence hand-picked
for it. The extraction codebook carries its own behavioural and zero-shot record
but no separate error analysis, because the only errors we can adjudicate are
the ones that reach a label.

That choice has a cost, and @sec-defect is what it bought. The defect lived in
the *interface* between the two stages, not inside either one, and no metric
either stage carries could have surfaced it. Scoring the composition means such
a defect shows up as a slightly worse number rather than as a diagnosis, and
here it did not show up at all, because the composition was scored on US chunks
where the defect is inert.

@tbl-c2-classes gives the four motivation labels and their definitions as
implemented. The order matters: the first two are responses to the economy and
therefore endogenous, and the last two are the pair the shock series keeps.

```{r}
#| label: tbl-c2-classes
#| tbl-cap: "**Motivation classes.** The Romer–Romer four-way taxonomy as implemented. The first two are endogenous to current output; the last two are exogenous and are the ones that survive into the shock series."
codebook_classes_table(here("prompts/c2b_classification.yml")) |>
  tt(width = c(0.26, 0.74)) |>
  tt_theme_report() |>
  style_tt(j = 2, align = "l")
```

#### Behavioural tests and zero-shot evaluation

The behavioural battery is the same one @tbl-hk-tests describes, and
@fig-c2-s1 shows the classification codebook clearing it. Test II carries more
weight here than it did for measure identification, because it is the
memorisation probe and this is the codebook whose ground truth is a published,
widely-cited act list.

```{r}
#| label: fig-c2-s1
#| fig-cap: "**Motivation classification, behavioural tests.** Tests I–III as percent correct; Test IV as Fleiss' kappa across label orderings. Test II is the memorisation probe for the classification codebook."
#| fig-height: 5
#| eval: !expr have_il
plot_s1_behavioral(latest(il$s1, "c2b"))
```

@fig-c2-s2 shows the zero-shot evaluation. Two of these metrics are the ones a
downstream user depends on. Exogenous precision, the share of acts handed over
as exogenous that really are, lands at 80.0% against an 85% gate, though on a
sample small enough that the interval runs from 63% to 96%. Sign accuracy on
truly-exogenous acts, at 91.3%, clears its 90% gate. The rest of the panel is
diagnostic: the four-way task is harder than the binary one, and exogenous
recall in particular is weak, which matters much less than precision because a
missed exogenous act shrinks the series while a wrongly-included one
contaminates it.

```{r}
#| label: fig-c2-s2
#| fig-cap: "**Motivation classification, zero-shot evaluation.** Point estimates with 95% bootstrap intervals; crosses mark the gate where one is defined. Exogenous precision and sign accuracy on true-exogenous acts are the two gates that matter downstream."
#| fig-height: 3.6
#| eval: !expr have_il
p <- plot_s2_metrics(latest(il$s2, "c2b"))
if (is.null(p)) cat("No S2 metrics available.\n") else p
```

@fig-c2-s3 reports the adherence tests, and it is the most consequential
exhibit in this section. It asks not how often the classifier is right but
whether it is reading our codebook to get there, by perturbing the codebook and
watching what happens to the output.

```{r}
#| label: fig-c2-s3
#| fig-cap: "**Motivation classification, codebook-adherence tests.** Tests V–VII against the full-codebook baseline. Test VI compares accuracy under informative and neutral label names; Test VII puts label names and definitions in conflict; the ablation strips codebook components one at a time."
#| fig-height: 5
#| eval: !expr have_il
plot_s3_behavioral(latest(il$s3, "c2b"), latest(il$s3_ablation, "c2b"))
```

The result is uncomfortable and we report it rather than bury it. When label
names and label definitions are put in conflict (Test VII), the
model follows the *names* almost always and the definitions almost never. Read
together with a small ablation penalty — stripping the codebook's definitions
and clarifications costs only a few points of accuracy, and even the fully
stripped condition retains most of the four-way task — the natural reading is
that the model arrives already knowing what "countercyclical" and "long-run"
mean in fiscal policy, from pretraining, and that our codebook is calibrating a
prior rather than teaching a construct.

That is good news for transfer and bad news for control. It is good news because
a prior learned from the general fiscal-policy literature is not tied to US
institutions, which is the property cross-country deployment needs. It is bad
news because a researcher who redefines a category away from its conventional
meaning should not assume the model will follow, and because agreement on US
data is weaker evidence than it appears. It is the sharpest reason the Malaysia
evidence, not the US benchmark, has to carry the argument.

There is a name for the property this result is about. @asirvatham_gpt_2026
separate two requirements an LLM measurement has to meet: **accuracy**, agreement
with a benchmark, and **directness**, the property that the rating comes from the
text being rated rather than from something merely correlated with it. Test VII is
a directness result, and it is the test we do worst on. Their diagnostic is
signal-stripping: remove the rated content and confirm the rating collapses. The
analogue here is to mask the stated motivation in a chunk and re-classify, and if
the exogeneity call survives, the model is reading the act's name, its era, and
its sponsor rather than its rationale. We have not run that test. It is the single
most informative check missing from this paper, and @tbl-coverage records it as
such.

The finding also puts the encouraging annotation literature in context. Where
@gilardi_chatgpt_2023 and @tornberg_large_2025 find models matching or beating
human coders, the constructs at issue are conventional ones a model has seen
defined many times over, and @ziems_can_2024 find the advantage narrows precisely
where labels are bespoke and taxonomic. A pretrained prior is the most plausible
reason our classifier does well on Romer and Romer's categories, which are
canonical in fiscal economics. It is equally a reason to expect less of any
category a researcher invents.

#### Manual error analysis and the label-noise correction

As with measure identification, the gate is a human reading every disagreement
between the model and the gold label and sorting it into the same six
categories. @tbl-c2-manual gives the distribution, adjudicated on the composed
extraction-and-classification chain rather than on either codebook alone.

```{r}
#| label: tbl-c2-manual
#| tbl-cap: "**Manual error analysis.** Distribution over the six-category framework for the motivation codebooks, adjudicated on the composed extraction-and-classification chain."
#| eval: !expr have_il
tt_manual_analysis_table(latest(il$s3_manual, "c2b"))
```

The label noise here has a different source than in @sec-c1-manual and is
harder to correct. Motivation gold labels come from Romer and Romer's own
classifications, which are act-level judgements published with their reasoning.
Where a chunk discusses an act only glancingly, or where the act's motivation is
genuinely mixed, the act-level label is not a defensible chunk-level target —
and unlike an index page, there is no mechanical signature that marks such a
case. Every Category B act here is of that kind: the act carries two motivations
in its own stated reasons, and the gold label follows a tie-breaking convention
rather than the codebook. Adjudicating those is itself an interpretive act, so
the Category B count here is less mechanical and more contestable than the
measure-side one. We again exclude only what the manual pass judged mislabelled
and keep the ambiguous cases in the denominator.

```{r}
#| label: tbl-c2-confusion
#| tbl-cap: "**Motivation classification, corrected confusion matrix.** Exogenous-versus-endogenous counts after excluding the acts the manual analysis judged mislabelled or undecidable in the gold standard."
#| eval: !expr have_il
tb <- tt_confusion_matrix(latest(il$s3_manual_bias_corrected, "c2b"))
if (is.null(tb)) cat("No corrected confusion matrix available.\n") else tb
```

```{r}
#| label: tbl-c2-gates
#| tbl-cap: "**Motivation classification, corrected performance.** Label-noise-adjusted rates against the two gates that bind: exogenous precision, and sign accuracy on acts that are truly exogenous."
#| eval: !expr have_il
tt_gate_metrics(
  latest(il$s3_manual_bias_corrected, "c2b"),
  targets = c(precision = 0.85, sign_accuracy = 0.90),
  labels  = c(precision = "Exogenous precision", recall = "Exogenous recall")
)
```

Exogenous precision is the quantity a downstream user depends on: it is the
share of the acts we hand over as exogenous that really are. It lands just below
its gate, on an effective sample small enough that the interval comfortably
contains the target — a result we report as borderline rather than as a pass.
Sign accuracy on truly-exogenous acts clears its gate. These two corrected
numbers, and the correction procedure that produces them, are the central
methodological output of the US benchmark.

## Malaysia {#sec-my-validation}

**Do the same codebooks, frozen, still work in a country with no labelled
data?**

This is the question the project exists to answer, and it cannot be answered the
way the last section was. There is no Malaysian equivalent of the Romer–Romer
act list, so there is nothing to compute an accuracy against. That is not a
temporary inconvenience: if a validated narrative series already existed for a
country, the pipeline would have little to offer it. The absence of ground truth
is the condition the method has to work under, not an obstacle on the way to
working under better conditions.

What replaces an accuracy rate is four separate lines of evidence, none
sufficient alone. The first is an **internal consistency test** that needs no
ground truth at all, exploiting a property of the corpus to check that the
pipeline reasons about fiscal substance rather than about English. The second is
a set of **worked disagreement cases**, printed in full so a reader can judge
the pipeline's reasoning directly rather than through a summary statistic. The
third is the **@vegh_how_2015 statutory-rate panel**, which is an external
referent for one narrow part of the output and no part of the motivation. The
fourth is
**expert agreement**, which is the only external check on the judgement itself
and is currently being collected.

The third deserves a word, because it qualifies the claim just made. No external
referent exists for *motivation*, which is the quantity the method is about. One
does exist for the *rate path*, which is a by-product the method has to get right
along the way, and it turns out to be checkable against published data for two
of the five instruments. That is a narrower thing than validation of the
pipeline, and it is more than nothing.

A fourth line exists in principle and we cannot yet use it.
@aruoba_identifying_nodate, identifying monetary policy shocks from narrative
records, validate not against an existing shock list but against forecast-error
predictability and against whether the recovered series produces
theory-consistent impulse responses. Neither test needs labels, which makes the
strategy the natural one for a country like Malaysia. Both need magnitude and
quarterly timing, which @tbl-coverage records as unbuilt. So the strategy is
available to this method and not yet to this dataset, and it is a further reason
@sec-coverage treats magnitude as a binding gap rather than a refinement.

### Cross-language (EN/BM) consistency

The test exploits an accident of the corpus. Ten Economic Reports were published
in both English and Bahasa Malaysia: the same fiscal year, the same underlying
acts, two languages. A pipeline that reasons about fiscal substance should
return comparable inventories from both. A pipeline that has learned English
surface patterns, or that is retrieving remembered facts cued by English
phrasing, should not.

This is a sharper instrument here than it would be in most applications, for two
reasons. Multilingual evaluations consistently find degradation on
lower-resource languages relative to English [@ahuja_mega_2023;
@rathje_gpt_2024], so equivalence is a real hypothesis and not a foregone
conclusion. And the paired-document design isolates the language effect from
item difficulty, since both versions describe identical events, which is exactly
the design the multilingual benchmarks adopt to separate the two
[@singh_indicgenbench_2024; @xuan_mmlu-prox_2025].

Because no ground truth exists on either side, the comparison is distributional:
measure counts, act counts, and the motivation label mix, rather than a matched
act-by-act agreement rate. That is a weaker test than matched agreement would
be, and we prefer it to the alternative, which would be to build the matching
with a third model and then validate two unvalidated things against each other.
Comparable distributions are consistent with language-robust reasoning without
proving it; sharply different ones would have been decisive evidence against.

```{r}
#| label: pending-malay
#| eval: !expr "!have_malay"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe Malaysia EN/BM consistency targets are unbuilt, so",
    "the cross-language figures cannot render. See malay_consistency.qmd.\n:::\n")
```

@fig-malay-c1 takes the first stage of the pipeline: how many distinct fiscal
measures the codebook surfaces from each report year, read in each language. The
question it answers is whether the scan sees the same amount of fiscal content
in a document regardless of the language it is written in.

```{r}
#| label: fig-malay-c1
#| fig-cap: "**Measure identification across languages.** Distinct fiscal measures surfaced per Economic Report year, by language. Comparable counts indicate the extraction step is language-robust."
#| fig-height: 4
#| eval: !expr "nrow(malay_er_c1_measures) > 0"
p <- plot_malay_c1_comparability(malay_er_c1_measures)
if (is.null(p)) cat("No C1 measure data available.\n") else p
```

@fig-malay-timeline carries the test through to the end of the chain, comparing
the act inventories the two language runs produce and the motivation labels
attached to them. This is the more demanding comparison, because it accumulates
whatever divergence the earlier stages introduced.

```{r}
#| label: fig-malay-timeline
#| fig-cap: "**Act inventory across languages.** Acts by year and motivation, faceted by language. Comparable structure across English and Bahasa Malaysia is the signal that the codebooks reason consistently across languages."
#| fig-height: 5
#| eval: !expr have_malay
p <- plot_malay_act_timeline(malay_inv, timing = "act_name")
if (is.null(p)) cat("No act inventory available.\n") else p
```

### Worked acts (disagreement cases)

Aggregate evidence has a limit: it can show that two runs agree without showing
whether either is reasoning well. So we do what @romer2010 do in their own
Exhibits 1 and 2, and print acts in full for the reader to judge.

The three below are chosen adversarially. Each is a case where the pipeline's
two independent reads, the preliminary narrative read formed during act
identification and the classification codebook's verdict, **disagree** about
exogeneity. These are not the pipeline at its best; they are the cases where it
cannot settle the question itself, and they are the ones an expert has to
adjudicate. Printing them is the point, because a method that hides its
disagreements is asking to be trusted rather than checked.

Read them with @sec-defect in mind. Two of the four labels that section
identifies as built on the wrong act's evidence are corporate-tax rows, and the
exhibits below are drawn from the June freeze, so a disagreement shown here may
be an artifact of that defect rather than a genuine interpretive tension. That
is itself instructive: without the evidence chain printed alongside the label, a
reader would have no way to tell the two apart.

```{r}
#| label: exhibit-cit
#| echo: false
#| output: asis
#| eval: !expr have_tax
tax_exhibit(
  tax_shocks, "MY-CIT-03",
  exhibit_label = "Exhibit 2",
  note = "The narrative pass read this rate cut as exogenous; the classifier read the surrounding crisis framing and called it countercyclical. Both readings are defensible from the corpus, which is what makes it an adjudication case."
)
```

```{r}
#| label: exhibit-spending
#| echo: false
#| output: asis
#| eval: !expr have_spending
spending_exhibit(
  spending_shocks, "MY-SPEND-04",
  exhibit_label = "Exhibit 3",
  note = "The narrative pass applied the two-condition spending screen and read the subsidy restructuring as endogenous; the classifier read it as deficit-driven, hence exogenous. The disagreement is about whether fiscal-pressure motives count as cyclical."
)
```

```{r}
#| label: exhibit-incentive
#| echo: false
#| output: asis
#| eval: !expr have_incentives
incentive_exhibit(
  incentive_shocks, "MY-INCENT-12",
  exhibit_label = "Exhibit 4",
  note = "An investment incentive announced inside a pandemic recovery package. The narrative pass weighted the long-run investment-promotion objective, the classifier weighted the crisis vehicle it arrived in."
)
```

### The Végh–Vuletin statutory-rate panel {#sec-rate-benchmark}

This section opened by saying no external referent exists for Malaysia. For
motivation that is true and will stay true. For the *rate path* it is not: the
statutory top rates the narrative pass reconstructs, shown as @fig-tax-rate-paths,
also appear in the hand-built cross-country rate panel @vegh_how_2015 assemble,
which covers Malaysia over this period. That comparison is the most direct
external validation the magnitude side admits. It is run here.

The benchmark is well chosen for more than convenience. It is the same series
whose procyclicality finding @sec-usecase takes as its starting point, so the
check and the use case are run against one another's data: we validate our rate
paths against the panel, then show what the panel's own cyclicality statistic
cannot resolve on the country it covers.

The check is narrow and worth stating precisely. It scores whether we recovered
the right rate, in the right year, on the two instruments that have a headline
rate. It says nothing about motivation, about spending, or about incentives,
because a rate panel carries a number per year and no record of why it moved.

```{r}
#| label: pending-concord
#| eval: !expr "!have_concord"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe rate-path concordance target is unbuilt, so",
    "the external benchmark cannot render. Run `tar_make()` on",
    "`rate_path_concordance`.\n:::\n")
```

```{r}
#| label: fig-rate-validation
#| fig-cap: "**Narrative rate paths against the Végh–Vuletin panel.** The thick grey line is their published series [@vegh_how_2015]; the thin dark line with markers is the path implied by the acts this pipeline identified. Exact agreement reads as the dark line sitting inside the grey one. The dark line starts where the corpus becomes dense, not where the rate series starts."
#| fig-width: 6.5
#| fig-height: 3.6
#| eval: !expr have_concord
p <- plot_rate_path_validation(rate_concord)
if (is.null(p)) cat("No concordance data available.\n") else p
```

@tbl-rate-concordance scores it event by event. Three patterns, and each says
something different.

**Where the corpus is dense, the paths agree exactly.** Every corporate rate
change the panel records from 1994 onward is matched by an act we identified,
at the same year and with the same change in percentage points. The personal
rate is comparable from 1999. This is a stronger result than a correlation
would be, because the match is on the level and the year, not on co-movement.

**Every miss predates the corpus.** The unmatched panel changes are all early:
the corpus holds a single document in each of several years before the mid-1990s
and only becomes dense afterwards, as @fig-corpus-coverage shows. That is a
statement about what we gave the pipeline to read, not about how it read. It is
also the cleanest available illustration of the point @sec-agentic makes about
recall: the ceiling is set by acquisition.

**Every extra is something the panel cannot represent.** The acts we identify
that the panel does not are not false positives. A one-off corporate surcharge,
a change to the imputation system, a cut to a middle band, a restructure that
leaves the top rate untouched: each moves real liabilities and none of them
moves a headline top rate, so a rate panel is blind to them by construction.
This is the asymmetry that motivates the whole exercise. The rate panel is the
better instrument for the narrow question of what the top rate was; the
narrative inventory is the only one of the two that can see the rest of the tax
policy.

```{r}
#| label: tbl-rate-concordance
#| tbl-cap: "**Event-level concordance.** Every rate change recorded by either source; the two numeric columns are the change in percentage points. Matched rows agree on year and magnitude. The timing row is one panel step the narrative pass resolves into two annual steps, agreeing on the level by the end of it."
#| eval: !expr have_concord
tb <- rate_concordance_table(rate_concord, detail = TRUE)
if (is.null(tb)) cat("No concordance data available.\n") else tb
```

Two limits on how far this evidence reaches. It covers corporate and personal
income tax only: the panel carries no consumption-tax rate for Malaysia, so the
GST introduction and repeal, which are the largest tax events in the inventory,
get no external check at all. And a matching rate path is consistent with, but
does not establish, a correct reading of the document, since a model could in
principle recover a well-known rate history without reading anything. The
cross-language test above is the evidence that bears on that, and it is
separate.

### The expert-agreement protocol

The external check is expert review, and its design has to answer a question
that does not arise when ground truth exists: what does agreement with a human
*mean*, when the human is not an oracle either?

The protocol asks regional fiscal-policy experts to code sampled acts
independently, using the same codebook definitions the model was given, and then
compares the two codings. Two headline targets are pre-registered: at least 80%
agreement on measure identification, and at least 70% on motivation. Those
thresholds are lower than an accuracy target would be, deliberately, because
they are **parity** targets rather than correctness targets. Experts disagree
with each other about motivation, sometimes substantially, and a model that
agrees with one expert as often as a second expert would is doing as well as the
task allows. Following @asirvatham_gpt_2026, the right comparison is therefore
model-versus-consensus against leave-one-out human-versus-consensus, with
expert-versus-expert agreement reported as the ceiling. Without that ceiling, a
70% figure is uninterpretable: it could be a weak model or a genuinely
contested construct, and only the human baseline separates them.

Three further features of the design are worth stating. Review is
**risk-stratified** rather than uniform: acts predicted exogenous in a recession
year are reviewed in full, because that is the cell where an error does the most
damage to any downstream estimate, as are acts carrying either of the two labels
the error analysis found most confusable. The remainder is spot-checked. The
protocol also asks a question the dataset cannot ask itself, namely which
significant acts are **missing**, which is the recall assessment @sec-agentic
defers to. And it separates genuine model error from irreducible ambiguity:
where two experts disagree with each other, the case is recorded as ambiguous
rather than scored against the model.

The first of those strata is more load-bearing than it looks. Acts predicted
exogenous in a downturn are precisely the cell the decomposition in
@sec-usecase rests on, and @sec-usecase-belief reports how few of them would
have to be relabelled to overturn it. So the review burden that section implies
is not open-ended: it is a specific, countable set of acts, and the protocol was
already routing them to full review before anyone had computed the number.

One gap in the protocol as currently written should be closed before it is run,
because @sec-usecase depends on it. The protocol selects acts for review by
purposive triage — every act predicted exogenous in a recession year, every act
carrying one of the two confusable motivation labels, and a convenience
spot-check of the rest. That is the right design for *finding* errors and the
wrong design for *correcting* an estimate: debiasing a downstream regression
against a validation set requires that set to be a probability sample, with
known inclusion probabilities [@angelopoulos_prediction-powered_2023;
@egami_using_2023]. The two purposes need either two samples or one design that
serves both.

# What is validated, and what is not {#sec-coverage}

The pipeline's stages rest on very different evidence, and the difference is
easy to lose once the output is a single tidy table. @fig-pipeline-unvalidated
inverts @fig-pipeline-validated: it lights the stages the US benchmark does
*not* cover, and @tbl-coverage states what each of them rests on instead.

```{r}
#| label: fig-pipeline-unvalidated
#| fig-cap: "**What the US benchmark does not cover.** The complement of @fig-pipeline-validated. These stages are attested by construction, by audit, or not yet at all."
#| fig-width: 6.5
#| fig-height: 4.2
plot_pipeline_flow("unvalidated", scale = 1.4)
```

```{r}
#| label: tbl-coverage
#| tbl-cap: "**Validation coverage.** What each stage of the pipeline rests on, and what is still missing. Nothing in the deliverable's exogeneity column is an adjudicated finding."
tibble::tribble(
  ~Stage,                    ~`Evidence we have`,                                                         ~`Evidence we lack`,
  "Corpus construction",     "Extraction QA; page and chunk counts reproducible from the pipeline",        "No recall audit over the universe of documents that exist",
  "Measure identification",  "Full behavioural, zero-shot and error-analysis record on 44 US acts",        "Deployment-corpus performance; the top-ranked-measure filter discards 84.5% of mentions",
  "Act identification",      "Per-instrument recall scorecards, frozen; human stamp on every event",       "No gold standard, so no recall or precision rate — only an audit of what was searched",
  "Act consolidation",       "Automated clustering built and scored against 49 US reference acts",         "Rejected for deployment, not validated for it",
  "Motivation and sign",     "Full record on US acts; label-noise-corrected gates",                        "Validated on single-act chunks, deployed on multi-measure speeches; no signal-stripping test of directness",
  "Magnitude",               "Traceable to a quoted rate or amount; CIT and PIT top-rate paths reproduce the Végh and Vuletin (2015) statutory-rate panel exactly over the corpus window", "No codebook, no behavioural tests; the external check reaches only headline top rates, not surcharges, band changes, consumption taxes or spending amounts",
  "Timing",                  "Announced and effective year, from the document; effective years match the Végh and Vuletin panel where both cover",  "No quarterly timing; the annual grain limits the downstream use case",
  "Exogeneity of the output", "Two independent preliminary reads per act, and their disagreements flagged", "Expert adjudication not yet run; the column is a model read"
) |>
  tt(width = c(0.20, 0.42, 0.38)) |>
  tt_theme_report() |>
  style_tt(j = 1:3, align = "l")
```

Three entries deserve emphasis because they are the ones most likely to be
over-read.

**The exogeneity column is an input, not a finding.** Every act carries two
preliminary reads — one from the narrative identification pass, one from the
validated classification codebook — and where they disagree the act is flagged
rather than resolved. Those disagreements are the worked exhibits in
@sec-my-validation. Adjudicating them is the expert's job and it has not
happened yet.

**Magnitude has a partial record and timing has none.** Both are extracted
agentically alongside each act, and the codebooks that would validate them are
not built. What magnitude now has is a scope-limited external check:
@sec-rate-benchmark shows the corporate and personal top-rate paths reproducing
the @vegh_how_2015 statutory-rate panel exactly wherever the corpus is dense. That
covers the rate-bearing acts on two instruments. It does not cover the
consumption taxes, the surcharges and band changes a top-rate series cannot
represent, or any spending amount, and those are most of the magnitude column.
Quarterly timing has no check of any kind.

**The deployment distribution was never validated.** This is the general form of
@sec-defect and the most transferable lesson here. The codebooks were
validated on US chunks that discuss one act each and deployed on Malaysian
budget speeches that discuss many. That gap produced a real defect that no US
test could have caught, and it is a gap any cross-country transfer of this kind
will have in some form.

The same accounting is worth doing for the closest related work, because the
comparison in @tbl-related-work was about design and this one is about evidence.
@tbl-rival-evidence sets it out. The instructive column is the last: on magnitude,
the dimension a multiplier estimate needs most, the four methods fail differently
rather than uniformly. @fritsch_high-frequency_nodate do recover percent-of-GDP
magnitudes and validate them against the Romer and Romer series, but for one
country and for an expectations series rather than realised liabilities.
@das_ai_2026 decline magnitude outright and are explicit about it.
@bhasin_identifying_2026 size their episodes and, where they diverge from the
benchmark, argue the benchmark is the inflated one rather than conceding an
accuracy limit, which is a defensible position and an unresolved one. Ours are
extracted from the announcing document and traceable to a quoted rate, and carry
no validation record at all. No method in this literature yet has a magnitude
series that is both externally validated and portable.

```{r}
#| label: tbl-rival-evidence
#| tbl-cap: "**What the evidence in each method supports.** The complement of @tbl-related-work: not what each design does, but what its validation licenses and what it does not. Read alongside @tbl-coverage, which does the same job stage by stage for this paper."
tibble::tribble(
  ~Method,             ~`Validation evidence`,                                                        ~`Scored against`,                                    ~`Not supported by that evidence`,
  "Das et al.",        "Concordance on overlapping country-years",                                    "Action-based consolidation datasets",                 "Magnitudes; countries with no benchmark series",
  "Bhasin et al.",     "Divergence from benchmark decomposed into sizing and event selection",         "OECD action-based benchmark",                         "Transfer beyond the OECD; which of the two series is closer to the truth",
  "Fritsch et al.",    "High correlation with the narrative tax-shock series",                         "Romer and Romer, United States",                      "Realised liabilities (the series is expectations); any non-US setting",
  "This paper",        "Behavioural, zero-shot and manual error analysis on 44 US acts; cross-language consistency; top-rate concordance with the Végh and Vuletin rate panel", "Romer and Romer labels; Végh and Vuletin Malaysian CIT and PIT statutory rates; no Malaysian motivation ground truth", "Magnitudes beyond the headline top rate; quarterly timing; the exogeneity column pending expert adjudication"
) |>
  tt(width = c(0.15, 0.29, 0.22, 0.34)) |>
  tt_theme_report() |>
  style_tt(j = 1:4, align = "l")
```

# A first use case: procyclicality decomposition {#sec-usecase}

This section is a demonstration, not an inferential claim. It exists to show
what a motivation-coded inventory lets a researcher ask that a statutory-rate
series does not, and the question is the one @kaminsky_when_2004 opened and
@frankel_graduation_2013 sharpened: when a country's fiscal record looks
procyclical, is that deliberate response to the cycle, or the accumulation of
measures taken for unrelated reasons that happened to land badly? A
motivation-coded inventory is the only kind of dataset that can tell those
apart, which is why @vegh_how_2015 could document the pattern and explicitly
decline to attribute it.

```{r}
#| label: pending-procyc
#| eval: !expr "!have_procyc"
#| echo: false
#| output: asis
cat("::: {.callout-warning}\nThe procyclicality targets are unbuilt, so the",
    "decomposition exhibits cannot render. Run `tar_make()` on `macro_panel`,",
    "`my_cycle` and `procyclicality_decomposition`.\n:::\n")
```

## What the statutory-rate measure can see {#sec-usecase-vv}

The standard measure of tax-policy cyclicality correlates the change in a
statutory rate with the cyclical component of output. Run on Malaysia it
returns nothing: `r if (have_procyc) sprintf("%.2f for the corporate rate and %.2f for the personal rate, neither remotely distinguishable from zero", vv_cross$corr_cit_cycle[vv_cross$ccode == "MYS"], vv_cross$corr_pit_cycle[vv_cross$ccode == "MYS"]) else "a correlation indistinguishable from zero"`.

The reason is not that Malaysian tax policy is acyclical. It is that a headline
statutory rate moves rarely, and a correlation computed on
`r if (have_procyc) vv_cross$n_cit_changes[vv_cross$ccode == "MYS"] else "a handful of"`
corporate rate changes in
`r if (have_procyc) vv_cross$n_years[vv_cross$ccode == "MYS"] else "forty-odd"`
years has almost no information in it. This is not a Malaysian peculiarity.
@fig-vv-context computes the same statistic for every country in the panel with
enough data, and the median one changed its corporate rate
`r if (have_procyc) stats::median(vv_cross$n_cit_changes[!is.na(vv_cross$corr_cit_cycle)]) else "a handful of"`
times in four decades. The measure is thin nearly everywhere, and Malaysia sits
close to the middle of the resulting spread.

```{r}
#| label: fig-vv-context
#| fig-cap: "**The statutory-rate measure is thin nearly everywhere.** Distribution across countries of the correlation between the annual change in the corporate statutory rate and the output gap. Malaysia is marked. The spread is centred near zero and is dominated by how few times a headline rate moves."
#| fig-width: 6.5
#| fig-height: 3.0
#| eval: !expr have_procyc
p <- plot_vv_context(vv_cross, highlight = "MYS")
if (is.null(p)) cat("No cross-country panel available.\n") else p
```

Two things follow. The measure is under-powered, which more data would fix. And
it is silent on attribution, which more data would not: a correlation between
rate changes and the cycle cannot distinguish a rate cut enacted *because*
output was strong from one enacted for unrelated reasons that happened to arrive
in a boom. That second limitation is the one the narrative inventory addresses.

## The fiscal record against the cycle {#sec-usecase-overlay}

@fig-procyclicality-overlay lays the inventory over Malaysia's business cycle.
The upper panel is the output gap; the lower panel is one mark per fiscal
action, placed at the year it took effect, above the line if it was
expansionary and below if contractionary, coloured by whether the codebook
judged its motivation exogenous.

Note that the signing axis here is the **fiscal impulse**, not the direction of
the change in liabilities. A spending increase and a tax cut are both
expansionary; the inventory figures earlier in the paper sign by liability
direction, which is the right axis for describing a tax system and the wrong one
for describing a stance.

```{r}
#| label: fig-procyclicality-overlay
#| fig-cap: "**The Malaysian fiscal record against the cycle.** Shaded bands are below-trend years, drawn behind both panels so a reader can drop from a downturn straight into the actions taken during it. Marks are stacked within a year rather than jittered, since the year is the claim."
#| fig-width: 6.5
#| fig-height: 4.8
#| eval: !expr have_procyc
p <- plot_procyclicality_overlay(my_cycle, procyc_events)
if (is.null(p)) cat("No procyclicality data available.\n") else p
```

## The decomposition {#sec-usecase-decomposition}

Crossing the impulse with the cycle state gives each action a **stance**. An
action is procyclical when it pushes the way the cycle is already going:
expansionary in a boom, contractionary in a bust. Crossing stance with the
motivation label gives @tbl-procyclicality, and with it the quantity
@vegh_how_2015 named as missing.

Of the `r if (have_procyc) pc_n else "N"` actions that carry both a stance and a
motivation, `r if (have_procyc) pc_proc else "M"` look procyclical on timing.
**`r if (have_procyc) pc_exo else "K"` of them carried a motivation unrelated to
the cycle.** Malaysia's procyclical-looking fiscal record is, on this reading,
overwhelmingly not a record of deliberate procyclical policy. It is long-run and
deficit-driven measures that happened to land against the cycle.

```{r}
#| label: tbl-procyclicality
#| tbl-cap: "**The decomposition.** Upper block: stance against motivation, all components. Lower block: the procyclical row under alternative cycle definitions, event timing, and component scope. The reported p is Fisher's exact test, which for a 2×2 with both margins fixed *is* the permutation test."
#| eval: !expr have_procyc
tb <- procyclicality_table(procyc_decomp)
if (is.null(tb)) cat("No decomposition available.\n") else tb
```

@fig-procyclicality-decomposition shows where each column's mass comes from, and
makes the split legible: the countercyclical column is dominated by acts the
codebook labelled countercyclical, which is close to a definitional result. The
procyclical column is long-run and deficit-driven measures, which is not.

```{r}
#| label: fig-procyclicality-decomposition
#| fig-cap: "**Stance by motivation.** The countercyclical column is mostly acts labelled countercyclical, which is near-definitional. The procyclical column is not, and it is the one the decomposition is about."
#| fig-width: 6.5
#| fig-height: 3.2
#| eval: !expr have_procyc
p <- plot_procyclicality_decomposition(procyc_events)
if (is.null(p)) cat("No decomposition available.\n") else p
```

@tbl-vv-comparison sets the two measures side by side. The point is not that the
narrative decomposition is more precise than the correlation. It is that they
answer different questions, and only one of them is the question about
attribution.

```{r}
#| label: tbl-vv-comparison
#| tbl-cap: "**What each measure supports.** The @vegh_how_2015 statutory-rate statistic and the narrative decomposition, on the same country and the same years."
#| eval: !expr have_procyc
tb <- vv_comparison_table(procyc_decomp, vv_cross, iso = "MYS")
if (is.null(tb)) cat("No comparison available.\n") else tb
```

## How much to believe it {#sec-usecase-belief}

Three reasons for caution, in descending order of how much they should worry a
reader.

**The labels are model output, and a p-value cannot speak to that.** At this
sample size the binding uncertainty is not sampling noise. So the quantity we
report instead is a **fragility index**: the number of exogenous acts in the
procyclical cell that would have to be relabelled endogenous before the
association stopped being significant. It is
`r if (have_procyc) pc_frag else "K"`, out of
`r if (have_procyc) pc_exo else "K"`. Set against that, the pipeline already
flags `r if (have_procyc) pc_flag else "some"` of the procyclical acts as cases
where its two independent reads disagree — below the breaking point, but not by
much. This is also the concrete cost of the defect in @sec-defect: four of the
51 motivation labels are known to be wrong in these exhibits, the same order of
magnitude as the fragility index, which is why the correction is a prerequisite
for the estimate rather than a tidying-up exercise.

**The result thins when spending is removed.** The last row of
@tbl-procyclicality drops the spending component and the exact test stops being
significant. That is expected and worth stating plainly: the stimulus packages
are both endogenous and countercyclical, so they supply most of the contrast.
What survives without them is the point estimate rather than the test
(`r if (have_procyc) sprintf("%d of the %d", pc_nospend$proc_exo, pc_nospend$proc_total) else "most"`
procyclical tax and incentive actions still carry an exogenous motivation), and
at this sample size a point estimate is the honest object anyway.

**Generated regressors.** Once an estimate is built on the labels, valid
inference is a measurement-error problem, not a classification-accuracy one
[@ludwig_large_2025; @battaglia_inference_2024]. The bias equals the correlation
of label error with the regressor, and the dangerous regressor is the business
cycle, the exact axis exogeneity turns on. So a downstream estimate must either
show mislabelling is uncorrelated with the cycle or scope the claim accordingly,
and would properly debias against a probability-sampled expert validation set
[@angelopoulos_prediction-powered_2023; @egami_using_2023]. The expert protocol
in @sec-my-validation already routes every act predicted exogenous in a
downturn to full review, which is exactly the fragility set; @sec-coverage notes
that it needs a probability sample as well as a purposive one.

One supporting diagnostic. Shifting the cycle series circularly through every
lag, which preserves its serial correlation and the events' clustering in time,
`r if (have_procyc) sprintf("puts the observed association ahead of all %d shifted alternatives", procyc_decomp$phase_perm$n_shifts) else "leaves the observed association ahead of the shifted alternatives"`.
Only as many distinct shifts exist as there are years, so this is a coarse check
rather than a second test, and it is reported as one.

**What is deliberately not here.** The natural next estimator is a local
projection of output on the signed exogenous events [@jorda_estimation_2005].
We do not run one. It would need the magnitude column, and @tbl-coverage records
that magnitudes carry no validation record beyond the headline statutory rate;
an impulse response built on an unvalidated regressor would look like the
paper's main result while being its weakest. It is the first thing a validated
magnitude codebook would unlock, and it is the reason @sec-coverage treats
magnitude as a binding gap rather than a refinement.

# Conclusion

The introduction set out four questions. This paper settles two of them in part,
leaves one alone, and delivers the fourth for one country.

On @cnj-1: **an LLM can reproduce narrative identification against Romer and
Romer's own labels for three of the five quantities the method
recovers.** Which passages describe a fiscal measure, what motivated
each act and whether that motivation was independent of current output, and the
direction of the change: each carries a full behavioural, zero-shot and
error-analysis record. Magnitude in percent of GDP and quarterly timing do not,
and we have been careful not to let the validated parts lend credibility to the
unvalidated ones. The most useful thing we learned here was not a score. It was
that a class of apparent model errors turned out to be errors in the benchmark,
which is a finding about evaluating codebooks against projected human labels
that any project at this scale will run into.

On @cnj-2: **the evidence that frozen codebooks transfer to a country with no
labelled data is favourable, and it is not an accuracy rate.** It cannot
be, because the absence of ground truth is the condition the method exists to
work under rather than an obstacle on the way to better conditions. What we have
instead is a bilingual corpus returning comparable inventories in both
languages, a set of disagreement cases printed in full for the reader to judge,
and statutory rate paths that reproduce the @vegh_how_2015 cross-country rate
panel exactly wherever the corpus is dense. None of the three is sufficient, and the one that
would be, expert adjudication, is running now. There is also a result here that
cuts against the paper's own interest: the ablation in @sec-c2 suggests the
classifier is running substantially on a pretrained prior rather than on our
codebook, which is why transfer is plausible and equally why demonstrating it on
US-adjacent material would prove little.

On @cnj-3: **we have nothing to report on comparable cross-country multipliers,
and the reason is now specific.** It needs magnitude, quarterly timing, and more
than one country. @sec-usecase declines to run the local projection that would be
the natural first step, because the magnitude column has no validation record
beyond the headline statutory rate. That is not a caution about ambition; it is
the concrete price of the next codebook.

On @cnj-4: **the Malaysia inventory is a dataset a local researcher can use, and
its provenance chain is the reason.** Every event traces to a quoted passage in a named primary-source
document. That is what let a reviewer's objection to two corporate-tax rows
surface a machinery defect that had silently mislabelled events across all three
components (@sec-defect), and it is what makes @sec-usecase's decomposition
checkable rather than merely reported. A dataset whose errors can be found is
worth more than one whose errors are invisible.

The use case in @sec-usecase is the clearest statement of what the exercise
buys. Malaysia's statutory-rate cyclicality is indistinguishable from zero, and
so is the median country's, because a headline rate moves a handful of times in
four decades and a correlation on those few movements cannot separate policy
responding to the cycle from policy that merely arrived during one. The
narrative inventory answers that second question directly, and its answer for
Malaysia is that the procyclical-looking record is overwhelmingly *not*
deliberate procyclical policy. We report that as a demonstration of what the
data supports, hedged by a fragility index rather than a p-value, because the
labels are model output and four relabellings would overturn it.

Which is the honest summary of the whole paper. The claim is not that the
pipeline replaces the expert. It is that it changes what the expert spends
months doing, and that the resulting dataset carries, alongside every number, an
account of how much weight that number will bear.

# Limitations and roadmap

@tbl-coverage states what is and is not evidenced. Three further limitations are
about the design rather than its current state.

**Transfer is a claim to defend, not to assume.** Fixed text measures notoriously
fail to transfer across domains [@gentzkow_text_2019], and cross-codebook
fine-tuning overfits [@halterman2025], so frozen country-agnostic codebooks have
to be shown to survive contact with a new country rather than assumed to. The
cross-language test and the regional extension are what put that claim at risk.
The finding in @sec-c2 that the classifier appears to be running on a
pretrained prior cuts both ways here: it is why transfer is plausible, and it is
why demonstrating transfer on US-adjacent material proves little.

**One country is one observation.** Malaysia was chosen for a stable political
window, a bilingual archive, and a documented fiscal record. Indonesia, Thailand,
the Philippines, and Vietnam follow the same pipeline, and the parts of the
method that turn out to be Malaysia-specific will only be visible then.

**The human stamp is load-bearing and does not scale linearly.** Act
identification and consolidation require a researcher to confirm the events and
the grain, per instrument, per country. The pipeline reduces months of work to
weeks; it does not remove the expert, and the claim in this paper is explicitly
about assistance, not automation.

<!-- # References {.unnumbered}

::: {#refs}
::: -->

::: {.content-visible when-format="typst"}
{{< appendix >}}
:::

::: {.content-hidden when-format="typst"}
# Appendix {.unnumbered}
:::

# Recall audits {#sec-appendix-recall}

The per-instrument recall audits, frozen with their datasets. As @sec-agentic
says, these record what the identification pass searched for and what it
returned; they are a transparency record, not a recall measurement, and the
recall assessment that matters is the expert review described in
@sec-my-validation. Rows naming a recall miss record an event the codebook scan
did not surface and the direct-reading pass recovered. Outcome text is truncated
for display.

```{r}
#| label: tbl-recall-cit
#| tbl-cap: "**Recall audit, corporate income tax.**"
#| eval: !expr have_tax
tb <- recall_scorecard_table(tax_shocks, instrument = "Corporate income tax")
if (is.null(tb)) cat("No recall scorecard available.\n") else tb
```

```{r}
#| label: tbl-recall-pit
#| tbl-cap: "**Recall audit, personal income tax.**"
#| eval: !expr have_tax
recall_scorecard_table(tax_shocks, instrument = "Personal income tax")
```

```{r}
#| label: tbl-recall-consumption
#| tbl-cap: "**Recall audit, consumption tax (GST/SST).**"
#| eval: !expr have_tax
recall_scorecard_table(tax_shocks, instrument = "Consumption tax (GST/SST)")
```

```{r}
#| label: tbl-recall-spending
#| tbl-cap: "**Recall audit, spending.**"
#| eval: !expr have_spending
recall_scorecard_table(spending = spending_shocks, instrument = "Spending")
```

```{r}
#| label: tbl-recall-incentive
#| tbl-cap: "**Recall audit, tax incentives.**"
#| eval: !expr have_incentives
recall_scorecard_table(incentive = incentive_shocks, instrument = "Tax incentives")
```

# How the US gold labels were built {#sec-appendix-ika}

Romer and Romer publish act-level classifications. The measure-identification
codebook reads chunks of documents. Evaluating one against the other requires
deciding, for every chunk in the US corpus, whether it is a positive: whether a
known act is substantively discussed in it. There is no published answer to that
question, so it has to be constructed, and this appendix records how, because
@sec-c1-manual shows the construction is the largest single source of error in
the US benchmark.

Two obvious approaches fail. Hand-coding twelve thousand chunks is infeasible.
Exact string matching of act names fails for a more interesting reason:
government PDFs from the 1940s to
the 1990s were digitised by OCR, which routinely breaks an act name across a
line, so "Revenue Act of 1962" appears in the extracted text as `Revenue Act
of\n1962` and matches nothing. Normalising whitespace recovers a large number of
matches, and for several acts it doubles the count, but it does not rescue acts
that are never named in the form the reference list uses.

The construction therefore assigns chunks to two tiers of positive, plus a
negative pool.

**Tier 1** chunks contain an act's own narrative passage verbatim. Romer and
Romer publish, with each act, the passages of presidential messages and
committee testimony that document its motivation. Searching the corpus for the
first 100 characters of each passage, with a 60-character fallback, locates the
chunk that passage came from. These are as close to certain positives as the
construction reaches, and there are few of them, because one act contributes
only as many Tier 1 chunks as it has published passages.

**Tier 2** chunks name a known act without containing its passage. Because act
names are unreliable strings, three matching strategies are used: whitespace-
normalised substring search; decomposition of compound names into searchable
subcomponents, which is what makes "Revenue Act of 1962" findable inside
"Expiration of Excess Profits Tax and of Temporary Income Tax Increases"; and a
small set of hand-written co-occurrence rules for acts that no substring reaches.
Deliberately, no year window and no keyword filter is applied. The tier is built
to maximise recall on the theory that downstream stages can filter.

**Negatives** are every remaining chunk, scored with a continuous relevance-key
density rather than excluded by a threshold. An earlier design excluded any
chunk containing fiscal vocabulary, which left 35 clean negatives out of
thousands and made precision essentially untestable.

The resulting split is heavily imbalanced: on the US corpus, Tier 1 accounts for
0.6% of chunks, Tier 2 for 12.2%, and negatives for 90.4%.

The weakness is in Tier 2, and it is structural rather than a tuning problem. An
act *name* appears in many places its substance does not: book indexes, tables
of contents, glossaries, running heads, and administrative appendices listing
every statute touched. Those chunks enter the evaluation carrying a positive
gold label and containing no fiscal content at all. When the codebook correctly
declines to call a table of contents a fiscal measure, the evaluation records a
false negative. That is the mechanism behind the Category B count in
@sec-c1-manual, and it is why the corrected gate, rather than the raw recall
figure, is the number to read.

# Codebooks {#sec-appendix-codebooks}

The codebooks are reproduced verbatim, as the model would read them.

## Measure identification

```{r}
#| label: appendix-c1
#| echo: false
#| output: asis
codebook_verbatim_md(here("prompts/c1_measure_id.yml"))
```

## Motivation evidence extraction

```{r}
#| label: appendix-c2a
#| echo: false
#| output: asis
codebook_verbatim_md(here("prompts/c2a_extraction.yml"))
```

## Motivation classification

```{r}
#| label: appendix-c2b
#| echo: false
#| output: asis
codebook_verbatim_md(here("prompts/c2b_classification.yml"))
```

