---
title: "C0 Act Aggregator — Method Comparison"
subtitle: "Aggregating C1 measure names into act-level canonical clusters"
date: today
date-format: long
execute:
  cache: true
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, stringdist, patchwork,
               scales, forcats, tidytext, uwot)

here::i_am("notebooks/c0_aggregator.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
source(here("R/c0_aggregator.R"))

tar_load(c(
  c0_us_measure_pool,
  c0_eval_gold_pairs,
  c0_jw_clusters,
  c0_jw_metrics,
  c0_us_embeddings,
  c0_us_gold_embeddings,
  c0_hdbscan_clusters,
  c0_hdbscan_metrics,
  c0_umap_grid,
  c0_hdbscan_umap_clusters,
  c0_hdbscan_umap_metrics,
  c0_f16_quantization_probe,
  c0_f16_quantization_probe_tier1,
  c0_fp32_reference_probe,
  c0_fp32_reference_probe_tier1
))

# Methods 2 + 3 (HDBSCAN, unblocked + year-blocked) are wired below.
# Methods 4 + 5 (hybrid LLM judge, stateful builder) remain Phase B-pending.
phase_b_pending <- FALSE
```

## Framing

C1 v0.7.0 extracts a `measure_name` string per chunk × measure, but the same fiscal act surfaces under many different forms ("ERTA", "Economic Recovery Tax Act of 1981", "Economic Recovery Tax Act (ERTA)", "the 1981 tax cut"). Downstream consumers — C2b motivation, deployment summaries, and any cross-country comparison — need a stable canonical mapping at the **act** level. The job of a "C0" Act Aggregator is to produce that mapping.

The Malaysia EN/BM consistency notebook ([malay_consistency.qmd](malay_consistency.qmd)) showed that JW single-linkage at threshold 0.15 fragments acts (especially BM-side, where "Subsidi Bahan Api" + "Subsidi Gula" split off from a single budget envelope), and that this fragmentation propagates downstream: C2b reads asymmetric evidence on the two language sides and apparent disagreement is partly clustering noise. So before designing a C0 codebook we need empirical evidence on which aggregation strategy holds up.

This notebook compares five methods on a label-anchored evaluation set (US C1 outputs joined to Romer & Romer's 40 gold-standard acts) and an out-of-distribution stress test (Malaysia EN/BM paired pairs).

**Scope.** Corpus-wide per country (one canonical list across all years of a corpus), not per-document and not cross-country. The C1 deployment runs per-country; C0 will too.

**Asymmetric failure cost.** Over-merging (collapse two real acts into one cluster) corrupts C2b inputs — bad. Under-merging (one act stays split across two clusters) just produces parallel-but-correct classifications — recoverable downstream. The headline pick should weight precision over recall.

**From asymmetric cost to precision over recall.** Pairwise F1 scores the clustering at the level of pairs of measure names: for each pair, the method either merges them or not, and gold either agrees or not. That gives the standard 2×2 table, with a false positive corresponding to an over-merge (method merges two names from different gold acts) and a false negative to an under-merge (method splits two names that share a gold act). Precision (TP / (TP + FP)) is therefore degraded by over-merges, and recall (TP / (TP + FN)) by under-merges. "Over-merging is the costly error" and "false positives are the costly error" are the same statement once the pairwise framing is adopted, so weighting precision over recall is the operational form of the asymmetric-cost argument.

**Gold partition caveat.** This framing assumes the gold partition is itself correct. If R&R's act list splits something a domain expert would have merged (or vice versa), the precision/recall labels inherit that. The Tier 1/Tier 2 leakage caveat below addresses a different concern (how chunks were tagged with gold acts); partition-correctness of the gold list itself is a separate assumption.

**What "win" means.** High pairwise F1 (the metric most sensitive to both kinds of error), high Adjusted Rand Index (partition-level agreement, chance-corrected), and a low over-merge:under-merge ratio. Bootstrap 95% CIs included but read as noise floors — pairs are not iid (they share names) so CIs are approximate.

## Descriptive baseline

@tbl-pool-summary sizes the canonicalization problem at the chunk-measure level. Its five rows describe the C1 v0.7.0 output pool feeding C0:

- **Total FISCAL_MEASURE rows** counts chunk × measure entries C1 flagged as FISCAL_MEASURE. Because v0.7.0 returns a `measures[]` array per chunk, one chunk can contribute multiple rows.
- **Distinct measure_name strings (any rank)** counts unique surface forms across the full pool, including secondary measures from chunks whose primary measure was something else. This is the maximum pool size C0 would face if it had to canonicalize every name C1 ever emits.
- **Distinct measure_name strings (rank == 1)** counts only the primary surface form per chunk. The codebook contract ([c1_measure_id.yml](../prompts/c1_measure_id.yml)) requires the first array element to be the most prominent measure discussed, so rank == 1 is C1's primary pick per chunk. This is the realistic pool size for a primary-measure-only deployment, which matches the current C2b assumption.
- **Years spanned** is the inclusive range of document years contributing chunks.
- **Distinct doc_id** is the number of source documents represented.

The gap between rows 2 and 3 quantifies how much of the surface-form pool comes from secondary mentions. A small gap means rank-1 carries most of the variance and C0 can be evaluated on the rank-1 subset without much loss. A large gap means non-primary measures contribute substantial unique forms, and C0's effective scope (primary-only vs all) becomes a load-bearing methodological choice. The eval set in §3 is built on rank == 1 only.

```{r}
#| label: tbl-pool-summary
#| tbl-cap: "Measure-name pool from c1_s2_results"

tibble::tibble(
  Metric = c("Total FISCAL_MEASURE rows",
             "Distinct measure_name strings (any rank)",
             "Distinct measure_name strings (rank == 1)",
             "Years spanned",
             "Distinct doc_id"),
  Value = c(
    sprintf("%d", nrow(c0_us_measure_pool)),
    sprintf("%d", dplyr::n_distinct(c0_us_measure_pool$measure_name)),
    sprintf("%d", c0_us_measure_pool |>
              dplyr::filter(measure_rank == 1L) |>
              dplyr::pull(measure_name) |> dplyr::n_distinct()),
    sprintf("%d-%d", min(c0_us_measure_pool$year, na.rm = TRUE),
            max(c0_us_measure_pool$year, na.rm = TRUE)),
    sprintf("%d", dplyr::n_distinct(c0_us_measure_pool$doc_id))
  )
) |> tt() |> tt_theme_report()
```

Surface-form variance per gold act — the magnitude of the canonicalization problem. For each gold act_name with at least one chunk in `c0_eval_gold_pairs`, count the distinct C1 surface forms (rank=1 only) that map to it.

```{r}
#| label: fig-variance-per-act
#| fig-cap: "Distinct C1 rank-1 surface forms per gold act (top 25 most fragmented)"
#| fig-width: 9
#| fig-height: 6

variance_per_act <- c0_eval_gold_pairs |>
  dplyr::distinct(measure_name, gold_act_name) |>
  dplyr::count(gold_act_name, name = "n_surface_forms") |>
  dplyr::arrange(dplyr::desc(n_surface_forms))

variance_per_act |>
  utils::head(25) |>
  dplyr::mutate(gold_act_name = forcats::fct_reorder(
    stringr::str_trunc(gold_act_name, 60), n_surface_forms)) |>
  ggplot(aes(n_surface_forms, gold_act_name)) +
  geom_col(fill = "#4477AA") +
  geom_text(aes(label = n_surface_forms), hjust = -0.2, size = 3) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(x = "Distinct rank-1 surface forms", y = NULL) +
  theme_minimal(base_size = 10)
```

@tbl-surface-forms-per-act drills into the same data as @fig-variance-per-act, using the full any-rank measure pool (rather than the rank-1 subset used by the figure). Gold acts appear as group headers most-fragmented first, annotated with the total number of distinct C1 surface forms that ever appeared in chunks tagged with that act at any rank. Under each group, the top five most-frequently-emitted surface forms with the count of chunk-rows in which C1 emitted them. The shape of the within-group distribution is what C0 has to recover: a dominant first form with a long tail rewards conservative merging because the dominant form is already a serviceable canonical; a flat distribution needs more aggressive merging to collapse parallel forms.

```{r}
#| label: tbl-surface-forms-per-act
#| tbl-cap: "Top 5 C1 surface forms per gold act (any rank). Group headers list the total number of distinct surface forms for the act; acts with fewer than five forms show all of them."

chunk_to_gold <- c0_eval_gold_pairs |>
  dplyr::distinct(chunk_id, gold_act_name)

any_rank_pool <- c0_us_measure_pool |>
  dplyr::inner_join(chunk_to_gold, by = "chunk_id",
                    relationship = "many-to-many")

top5_per_act <- any_rank_pool |>
  dplyr::count(gold_act_name, measure_name, name = "n_appearances") |>
  dplyr::group_by(gold_act_name) |>
  dplyr::arrange(dplyr::desc(n_appearances), measure_name, .by_group = TRUE) |>
  dplyr::slice_head(n = 5L) |>
  dplyr::ungroup()

act_summary_anyrank <- any_rank_pool |>
  dplyr::group_by(gold_act_name) |>
  dplyr::summarize(
    n_surface_forms = dplyr::n_distinct(measure_name),
    .groups = "drop"
  ) |>
  dplyr::arrange(dplyr::desc(n_surface_forms), gold_act_name)

# Interleave group-header rows (no count) with indented data rows so we get
# precise control over bolding and indentation.
indent <- strrep(" ", 4L)
rows <- list()
for (i in seq_len(nrow(act_summary_anyrank))) {
  act     <- act_summary_anyrank$gold_act_name[i]
  n_forms <- act_summary_anyrank$n_surface_forms[i]
  forms   <- top5_per_act |> dplyr::filter(gold_act_name == act)

  rows[[length(rows) + 1L]] <- tibble::tibble(
    `Surface form` = sprintf("%s (%d surface forms)", act, n_forms),
    Appearances   = ""
  )
  rows[[length(rows) + 1L]] <- tibble::tibble(
    `Surface form` = paste0(indent, forms$measure_name),
    Appearances   = as.character(forms$n_appearances)
  )
}
display_df <- dplyr::bind_rows(rows)

header_positions <- which(display_df$Appearances == "")

display_df |>
  tt() |>
  tinytable::format_tt(escape = TRUE) |>
  tt_theme_report() |>
  tinytable::style_tt(i = header_positions, bold = TRUE) |>
  tinytable::style_tt(j = 1, align = "l") |>
  tinytable::style_tt(j = 2, align = "r")
```

```{r}
#| label: tbl-variance-summary
#| tbl-cap: "Distribution of surface-form fragmentation across the 39 gold acts"

variance_per_act |>
  dplyr::summarize(
    `Acts evaluated` = dplyr::n(),
    `Median surface forms / act` = stats::median(n_surface_forms),
    `Mean surface forms / act`   = round(mean(n_surface_forms), 2),
    `Max surface forms / act`    = max(n_surface_forms),
    `Acts with ≥ 5 forms`        = sum(n_surface_forms >= 5L),
    `Acts with exactly 1 form`   = sum(n_surface_forms == 1L)
  ) |>
  tidyr::pivot_longer(everything(), names_to = "Metric", values_to = "Value") |>
  tt() |> tt_theme_report()
```

The headline: even with R&R's well-known and well-named acts, the median act surfaces under multiple forms; the tail is heavy. A C0 that returns one cluster per surface form (i.e., no merging) is useless; a C0 that merges everything is also useless. The methods below try to land somewhere useful.

## Eval set construction

The eval set projects the chunk-tier system through C1's LLM output along the chain `aligned_data → c1_chunk_data → c1_s2_results → c0_eval_gold_pairs`:

1. **`c1_chunk_data`** (built by `R/identify_chunk_tiers.R`) emits `(chunk_id, doc_id, act_name, tier)` rows by string-matching gold passages and act-name subcomponents against chunk text. Tier 1 marks chunks containing a verbatim gold passage; Tier 2 marks chunks matching an act name via subcomponent decomposition or co-occurrence rules. Chunks matching multiple acts within the same tier are row-replicated, one row per (chunk × act). The Tier 1 / Tier 2 exclusion is chunk-level, not (chunk, act)-level: any chunk that is Tier 1 for *some* act is barred from Tier 2 consideration for *every* act, so Tier 2 evidence on mixed-tier multi-act chunks is dropped at the source.

2. **`c1_s2_results`** runs the C1 LLM on these tagged chunks via leave-one-act-out cross-validation; each (chunk × gold act) pair is its own fold with its own few-shot set. The LLM's emitted `measures[]` array is exploded into one row per (chunk × gold act × measure_rank), so a chunk tagged with two gold acts is classified twice and may carry different `measure_name` strings under each gold tag.

3. **`build_c0_eval_gold_pairs()`** (in `R/c0_aggregator.R`) keeps only rows with `measure_rank == 1`, `pred_label == FISCAL_MEASURE`, and `tier ∈ {1, 2}`; renames `act_name → gold_act_name`; and flags surface forms paired with >1 distinct gold act as `ambiguous` for the evaluator to majority-vote.

Two consequences carry into every downstream metric. (i) The rank-1 filter discards non-primary measures the LLM emitted, so a chunk discussing N gold acts contributes at most N rows with only the rank-1 surface form for each. (ii) Chunks the LLM mislabeled `NOT_FISCAL_MEASURE` drop out entirely, silently truncating multi-act representation whenever the LLM's per-fold output is inconsistent.

```{r}
#| label: tbl-gold-construction
#| tbl-cap: "Eval-set composition. Rows count `(measure_name, gold_act_name, chunk_id)` tuples; distinct surface forms count unique `measure_name` strings — the population the clusterer canonicalizes. The two indented rows partition the Tier 1 + Tier 2 pool by whether each surface form maps to one (unambiguous) or more (ambiguous) gold acts. Percentages are shares of the Tier 1 + Tier 2 totals."

t1_slice    <- c0_eval_gold_pairs |> dplyr::filter(tier == 1L)
unamb_slice <- c0_eval_gold_pairs |> dplyr::filter(!ambiguous)
amb_slice   <- c0_eval_gold_pairs |> dplyr::filter(ambiguous)

n_rows_all  <- nrow(c0_eval_gold_pairs)
n_names_all <- dplyr::n_distinct(c0_eval_gold_pairs$measure_name)

fmt_pct <- function(x, total) {
  sprintf("%d (%d%%)", x, round(100 * x / total))
}

tibble::tibble(
  Slice = c(
    "Tier 1 (verbatim passage match)",
    "Tier 1 + Tier 2 (full eval pool)",
    "    ↳ unambiguous names (1 gold act per name)",
    "    ↳ ambiguous names (≥2 gold acts per name)"
  ),
  `Chunk × gold-act rows` = c(
    as.character(nrow(t1_slice)),
    as.character(n_rows_all),
    fmt_pct(nrow(unamb_slice), n_rows_all),
    fmt_pct(nrow(amb_slice),   n_rows_all)
  ),
  `Distinct surface forms` = c(
    as.character(dplyr::n_distinct(t1_slice$measure_name)),
    as.character(n_names_all),
    fmt_pct(dplyr::n_distinct(unamb_slice$measure_name), n_names_all),
    fmt_pct(dplyr::n_distinct(amb_slice$measure_name),   n_names_all)
  ),
  `Gold acts covered` = c(
    as.character(dplyr::n_distinct(t1_slice$gold_act_name)),
    as.character(dplyr::n_distinct(c0_eval_gold_pairs$gold_act_name)),
    "—",
    "—"
  )
) |>
  tt() |>
  tt_theme_report() |>
  tinytable::style_tt(j = 1,   align = "l") |>
  tinytable::style_tt(j = 2:4, align = "r")
```

**On ambiguity.** Roughly 60% of the Tier 1+2 rows have a `measure_name` that's been paired with more than one `gold_act_name` across chunks. Most cases trace to Tier 2's permissive matching: a chunk may be tagged with act X because act X is mentioned somewhere, but the most-prominent measure C1 extracts is a different act. The evaluator handles this by majority-voting per `measure_name` within each bootstrap resample (so the contingency table operates on unique names). Tier 1 is shown as the cleaner anchor; Tier 1+2 is shown for statistical power.

**Gold-set leakage caveat.** The Tier 1/Tier 2 tags themselves were assigned by fuzzy-matching the gold `act_name` against the chunk text, so methods that exploit surface similarity (especially JW) will look artificially strong on this evaluation. Treat the per-act recall column as a complement — methods that nail F1 but miss specific acts may be exploiting the leakage.

## Method 1 — JW single-linkage

Before reading the threshold sweep, @fig-jw-within-act shows what JW is being asked to recover. Each row is a gold act with ≥2 surface forms in the eval pool; the box and jittered points show the pairwise JW distances among the surface forms tagged to that act. Vertical dashes mark the swept thresholds (0.10–0.30). An act whose distribution sits entirely left of a threshold line is fully recoverable at that threshold; an act whose distances spread past 0.30 cannot be merged without dragging in surface forms from neighbouring acts (i.e., over-merge). Ambiguous names are kept under every gold act they're tagged to, matching the descriptive intent of the figure rather than the evaluator's majority-vote.

```{r}
#| label: fig-jw-within-act
#| fig-cap: "Within-gold-act pairwise JW distances. Vertical dashes mark the swept thresholds. Acts with a single surface form contribute no pairs and are omitted."
#| fig-width: 10
#| fig-height: 8

within_act_jw <- function(pool) {
  pool |>
    dplyr::distinct(measure_name, gold_act_name) |>
    dplyr::group_by(gold_act_name) |>
    dplyr::filter(dplyr::n() >= 2L) |>
    dplyr::group_modify(~ {
      names <- .x$measure_name
      d <- stringdist::stringdistmatrix(names, names, method = "jw")
      tibble::tibble(jw_distance = d[upper.tri(d)])
    }) |>
    dplyr::ungroup()
}

tier1_pairs  <- c0_eval_gold_pairs |>
  dplyr::filter(tier == 1L) |>
  within_act_jw()
tier12_pairs <- within_act_jw(c0_eval_gold_pairs)

plot_within <- function(df, title) {
  df |>
    dplyr::mutate(gold_act_name = forcats::fct_reorder(
      stringr::str_trunc(gold_act_name, 50),
      jw_distance, .fun = stats::median)) |>
    ggplot(aes(jw_distance, gold_act_name)) +
    geom_vline(xintercept = c(0.10, 0.15, 0.20, 0.25, 0.30),
               linetype = "dashed", colour = "grey60", linewidth = 0.3) +
    geom_boxplot(outlier.shape = NA, fill = "#4477AA", alpha = 0.3,
                 linewidth = 0.3) +
    geom_jitter(width = 0, height = 0.15, size = 0.7, alpha = 0.6,
                colour = "#4477AA") +
    scale_x_continuous(limits = c(0, NA),
                       breaks = c(0, 0.10, 0.15, 0.20, 0.25, 0.30, 0.5)) +
    labs(title = title, x = "Within-act pairwise JW distance", y = NULL) +
    theme_minimal(base_size = 9)
}

p_a <- plot_within(tier1_pairs,  "A. Tier 1 only")
p_b <- plot_within(tier12_pairs, "B. Tier 1 + Tier 2")

p_a + p_b + patchwork::plot_layout(ncol = 2)
```

**What this means for Method 1.** The median within-act JW distance is ~0.47 on both tiers, while the loosest threshold swept is 0.30. The typical pair of surface forms that *ought* to be merged is therefore geometrically out of reach for JW at any threshold the table considers; only a small minority of pairs (near-duplicates differing in spacing, punctuation, or abbreviation expansion) sit below 0.30. The implication is unavoidable: **JW single-linkage cannot canonicalize this pool on its own.** The threshold sweep in @tbl-jw-metrics is best read as quantifying the asymmetric precision/recall cost given that fundamental ceiling, not as a search for a workable operating point. A method that can bridge `ERTA` ↔ `Economic Recovery Tax Act of 1981` — semantic, not lexical — is required, which is what Phase B is for.

@fig-jw-name-to-gold checks the same geometry from a different angle: instead of pairwise distances among surface forms, it measures each surface form's JW distance to its gold-act label. If C0 had the gold canonical list as an oracle and tried to map surface forms to it by JW, the swept thresholds tell us how many surface forms it would catch. The answer mirrors @fig-jw-within-act: median ~0.41 on both tiers, well above 0.30. A handful of exact matches sit at 0 (surface form is the gold label string verbatim), but most surface forms are too lexically distant from their canonical to be matched without semantic reasoning.

```{r}
#| label: fig-jw-name-to-gold
#| fig-cap: "Surface-name → gold-label JW distance. Same panels, threshold lines, and visual style as @fig-jw-within-act; each point is one (measure_name, gold_act_name) pair."
#| fig-width: 10
#| fig-height: 8

name_to_gold_jw <- function(pool) {
  pool |>
    dplyr::distinct(measure_name, gold_act_name) |>
    dplyr::mutate(jw_distance = stringdist::stringdist(
      measure_name, gold_act_name, method = "jw"))
}

tier1_to_gold  <- c0_eval_gold_pairs |>
  dplyr::filter(tier == 1L) |>
  name_to_gold_jw()
tier12_to_gold <- name_to_gold_jw(c0_eval_gold_pairs)

plot_to_gold <- function(df, title) {
  df |>
    dplyr::mutate(gold_act_name = forcats::fct_reorder(
      stringr::str_trunc(gold_act_name, 50),
      jw_distance, .fun = stats::median)) |>
    ggplot(aes(jw_distance, gold_act_name)) +
    geom_vline(xintercept = c(0.10, 0.15, 0.20, 0.25, 0.30),
               linetype = "dashed", colour = "grey60", linewidth = 0.3) +
    geom_boxplot(outlier.shape = NA, fill = "#CC6677", alpha = 0.3,
                 linewidth = 0.3) +
    geom_jitter(width = 0, height = 0.15, size = 0.7, alpha = 0.6,
                colour = "#CC6677") +
    scale_x_continuous(limits = c(0, NA),
                       breaks = c(0, 0.10, 0.15, 0.20, 0.25, 0.30, 0.5)) +
    labs(title = title, x = "Surface-name → gold-label JW distance", y = NULL) +
    theme_minimal(base_size = 9)
}

p_a <- plot_to_gold(tier1_to_gold,  "A. Tier 1 only")
p_b <- plot_to_gold(tier12_to_gold, "B. Tier 1 + Tier 2")

p_a + p_b + patchwork::plot_layout(ncol = 2)
```

The within-act figures pin down what JW can mechanically do; the tables and figures below quantify the consequences on the precision-over-recall lens the Framing committed to. Each row of the ladders is one variant (threshold × year-blocking choice); each cell carries a 95% bootstrap CI from resampling gold rows.

- **Precision** is the share of predicted-merge pairs that share a gold act. It is degraded by over-merging and is the column the asymmetric-cost argument weights most.
- **Recall** is the share of gold-merge pairs the method captures. It is degraded by under-merging, the recoverable failure mode.
- **F1** is their harmonic mean. Useful single-number summary, but symmetric in P and R, so it under-weights the cost asymmetry the C2b consumer cares about.
- **ARI** (Adjusted Rand Index) is partition-level chance-corrected agreement with the gold partition. It complements F1: pairwise F1 can be inflated by near-singleton partitions that score well on rare-pair accounting, and ARI catches that.
- **Purity** is the mean within-cluster plurality-class share. The most precision-flavoured of the three: it ignores under-merges entirely and trivially reaches 1.0 in the one-name-per-cluster limit, so read it together with `n clusters`.
- **Over:Under** is the ratio of over-merge pair count to under-merge pair count. The most direct asymmetric-cost summary: values below 1 mean the variant makes fewer over-merges than under-merges (precision-friendly); large values mean precision has been sacrificed to capture recall.
- **n clusters** is the partition size. There are ≈39 gold acts in the eval, so healthy variants sit in roughly that order of magnitude; far below means severe over-merging, far above means failure to merge anything.

CIs are wide because the eval is small (≈27 distinct names in Tier 1, ≈50 in Tier 1+2) and pairs share names so are not iid; read them as noise floors rather than formal confidence intervals.

The figures below give the same data three additional ways. @fig-jw-f1 traces Precision, Recall, and F1 across the threshold sweep so the reader can see how the P/R trade-off evolves as JW relaxes; @fig-jw-overunder is the raw-counts log-log scatter showing the failure-mode geometry; @fig-jw-pr is the same scatter on the bounded Precision/Recall axes, where the upper-left corner is the asymmetric-cost ideal.

```{r}
#| label: prep-jw-ladder

shape_jw_ladder <- function(eval_tier) {
  format_method_ladder(c0_jw_metrics, eval_tier = eval_tier) |>
    dplyr::mutate(
      `JW threshold` = as.numeric(
        stringr::str_extract(variant_id, "(?<=t)\\d\\.\\d{2}")),
      blocking = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                                "Year ±2", "Unblocked")
    ) |>
    dplyr::arrange(blocking, `JW threshold`) |>
    dplyr::select(-variant_id)
}

render_jw_ladder <- function(ladder) {
  panel_starts <- ladder |>
    dplyr::mutate(.row = dplyr::row_number()) |>
    dplyr::group_by(blocking) |>
    dplyr::summarize(start = min(.row), .groups = "drop") |>
    dplyr::arrange(start)
  group_index <- stats::setNames(as.list(panel_starts$start),
                                 panel_starts$blocking)

  display <- ladder |>
    dplyr::select(-blocking) |>
    dplyr::relocate(`JW threshold`) |>
    dplyr::mutate(`JW threshold` = sprintf("%.2f", `JW threshold`))

  display |>
    tinytable::tt() |>
    tt_theme_report() |>
    tinytable::group_tt(i = group_index)
}

ladder_t1  <- shape_jw_ladder("tier1")
ladder_t12 <- shape_jw_ladder("tier12")
```

```{r}
#| label: tbl-jw-metrics
#| tbl-cap: "JW single-linkage, Tier 1 evaluation (clean, n≈27 names). Cells: point estimate [95% bootstrap CI]. Over:Under ratio = over-merge pair count / under-merge pair count; lower is more precision-friendly."

render_jw_ladder(ladder_t1)
```

```{r}
#| label: tbl-jw-metrics-t12
#| tbl-cap: "JW single-linkage, Tier 1 + Tier 2 evaluation (more power, more noise). Same column definitions as @tbl-jw-metrics."

render_jw_ladder(ladder_t12)
```

```{r}
#| label: fig-jw-f1
#| fig-cap: "Pairwise Precision, Recall, and F1 across JW thresholds × year-blocking, both eval tiers. Ribbons are 95% bootstrap CIs."
#| fig-width: 10
#| fig-height: 6

c0_jw_metrics |>
  dplyr::filter(metric %in% c("pairwise_precision", "pairwise_recall",
                              "pairwise_f1")) |>
  dplyr::mutate(
    threshold = as.numeric(stringr::str_extract(variant_id, "(?<=t)\\d\\.\\d{2}")),
    blocking  = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                               "year ±2", "unblocked"),
    metric    = forcats::fct_recode(metric,
                                    Precision = "pairwise_precision",
                                    Recall    = "pairwise_recall",
                                    F1        = "pairwise_f1") |>
                forcats::fct_relevel("Precision", "Recall", "F1")
  ) |>
  ggplot(aes(threshold, value, colour = blocking, fill = blocking)) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  facet_grid(eval_tier ~ metric) +
  scale_colour_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "JW threshold (merge if distance ≤ threshold)",
       y = NULL,
       colour = NULL, fill = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

```{r}
#| label: fig-jw-overunder
#| fig-cap: "Over-merge vs under-merge counts across JW variants, Tier 1+2 evaluation. Up-and-right is precision-loving; down-and-left is balanced."
#| fig-width: 8
#| fig-height: 4

c0_jw_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier12") |>
  dplyr::mutate(
    threshold = as.numeric(stringr::str_extract(variant_id, "(?<=t)\\d\\.\\d{2}")),
    blocking  = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                               "year ±2", "unblocked")
  ) |>
  ggplot(aes(under_merge_count, over_merge_count, colour = blocking,
             label = sprintf("t=%.2f", threshold))) +
  geom_point(size = 2) +
  geom_text(size = 3, hjust = -0.2, vjust = -0.4,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_log10() + scale_y_log10() +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Under-merge pair count (log)",
       y = "Over-merge pair count (log)",
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

```{r}
#| label: fig-jw-pr
#| fig-cap: "Pairwise Precision vs Recall across JW variants, Tier 1+2 evaluation. Upper-left is the precision-favouring corner the asymmetric-cost framing prefers; upper-right is the unreachable ideal."
#| fig-width: 8
#| fig-height: 4

c0_jw_metrics |>
  dplyr::filter(metric %in% c("pairwise_precision", "pairwise_recall"),
                eval_tier == "tier12") |>
  dplyr::select(variant_id, metric, value) |>
  tidyr::pivot_wider(names_from = metric, values_from = value) |>
  dplyr::mutate(
    threshold = as.numeric(stringr::str_extract(variant_id, "(?<=t)\\d\\.\\d{2}")),
    blocking  = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                               "year ±2", "unblocked")
  ) |>
  ggplot(aes(pairwise_recall, pairwise_precision, colour = blocking,
             label = sprintf("t=%.2f", threshold))) +
  geom_point(size = 2) +
  geom_text(size = 3, hjust = -0.2, vjust = -0.4,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Pairwise Recall",
       y = "Pairwise Precision",
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

### Reading the JW results together

Three patterns hold up across the ladder, the metric grid, and the two scatters.

First, JW's precision ceiling is meaningful but its recall floor is rock-bottom. At the tightest year-blocked thresholds (Tier 1, t = 0.10–0.20, Year ±2), Precision reaches 0.50 with Over:Under at 0.10 — a precision-friendly operating point. But Recall pins at ≈0.09 for the same variant, because the within-act distance figure already established the geometric ceiling: most of the surface-form pairs that *ought* to merge sit beyond 0.30, geometrically out of reach for JW at any threshold considered here.

Second, relaxing the threshold trades precision for recall along a roughly monotonic front in @fig-jw-pr, but the trade is bad: pairwise F1 stays below ≈0.15 across the entire sweep, and Over:Under explodes to 84× at threshold 0.30 unblocked. There is no operating point on the front that delivers acceptable recall without catastrophic over-merging.

Third, year-blocking shifts the curve in the right direction at every threshold — lower Over:Under, tighter precision CIs, negligible recall cost. This is a structural finding rather than a JW-specific one: the year-window prior will be worth carrying into the embedding and hybrid methods in Phase B.

### Bottom line for Method 1

JW single-linkage cannot canonicalise this measure pool on its own. The geometric ceiling from the within-act figures is binding, and the eval confirms it: pairwise F1 ≤ 0.15 anywhere on the precision-friendly side of the front, and any threshold loose enough to lift Recall meaningfully above 0.30 also lifts Over:Under into double digits. The gap between *ERTA* and *Economic Recovery Tax Act of 1981* is semantic, not lexical, and JW has no machinery to bridge it.

What JW remains useful for is the role its mechanics actually support: a precision-friendly **blocker** — a cheap pre-filter that catches near-duplicate surface forms (spacing, punctuation, abbreviation expansion) at near-zero false-positive cost. Method 4's hybrid pipeline below uses exactly this property: JW (or a cheap analog) prunes the pair space, and a semantic verifier handles the merges that surface similarity cannot. The load-bearing canonicalisation work, though, has to come from a method that understands meaning — Methods 2–5 are the candidate replacements.

## Embedding-geometry probe (F16 vs FP32)

Method 1 closed on the observation that the gap to canonicalise this pool is semantic, not lexical. The next three sections test whether multilingual embeddings can bridge it. Before reading the eval, one preliminary: the Methods 2/3 embeddings come from a community F16 port of `intfloat/multilingual-e5-large-instruct` (`jeffh/intfloat-multilingual-e5-large-instruct:f16`) running locally under Ollama. F16 compresses the reference FP32 weights and could in principle distort the embedding geometry that HDBSCAN relies on. Two cheap probes on the gold eval pool check that the port preserves the semantic structure HDBSCAN needs.

The probes run on a 2×2 grid: **precision** (F16 local vs FP32 reference via DeepInfra) × **eval slice** (full unambiguous pool vs Tier 1 only). The FP32 reference disentangles two distinct failure modes the F16-only probe couldn't separate: "F16 quantization broke the geometry" (FP32 passes where F16 fails) vs "the model itself can't do this task" (both precisions fail). The Tier 1 slice in turn strips Tier 2's permissive IKA tagging, which §"Eval-set noise probe" below shows is genuinely noisy. Reading the four cells together discriminates quantization noise, model-capacity ceiling, and label noise.

```{r}
#| label: tbl-f16-probe
#| tbl-cap: "Embedding-geometry probe on a 2×2 (precision × eval slice) grid. Top-1 NN same-act rate is the share of eval names whose nearest-neighbour by cosine shares a gold act (restricted to acts with ≥2 members so the metric is well-defined). Separation = median within-act cosine − median between-act cosine; positive means the embedding preserves the act-level structure. F16 rows come from the local Ollama port (`jeffh/intfloat-multilingual-e5-large-instruct:f16`); FP32 rows from DeepInfra-hosted `intfloat/multilingual-e5-large-instruct`. Same instruction prefix on both."

dplyr::bind_rows(
  dplyr::mutate(c0_f16_quantization_probe,       Precision = "F16 (local Ollama)",  `Eval slice` = "Full pool (Tier 1 + 2)"),
  dplyr::mutate(c0_f16_quantization_probe_tier1, Precision = "F16 (local Ollama)",  `Eval slice` = "Tier 1 only"),
  dplyr::mutate(c0_fp32_reference_probe,         Precision = "FP32 (DeepInfra)",    `Eval slice` = "Full pool (Tier 1 + 2)"),
  dplyr::mutate(c0_fp32_reference_probe_tier1,   Precision = "FP32 (DeepInfra)",    `Eval slice` = "Tier 1 only")
) |>
  dplyr::transmute(
    Precision,
    `Eval slice`,
    `Top-1 NN same-act rate`    = sprintf("%.3f", top1_nn_same_act_rate),
    `Median within-act cosine`  = sprintf("%.3f", median_within_act_cosine),
    `Median between-act cosine` = sprintf("%.3f", median_between_act_cosine),
    `Separation`                = sprintf("%.3f", separation),
    `Eval names`                = n_eval_names,
    `Eval acts`                 = n_eval_acts
  ) |>
  tt() |> tt_theme_report()
```

```{r}
#| label: f16-verdict
#| echo: false
#| results: asis

slice_pass <- function(p) {
  sep_ok <- !is.na(p$separation) && p$separation >= 0.05
  top_ok <- !is.na(p$top1_nn_same_act_rate) && p$top1_nn_same_act_rate >= 0.40
  list(sep_ok = sep_ok, top_ok = top_ok, pass = sep_ok && top_ok)
}

p_f16_full  <- c0_f16_quantization_probe
p_f16_t1    <- c0_f16_quantization_probe_tier1
p_fp32_full <- c0_fp32_reference_probe
p_fp32_t1   <- c0_fp32_reference_probe_tier1

v_f16_t1  <- slice_pass(p_f16_t1)
v_fp32_t1 <- slice_pass(p_fp32_t1)

fmt_slice <- function(label, p, v) {
  sprintf("%s: separation %.3f (floor 0.05), top-1 NN rate %.3f (floor 0.40) — %s",
          label, p$separation %||% NA_real_,
          p$top1_nn_same_act_rate %||% NA_real_,
          if (v$pass) "**PASS**" else "**FAIL**")
}

cat(fmt_slice("F16, Tier 1",  p_f16_t1,  v_f16_t1),  "  \n")
cat(fmt_slice("FP32, Tier 1", p_fp32_t1, v_fp32_t1), "\n\n")

cat("Reading Tier 1 vs Tier 1 keeps the gold-partition noise floor constant on both rows so the F16↔FP32 contrast is apples-to-apples.\n\n")

if (!v_f16_t1$pass && v_fp32_t1$pass) {
  cat("**Verdict — F16 quantization is the culprit.** The FP32 reference clears the floors that F16 misses, so the embedding geometry HDBSCAN needs exists at full precision and is being degraded by the F16 port. The diagnostic conclusion is that migrating Methods 2 + 3 to FP32 is worth a separate plan. Until that lands, Methods 2/3 read below remain on F16 and should be interpreted as a lower bound on what the embedding family can achieve.\n")
} else if (!v_f16_t1$pass && !v_fp32_t1$pass) {
  cat("**Verdict — embedding-model ceiling, not quantization.** Both F16 and FP32 fail the same floors, so the issue is not F16 specifically — `multilingual-e5-large-instruct` itself does not produce the relative cosine spread the floors require on this pool. Two interpretations to weigh: (i) E5-instruct embeddings live in a tight band of the unit sphere by design, and the 0.05/0.40 floors may be inappropriate for this family; (ii) the model is genuinely under-resolving fiscal-act semantics and a stronger or differently-tuned embedding family is needed. Either way, the FP32 migration is *not* the fix. Read Methods 2/3 below as the empirical end-to-end test — if HDBSCAN's Tier 1 numbers are competitive with JW, the floors were the over-strict reading.\n")
} else if (v_f16_t1$pass && v_fp32_t1$pass) {
  cat("**Verdict — F16 was fine all along.** Both precisions clear the floors on Tier 1, so the earlier full-pool failure was carried by Tier 2 label noise rather than quantization or model capacity. Methods 2/3 read below on F16 with no caveat.\n")
} else {
  cat("**Verdict — anomalous.** F16 passes Tier 1 but FP32 fails; this is unexpected since FP32 has strictly more precision. Inspect the raw probe outputs, confirm the FP32 endpoint returned full-rank vectors, and check the instruction-prefix handling before drawing conclusions.\n")
}
```

E5-family embeddings always live in a narrow band of the unit sphere, so absolute cosines aren't crisp — the probe targets the *relative* gap between within-act and between-act pairs, which is the property HDBSCAN cares about. Methods 2 and 3 below continue to read `c0_us_embeddings` (the F16 matrix); the FP32 numbers above are diagnostic only.

## Eval-set noise probe: surface-year vs doc-year mismatch

An alternative explanation for a soft F16 verdict — and a noise source the within-act geometry from §Method 1 already hinted at — is that the gold pairs themselves are temporally noisy. If C1 sometimes catches contemporary discussions of acts passed decades earlier (e.g. a 2010 ERP paragraph referring back to the Revenue Act of 1964), the resulting `(measure_name, gold_act_name)` pair will look lexically distant — not because the model misclassified, but because the surface form drifts as the act becomes historical context rather than current policy. @fig-eval-noise-temporal probes this directly: for each eval tuple whose surface form contains a 4-digit year, plot the gap between the document's publication year and the year referenced in the surface form against the JW distance to the gold label.

```{r}
#| label: fig-eval-noise-temporal
#| fig-cap: "Temporal mismatch in eval pairs. Each point is one (measure_name, gold_act_name, chunk_id) tuple whose surface form contains a 4-digit year. x = doc publication year − year referenced in the surface form (large positive ⇒ retrospective reference); y = JW distance from surface form to gold label. Tuples whose surface form has no extractable year are omitted."
#| fig-width: 8
#| fig-height: 5
year_pattern <- "\\b(19|20)\\d{2}\\b"

chunk_doc_year <- c0_us_measure_pool |>
    dplyr::distinct(chunk_id, doc_year = year)

noise_diag <- c0_eval_gold_pairs |>
    dplyr::distinct(measure_name, gold_act_name, chunk_id, tier) |>
    dplyr::left_join(chunk_doc_year, by = "chunk_id") |>
    dplyr::mutate(
        surface_year = as.integer(stringr::str_extract(measure_name, year_pattern)),
        year_diff = doc_year - surface_year,
        jw_distance = stringdist::stringdist(measure_name, gold_act_name,
            method = "jw"
        ),
        tier_label = factor(tier,
            levels = c(1L, 2L),
            labels = c("Tier 1", "Tier 2")
        )
    ) |>
    dplyr::filter(!is.na(surface_year), !is.na(doc_year))

ggplot(noise_diag, aes(year_diff, jw_distance, colour = tier_label)) +
    geom_hline(
        yintercept = c(0.10, 0.15, 0.20, 0.25, 0.30),
        linetype = "dashed", colour = "grey60", linewidth = 0.3
    ) +
    geom_vline(
        xintercept = 0, linetype = "dotted",
        colour = "grey40", linewidth = 0.3
    ) +
    geom_point(alpha = 0.5, size = 1.6) +
    geom_smooth(aes(colour = tier_label),
        se = FALSE,
       linewidth = 0.4, linetype = "dashed"
    ) +
    scale_colour_brewer(palette = "Set1") +
    labs(
        x = "Doc year − surface-referenced year",
        y = "JW distance (surface name → gold label)",
        colour = NULL
    ) +
    theme_minimal(base_size = 10) +
    theme(legend.position = "top")
```

@fig-eval-noise-temporal-cosine repeats the panel on the semantic axis: y is cosine distance between the surface-name and gold-label embeddings (E5-multilingual F16, L2-normalised — so cosine distance = 1 − dot product, on the same scale as the F16 probe). The interpretation flips relative to JW: if retrospective references drift *only* lexically (paraphrase) but stay semantically anchored, the cloud should stay flat as x grows; if they drift semantically as well (different terminology or focus when describing the act decades later), the cloud should rise. Reading the two figures together separates the two failure modes.

```{r}
#| label: fig-eval-noise-temporal-cosine
#| fig-cap: "Same panel as @fig-eval-noise-temporal but with cosine distance (E5-multilingual F16 embeddings, L2-normalised) on y. Pairs are dropped when either the surface form or the gold label is absent from its respective embedding matrix."
#| fig-width: 8
#| fig-height: 5

surface_in_emb <- intersect(noise_diag$measure_name,
                            rownames(c0_us_embeddings))
gold_in_emb    <- intersect(noise_diag$gold_act_name,
                            rownames(c0_us_gold_embeddings))

noise_diag_cos <- noise_diag |>
  dplyr::filter(measure_name %in% surface_in_emb,
                gold_act_name %in% gold_in_emb)

S <- c0_us_embeddings[noise_diag_cos$measure_name, , drop = FALSE]
G <- c0_us_gold_embeddings[noise_diag_cos$gold_act_name, , drop = FALSE]
noise_diag_cos$cos_distance <- 1 - rowSums(S * G)

ggplot(noise_diag_cos, aes(year_diff, cos_distance, colour = tier_label)) +
  geom_vline(xintercept = 0, linetype = "dotted",
             colour = "grey40", linewidth = 0.3) +
  geom_point(alpha = 0.5, size = 1.6) +
  geom_smooth(se = FALSE, linewidth = 0.4, linetype = "dashed") +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Doc year − surface-referenced year",
       y = "Cosine distance (surface name → gold label)",
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

Finally, @fig-jw-vs-cosine puts the two distance measures head-to-head on the full eval pool (no year filter): one point per distinct `(measure_name, gold_act_name)` pair, x is the JW distance, y is the cosine distance. The correlation answers whether the embedding is buying us anything JW couldn't already give us — a strong positive correlation would mean E5 is essentially re-deriving lexical similarity, while a weak correlation means the two distance measures disagree on which pairs are similar, and there is genuine semantic signal in the embedding above and beyond surface overlap. Both Pearson (linear) and Spearman (rank) correlations are reported because the JW–cosine relationship is plausibly nonlinear.

```{r}
#| label: fig-jw-vs-cosine
#| fig-cap: "JW distance vs cosine distance for every distinct (measure_name, gold_act_name) pair in the eval pool. Colour marks the cleanest tier in which the pair appears (Tier 1 if at least one chunk pairs the surface form with the gold act via verbatim match). Pairs whose surface form or gold label is missing from its embedding matrix are dropped. Annotated correlations include the full pool; the dashed line is an unweighted linear fit."
#| fig-width: 8
#| fig-height: 5

pair_min_tier <- c0_eval_gold_pairs |>
  dplyr::group_by(measure_name, gold_act_name) |>
  dplyr::summarize(tier = min(tier, na.rm = TRUE), .groups = "drop") |>
  dplyr::mutate(tier_label = factor(tier, levels = c(1L, 2L),
                                    labels = c("Tier 1", "Tier 2")))

distance_pairs <- pair_min_tier |>
  dplyr::filter(measure_name  %in% rownames(c0_us_embeddings),
                gold_act_name %in% rownames(c0_us_gold_embeddings)) |>
  dplyr::mutate(jw_distance = stringdist::stringdist(
    measure_name, gold_act_name, method = "jw"))

S <- c0_us_embeddings[distance_pairs$measure_name, , drop = FALSE]
G <- c0_us_gold_embeddings[distance_pairs$gold_act_name, , drop = FALSE]
distance_pairs$cos_distance <- 1 - rowSums(S * G)

cor_pearson  <- stats::cor(distance_pairs$jw_distance,
                           distance_pairs$cos_distance, method = "pearson")
cor_spearman <- stats::cor(distance_pairs$jw_distance,
                           distance_pairs$cos_distance, method = "spearman")

ggplot(distance_pairs, aes(jw_distance, cos_distance, colour = tier_label)) +
  geom_point(alpha = 0.5, size = 1.6) +
  geom_smooth(method = "lm", se = TRUE,
              linewidth = 0.4, linetype = "dashed") +
  annotate("label", x = Inf, y = -Inf,
           label = sprintf("Pearson  r = %.2f\nSpearman ρ = %.2f\nn = %d pairs",
                           cor_pearson, cor_spearman, nrow(distance_pairs)),
           hjust = 1.05, vjust = -0.2, size = 3.2,
           label.size = 0, fill = scales::alpha("white", 0.7)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "JW distance (surface name → gold label)",
       y = "Cosine distance (surface name → gold label)",
       colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

## Method 2 — Multilingual embedding + HDBSCAN

Switching from string distance to embedding distance is what the within-act geometry from §Method 1 demanded. Method 2 embeds each measure name with `intfloat/multilingual-e5-large-instruct`, L2-normalises the vectors so cosine ranking equals euclidean ranking, then runs HDBSCAN over the resulting distance matrix. Three `min_cluster_size` values are swept — 2 (most permissive), 3, and 5 (close to the median gold-act size). Singleton "noise" points HDBSCAN refuses to merge are kept as their-own-clusters so the pairwise scoring treats them as under-merges, not over-merges (the recoverable failure mode the Framing committed to).

```{r}
#| label: prep-hdb-ladder

shape_hdb_ladder <- function(eval_tier, blocking) {
  format_method_ladder(c0_hdbscan_metrics, eval_tier = eval_tier) |>
    dplyr::mutate(
      `min_cluster_size` = as.integer(
        stringr::str_extract(variant_id, "(?<=_m)\\d+(?=_)")),
      .blocking = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                                 "yw2", "unblocked")
    ) |>
    dplyr::filter(.blocking == blocking) |>
    dplyr::arrange(`min_cluster_size`) |>
    dplyr::select(-variant_id, -.blocking)
}

render_hdb_ladder <- function(ladder) {
  ladder |>
    dplyr::relocate(`min_cluster_size`) |>
    tinytable::tt() |>
    tt_theme_report()
}

ladder_hdb_t1   <- shape_hdb_ladder("tier1",  "unblocked")
ladder_hdb_t12  <- shape_hdb_ladder("tier12", "unblocked")
ladder_hdb_yw_t1  <- shape_hdb_ladder("tier1",  "yw2")
ladder_hdb_yw_t12 <- shape_hdb_ladder("tier12", "yw2")
```

```{r}
#| label: tbl-hdb-metrics
#| tbl-cap: "HDBSCAN unblocked, Tier 1 evaluation. Same metric definitions as @tbl-jw-metrics."

render_hdb_ladder(ladder_hdb_t1)
```

```{r}
#| label: tbl-hdb-metrics-t12
#| tbl-cap: "HDBSCAN unblocked, Tier 1 + Tier 2 evaluation."

render_hdb_ladder(ladder_hdb_t12)
```

```{r}
#| label: fig-hdb-pr
#| fig-cap: "HDBSCAN Precision vs Recall across the (min_cluster_size × year-blocking) grid, Tier 1+2 evaluation. Upper-left is the precision-favouring corner the asymmetric-cost framing prefers."
#| fig-width: 8
#| fig-height: 4

c0_hdbscan_metrics |>
  dplyr::filter(metric %in% c("pairwise_precision", "pairwise_recall"),
                eval_tier == "tier12") |>
  dplyr::select(variant_id, metric, value) |>
  tidyr::pivot_wider(names_from = metric, values_from = value) |>
  dplyr::mutate(
    min_cluster_size = as.integer(stringr::str_extract(variant_id, "(?<=_m)\\d+(?=_)")),
    blocking         = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                                      "year ±2", "unblocked")
  ) |>
  ggplot(aes(pairwise_recall, pairwise_precision, colour = blocking,
             label = sprintf("m=%d", min_cluster_size))) +
  geom_point(size = 2) +
  geom_text(size = 3, hjust = -0.2, vjust = -0.4,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Pairwise Recall", y = "Pairwise Precision", colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

```{r}
#| label: fig-umap-clusters
#| fig-cap: "UMAP projection of E5 embeddings. Left: coloured by predicted HDBSCAN cluster (best unblocked variant by Tier 1 F1); singleton/noise names in grey. Right: coloured by gold act for names in the unambiguous eval pool; names outside the eval pool in grey. Same 2D coordinates in both panels."
#| fig-width: 10
#| fig-height: 5

set.seed(20260523L)
umap_xy <- uwot::umap(c0_us_embeddings,
                      n_neighbors = 15L, min_dist = 0.1,
                      metric = "cosine", verbose = FALSE)
umap_df <- tibble::tibble(
  measure_name = rownames(c0_us_embeddings),
  umap_x = umap_xy[, 1L],
  umap_y = umap_xy[, 2L]
)

best_hdb_variant_f1 <- c0_hdbscan_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier1",
                stringr::str_detect(variant_id, "unblocked")) |>
  dplyr::slice_max(value, n = 1L, with_ties = FALSE) |>
  dplyr::pull(variant_id)

hdb_best <- c0_hdbscan_clusters |>
  dplyr::filter(variant_id == best_hdb_variant_f1) |>
  dplyr::transmute(
    measure_name,
    cluster_label = dplyr::if_else(n_members >= 2L,
                                   sprintf("c%d", cluster_id),
                                   NA_character_)
  )

gold_unamb <- c0_eval_gold_pairs |>
  dplyr::filter(!ambiguous) |>
  dplyr::distinct(measure_name, gold_act_name)

p_pred <- umap_df |>
  dplyr::left_join(hdb_best, by = "measure_name") |>
  ggplot(aes(umap_x, umap_y, colour = cluster_label)) +
  geom_point(size = 1.4, alpha = 0.85) +
  scale_colour_discrete(na.value = "grey75", guide = "none") +
  labs(title = sprintf("Predicted HDBSCAN clusters (variant %s)",
                       best_hdb_variant_f1),
       x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(axis.text = element_blank(), panel.grid.minor = element_blank())

p_gold <- umap_df |>
  dplyr::left_join(gold_unamb, by = "measure_name") |>
  dplyr::mutate(gold_label = stringr::str_trunc(gold_act_name, 30)) |>
  ggplot(aes(umap_x, umap_y, colour = gold_label)) +
  geom_point(size = 1.4, alpha = 0.85) +
  scale_colour_discrete(na.value = "grey75", guide = "none") +
  labs(title = "Gold act (unambiguous eval names)",
       x = NULL, y = NULL) +
  theme_minimal(base_size = 9) +
  theme(axis.text = element_blank(), panel.grid.minor = element_blank())

p_pred + p_gold + patchwork::plot_layout(ncol = 2)
```

### Reading the HDBSCAN results

The headline comparison is @fig-hdb-pr against @fig-jw-pr: if HDBSCAN's frontier sits up and to the right of JW's, embeddings have bought us what JW couldn't. The UMAP scatter is diagnostic — predicted clusters that overlap a single gold colour on the right panel are correct merges; predicted clusters spanning multiple gold colours are over-merges; gold colours fragmented across multiple predicted clusters are under-merges. The Synthesis section at the bottom of the notebook applies the same precision-friendly shortlist rule used for JW to pick a candidate variant.

## Method 3 — Year-blocked variants

Method 3 reruns the same HDBSCAN sweep with the year-window mask from §Method 1: pairs whose chunk years differ by more than 2 are forced to maximum distance and so cannot merge. The prior is that the same act tends to surface in a narrow temporal window, so the mask should reduce cross-decade over-merges without sacrificing recall.

```{r}
#| label: tbl-hdb-yw2-metrics
#| tbl-cap: "HDBSCAN year-blocked (±2), Tier 1 evaluation."

render_hdb_ladder(ladder_hdb_yw_t1)
```

```{r}
#| label: tbl-hdb-yw2-metrics-t12
#| tbl-cap: "HDBSCAN year-blocked (±2), Tier 1 + Tier 2 evaluation."

render_hdb_ladder(ladder_hdb_yw_t12)
```

The year-blocked points are shown alongside the unblocked points in @fig-hdb-pr (different colour). The JW year-blocking finding from §Method 1 was modestly precision-friendly with negligible recall cost; if the embedding equivalent shifts the front in the same direction, the temporal prior is structural rather than JW-specific and should carry into Methods 4–5 when they land.

## Methods 2 + 3 with UMAP dimensionality reduction

The F16 and FP32 embedding-geometry probes above both failed the `separation ≥ 0.05` and `top-1 NN same-act rate ≥ 0.40` floors on the gold pool. The verdict text considered two interpretations — F16 quantization noise, and the embedding-model ceiling — but did not weigh a third: **high-dimensional distance concentration**. `intfloat/multilingual-e5-large-instruct` returns 1024-dimensional vectors, and the unblocked HDBSCAN above operates on the full-dimensional cosine distance matrix. Density-based clustering is known to degrade above k≈50 dim through hubness and farthest-to-nearest ratio compression; the standard fix in the BERTopic / Top2Vec neighbourhood is to interpose UMAP between the embedding and the clusterer, collapsing the ambient dim to ~5–10 and (optionally) packing within-cluster points tightly via `min_dist = 0`. This section ports that recipe into Methods 2 + 3 and reads the with-vs-without delta against the no-UMAP baseline above.

The sweep grid is the cartesian product of three UMAP hyperparameters with the existing HDBSCAN grid:

- `n_neighbors ∈ {5, 15, 30}` — uwot default 15; the small end emphasises local structure, the large end emphasises global structure.
- `n_components ∈ {2, 5, 10}` — uwot default 2; BERTopic's standard is 5; 10 is the upper end of clustering practice.
- `min_dist ∈ {0.0, 0.01, 0.1}` — uwot default 0.01; 0.0 is BERTopic's tight-packing default; 0.1 matches the existing diagnostic UMAP plot at @fig-umap-clusters.
- Inner sweep matches the no-UMAP variants: `min_cluster_size ∈ {2, 3, 5}` and `year_window ∈ {unblocked, ±2}`.

That gives 27 UMAP cells × 6 HDBSCAN cells = 162 variants. Metric `"cosine"` is fixed (the input embeddings are L2-normalised); UMAP is reduced once per UMAP cell and reused across the inner HDBSCAN sub-cells. The seed is fixed at the target level (`run_hdbscan_umap_clusters_grid(seed = 20260526L)`) so the entire sweep is deterministic under `tar_make()`. Distance metric inside HDBSCAN switches from cosine on the raw embeddings to Euclidean on the UMAP output (UMAP does not preserve L2-normalisation, so cosine on the reduced space is not meaningful); the year-window mask uses a `max(d) + 1` sentinel instead of the cosine `1.0` cap.

```{r}
#| label: prep-hdb-umap

# Parse the four sweep dimensions out of the variant_id once so every chunk
# below can read them by name.
parse_umap_variant <- function(metrics_or_clusters_df) {
  metrics_or_clusters_df |>
    dplyr::mutate(
      umap_n_neighbors  = as.integer(stringr::str_extract(variant_id,
                                                          "(?<=nn)\\d+")),
      umap_n_components = as.integer(stringr::str_extract(variant_id,
                                                          "(?<=nc)\\d+")),
      umap_min_dist     = as.numeric(stringr::str_extract(variant_id,
                                                          "(?<=md)\\d+\\.\\d+")),
      min_cluster_size  = as.integer(stringr::str_extract(variant_id,
                                                          "(?<=_m)\\d+(?=_)")),
      blocking          = dplyr::if_else(stringr::str_detect(variant_id, "yw2"),
                                         "year ±2", "unblocked")
    )
}

hdb_umap_metrics_parsed <- parse_umap_variant(c0_hdbscan_umap_metrics)
```

```{r}
#| label: prep-hdb-umap-shortlist

# Per-(blocking, min_cluster_size) winner using the precision-friendly rule:
# lowest over_merge_count first, then highest F1. Returns variant_ids to join
# back against the formatted ladder for display.
pick_umap_winners <- function(eval_tier_value) {
  c0_hdbscan_umap_metrics |>
    dplyr::filter(metric == "pairwise_f1",
                  eval_tier == eval_tier_value) |>
    parse_umap_variant() |>
    dplyr::group_by(blocking, min_cluster_size) |>
    dplyr::arrange(over_merge_count, dplyr::desc(value),
                   .by_group = TRUE) |>
    dplyr::slice_head(n = 1L) |>
    dplyr::ungroup() |>
    dplyr::select(variant_id, blocking, min_cluster_size,
                  umap_n_neighbors, umap_n_components, umap_min_dist)
}

assemble_umap_shortlist <- function(eval_tier_value) {
  winners <- pick_umap_winners(eval_tier_value)
  formatted <- format_method_ladder(c0_hdbscan_umap_metrics,
                                    eval_tier = eval_tier_value)
  winners |>
    dplyr::inner_join(formatted, by = "variant_id") |>
    dplyr::arrange(blocking, min_cluster_size) |>
    dplyr::transmute(
      `Blocking`       = blocking,
      `min_cluster_size`,
      `n_neighbors`    = umap_n_neighbors,
      `n_components`   = umap_n_components,
      `min_dist`       = sprintf("%.3f", umap_min_dist),
      Precision, Recall, F1, ARI, `Over:Under`, `n clusters`
    )
}
```

```{r}
#| label: tbl-hdb-umap-best
#| tbl-cap: "Best UMAP cell per (min_cluster_size × year-blocking) sub-cell, Tier 1 evaluation. Within each sub-cell we sort by `over_merge_count` ascending then F1 descending (same shortlist rule as @tbl-hdb-pick) and report the winner. Cells: point estimate [95% bootstrap CI]."

assemble_umap_shortlist("tier1") |> tt() |> tt_theme_report()
```

```{r}
#| label: tbl-hdb-umap-best-t12
#| tbl-cap: "Same shortlist rule on Tier 1 + Tier 2 evaluation (more power, more noise)."

assemble_umap_shortlist("tier12") |> tt() |> tt_theme_report()
```

```{r}
#| label: fig-hdb-umap-pr
#| fig-cap: "Pairwise Precision vs Recall, UMAP-reduced HDBSCAN, Tier 1+2 evaluation. Each small point is one (UMAP cell × HDBSCAN cell) variant, faceted by `n_components` to show how the BERTopic-style choice of intermediate dimensionality moves the front. The labelled large diamond is the best no-UMAP HDBSCAN variant from §Method 2/3 (same shortlist rule), included as a reference. Upper-left is the precision-favouring corner the asymmetric-cost framing prefers."
#| fig-width: 10
#| fig-height: 5

baseline_best_id <- c0_hdbscan_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier12") |>
  dplyr::slice_min(over_merge_count, n = 1L, with_ties = TRUE) |>
  dplyr::slice_max(value, n = 1L, with_ties = FALSE) |>
  dplyr::pull(variant_id)

baseline_pr <- c0_hdbscan_metrics |>
  dplyr::filter(variant_id == baseline_best_id,
                metric %in% c("pairwise_precision", "pairwise_recall"),
                eval_tier == "tier12") |>
  dplyr::select(metric, value) |>
  tidyr::pivot_wider(names_from = metric, values_from = value)

umap_pr <- hdb_umap_metrics_parsed |>
  dplyr::filter(metric %in% c("pairwise_precision", "pairwise_recall"),
                eval_tier == "tier12") |>
  dplyr::select(variant_id, metric, value,
                umap_n_neighbors, umap_n_components, umap_min_dist,
                min_cluster_size, blocking) |>
  tidyr::pivot_wider(names_from = metric, values_from = value)

ggplot(umap_pr, aes(pairwise_recall, pairwise_precision,
                    colour = factor(umap_n_components),
                    shape = blocking)) +
  geom_point(size = 1.8, alpha = 0.75) +
  geom_point(data = baseline_pr,
             aes(pairwise_recall, pairwise_precision),
             inherit.aes = FALSE,
             shape = 18, size = 5, colour = "black") +
  geom_text(data = baseline_pr,
            aes(pairwise_recall, pairwise_precision,
                label = "no-UMAP baseline"),
            inherit.aes = FALSE,
            hjust = -0.15, vjust = -0.3, size = 3) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(x = "Pairwise Recall", y = "Pairwise Precision",
       colour = "n_components", shape = NULL) +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

```{r}
#| label: fig-umap-separation-probe
#| fig-cap: "Embedding-geometry probe repeated on the UMAP-reduced space. For each UMAP cell we compute, on the gold eval pool (Tier 1 only — strips Tier 2 label noise to match the F16/FP32 probe at @tbl-f16-probe), the median Euclidean distance among pairs sharing a gold act vs the median across pairs from different gold acts. `separation = between − within`; positive means the UMAP reduction preserves act-level structure. Each point is one UMAP cell; colour marks `n_components`. If separation rises materially above 0 across cells (the no-UMAP probe failed even after L2-normalisation), high-D distance concentration was the binding constraint and the no-UMAP verdict's third hypothesis is the live one."
#| fig-width: 9
#| fig-height: 5

eval_tier1_pool <- c0_eval_gold_pairs |>
  dplyr::filter(tier == 1L, !ambiguous) |>
  dplyr::distinct(measure_name, gold_act_name)

within_between_separation <- function(reduced, eval_pool) {
  eval_in <- eval_pool |>
    dplyr::filter(measure_name %in% rownames(reduced))
  acts_with_pairs <- eval_in |>
    dplyr::count(gold_act_name) |>
    dplyr::filter(n >= 2L) |>
    dplyr::pull(gold_act_name)
  eval_in <- eval_in |>
    dplyr::filter(gold_act_name %in% acts_with_pairs)
  if (nrow(eval_in) < 4L) return(tibble::tibble(
    median_within = NA_real_, median_between = NA_real_,
    separation = NA_real_, n_names = nrow(eval_in)))

  names_vec <- eval_in$measure_name
  acts_vec  <- eval_in$gold_act_name
  d <- as.matrix(stats::dist(reduced[names_vec, , drop = FALSE],
                             method = "euclidean"))
  same_act <- outer(acts_vec, acts_vec, FUN = "==")
  upper <- upper.tri(d)
  within  <- d[upper & same_act]
  between <- d[upper & !same_act]
  mw <- stats::median(within,  na.rm = TRUE)
  mb <- stats::median(between, na.rm = TRUE)
  tibble::tibble(median_within = mw, median_between = mb,
                 separation = mb - mw, n_names = nrow(eval_in))
}

umap_seed <- 20260526L
sep_by_cell <- purrr::pmap_dfr(c0_umap_grid,
  function(n_neighbors, n_components, min_dist) {
    reduced <- umap_reduce_embeddings(
      c0_us_embeddings,
      n_neighbors  = n_neighbors,
      n_components = n_components,
      min_dist     = min_dist,
      seed         = umap_seed
    )
    res <- within_between_separation(reduced, eval_tier1_pool)
    dplyr::bind_cols(
      tibble::tibble(n_neighbors  = as.integer(n_neighbors),
                     n_components = as.integer(n_components),
                     min_dist     = min_dist),
      res
    )
  })

ggplot(sep_by_cell,
       aes(min_dist, separation,
           colour = factor(n_components),
           shape  = factor(n_neighbors))) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey40") +
  geom_point(size = 2.4, alpha = 0.85,
             position = position_jitter(width = 0.005, height = 0,
                                        seed = 20260526L)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(x = "UMAP min_dist", y = "Separation (between − within, Euclidean)",
       colour = "n_components", shape = "n_neighbors") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "top")
```

### Reading the UMAP results

The UMAP variant set is large (27 cells), so the precision-friendly shortlist tables above are the primary read. The asymmetric-cost lens applies unchanged: among UMAP cells with low `over_merge_count`, pick the F1 winner; compare against the no-UMAP baseline diamond in @fig-hdb-umap-pr. Three patterns to look for:

- **UMAP shifts the front up-and-right of the no-UMAP diamond.** If the cloud of UMAP variants in @fig-hdb-umap-pr sits at meaningfully higher precision AND comparable-or-higher recall, BERTopic-style dim-reduction is the load-bearing improvement and C0 should deploy with UMAP.
- **UMAP recovers separation above the F16/FP32 probe floor.** Positive `separation` values in @fig-umap-separation-probe confirm high-D distance concentration was the binding constraint that the §"Embedding-geometry probe" verdict missed. Negative or near-zero values across all cells means dim-reduction is *not* the fix and the writeup needs to escalate to a different embedding family (the "model ceiling" branch of the existing verdict becomes the live interpretation).
- **`n_components` is the load-bearing UMAP hyperparameter.** Theory says intermediate dim (5–10) should beat both extremes (2 = throws away too much, 30+ = retains the concentration problem). If the facets in @fig-hdb-umap-pr are flat in `n_components`, the UMAP improvement is robust to that choice. If they aren't, that's the parameter the synthesis section needs to fix at the winning value.

## Method 4 — Hybrid (embedding nearest-neighbours + LLM pairwise judge)

::: {.callout-warning title="Phase B pending"}
The hybrid pipeline (embedding top-k → Haiku YES/NO verifier → union-find) lands once Phase B is built. Expected to dominate the cost/quality frontier: cheap blocker + targeted LLM verification only on the candidate pairs that survive embedding similarity.
:::

## Method 5 — Stateful LLM canonical-list build

::: {.callout-warning title="Phase B pending"}
The stateful walker (Haiku reads each doc's measure names, matches to existing canonicals or proposes new ones) lands once Phase B is built. Will include a multi-seed shuffle probe for order sensitivity.
:::

## Method comparison ladder

```{r}
#| label: tbl-ladder
#| tbl-cap: "Method comparison — best variant per method by Tier 1 F1, Tier 1 evaluation. Methods 4 + 5 are still Phase B-pending."

# Pick best by numeric F1 (read from the raw metrics tibble, not the
# formatted ladder cells where slicing is lexicographic).
best_jw_id  <- c0_jw_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier1") |>
  dplyr::slice_max(value, n = 1L, with_ties = FALSE) |>
  dplyr::pull(variant_id)
best_hdb_id <- c0_hdbscan_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier1") |>
  dplyr::slice_max(value, n = 1L, with_ties = FALSE) |>
  dplyr::pull(variant_id)

best_hdb_umap_id <- c0_hdbscan_umap_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier1") |>
  dplyr::slice_max(value, n = 1L, with_ties = FALSE) |>
  dplyr::pull(variant_id)

best_jw_row <- format_method_ladder(c0_jw_metrics, eval_tier = "tier1") |>
  dplyr::filter(variant_id == best_jw_id) |>
  dplyr::transmute(
    Method  = "JW (best variant)",
    Variant = best_jw_id,
    Precision, Recall, F1, ARI, Purity, `Over:Under`, `n clusters`
  )
best_hdb_row <- format_method_ladder(c0_hdbscan_metrics, eval_tier = "tier1") |>
  dplyr::filter(variant_id == best_hdb_id) |>
  dplyr::transmute(
    Method  = "HDBSCAN (best variant)",
    Variant = best_hdb_id,
    Precision, Recall, F1, ARI, Purity, `Over:Under`, `n clusters`
  )
best_hdb_umap_row <- format_method_ladder(c0_hdbscan_umap_metrics,
                                          eval_tier = "tier1") |>
  dplyr::filter(variant_id == best_hdb_umap_id) |>
  dplyr::transmute(
    Method  = "HDBSCAN + UMAP (best variant)",
    Variant = best_hdb_umap_id,
    Precision, Recall, F1, ARI, Purity, `Over:Under`, `n clusters`
  )

dplyr::bind_rows(best_jw_row, best_hdb_row, best_hdb_umap_row) |>
  tt() |> tt_theme_report()
```

Three rows now — JW, no-UMAP HDBSCAN, and UMAP+HDBSCAN — all picked by Tier 1 F1 within their own method. Reading across the rows: if UMAP+HDBSCAN dominates the no-UMAP row on F1 *and* sits at lower Over:Under, BERTopic-style dim-reduction is the load-bearing improvement. If UMAP+HDBSCAN wins on F1 but trails on Over:Under, embeddings + UMAP bought recall at the cost of the precision-friendly profile the Framing demands — the Synthesis shortlist below will pick differently. Methods 4 + 5 will extend this table once they land.

## Out-of-distribution: Malaysia EN/BM

::: {.callout-warning title="Phase B pending"}
Once Phase B is built, the best variant from each method is applied to the Malaysia EN+BM measure pool (`malay_er_c1_measures`); we report the fraction of `malay_er_curated_matches` EN↔BM pairs that land in the same predicted cluster. Methods that rely on surface similarity (JW) are expected to fail here because the language families are different; multilingual embedding methods should hold up.
:::

## Cost diagnostics

```{r}
#| label: tbl-cost
#| tbl-cap: "Per-method cost. Methods 1–3 are zero-cost (pure-R or local Ollama). Methods 4–5 will populate from `logs/api_calls.csv` once wired."

tibble::tibble(
  Method = c("Method 1: JW single-linkage",
             "Method 2: E5-multilingual + HDBSCAN",
             "Method 3: E5-multilingual + HDBSCAN (year-blocked)"),
  `Tokens (input)`          = c("0", "≈4K", "≈4K"),
  `Tokens (output)`         = c("0", "0", "0"),
  `Pilot cost (USD)`        = c("0.00", "0.00", "0.00"),
  `Projected SEA cost (USD)`= c("0.00", "0.00", "0.00"),
  Notes = c("Pure-R, no API calls",
            "Local Ollama (jeffh/intfloat-multilingual-e5-large-instruct:f16); embeddings cached in `c0_us_embeddings`",
            "Reuses Method 2 embeddings; year-blocking applied at the distance-matrix step")
) |> tt() |> tt_theme_report()
```

::: {.callout-warning title="Methods 4 + 5 still pending"}
Phase B cost diagnostics for Method 4 (Haiku pairwise-judge) and Method 5 (Haiku stateful-builder) will populate from `logs/api_calls.csv` filtered by method tag once those targets run.
:::

## Synthesis

::: {.callout-note title="Methods 4 + 5 still pending"}
A final method + parameter recommendation waits for all five methods to be evaluated. Methods 1–3 (JW, HDBSCAN unblocked, HDBSCAN year-blocked) are now wired; Methods 4 (hybrid LLM judge) and 5 (stateful builder) remain to land.

Until then, the shortlist below applies the precision-friendly rule from the Framing to each finished method in isolation: among variants with the lowest over-merge count, pick the variant with the highest F1. Compare the JW pick against the HDBSCAN pick to see whether the semantic bridge embeddings provide outweighs whatever lexical-only precision JW retains.
:::

### Reading the JW results in isolation

```{r}
#| label: tbl-jw-pick
#| tbl-cap: "JW variant picked under the asymmetric-cost rule (Tier 1+2 eval): best F1 among variants with low over-merge"

jw_pick <- c0_jw_metrics |>
  dplyr::filter(metric == "pairwise_f1", eval_tier == "tier12") |>
  dplyr::arrange(over_merge_count, dplyr::desc(value))

jw_pick |>
  dplyr::transmute(
    variant_id,
    F1 = sprintf("%.3f", value),
    `CI` = sprintf("[%.3f, %.3f]", ci_lo, ci_hi),
    `Over-merge` = over_merge_count,
    `Under-merge` = under_merge_count,
    `n clusters` = n_predicted_clusters
  ) |>
  utils::head(5) |>
  tt() |> tt_theme_report()
```

This shortlist is intentionally sorted by `over_merge_count` ascending — the column the C0 user should weight most. The top row is the most precision-friendly JW variant; the F1 column says whether that precision came at an unacceptable recall hit.

### Reading the HDBSCAN results in isolation

```{r}
#| label: tbl-hdb-pick
#| tbl-cap: "HDBSCAN variant picked under the asymmetric-cost rule (Tier 1+2 eval): best F1 among variants with low over-merge. Union of no-UMAP variants (from `c0_hdbscan_metrics`) and UMAP-reduced variants (from `c0_hdbscan_umap_metrics`); `Method` marks which family each row comes from."

dplyr::bind_rows(
  c0_hdbscan_metrics |>
    dplyr::filter(metric == "pairwise_f1", eval_tier == "tier12") |>
    dplyr::mutate(Method = "HDBSCAN"),
  c0_hdbscan_umap_metrics |>
    dplyr::filter(metric == "pairwise_f1", eval_tier == "tier12") |>
    dplyr::mutate(Method = "HDBSCAN + UMAP")
) |>
  dplyr::arrange(over_merge_count, dplyr::desc(value)) |>
  dplyr::transmute(
    Method,
    variant_id,
    F1            = sprintf("%.3f", value),
    `CI`          = sprintf("[%.3f, %.3f]", ci_lo, ci_hi),
    `Over-merge`  = over_merge_count,
    `Under-merge` = under_merge_count,
    `n clusters`  = n_predicted_clusters
  ) |>
  utils::head(8) |>
  tt() |> tt_theme_report()
```

Same sort rule as the JW shortlist. Reading across the three families (JW above, no-UMAP HDBSCAN, UMAP+HDBSCAN): the deployment pick is the top row of this table. If a UMAP+HDBSCAN variant displaces no-UMAP HDBSCAN at the top, deploy with UMAP for Malaysia. If both no-UMAP rows trail JW on over-merge, the precision-friendly C0 user keeps JW as the conservative default and treats the embedding methods as recall-extenders — Method 4's hybrid pipeline is the natural reconciliation.

