---
title: "C1: Measure Identification"
subtitle: "H&K Stages S0-S3 — final codebook (v0.7.0) and development history"
date: today
date-format: long
execute:
  cache: true
---

```{r}
#| label: setup
#| cache: false
#| message: false

pacman::p_load(targets, tidyverse, here, tinytable, patchwork, scales)

here::i_am("notebooks/c1_measure_id.qmd")
tar_config_set(store = here("_targets"))

source(here("R/tt_theme.R"))
source(here("R/iteration_reporting.R"))
source(here("R/codebook_stage_0.R"))

# Parsed development history (all codebooks); live final-iteration targets.
il          <- tar_read(iteration_logs)
c1_codebook <- tar_read(c1_codebook)
c1_neg      <- tar_read(c1_negative_examples)
aligned     <- tar_read(aligned_data)
s2_eval     <- tar_read(c1_s2_eval)
s3          <- tar_read(c1_s3_results)

# Select the latest formal (Claude) C1 iteration within a parsed-log tibble.
# Recomputed from the log on every render, so adding an iteration re-points the
# figures/tables automatically.
latest <- function(tbl) {
  tbl |>
    dplyr::filter(codebook == "c1", grepl("claude", model)) |>
    dplyr::filter(iteration == max(iteration))
}
```

C1 implements R&R Step 2 (Measure Identification): scanning a document chunk and
deciding whether any portion of it substantively discusses a specific fiscal
measure. The current deliverable is **v0.7.0**, accepted for deployment at S3
iteration 33. This notebook has two parts: the full final-iteration battery
(mimicking the Halterman & Keith presentables) and the development history that
produced it. The cross-codebook rollup lives in
[iteration_summary.qmd](iteration_summary.qmd).

# Part 1: The Final Codebook (v0.7.0)

## S0: Codebook and Prompt

C1 is a binary classification task with two classes:

- `FISCAL_MEASURE`: any portion of the chunk substantively discusses a specific
  enacted, proposed, or under-consideration fiscal action.
- `NOT_FISCAL_MEASURE`: everything else (general economic commentary, aggregates,
  administrative procedure, bare name-drops).

Under v0.7.0 the positive class additionally returns a structured `measures[]`
array (one entry per distinct act, with a per-measure country tag and three
relevance flags), which is what downstream codebooks consume.

```{r}
#| label: tbl-s0-summary
#| tbl-cap: "C1 codebook structure (live `c1_codebook`)."

tibble(
  Field = c("Name", "Version", "Classes", "Clarifications per class",
            "Neg. clarifications per class"),
  Value = c(
    c1_codebook$name,
    c1_codebook$version,
    as.character(length(c1_codebook$classes)),
    paste(map_int(c1_codebook$classes, ~length(.x$clarification)), collapse = ", "),
    paste(map_int(c1_codebook$classes, ~length(.x$negative_clarification)), collapse = ", ")
  )
) |>
  tt() |>
  tt_theme_report()
```

### Evaluation negatives

The zero-shot evaluation draws stratified hard negatives from `us_body`,
designed to test the most common confusion patterns.

```{r}
#| label: tbl-s0-negatives
#| tbl-cap: "Stratified hard-negative pool for evaluation."

c1_neg |>
  count(negative_type) |>
  mutate(pct = sprintf("%.1f", n / sum(n) * 100)) |>
  rename(Type = negative_type, Count = n, `%` = pct) |>
  tt() |>
  tt_theme_report()
```

```{r}
#| label: tbl-s0-data
#| tbl-cap: "Evaluation data sources."

n_passages <- sum(aligned$n_passages)
n_neg <- nrow(c1_neg)

tibble(
  Source = c("Positive (aligned_data passages)",
             "Negative (stratified hard negatives)", "Total"),
  Count = c(n_passages, n_neg, n_passages + n_neg)
) |>
  tt() |>
  tt_theme_report()
```

### The full prompt

This is the assembled system prompt the model saw for the US evaluation run,
with the `{country}` / `{country_iso}` tokens resolved (here, the United States).
At deployment the same template is re-rendered per country.

:::{.callout-note .column-page }

## Prompt exccerpt from C1

```{r}
#| echo: false



prompt_text <- construct_codebook_prompt(
  c1_codebook,
  country_iso = "US",
  country     = "united states"
)

# cat("```\n", prompt_text, "\n```\n", sep = "")

cat(prompt_text)
```

:::

## S1: Behavioral Tests (H&K Figure 3)

S1 checks that the model emits legal labels, can recover the class definitions
from memory, and is insensitive to class ordering. Test III (in-context example
recovery) is **skipped** for C1: v0.7.0 carries no in-codebook examples.



```{r}
#| label: fig-s1-behavioral
#| fig-cap: "S1 behavioral tests for C1 v0.7.0. Tests I/II shown as percent correct; Test IV shown as Fleiss' kappa with Landis (1977) reference bands."
#| fig-height: 5
#| warning: false

plot_s1_behavioral(latest(il$s1))
```

```{r}
#| label: tbl-s1-summary
#| tbl-cap: "S1 results for the latest C1 iteration."

latest(il$s1) |>
  filter(!skipped) |>
  transmute(
    Test = test,
    Value = if_else(grepl("IV", test),
                    sprintf("%.1f%% change", value * 100),
                    sprintf("%.0f%%", value * 100)),
    Threshold = if_else(grepl("IV", test),
                        sprintf("< %.0f%%", threshold * 100),
                        sprintf("≥ %.0f%%", threshold * 100)),
    Status = if_else(pass, "Pass", "Fail")
  ) |>
  tt() |>
  tt_theme_report()
```


[@tbl-s1-summary should have neat labels for test names, as those shown in @fig-s1-behavioral.]{.aside}

## S2: Zero-Shot Evaluation

Tier 1 chunks carry verbatim R&R passages (high-confidence labels); Tier 2
chunks are name-matched via the IKA heuristic (noisier). Tier 1 recall is the
metric we trust most.

```{r}
#| label: tbl-s2-metrics
#| tbl-cap: "S2 zero-shot metrics with 95% bootstrap CIs (latest C1 iteration)."

tt_s2_metrics_table(latest(il$s2))
```

### Confusion matrix

```{r}
#| label: tbl-s2-confusion
#| tbl-cap: "S2 confusion matrix (live `c1_s2_eval`). Rows = predicted, columns = true."

as.data.frame.matrix(s2_eval$confusion_matrix) |>
  rownames_to_column("Predicted") |>
  tt() |>
  tt_theme_report()
```

### Per-act recall

Acts with passage-level recall below 1 may indicate codebook weaknesses or
(more often at Tier 2) noisy labels.

```{r}
#| label: tbl-s2-act-recall
#| tbl-cap: "Acts with imperfect passage-level recall (live `c1_s2_eval$act_recall`)."

s2_eval$act_recall |>
  filter(recall < 1) |>
  arrange(recall) |>
  mutate(recall = sprintf("%.2f", recall)) |>
  rename(Act = act_name, Year = year, Chunks = n_chunks,
         `T1` = n_tier1, `T2` = n_tier2, Correct = n_correct, Recall = recall) |>
  tt() |>
  tt_theme_report()
```

### Multi-measure act recall

A v0.7.0-specific secondary diagnostic: for chunks discussing multiple acts, did
the model list each act?

```{r}
#| label: tbl-s2-multimeasure
#| tbl-cap: "Per-act recall on multi-act chunks (live `c1_s2_eval$multi_measure_act_recall`)."

if (!is.null(s2_eval$multi_measure_act_recall) &&
    nrow(s2_eval$multi_measure_act_recall) > 0) {
  s2_eval$multi_measure_act_recall |>
    arrange(recall) |>
    mutate(recall = sprintf("%.2f", recall)) |>
    rename(Act = act_name, Chunks = n_chunks,
           Recalled = n_recalled, Recall = recall) |>
    head(20) |>
    tt() |>
    tt_theme_report()
} else {
  cat("No multi-measure act-recall diagnostic available for this run.")
}
```

### Error examples

```{r}
#| label: tbl-s2-errors
#| tbl-cap: "Sample of S2 misclassifications (live `c1_s2_eval$error_analysis`)."

if (nrow(s2_eval$error_analysis) > 0) {
  s2_eval$error_analysis |>
    select(act_name, year, text_type, true_label, pred_label) |>
    slice_head(n = 15) |>
    rename(Act = act_name, Year = year, Type = text_type,
           True = true_label, Predicted = pred_label) |>
    tt() |>
    tt_theme_report()
} else {
  cat("No errors in S2 evaluation.")
}
```

## S3: Error Analysis (H&K Figure 4)

S3 perturbs the codebook to probe robustness: Test V (exclusion-criteria
minimal pairs), Test VI (generic labels), Test VII (swapped label names).

```{r}
#| label: fig-s3-behavioral
#| fig-cap: "S3 behavioral tests for C1 v0.7.0 (Tests V–VII), with the full-codebook baseline F1 as a reference line."
#| fig-height: 5
#| warning: false

plot_s3_behavioral(latest(il$s3), latest(il$s3_ablation))
```

### Ablation (H&K Table 4)

Components ranked by the accuracy/F1 drop when removed. Larger drops indicate
more load-bearing components.

```{r}
#| label: tbl-s3-ablation
#| tbl-cap: "Component ablation for the latest C1 iteration."

tt_ablation_table(latest(il$s3_ablation))
```

### Multi-measure schema diagnostics

Three v0.7.0 schema checks (live `c1_s3_results`): are we over-listing measures,
mis-tagging country of enactment, or under-listing on multi-act chunks?

```{r}
#| label: tbl-s3-overlisting
#| tbl-cap: "Over-listing: distribution of measures listed per FISCAL_MEASURE chunk."

ov <- s3$overlisting
tibble(
  Quantity = c("Mean measures / chunk", "Median (p50)", "p90",
               "Max", "Chunks with ≥3 measures", "FISCAL_MEASURE chunks"),
  Value = c(sprintf("%.2f", ov$mean_n_measures), sprintf("%.0f", ov$p50),
            sprintf("%.0f", ov$p90), as.character(ov$max_n_measures),
            as.character(ov$n_tail), as.character(ov$n_chunks))
) |>
  tt() |>
  tt_theme_report()
```

```{r}
#| label: tbl-s3-country
#| tbl-cap: "Country-of-enactment tag distribution across listed measures."

cd <- s3$country_distribution
tibble(
  Quantity = c("Domestic (US) measures", "OTHER (foreign) measures",
               "Chunks containing any OTHER measure"),
  Value = c(as.character(cd$n_domestic), as.character(cd$n_other),
            as.character(cd$n_chunks_with_other))
) |>
  tt() |>
  tt_theme_report()
```

```{r}
#| label: tbl-s3-underlisting
#| tbl-cap: "Under-listing: per-act recall on chunks expected to name multiple acts."

ul <- s3$under_listing
if (!is.null(ul)) {
  tibble(
    Quantity = c("Multi-act chunks tested", "Mean per-act recall",
                 "Chunks with full recall"),
    Value = c(as.character(ul$n_total), sprintf("%.2f", ul$mean_recall),
              as.character(ul$n_full_recall))
  ) |>
    tt() |>
    tt_theme_report()
} else {
  cat("No multi-act chunks in the S3 baseline set; under-listing not evaluated.")
}
```

## Manual Error Analysis (H&K Table 5)

The actual S3 gate is a human read of all 40 baseline chunks against the H&K
six-category framework (A–F). This is the headline result for v0.7.0.

```{r}
#| label: tbl-s3-manual
#| tbl-cap: "Manual error-analysis category distribution (latest C1 iteration)."

tt_manual_analysis_table(latest(il$s3_manual))
```

After excluding Category B (incorrect gold standard), the label-noise-adjusted
performance is:

```{r}
#| label: tbl-s3-biascorrected
#| tbl-cap: "Bias-corrected gate metrics (Category B excluded)."

tt_bias_corrected_table(latest(il$s3_manual_bias_corrected))
```

The two Tier 1 false negatives in v0.7.0 (a Category E pair) are genuine
semantic errors: the model recognized the relevant material but applied a
stricter "substantive" bar than the codebook authorizes. The hypothesized cause
is that the multi-measure schema raises the implicit threshold for what counts
as a measure worth listing. The gaps were judged non-blocking and recorded for a
future C1 revision; v0.7.0 was accepted for deployment.

# Part 2: Development History

Everything below is sourced from `prompts/iterations/c1.yml` via the parsed
`iteration_logs` target, so editing the log re-renders these figures and the
narrative.

## Timeline

```{r}
#| label: tbl-timeline
#| tbl-cap: "C1 iteration timeline (all runs, including exploratory non-Claude models)."

il$meta |>
  filter(codebook == "c1") |>
  arrange(iteration) |>
  transmute(
    Iter = iteration,
    Version = codebook_version,
    Stage = toupper(stage),
    Date = format(date, "%Y-%m-%d"),
    Model = str_remove(model, "-20251001$"),
    Pass = case_when(overall_pass ~ "✓", !overall_pass ~ "✗", TRUE ~ "—")
  ) |>
  tt() |>
  tt_theme_report()
```

## Performance trajectory

The S2 zero-shot trajectory across formal (Claude) iterations. The decisive jump
came with the v0.3.0 "detection reframe" (recall 29% → 81%), and the v0.4.0
relevance-filter reframe pushed Tier 1 recall to ceiling.

```{r}
#| label: fig-recall-trajectory
#| fig-cap: "C1 Tier 1 recall across formal iterations."
#| fig-height: 3.5
#| warning: false

plot_metric_trajectory(il$s2, "tier1_recall", codebook_filter = "c1")
```

```{r}
#| label: fig-combined-trajectory
#| fig-cap: "C1 combined recall across formal iterations."
#| fig-height: 3.5
#| warning: false

plot_metric_trajectory(il$s2, "combined_recall", codebook_filter = "c1")
```

```{r}
#| label: fig-precision-trajectory
#| fig-cap: "C1 precision across formal iterations."
#| fig-height: 3.5
#| warning: false

plot_metric_trajectory(il$s2, "precision", codebook_filter = "c1")
```

## What we learned, by version

For each codebook version, the concluding iteration's interpretation and
decision (verbatim from the log).

```{r}
#| label: version-narrative
#| echo: false
#| results: asis

version_story <- il$meta |>
  filter(codebook == "c1", !is.na(codebook_version)) |>
  group_by(codebook_version) |>
  slice_max(iteration, n = 1, with_ties = FALSE) |>
  ungroup() |>
  arrange(iteration)

for (i in seq_len(nrow(version_story))) {
  r <- version_story[i, ]
  cat(sprintf("\n### v%s — concluded at iteration %d (%s, %s)\n\n",
              r$codebook_version, r$iteration, toupper(r$stage),
              format(r$date, "%Y-%m-%d")))
  if (!is.na(r$interpretation) && nzchar(trimws(r$interpretation))) {
    cat("**What we found.** ", trimws(r$interpretation), "\n\n", sep = "")
  }
  if (!is.na(r$decision) && nzchar(trimws(r$decision))) {
    cat("**Decision.** ", trimws(r$decision), "\n\n", sep = "")
  }
}
```
