---
title: "Deployment Headline: Cross-Country Fiscal-Shock Inventory"
subtitle: "Full C1 → C0 → C2 deployment pipeline, by country"
execute:
  cache: true
---

```{r}
#| label: setup
#| cache: false


pacman::p_load(targets, tidyverse, tinytable, here, stringr, purrr, fixest)

here::i_am("notebooks/deployment.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R")) # pretty_motivation() + motivation palette
source(here("R/deployment_report.R"))
theme_set(theme_minimal())

tar_load(c(
    country_chunks,
    country_c1_predictions,
    country_c1_measures,
    country_c0_clusters,
    country_c0_acts,
    country_c2b,
    country_measure_pool
))

# Positionally zip the country identity onto every branched target, then flatten.
chunks <- bind_country(country_chunks, country_chunks)
c1_raw <- bind_country(country_c1_predictions, country_chunks)
c1_meas <- bind_country(country_c1_measures, country_chunks)

m <- compute_deployment_tallies(
    c0_acts      = bind_country(country_c0_acts, country_chunks),
    c0_clusters  = bind_country(country_c0_clusters, country_chunks),
    c2b          = bind_country(country_c2b, country_chunks),
    measure_pool = bind_country(country_measure_pool, country_chunks)
)

inv <- m$inventory
n_country <- dplyr::n_distinct(chunks$country_iso)
inv_pending <- nrow(inv) == 0L
```

## Framing

This is the deployment headline: the end-to-end output of the validated **C1 → C0 → C2** pipeline run on each deployment country's full document corpus. It presents the same figures and tables as the Malaysia EN/BM consistency report, but with the comparison axis swapped from **language** to **country**.

The pipeline runs independently per country. C1 identifies fiscal measures, C0 aggregates measure names into acts, and C2 classifies each act's motivation, sign, and exogeneity. Today only Malaysia is deployed, so every comparison figure shows a single country; the notebook auto-extends to a panel per country as Indonesia, Thailand, the Philippines, and Vietnam are onboarded.

::: {.column-margin}
This is a deployment inventory, not a validity test. There are no ground-truth labels for any deployment country; the act-level outputs are inputs requiring expert validation (Phase 2 success criteria: ≥80% expert agreement on C1, ≥70% on C2), not research findings.
:::

Each country's acts are clustered within that country only. We do **not** attempt to merge acts across countries: a Malaysian GST and a future Indonesian GST are distinct national measures, so there is no cross-country analog of the consistency report's joint EN/BM merge probe. The per-country C0 cluster summary below takes that slot.

### Scope

@fig-scope shows each country's corpus extent: pages and chunks per document year, summed across that year's documents. As more countries are added, comparable totals indicate comparable corpus depth; large gaps flag coverage asymmetries that bound everything downstream.

```{r}
#| label: fig-scope
#| fig-cap: "**Corpus scope.** Pages and chunks per country and document year, summed across that year's documents."
#| fig-width: 8
#| fig-height: 5

ps <- plot_deployment_scope(chunks)
if (is.null(ps)) cat("No chunk data available yet.\n") else print(ps)
```

[@fig-scope looks good, but the x axis should only display breaks every 5 years. So do all figures that plot years in the x axis.]{.aside}

[We need a table that shows, per country (rows) the headline tally of Documents, Pages, Chunks, Surfaced fiscal measures (only rank 1), Clustered fiscal measures, Exogenous fiscal measures. Also include the years range we're analysing in each country in the first row.]{.aside}

## C1-step comparability

Before aggregation, what set of raw fiscal measures does C1 surface for each country? This is where real extraction differences live, independent of how acts are later aggregated. @fig-c1-comparability counts the distinct fiscal measures C1 produced per year, by country.

```{r}
#| label: fig-c1-comparability
#| fig-cap: "**Distinct C1 fiscal measures per year, by country (pre-aggregation).** The raw extraction load that bounds any downstream aggregation."
#| fig-width: 8
#| fig-height: 3.5

pc <- plot_deployment_c1_comparability(c1_meas)
if (is.null(pc)) cat("No measures available yet.\n") else print(pc)
```

### Measure density and the rank-1 filter

The comparability count above is a post-filter count: the deployment filter keeps only the most-prominent (rank-1) motivated measure per chunk. Before that cut, C1 surfaces several measures per chunk. @fig-c1-density shows that raw load per country-year, and @fig-c1-fiscal-rate shows the share of chunks whose rank-1 measure is a motivated fiscal measure. Measure-dense years carry several times the typical load, and the rank-1 cut discards the rest.

```{r}
#| label: fig-c1-density
#| fig-cap: "**Motivated fiscal measures C1 surfaces per year, before the rank-1 deployment filter.** Dense years carry the heaviest load, which is what the rank-1 cut discards."
#| fig-width: 8
#| fig-height: 3.5

pd <- plot_deployment_c1_density(c1_raw)
if (is.null(pd)) cat("No motivated measures available yet.\n") else print(pd)
```

```{r}
#| label: fig-c1-fiscal-rate
#| fig-cap: "**Share of each year's chunks whose most-prominent (rank-1) measure is a motivated fiscal measure, by country.**"
#| fig-width: 8
#| fig-height: 3.5

pr <- plot_deployment_c1_fiscal_rate(c1_raw, chunks)
if (is.null(pr)) cat("No chunk data available yet.\n") else print(pr)
```


[Let's remove altogather the Rank-1 measurements and mentions to it. Always display Surfaced fiscal measures.]{.aside}

## C0 aggregation diagnostics

C0 collapses C1 measure names into act-level clusters. The figures below describe that aggregation per country.

### Compression

@fig-c0-perdoc shows the compression ratio, how many surfaced measures collapse into each distinct act, by year and country, averaged across the year's documents.

```{r}
#| label: fig-c0-perdoc
#| fig-cap: "**C0 compression (measures collapsed into distinct acts) per year and country.** Higher bars mean more measure variants per canonical act."
#| fig-width: 8
#| fig-height: 3.5

pd0 <- plot_deployment_c0_perdoc(m$per_doc)
if (is.null(pd0)) cat("No per-document aggregation available yet.\n") else print(pd0)
```

[@fig-c0-perdoc should show merge rate (surfaced / clusters) - 1 instead]{.aside}

### Per-country cluster summary

How many distinct acts does C0 produce for each country, and how many measure variants does each act absorb on average? @tbl-c0-summary is the per-country analog of the consistency report's joint merge-rate probe.

```{r}
#| label: tbl-c0-summary
#| tbl-cap: "Per-country C0 cluster summary: distinct acts, measure variants, and cluster sizes"

if (nrow(m$c0_summary) == 0L) {
  tibble(Note = "C0 clusters not yet built") |> tt() |> tt_theme_report()
} else {
  m$c0_summary |>
    transmute(
      Country = str_to_title(country),
      `Acts (clusters)` = n_clusters,
      `Measure variants` = n_variants,
      `Mean variants / act` = round(mean_members, 2),
      `Largest cluster` = max_members
    ) |>
    tt() |>
    tt_theme_report()
}
```

[@tbl-c0-summary should show the following columns per country: Measure variants (surfaced names), Acts (clusters), Merge rate, N singletons (Clusters with one distinct name), Largest cluster]{.aside}

```{r}
#| label: fig-enacted
#| fig-cap: |
#|    **Acts by their *enacted* status.** An act is said to be enacted if C2b found sufficient evidence in that regard.

country_c2b[[1]] |>
  drop_na(enacted) |>
  ggplot(aes(x = enacted)) +
  geom_bar()

```


[Please make @fig-enacted show neat enacted labels and axis labels]{.aside}

```{r}
#| label: fig-enacted-evidence
#| fig-cap: Probability of enacted by amount of evidence


country_c2b[[1]] |>
  mutate(enacted = as.numeric(enacted)) |>
  ggplot(aes(n_evidence_items, enacted)) +
  geom_smooth(se = F) +
  scale_x_log10() +
  labs(
    x = "Evidence items (log scale)",
    y = "P(enacted)"
  )

```

## Identification by frequency

Notice that the language used by @romer to define what constitutes as a "significant change to fiscal liabilities" over-identifies seemingly innocuous fiscal changes when operationalizing it through the LLM chain we developed. We showed whilst [developing C0](/notebooks/c0_aggregator.qmd) that we can recover @romer 's taxonomy of US fiscal changes by selecting a threshold of the most discussed acts in the corpus.

This section asks what is a sensible threshold for selecting meaningful fiscal changes that surfaced from our pipeline, that we know are over-identified. While RR's taxonomy included 49 distinct fiscal changes spanning 58 years (from 1945 to 2003), the raw acts aggregated by clusters found 805 such acts in 42 years in Malaysia (from 1981 to 2023). Thus we say RR's language, as operationalized by our pipeline, over-identifies fiscal changes.

To sift through the most relevant changes, without yet knowing the fiscal magnitude of each, we recur to a frequentist approach. We use two raw variables to map on to a relevance index. First, more relevant acts should have produced more **evidence items**, as surfaced by C2a. Moreover, more relevant acts should also be discussed in **more parts of the corpus**. So, we will want the acts for which we have a lot of evidence to extract their motivation from, and that were more widely discussed in the government documents.


Notice in @fig-relevance-scatter that, some acts have just one piece of evidence extracted from one chunk, in all likelihood, these acts are false positives (e.g. insignificant acts that RR would not have identified as significant enough); these acts should score the lowest with our relevance measure. On the other extreme, a single act surfaced some 250 evidence items from almost 50 different chunks; this act is most certainly important enough that RR would have identified it, so it should get the highest available score. Consider the following Cobb-Douglass relevance function

$$
\text{relevance} = \text{evidence items} \cdot \text{chunks addressed}
$$ 



Then, how is *relevance* accumulated by act? @fig-acts-chosen shows that 23 acts accumulate 80% of the relevance, as defined above. 





```{r}
docs_start <- m$per_doc$year |> min()
docs_end <- m$per_doc$year |> max()

headline <- country_c2b[[1]] |>
  filter(enacted, between(year, docs_start, docs_end)) |>
  arrange(desc(n_evidence_items), desc(n_chunks)) |>
  mutate(
    relevance = n_evidence_items * n_chunks
  ) |>
  arrange(desc(relevance)) |>
  mutate(
    relevance_per = relevance / sum(relevance),
    cum_relevance = cumsum(relevance_per),
    acts_chosen = row_number()
    )
```


:::: {layout-ncol=2 .column-page-right}

```{r}
#| label: fig-relevance-scatter
#| fig-cap: |
#|    **A measure of relevance**. Relevance as a mapping of evidence items and chunks were the act was found. 
#| fig-cap-location: bottom


headline |>
  ggplot(aes(n_chunks, n_evidence_items)) +
  geom_jitter(aes(size = relevance_per, color = (cum_relevance <= 0.80))) +
  scale_x_log10(n.breaks = 10) +
  scale_y_log10(n.breaks = 10) +
  scale_size_continuous(labels = scales::percent) +
  labs(
    x = "Chunks (log scale)",
    y = "Evidence items (log scale)",
    size = "% Relevance",
    color = "Chosen"
  ) 

```

```{r}
#| label: fig-acts-chosen
#| fig-cap: |
#|    **Choosing acts according to the accumulated evidence items**. 23 acts accumulate 80% of the relevance.
#| fig-cap-location: bottom

 

headline |>
  add_row(acts_chosen = 0, cum_relevance = 0) |>
  ggplot(aes(cum_relevance, acts_chosen)) +
  geom_step() +
  scale_y_log10(n.breaks = 20) +
  scale_x_continuous(labels = scales::percent, n.breaks = 6, limits = c(0, 1)) +
  geom_vline(xintercept = 0.8, linetype = "dashed", color = "darkgrey") +
  labs(
    x = "% relevance accumulated",
    y = "Acts chosen (log scale)"
  )


```


::::


```{r}
#| label: fig-chosen-count
#| fig-cap: |
#|    **Timeline of un/chosen acts**.

headline |>
  ggplot(aes(year, fill = (cum_relevance <= 0.8))) +
  geom_bar() +
  scale_y_log10(n.breaks = 6) +
  labs(
    x = "Year",
    y = "Acts (log scale)",
    fill = "Chosen"
  ) 
```


@tbl-top-20 shows the chosen acts. For example, the **Goods and Services Tax (GST)** pools 230 evidence items drawn from 46 chunks spanning 7 EN/BM surface forms that C0 merged into one act; C2b read that consolidated evidence and classified the act's motivation as deficit-driven (revenue increase).


```{r}
#| label: tbl-top-20
#| tbl-cap: |
#|    Acts that accumulate 80% of relevance.
#| tbl-column: body-outset

headline |>
  arrange(desc(relevance)) |>
  filter(cum_relevance <= 0.8) |>
  mutate(cum_relevance = scales::percent(cum_relevance, accuracy = 0.1)) |>
  select(
    `Act name` = act_name,
    Year = year,
    Motivation = pred_label,
    `Effect` = pred_sign,
    `Evidence items` = n_evidence_items,
    `Chunks` = n_chunks,
    `Accumulated relevance` = cum_relevance
    ) |>
  tt()
```


[Add a footnote to @tbl-top-20 explaining how the year variable was computed. The footnote should also describe what evidence items are, and what the effect variable is describing (see prompts/c2b_classification.yml). Also, note that using tt_theme_report() on this table showed horizontal lines between each row; this should not happen, horizontal lines should only indicate the start of the table, the end of the header and the end of the table. Also, remove the [c77] codes from the act names using regex; they are a pipeline mechanism and have no real-world meaning.]{.aside}

### Act inventory marginals

[Everything in this section and below should be filtered to the chosen acts with the 80% rule defined above.]{.aside}

Pooling all of a country's documents, @tbl-c0-counts reports the act count and the exogenous/endogenous split, and @fig-c0-labels shows the motivation-label marginal distribution.

```{r}
#| label: tbl-c0-counts
#| tbl-cap: "Per-country act inventory: counts and exogeneity split"

if (inv_pending) {
  tibble(Note = "C2 inventory not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_country$counts |>
    transmute(
      Country = str_to_title(country),
      `C0 acts` = n_acts,
      Exogenous = n_exogenous,
      Endogenous = n_endogenous
    ) |>
    tt() |>
    tt_theme_report()
}
```

```{r}
#| label: fig-c0-labels
#| fig-cap: "**Motivation-label marginal distribution by country.** How many acts each country assigns to each motivation category."
#| fig-width: 8
#| fig-height: 3.5

if (inv_pending) {
  cat("C2 inventory not yet built.\n")
} else {
  pl <- plot_deployment_c0_labels(m$per_country$labels)
  if (is.null(pl)) cat("No motivation labels available yet.\n") else print(pl)
}
```




## Headline: act inventory timeline

The pipeline produces, per country, a set of acts each carrying a C2 motivation label, sign, and exogeneity. The headline is a timeline: for each year and country, acts are counted into motivation × sign cells and drawn as a diverging stacked bar. Increases grow the bar upward, decreases downward, no-change acts sit at the baseline, colour encodes motivation, faceted by country.

The two figures differ only in the timing source. The primary one dates each act by its **source-document year** (always present); the secondary one by the year **extracted from the act name** (language-invariant numerals, but sparse: most acts carry no name-embedded year).

### Act-name year multiset

```{r}
#| label: fig-act-years
#| fig-cap: "**Act-name-year multiset by country** (years parsed from act names). Sparse: most acts carry no name-embedded year."
#| fig-width: 8
#| fig-height: 3.5

if (inv_pending || nrow(m$per_country$act_years) == 0L) {
  cat("No act-name years available yet.\n")
} else {
  pay <- plot_deployment_act_years(m$per_country$act_years)
  if (is.null(pay)) cat("No act-name years available yet.\n") else print(pay)
}
```

[There is no need for @fig-act-years, since above we show the distribution of acts by year and chosen status.]{.aside}

### Timeline by source-document year

```{r}
#| label: fig-timeline-docyear
#| fig-cap: "**Act counts by modal source-document year, by country.** Diverging stacked bars: increases grow upward, decreases downward, no-change acts sit at the baseline; colour = motivation."
#| fig-width: 9
#| fig-height: 7

p1 <- plot_deployment_act_timeline(inv, timing = "doc_year")
if (is.null(p1)) cat("Inventory not yet built.\n") else print(p1)
```

### Timeline by act-name year

```{r}
#| label: fig-timeline-actname
#| fig-cap: "**Act counts by year extracted from the act name, by country.** Same encodings as above; covers only the acts that carry a name-embedded year."
#| fig-width: 9
#| fig-height: 7

p2 <- plot_deployment_act_timeline(inv, timing = "act_name")
if (is.null(p2)) cat("No acts carry a name-embedded year yet.\n") else print(p2)
```


[Collapse @fig-timeline-actname and @fig-timeline-docyear to a single plot that shows only chosen acts by the single year variable we have been using above.]{.aside}

## Interpretation

Read the evidence in pipeline order:

- **C1 step.** The comparability, density, and fiscal-rate figures describe what raw material the extractor surfaces per country. Any extraction gap here bounds everything downstream and is not fixable by the aggregator.
- **C0 step.** The compression figure and cluster summary describe how aggressively C0 collapses measure variants into acts. A country whose corpus repeats the same handful of acts across many documents compresses more.
- **C2 / headline.** The timelines are the deliverable: each country's acts placed by year, direction, and motivation. They are the starting point for expert review, not a finding.

## Implications

The same pipeline, codebooks, and figures apply to every country: onboarding the next one adds a corpus and a configuration entry, and this notebook gains a panel. The act inventory each country produces is the artifact an expert reviews against the Phase 2 agreement targets; comparable structure across countries (similar compression, plausible motivation marginals, no degenerate years) is the signal that the country-agnostic codebooks transferred.
