---
title: "Manual Identification of Malaysia's Tax-Incentive Shocks"
subtitle: "AI-assisted narrative pass over the deployment corpus (Klemm & Van Parys 2012-style incentive side)"
number-sections: true
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, qs2)

here::i_am("notebooks/incentives_identification.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
set_theme(theme_minimal())

# Frozen, human-stamped incentive shock dataset (produced by the
# /identify-incentives skill). Loaded here so both the recall scorecard
# (@sec-recall) and the structured dataset (@sec-frozen) render from the single
# source of truth when it exists.
frozen_path <- here("data", "validated", "MY_INCENTIVE_shocks.qs")
have_frozen <- file.exists(frozen_path)
if (have_frozen) inc_frozen <- qs2::qs_read(frozen_path)
```

## Purpose

This note documents a focused, **AI-assisted manual identification** of changes to
Malaysia's *tax incentives and holidays* — tax holidays (Pioneer Status), investment
and reinvestment allowances, concessionary investor rates (Labuan, Principal Hub,
sectoral special rates), free/export/enterprise zones, sectoral and R&D incentives,
consumption-side holidays, and personal-income reliefs — run directly over the
deployment corpus. It is the incentive-side analogue of
`notebooks/spending_identification.qmd`, and it deliberately follows the spirit of
the narrative-identification literature [@romer2010; @klemm_empirical_2012]: close reading of
source documents, with the LLM pipeline as a retrieval-and-extraction aide, not as
the adjudicator.

The method mirrors our `/manual-analysis` protocol. Like the spending side and
**unlike** the statutory-rate side, **C1 does not pre-screen incentives** (C1 is
scoped to standing tax rates/liabilities), so identification is driven by **direct
reading of the corpus** (`country_body` / `country_chunks`), not the C1 measure
pool. The seed of the dataset is the set of concessionary regimes the
`/identify-cit` pass deliberately set aside. A human reads the surfaced evidence and
makes the final call on *what counts as an incentive change*, *how scattered surface
forms consolidate*, *whether the event belongs here or in the tax/spending sets*,
and *whether each change is exogenous*. Claude pattern-matches against the source
text; the researcher confers meaning.

::: {.callout-important}
The unit is one row per **legislative change** to an incentive (a new scheme, a
rate/duration change, a sectoral extension, a repeal) — the standing existence of a
long-running scheme is not a row. Each event lives in exactly one dataset:
incentives are *targeted, time-bounded, or base/carve-out-rate* measures; universal,
permanent statutory-rate changes belong to the CIT/PIT/consumption sets
(ownership-by-mechanism).
:::

::: {.callout-important}
The exogenous/endogenous column in @tbl-exogeneity is a **preliminary suggestion
with its supporting quote, pending expert adjudication**. Incentives skew exogenous
(FDI-attraction / industrial-policy / structural-competitiveness motives are
long-run, non-cyclical); only an explicit cyclical/crisis rationale reads as
endogenous. It is not a research finding, and it is carried *alongside* — never
replacing — C2b's downstream label.
:::

## What the corpus surfaced, and what C1 missed {#sec-recall}

Because C1 is tax-rate-scoped, the keyword scan of the C1 pool is expected to
surface almost no incentive events — that is itself a finding. The events below are
recovered by **reading the anchor documents directly** (Budget Speeches' incentive
announcements, Economic Reports, MIDA / investment-authority materials, the
Promotion of Investments Act and Income Tax Act incentive schedules). The mandatory
recall scorecard records what was searched and missed.

```{r}
#| label: tbl-recall
#| tbl-cap: "**Recall scorecard**. How each headline incentive change fared through the identification pass. The near-empty C1 floor is logged as a finding; 'Recovered manually' marks events recovered by direct reading."

recall <- if (have_frozen) {
  inc_frozen$recall_scorecard[[1]] |>
    dplyr::rename(Stage = stage, Outcome = outcome)
} else {
  tibble::tribble(
    ~Stage,                          ~Outcome,
    "C1 keyword scan",               "Expected near-empty: C1 is scoped to standing tax rates, surfaces ~0 incentive measures",
    "Direct corpus sweep",           "Primary method — read Budget Speech incentive announcements / Economic Report / MIDA / PIA sections directly",
    "Seed list (CIT carve-outs)",    "Labuan 3%; Principal Hub 10% (2019); biomass (2001); pharma/vaccine 0–10% (2021); manufacturing-relocation 15% (2023); DEB 5pp (1981)",
    "Recall checkpoint — PIA 1986",  "Promotion of Investments Act 1986 (Pioneer Status / ITA framework) — recovered from source",
    "Recall checkpoint — zones",     "MSC / Multimedia Super Corridor and Labuan offshore regime changes — recovered from source"
  )
}

recall |>
  tt() |>
  tt_theme_report()
```

## Headline incentive changes {#sec-headline}

```{r}
#| label: tbl-headline
#| tbl-cap: "**Headline incentive changes**. Identified changes to tax incentives, with mechanism category, underlying base, direction, rate (where applicable), magnitude, and the budget/statute that announced each. Renders from the frozen dataset once the /identify-incentives skill has stamped it."
#| eval: !expr have_frozen

inc_frozen |>
  dplyr::select(shock_id, act_label,
                Category = incentive_category, Base = tax_type,
                Direction = direction,
                `Δpp` = delta_pp, Magnitude = magnitude_note,
                Announced = announced_year, Effective = effective_year) |>
  tt() |>
  tt_theme_report()
```

::: {.content-visible when-format="html"}
```{r}
#| label: headline-pending
#| eval: !expr "!have_frozen"
#| echo: false
#| output: asis

cat("::: {.callout-note}\nThe frozen incentive dataset has not been produced yet.",
    "Run the `/identify-incentives` skill to identify, consolidate, and stamp the",
    "incentive changes; this section will then render the headline table.\n:::\n")
```
:::

## Motivation evidence and preliminary exogeneity {#sec-exogeneity}

For each incentive change the corpus carries motivation language. @tbl-exogeneity
pairs each act with the most diagnostic quote and a **preliminary**
exogenous/endogenous read. The default prior is exogenous (FDI-attraction,
industrial policy, structural competitiveness, or tax-competition reactivity are all
non-cyclical); only an explicit cyclical/crisis rationale reads as endogenous
("acknowledgment ≠ endogeneity").

```{r}
#| label: tbl-exogeneity
#| tbl-cap: "**Motivation evidence, pending adjudication**. The stated rationale for each incentive change and a preliminary exogenous/endogenous read. Suggestions anchored in quoted text, not settled classifications; the exogeneity calls are the object of expert adjudication, and C2b assigns the downstream label."
#| eval: !expr have_frozen

inc_frozen |>
  dplyr::select(shock_id,
                `Preliminary read` = exogenous_preliminary,
                `Motivation quote (source text)` = exogeneity_quote) |>
  tt() |>
  tt_theme_report()
```

## Structured frozen dataset {#sec-frozen}

The narrative identification above is the human-curated content; the
machine-readable form is frozen by the `/identify-incentives` skill to
`data/validated/MY_INCENTIVE_shocks.qs`, following the contract in
`docs/phase_1/incentive_shock_schema.md` (one row per announced incentive change,
with a `member_chunks` manifest linking each shock to its corpus chunks and a
mandatory recall scorecard). That frozen file — not this notebook's tribbles — is
the reference input the incentive pipeline binds and feeds to the frozen C2a/C2b for
motivation and exogeneity (`tar_read(incentive_shocks)`).

```{r}
#| label: tbl-frozen-manifest
#| tbl-cap: "**Member-chunk manifest**. Number of corpus chunks attributed to each incentive change, and how many were recovered (present in the corpus but not surfaced by C1 — for incentives, expected to be most). The recovered chunks drive the downstream C2a re-run."
#| eval: !expr have_frozen

inc_frozen |>
  dplyr::mutate(
    `Member chunks`    = purrr::map_int(member_chunks, nrow),
    `Recovered chunks` = purrr::map_int(recovered_chunks, nrow)
  ) |>
  dplyr::select(shock_id, act_label, `Member chunks`, `Recovered chunks`) |>
  tt() |>
  tt_theme_report()
```

## References {.unnumbered}

::: {#refs}
:::
