---
title: "Malaysia EN/BM Cross-Language Consistency Test"
subtitle: "Full C1 → C0 → C2 pipeline on parallel Economic Report pairs, 2014-2020 + 2022"
date: today
date-format: long
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, stringr)

here::i_am("notebooks/malay_consistency.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))

tar_load(c(
  malay_er_pair_years,
  malay_er_chunks,
  malay_er_c1_measures,
  malay_er_measure_pool,
  malay_er_c0_perdoc,
  malay_er_c0_perlang,
  malay_er_c0_joint,
  malay_er_consistency_metrics
))

m <- malay_er_consistency_metrics
inv <- m$inventory
inv_pending <- nrow(inv) == 0L
c0_pending  <- nrow(malay_er_c0_perlang) == 0L
```

## Framing

The C1 → C0 → C2 codebook pipeline was validated on US English documents. Malaysia has 38 BM-only documents in its ready corpus (primarily older Budget Speeches and Economic Reports). Before trusting LLM outputs on those, we need evidence that the pipeline produces equivalent fiscal reasoning regardless of input language.

This is a **consistency test, not a validity test**. We have no labels for Malaysia. EN outputs are treated as the reference because the codebooks crossed S3 gates on EN-only US data. Passing this test is necessary-but-not-sufficient for extending deployment to the **BM-only-document slice (38 of 99 ready documents)**; it does not gate the broader Malaysia deployment, and it is **not a substitute for expert agreement on the EN-side outputs** (Phase 2's success criteria: ≥80% on C1, ≥70% on C2 against expert review).

The test runs the **full pipeline independently on each language** and compares the outputs with **data/statistical evidence only**. It slices `country_chunks` to Economic Reports with parallel EN+BM coverage, runs its own C1 → C2a chain, then aggregates measure names into acts with the **C0 act aggregator** (M5 LLM canonical clustering, `prompts/c0_canonicalize.yml`, Haiku) at three scopes:

- **per-document** — aggregate within each language × year document (granular aggregation capability/limits);
- **per-language** — pool all EN documents / all BM documents (deployment-realistic; this scope feeds C2 and the headline timeline);
- **joint EN+BM** — pool both languages together (probes whether C0 can co-cluster the EN and BM surface forms of the same act).

There is **no auxiliary-API matching step**. The previous iteration used a Sonnet matcher to propose EN ↔ BM cluster pairs, but that injected an unverified LLM judgment into the headline (foreign comparators were matched to each other). We removed it: comparison is now distributional (tallies of act counts, exogenous/endogenous splits, motivation-label marginals, act-name-year multisets) plus **two timeline figures** that place each side's acts by timing, direction, and exogeneity, leaving paired interpretation to the reader.

::: {.callout-warning}
This test only speaks to **modern, professionally-translated MoF prose** (Economic Reports 2014-2020, 2022). It does NOT speak to pre-2009 Budget Speech BM, older Economic Report BM (pre-2014), or crisis-booklet BM. Those are different vintages and registers; deployment to them requires separate evidence.
:::

```{r}
#| label: scope-summary
#| results: 'asis'

cat(sprintf(
  "**Scope:** %d pair years × 2 languages = %d documents; %d total chunks.\n\n",
  length(malay_er_pair_years),
  2L * length(malay_er_pair_years),
  nrow(malay_er_chunks)
))
cat(sprintf("**Pair years:** %s\n\n",
            paste(malay_er_pair_years, collapse = ", ")))
```

## C1-step comparability

Before aggregation, does C1 surface a comparable set of raw fiscal measures on each side? This is the step where real extraction asymmetries live (independent of how acts are later aggregated). The table counts the distinct `measure_name` strings C1 produced per document, by language.

```{r}
#| label: tbl-c1-comparability
#| tbl-cap: "Distinct C1 measure names per year, by language (pre-aggregation)"

c1m_wide <- malay_er_c1_measures |>
  mutate(doc_language = if_else(str_detect(doc_id, "-BM$"), "bm", "en")) |>
  distinct(year, doc_language, measure_name) |>
  count(year, doc_language, name = "n_measures") |>
  pivot_wider(names_from = doc_language, values_from = n_measures,
              names_prefix = "n_", values_fill = 0L)

if (!"n_en" %in% names(c1m_wide)) c1m_wide$n_en <- 0L
if (!"n_bm" %in% names(c1m_wide)) c1m_wide$n_bm <- 0L

c1m_wide |>
  mutate(`Signed drift (BM − EN)` = n_bm - n_en) |>
  arrange(year) |>
  rename(Year = year, `n EN` = n_en, `n BM` = n_bm) |>
  tt() |>
  tt_theme_report()
```

Two C1-level residuals carry over from the prior iteration's analysis and are **not** aggregation artifacts: (i) some EN/BM years contain **foreign-comparator** measures (Japan, India, Australia stimulus packages; Philippines TRAIN Act) that should be filtered upstream by country-of-enactment tagging in C1; and (ii) **2018 BM** extracted suspiciously few measures from a full Economic Report — a recall asymmetry to investigate at C1, not C0. These belong to the C1 step and bound how comparable any downstream aggregation can be.

## C0 aggregation diagnostics

C0 collapses C1 measure names into act-level clusters. The three scopes test different properties of that aggregation.

### Per-document: granular aggregation symmetry

Per document, the count of distinct C1 measures, the count of C0 acts, and the compression ratio (measures → acts). A symmetric aggregator should compress comparably on both sides; the previous JW clusterer fragmented BM more than EN.

```{r}
#| label: tbl-c0-perdoc
#| tbl-cap: "C0 per-document aggregation: measures, acts, and compression by language"

if (c0_pending || nrow(m$per_doc) == 0L) {
  tibble(Note = "C0 per-doc target not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_doc |>
    transmute(
      Year = year,
      Language = toupper(doc_language),
      `C1 measures` = n_measures,
      `C0 acts` = n_acts,
      `Compression` = sprintf("%.2f×", compression)
    ) |>
    arrange(Year, Language) |>
    tt() |>
    tt_theme_report()
}
```

### Per-language: act inventory marginals

Pooling all documents per language, how do the two full-pipeline runs compare on act count and the exogenous/endogenous split? Identical marginals are necessary-but-not-sufficient for per-act consistency (the timeline below addresses the rest), but a large marginal gap is itself a finding.

```{r}
#| label: tbl-c0-perlang-counts
#| tbl-cap: "Per-language act inventory: counts and exogeneity split"

if (inv_pending) {
  tibble(Note = "C2 inventory not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_language$counts |>
    transmute(
      Language = toupper(side),
      `C0 acts` = n_acts,
      Exogenous = n_exogenous,
      Endogenous = n_endogenous
    ) |>
    tt() |>
    tt_theme_report()
}
```

```{r}
#| label: tbl-c0-perlang-labels
#| tbl-cap: "Motivation-label marginal distribution by language. A BM-side over-representation of a structural label (e.g. LONG_RUN) surfaces here as a marginal shift — no act-by-act pairing required."

if (inv_pending) {
  tibble(Note = "C2 inventory not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_language$labels |>
    mutate(side = toupper(side)) |>
    pivot_wider(names_from = side, values_from = n, values_fill = 0L) |>
    rename(`Motivation label` = pred_label) |>
    tt() |>
    tt_theme_report()
}
```

### Joint EN+BM: cross-language merge rate

Run on the combined EN+BM pool, how often does C0 place an EN surface form and a BM surface form of the same act into one cluster? This is the direct, no-API probe of whether the current `c0_canonicalize` prompt bridges languages (e.g. `Subsidi Bahan Api` ↔ `fuel subsidy`). A low merge rate means C0 treats the languages as disjoint and cross-language aggregation needs a translation step.

```{r}
#| label: tbl-c0-joint
#| tbl-cap: "Joint-pool clustering: mixed (EN+BM) vs single-language clusters"

if (c0_pending || nrow(m$joint) == 0L || m$joint$n_clusters == 0L) {
  tibble(Note = "C0 joint target not yet built") |> tt() |> tt_theme_report()
} else {
  m$joint |>
    transmute(
      `Total clusters` = n_clusters,
      `Mixed (EN+BM)` = n_mixed,
      `EN-only` = n_en_only,
      `BM-only` = n_bm_only,
      `Merge rate` = sprintf("%.1f%%", 100 * merge_rate)
    ) |>
    tt() |>
    tt_theme_report()
}
```

When the merge rate is non-trivial, audit a few mixed clusters by eye (do the EN and BM members really name the same act?). This eyeball check is the human-verified replacement for the dropped Sonnet matcher.

## Headline: final-output consistency

The pipeline runs end-to-end on each language and produces a set of acts, each carrying a C2 motivation label, sign, and exogeneity. The question is whether the two runs land on the same fiscal reasoning. Pure marginals cannot answer this (identical totals can hide compensating per-act errors), so the headline is **two timeline figures**: each act is a point placed on the year axis, coloured by motivation, shaped by sign, drawn solid if exogenous and hollow if endogenous, faceted by language. The reader compares the EN and BM rows directly.

The two figures differ only in the timing source: the first dates each act by the year **extracted from its name** (language-invariant numerals; partial coverage), the second by the **source-document year** (always present).

### Act-name year multiset

A quick deterministic check before the plots: do the two sides surface acts carrying the same set of name-embedded years?

```{r}
#| label: tbl-act-years
#| tbl-cap: "Act-name-year multiset by language (years parsed from act names)"

if (inv_pending || nrow(m$per_language$act_years) == 0L) {
  tibble(Note = "No act-name years available yet") |> tt() |> tt_theme_report()
} else {
  m$per_language$act_years |>
    mutate(side = toupper(side)) |>
    pivot_wider(names_from = side, values_from = n, values_fill = 0L) |>
    rename(`Act-name year` = act_name_year) |>
    arrange(`Act-name year`) |>
    tt() |>
    tt_theme_report()
}
```

### Timeline by act-name year

```{r}
#| label: fig-timeline-actname
#| fig-cap: "EN- vs BM-side acts placed by the year extracted from the act name. Colour = motivation; shape = sign (▲ increase, ▼ decrease, ● no change); solid = exogenous, faded = endogenous."
#| fig-width: 9
#| fig-height: 7
#| warning: false

p1 <- plot_malay_act_timeline(inv, timing = "act_name")
if (is.null(p1)) cat("No acts carry a name-embedded year yet.\n") else print(p1)
```

### Timeline by source-document year

```{r}
#| label: fig-timeline-docyear
#| fig-cap: "EN- vs BM-side acts placed by the modal source-document year. Same encodings as above."
#| fig-width: 9
#| fig-height: 7
#| warning: false

p2 <- plot_malay_act_timeline(inv, timing = "doc_year")
if (is.null(p2)) cat("Inventory not yet built.\n") else print(p2)
```

## Interpretation

Read the evidence in pipeline order:

- **C1 step.** Any large pre-aggregation drift, foreign-comparator contamination, or the 2018 BM recall gap is a C1 issue and bounds everything downstream — it is not fixable by the aggregator.
- **C0 step.** The per-document compression table closes the loop opened last iteration: does C0 produce **symmetric** act-level aggregation across languages where the JW clusterer fragmented BM more than EN? The joint merge rate separately tells us whether C0 bridges languages at all or treats them as disjoint vocabularies.
- **C2 / headline.** The two timelines are where cross-language consistency of the *full* pipeline is read. Agreement looks like the same years lighting up with the same colours and shapes on both facets; systematic colour shifts (e.g. BM reaching for a structural label) or sign flips on the same act are the disagreements worth chasing. The act-name-year multiset and label-marginal tables are the statistical companions to the visual read.

::: {.callout-note title="Methodological caveat: translation, not independent source"}
The BM Economic Reports are MoF translations of the EN versions, not independent fiscal-policy statements. Two consequences: **correlated errors are invisible** (if Haiku misreads a passage the same way in both languages because the translation preserved the misleading phrasing, this test records agreement); and **some disagreements are translation artifacts** (the translator's word choice, not the model's reasoning). The motivation language R&R rely on — "the actual words used by policymakers" — only fully applies to the EN side. Catching correlated errors requires expert validation on the EN side.
:::

## Implications for deployment scope

This test, conditional on its results, speaks to a narrow population of BM documents:

- **Speaks to:** professional MoF translations of modern Economic Reports (2014+), commissioned alongside the EN versions and expected to be high-quality.
- **Does NOT speak to:** older Economic Report BM (1995-2013, 19 documents — MoF Bahasa originals, not parallel translations), all Budget Speech BM (9 documents, 2000-2008 + 2012 + 2015 — no EN parallels, different register), or crisis-booklet BM (4 documents — branded marketing register).

The 2024 and 2025 EN+BM ER pairs exist on disk but fall outside the current Malaysia pilot window (`pilot_year_max = 2022`). Extending the test to those years requires only an update to `build_country_configs()`; the consistency-test sub-pipeline auto-includes any new pair years from `MANIFEST.csv`.
