---
title: "The Malaysia Statutory Fiscal-Change Dataset"
subtitle: "An LLM-assisted, expert-validated dataset of corporate, personal, and consumption tax changes, discretionary government spending shocks, and tax-incentive changes (1990–2023)"
number-sections: true
---

```{r}
#| label: setup
#| cache: false
#| message: false

pacman::p_load(targets, tidyverse, tinytable, here, scales)

here::i_am("notebooks/malaysia_dataset.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))   # pretty_motivation() + motivation palette
source(here("R/tax_shock_report.R"))
source(here("R/spending_shock_report.R"))   # reuses the tax helpers' internals
source(here("R/incentive_shock_report.R"))  # reuses the tax helpers' internals
set_theme(theme_minimal())

tax_shocks       <- tryCatch(tar_read(tax_shocks), error = function(e) tibble::tibble())
have_data        <- nrow(tax_shocks) > 0
spending_shocks  <- tryCatch(tar_read(spending_shocks), error = function(e) tibble::tibble())
have_spending    <- nrow(spending_shocks) > 0
incentive_shocks <- tryCatch(tar_read(incentive_shocks), error = function(e) tibble::tibble())
have_incentives  <- nrow(incentive_shocks) > 0
```

This page accompanies the dataset of **statutory fiscal changes for Malaysia**
built for *"Scaling Narrative Fiscal Shock Identification with LLMs"* (Degetau &
Samano, World Bank). It is written for the reviewer: it shows the headline picture,
lets you download the data, walks several acts in full, and explains how the
dataset was produced. It covers both sides of the budget: **statutory tax changes**
and **discretionary government spending shocks**.

Each row is **one narratively-announced act × instrument**. On the tax side an
instrument is a corporate income tax (CIT), personal income tax (PIT), or broad
consumption tax (CONSUMPTION = VAT/GST/SST) change, carrying the statutory rate
change. On the spending side it is a discretionary expenditure measure — a
stimulus or relief package, a subsidy-policy change, or a large allocation —
carrying a magnitude in ringgit rather than a rate. Both carry a motivation label
and an exogenous/endogenous read produced with the validated C2b codebook. The
unit of observation and the motivation taxonomy follow the narrative approach of
@romer2010; the LLM pipeline that produces them is validated with the framework of
@halterman2025. Spending exogeneity additionally carries a preliminary read from
the two-condition screen of @das_ai_2026 (non-cyclical motive and no contemporaneous
macro rationale), kept alongside the C2b verdict.

::: {.callout-important}
**Exogeneity here is a preliminary input, not a finding.** The exogenous/endogenous
classification is the LLM's read of the narrative record and is provided to *assist*
an expert, not to replace one. It is pending expert adjudication. Throughout the
project we keep the human in the loop: the model proposes, the expert disposes. The
two worked examples below were both chosen because the model's read and the
preliminary narrative read disagree — exactly the cases an expert must settle.
:::

```{r}
#| label: pending
#| eval: !expr "!have_data"
#| echo: false
#| output: asis

cat("::: {.callout-warning}\nThe `tax_shocks` target is empty or unbuilt, so the",
    "figures, download, and exhibits below cannot render. Build the deliverable",
    "(see the pipeline notebooks) and re-render this page.\n:::\n")
```

# Statutory tax changes

## Tax changes at a glance {#sec-inventory}

```{r}
#| label: tbl-inventory
#| tbl-cap: "**Statutory tax-change inventory.** One row per announced act × tax type. Δpp is the change in the headline statutory rate; motivation and the exogenous/endogenous read are from the validated C2b codebook (v0.9.1)."
#| eval: !expr have_data

tax_inventory_table(tax_shocks)
```

```{r}
#| label: fig-rate-paths
#| fig-cap: "**Statutory rate paths by tax type.** Step path of the headline statutory rate implied by the rate-bearing acts. Non-rate acts (one-off levies, regime introductions without a standing rate) are omitted from the path."
#| eval: !expr have_data

plot_tax_rate_paths(tax_shocks)
```

```{r}
#| label: fig-timeline
#| fig-cap: "**Tax-change timeline.** Acts per year × tax type as a diverging stacked bar: hikes grow the bar upward, cuts downward. Colour denotes the C2b motivation. Faceted by tax type."
#| eval: !expr have_data

plot_tax_timeline(tax_shocks)
```

## Download the tax dataset {#sec-download}

A clean, flat version of the dataset (one row per act × tax type, no nested
provenance columns) is available in three formats. The Stata file carries variable
labels; the Excel workbook includes a second sheet with the data dictionary
reproduced below.

```{r}
#| label: download-check
#| echo: false
#| eval: !expr have_data
#| output: asis

# The files are written reproducibly by the `tax_shocks_clean_files` target
# (R/tax_shock_report.R). We only assert they exist; the links below are
# relative paths the published site resolves.
ok <- tryCatch(all(file.exists(tar_read(tax_shocks_clean_files))),
               error = function(e) FALSE)
if (!ok) cat("::: {.callout-warning}\nDownload files not found —",
             "run `tar_make(tax_shocks_clean_files)`.\n:::\n")
```

- [Download CSV](../data/validated/MY_tax_shocks_clean.csv)
- [Download Excel (.xlsx, with data dictionary)](../data/validated/MY_tax_shocks_clean.xlsx)
- [Download Stata (.dta, labelled)](../data/validated/MY_tax_shocks_clean.dta)

```{r}
#| label: tbl-dictionary
#| tbl-cap: "**Data dictionary.** Columns of the downloadable dataset."

tax_export_dictionary_table()
```

## Two tax acts in full {#sec-exhibits}

To show how the dataset is built act-by-act, we reproduce two narrative summaries
in the style of @romer2010 (their Exhibits 1 and 2, p. 772): a header with the
classification, the identification and motivation reasoning, the most diagnostic
source quotation, and links to the source documents. We deliberately picked two
acts where the model's classification and the preliminary narrative read **diverge**
— the methodologically consequential cases.

### An endogenous act: the 1998 corporate tax cut {#sec-exhibit-cit}

The Budget 1998 corporate tax cut (30% → 28%, announced October 1997) is an
**omnibus** budget act: alongside the headline CIT cut it bundled import-duty
changes and productivity incentives. Asked for the sign of the effect on tax
liabilities, the model returned `"mixed"` — a value outside the permitted
`{increase, decrease, no_change}` set, because the package has no single sign. The
pipeline degrades gracefully: it keeps the otherwise-valid classification
(countercyclical, endogenous) and records the signed effect as `NA`, preserving the
raw `"mixed"` for the audit trail. The act also illustrates the preliminary-vs-model
tension: read narrowly, the rate cut reads as a long-run competitiveness move
(exogenous); read in its Asian-crisis context, the budget reads as cyclical
stabilization (endogenous). This is the kind of call we hand to an expert.

```{r}
#| label: exhibit-cit
#| echo: false
#| eval: !expr have_data
#| output: asis

tax_exhibit(tax_shocks, "MY-CIT-03", exhibit_label = "Exhibit 1")
```

### An exogenous act: the 2000 personal income tax cut {#sec-exhibit-pit}

The Budget 2000 across-the-board one-percentage-point cut to personal income tax
illustrates the opposite tension. Coming shortly after the 1997–98 crisis, the
preliminary narrative read flags it as part of a demand-recovery package
(endogenous). The model, weighing the budget's stated rationale — global
competition (WTO, AFTA), R&D, human-capital and competitiveness — classifies it as
a long-run, exogenous change. Again, the two reads are kept side by side for the
expert.

```{r}
#| label: exhibit-pit
#| echo: false
#| eval: !expr have_data
#| output: asis

tax_exhibit(tax_shocks, "MY-PIT-01", exhibit_label = "Exhibit 2")
```

# Government spending shocks

The dataset also covers **discretionary government spending shocks**: major
spending actions announced in budgets and crisis responses — stimulus and relief
packages, subsidy-policy changes, and large allocations. Unlike a tax change there
is no statutory rate, so the magnitude is recorded in ringgit (or as a share of
expenditure) as free text rather than as a rate change, and the direction is an
increase or decrease in spending. The motivation and the exogenous/endogenous read
again come from the validated C2b codebook; a preliminary read from the
@das_ai_2026 two-condition screen is carried alongside. Routine baseline
development expenditure (for example the standing Five-Year Plan framework) is
deliberately excluded — only discretionary changes count as shocks.

```{r}
#| label: pending-spending
#| eval: !expr "!have_spending"
#| echo: false
#| output: asis

cat("::: {.callout-warning}\nThe `spending_shocks` target is empty or unbuilt, so",
    "the spending figures, download, and exhibit below cannot render. Build the",
    "deliverable and re-render this page.\n:::\n")
```

## Spending shocks at a glance {#sec-spending-inventory}

```{r}
#| label: tbl-spending-inventory
#| tbl-cap: "**Government spending-shock inventory.** One row per announced spending act. The exogenous/endogenous read is from the validated C2b codebook (v0.9.1); category, magnitude, and motivation appear in the exhibit and downloads."
#| eval: !expr have_spending

spending_inventory_table(spending_shocks)
```

```{r}
#| label: fig-spending-timeline
#| fig-cap: "**Spending-shock timeline.** Acts per year as a diverging stacked bar: spending increases grow the bar upward, decreases downward. Colour denotes the C2b exogenous/endogenous classification. A single panel (no per-category facet): the spending mix is dominated by crisis-response packages."
#| eval: !expr have_spending

plot_spending_timeline(spending_shocks)
```


## Download the spending dataset {#sec-spending-download}

The spending dataset is available in the same three formats as the tax dataset
(one row per spending act, no nested provenance columns). The Stata file carries
variable labels; the Excel workbook includes a second sheet with the data
dictionary reproduced below.

```{r}
#| label: spending-download-check
#| echo: false
#| eval: !expr have_spending
#| output: asis

# Written reproducibly by the `spending_shocks_clean_files` target
# (R/spending_shock_report.R). We only assert they exist; the links below are
# relative paths the published site resolves.
ok <- tryCatch(all(file.exists(tar_read(spending_shocks_clean_files))),
               error = function(e) FALSE)
if (!ok) cat("::: {.callout-warning}\nDownload files not found —",
             "run `tar_make(spending_shocks_clean_files)`.\n:::\n")
```

- [Download CSV](../data/validated/MY_spending_shocks_clean.csv)
- [Download Excel (.xlsx, with data dictionary)](../data/validated/MY_spending_shocks_clean.xlsx)
- [Download Stata (.dta, labelled)](../data/validated/MY_spending_shocks_clean.dta)

```{r}
#| label: tbl-spending-dictionary
#| tbl-cap: "**Spending data dictionary.** Columns of the downloadable spending dataset."

spending_export_dictionary_table()
```

## A spending act in full {#sec-spending-exhibit}

As on the tax side, we reproduce one spending act in full, again choosing a case
where the model's classification and the preliminary narrative read **diverge**.

The June 2008 fuel-subsidy restructuring cut a major subsidy in response to a
global crude-oil price spike that had made the subsidy bill fiscally
unsustainable. The preliminary @das_ai_2026 screen, seeing a contemporaneous macro
trigger, reads it as endogenous; the C2b model, weighing the deficit-reduction
rationale, reads it as exogenous and deficit-driven. The two reads are kept side by
side for the expert.

```{r}
#| label: spending-exhibit
#| echo: false
#| eval: !expr have_spending
#| output: asis

spending_exhibit(spending_shocks, "MY-SPEND-04", exhibit_label = "Exhibit 3")
```

# Tax incentives and holidays

The dataset also covers **changes to tax incentives and holidays**: tax holidays
(Pioneer Status), investment and reinvestment allowances, concessionary investor
rates (Labuan, Principal Hub, sectoral special rates), free/export/enterprise zones,
and consumption- and personal-side reliefs. These differ from a statutory rate
change in being *targeted* (conditional on investor, sector, region, or activity),
often *time-bounded*, and frequently operating through the tax base or a carve-out
rate rather than the standing rate — so each event lives here rather than in the
tax tables (ownership-by-mechanism). The unit is one row per legislative *change*;
the standing existence of a long-running scheme is not a shock. The motivation and
the exogenous/endogenous read again come from the validated C2b codebook; a
preliminary read is carried alongside. Incentives skew exogenous — FDI-attraction,
industrial policy, and structural competitiveness are long-run motives.

```{r}
#| label: pending-incentives
#| eval: !expr "!have_incentives"
#| echo: false
#| output: asis

cat("::: {.callout-warning}\nThe `incentive_shocks` target is empty or unbuilt, so",
    "the incentive figures, download, and exhibit below cannot render. Build the",
    "deliverable and re-render this page.\n:::\n")
```

## Incentive changes at a glance {#sec-incentive-inventory}

```{r}
#| label: tbl-incentive-inventory
#| tbl-cap: "**Tax-incentive shock inventory.** One row per announced incentive change. The exogenous/endogenous read is from the validated C2b codebook (v0.9.1); direction, rate detail, magnitude, and motivation appear in the exhibit and downloads."
#| eval: !expr have_incentives

incentive_inventory_table(incentive_shocks)
```

```{r}
#| label: fig-incentive-timeline
#| fig-cap: "**Incentive-shock timeline.** Acts per year as a diverging stacked bar: new or expanded incentives grow the bar downward, repeals/scale-backs upward. Colour denotes the C2b exogenous/endogenous classification. A single panel (no per-base facet): incentives are dominated by the corporate base."
#| eval: !expr have_incentives

plot_incentive_timeline(incentive_shocks)
```


## Download the incentive dataset {#sec-incentive-download}

The incentive dataset is available in the same three formats as the tax and spending
datasets (one row per incentive change, no nested provenance columns). The Stata file
carries variable labels; the Excel workbook includes a second sheet with the data
dictionary reproduced below.

```{r}
#| label: incentive-download-check
#| echo: false
#| eval: !expr have_incentives
#| output: asis

# Written reproducibly by the `incentive_shocks_clean_files` target
# (R/incentive_shock_report.R). We only assert they exist; the links below are
# relative paths the published site resolves.
ok <- tryCatch(all(file.exists(tar_read(incentive_shocks_clean_files))),
               error = function(e) FALSE)
if (!ok) cat("::: {.callout-warning}\nDownload files not found —",
             "run `tar_make(incentive_shocks_clean_files)`.\n:::\n")
```

- [Download CSV](../data/validated/MY_incentive_shocks_clean.csv)
- [Download Excel (.xlsx, with data dictionary)](../data/validated/MY_incentive_shocks_clean.xlsx)
- [Download Stata (.dta, labelled)](../data/validated/MY_incentive_shocks_clean.dta)

```{r}
#| label: tbl-incentive-dictionary
#| tbl-cap: "**Incentive data dictionary.** Columns of the downloadable incentive dataset."

incentive_export_dictionary_table()
```

## An incentive change in full {#sec-incentive-exhibit}

As on the tax and spending sides, we reproduce one incentive change in full once the
dataset is stamped, again favouring a case where the model's classification and the
preliminary narrative read can be compared side by side for the expert.

```{r}
#| label: incentive-exhibit
#| echo: false
#| eval: !expr have_incentives
#| output: asis

incentive_exhibit(incentive_shocks, "MY-INCENT-01", exhibit_label = "Exhibit 4")
```

# How the dataset was built {#sec-process}

The dataset is the tail of a validated, country-agnostic pipeline. Each stage has
its own notebook documenting its construction and validation:

1. **Corpus.** Primary-source fiscal documents (Economic Reports, Budget Speeches,
   Bank Negara annual reports, 1990–2023) are extracted and quality-checked
   ([corpus verification](verify_country_body.qmd)).
2. **Measure identification (C1).** An LLM codebook surfaces candidate fiscal
   measures from each document chunk, validated against US ground truth with the
   @halterman2025 framework ([C1 evaluation](c1_measure_id.qmd)).
3. **Act aggregation (C0).** Surface forms of the same act, scattered across
   documents and languages, are collapsed into canonical acts
   ([C0 method comparison](c0_aggregator.qmd)).
4. **Motivation & exogeneity (C2).** A two-stage codebook extracts motivation
   evidence (C2a) and classifies each act's motivation and exogeneity (C2b), using
   the @romer2010 taxonomy (spending-driven, countercyclical, deficit-driven,
   long-run) ([cross-country deployment](deployment.qmd)).
5. **Statutory tax-shock identification.** For each tax instrument, an AI-assisted
   narrative pass reads the surfaced evidence and the source documents directly,
   recovers misses, and consolidates events at the announced-act grain, with a
   recall scorecard and a human stamp
   ([CIT](cit_identification.qmd) · [PIT](pit_identification.qmd) ·
   [consumption](vat_identification.qmd)).
6. **Spending-shock identification.** Because C1 is tax-scoped, the spending side
   is identified by reading the corpus directly for discretionary spending actions,
   consolidated the same way with a recall scorecard and a human stamp; preliminary
   exogeneity uses the @das_ai_2026 two-condition screen
   ([spending identification](spending_identification.qmd)).
7. **Tax-incentive identification.** Changes to selective tax-relief instruments
   (holidays, allowances, concessionary rates, zones, reliefs) are likewise
   identified by reading the corpus directly, consolidated at the legislative-change
   grain with a recall scorecard and a human stamp; each event is kept disjoint from
   the statutory-rate and spending sets (ownership-by-mechanism)
   ([incentive identification](incentives_identification.qmd)).
8. **Binding & enrichment.** The frozen per-instrument, spending, and incentive
   datasets are bound and each act is enriched with the C2b motivation and exogeneity
   classification ([tax deliverable assembly](tax_shocks.qmd)).

The result is the dataset on this page: a transparent, reproducible starting point
for expert validation, not a finished verdict. Reviewer disagreements — especially
on the exogeneity column — are the intended next input, in keeping with the
narrative tradition of @romer2010, where every classification is documented so
others can check it.

### References {.unnumbered}