---
title: "Malaysia EN/BM Cross-Language Consistency Test"
subtitle: "Full C1 → C0 → C2 pipeline on parallel Economic Report EN+BM pairs, 2014–2025"
date: today
date-format: long
execute: 
  cache: true
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, stringr)

here::i_am("notebooks/malay_consistency.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))

tar_load(c(
  malay_er_pair_years,
  malay_er_chunks,
  malay_er_c1,
  malay_er_c1_measures,
  malay_er_measure_pool,
  malay_er_c0_perdoc,
  malay_er_c0_perlang,
  malay_er_c0_joint,
  malay_er_consistency_metrics
))

m <- malay_er_consistency_metrics
inv <- m$inventory
inv_pending <- nrow(inv) == 0L
c0_pending  <- nrow(malay_er_c0_perlang) == 0L
```

## Framing


In this section, we seek to answer the question: Can our pipeline produce equivalent results with two equivalent inputs that are in two different languages? And more specifically, will including fiscal documents in other languages (Malay) yield results that are comparable to the ones we obtained in English during the development phase?

The C1 → C0 → C2 codebook pipeline was validated on US English documents. Malaysia has 38 BM-only documents in its ready corpus (primarily older Budget Speeches and Economic Reports). Before trusting LLM outputs on those, we need evidence that the pipeline produces **equivalent fiscal reasoning regardless of input language**.

This is a consistency test, not a validity test. We have no labels for Malaysia. EN outputs are treated as the reference because the codebooks crossed S3 gates on EN-only US data. Passing this test is necessary-but-not-sufficient for extending deployment to the **BM-only-document slice (38 of 99 ready documents)**; it does not gate the broader Malaysia deployment, and it is **not a substitute for expert agreement on the EN-side outputs** (Phase 2's success criteria: ≥80% on C1, ≥70% on C2 against expert review).

The test runs the **full pipeline independently on each language** and compares the outputs within each step of the process. It slices `country_chunks` to Economic Reports with parallel EN+BM coverage, runs its own C1 → C2a chain, then aggregates measure names into acts with the **C0 act aggregator** (M5 LLM canonical clustering, `prompts/c0_canonicalize.yml`, Haiku) at three scopes:

[Because these notebooks will be the building blocks for the final paper, we want to write them for external audiences. So, in general, we do not want to mention code elements or files.]{.aside}

- **per-document** — aggregate within each language × year document (granular aggregation capability/limits);
- **per-language** — pool all EN documents / all BM documents (deployment-realistic; this scope feeds C2 and the headline timeline);
- **joint EN+BM** — pool both languages together (probes whether C0 can co-cluster the EN and BM surface forms of the same act).


::: {.column-margin}
This test only speaks to **modern, professionally-translated MoF prose** (Economic Reports 2014–2025; see the 2023 edition caveat under C1-step comparability). It does NOT speak to pre-2009 Budget Speech BM, older Economic Report BM (pre-2014), or crisis-booklet BM. Those are different vintages and registers; deployment to them requires separate evidence.
:::

```{r}
#| label: scope-summary
#| results: 'asis'

cat(sprintf(
  "**Scope:** %d pair years × 2 languages = %d documents; %d total chunks.\n\n",
  length(malay_er_pair_years),
  2L * length(malay_er_pair_years),
  nrow(malay_er_chunks)
))
cat(sprintf("**Pair years:** %s\n\n",
            paste(malay_er_pair_years, collapse = ", ")))
```


[We should include a Scope subsection that shows a figure with the number of pages and chunks in each document-year-language to let the reader understand the extend to which we're comparing the same document. This will obviate the need for a "pair years/scope" line as above.]{.aside}


## C1-step comparability

Before aggregation, does C1 surface a comparable set of raw fiscal measures on each side? This is the step where real extraction asymmetries live (independent of how acts are later aggregated). @tbl-c1-comparability counts the distinct `measure_name` strings C1 produced per document, by language.

```{r}
#| label: tbl-c1-comparability
#| tbl-cap: "Distinct C1 measure names per year, by language (pre-aggregation)"

c1m_wide <- malay_er_c1_measures |>
  mutate(doc_language = if_else(str_detect(doc_id, "-BM$"), "bm", "en")) |>
  distinct(year, doc_language, measure_name) |>
  count(year, doc_language, name = "n_measures") |>
  pivot_wider(names_from = doc_language, values_from = n_measures,
              names_prefix = "n_", values_fill = 0L)

if (!"n_en" %in% names(c1m_wide)) c1m_wide$n_en <- 0L
if (!"n_bm" %in% names(c1m_wide)) c1m_wide$n_bm <- 0L

c1m_wide |>
  mutate(`Signed drift (BM − EN)` = n_bm - n_en) |>
  arrange(year) |>
  rename(Year = year, `n EN` = n_en, `n BM` = n_bm) |>
  tt() |>
  tt_theme_report()
```


[@tbl-c1-comparability should be a plot.]{.aside}

Two C1-level residuals carry over from the prior iteration's analysis and are **not** aggregation artifacts: (i) some EN/BM years contain **foreign-comparator** measures (Japan, India, Australia stimulus packages; Philippines TRAIN Act) that should be filtered upstream by country-of-enactment tagging in C1; and (ii) **2018 BM** extracted suspiciously few measures from a full Economic Report, a recall asymmetry to investigate at C1, not C0. These belong to the C1 step and bound how comparable any downstream aggregation can be.

::: {.callout-warning .column-margin}
The 2023 EN and BM documents are **different editions of the 2023 budget cycle**, not parallel translations. The BM side is the Oct-2022 full *Tinjauan Ekonomi 2023*; the EN side is the Feb-2023 re-tabled-budget *Updates on Economic and Fiscal Outlook and Revenue Estimates 2023* — a condensed Outlook plus a bundled Federal Government Revenue Estimates / Treasury Memorandum block that has no BM counterpart here (or in any 2014–2022 EN file). Expect **EN-heavy C1 measure counts for 2023**; the drift on that row reflects document-edition asymmetry, not a language/reasoning inconsistency. See the 2026-06-05 provenance note in `data/manual/malaysia/README.md`.
:::

### Why the drift varies by year

@tbl-c1-comparability drift is not uniform across years; it concentrates in 2014 (the GST-announcement Budget). Two structural features of each year's Economic Report drive it, and both peak in 2014.

First, **measure density**. The `malay_er_c1_measures` count is a post-filter count: the deployment filter (`filter_c1_measures()`) keeps only the most-prominent (rank-1) motivated measure per chunk. Before that cut, C1 surfaces several measures per chunk, and 2014 carries roughly two to three times the motivated-fiscal-measure load of a typical year. The denser the report, the more measures the rank-1 cut discards, and the more room chunk-boundary differences between the parallel EN and BM documents have to change *which* measure survives on each side.

```{r}
#| label: fig-c1-density
#| fig-cap: "Motivated fiscal measures C1 surfaces per year by language, before the rank-1 deployment filter. 2014 carries the heaviest load of any pilot year, which is what the rank-1 cut discards most and most asymmetrically."
#| fig-width: 8
#| fig-height: 3.5
#| warning: false

pd <- plot_malay_c1_density(malay_er_c1)
if (is.null(pd)) cat("No motivated measures available yet.\n") else print(pd)
```

Second, **chunk-level classification asymmetry**. With near-identical chunk counts across the two parallel documents, BM flags a higher share of its 2014 chunks as carrying a motivated rank-1 fiscal measure than EN does. Since each such chunk contributes exactly one measure to Table 1, this EN/BM gap *is* the drift, and it is widest in 2014.

```{r}
#| label: fig-c1-fiscal-rate
#| fig-cap: "Share of each year's chunks whose most-prominent (rank-1) measure is a motivated fiscal measure, by language. The EN/BM gap, the per-chunk asymmetry the drift reflects, is widest in 2014."
#| fig-width: 8
#| fig-height: 3.5
#| warning: false

pr <- plot_malay_c1_fiscal_rate(malay_er_c1, malay_er_chunks)
if (is.null(pr)) cat("No chunk data available yet.\n") else print(pr)
```

## C0 aggregation diagnostics

C0 collapses C1 measure names into act-level clusters. The three scopes test different properties of that aggregation.

### Per-document: granular aggregation symmetry

Per document, the count of distinct C1 measures, the count of C0 acts, and the compression ratio (surfaced act names → distinct acts). A symmetric aggregator should compress comparably on both sides; the previous JW clusterer fragmented BM more than EN.

```{r}
#| label: tbl-c0-perdoc
#| tbl-cap: "C0 per-document aggregation: measures, acts, and compression by language"

if (c0_pending || nrow(m$per_doc) == 0L) {
  tibble(Note = "C0 per-doc target not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_doc |>
    transmute(
      Year = year,
      Language = toupper(doc_language),
      `C1 measures` = n_measures,
      `C0 acts` = n_acts,
      `Compression` = sprintf("%.2f×", compression)
    ) |>
    arrange(Year, Language) |>
    tt() |>
    tt_theme_report()
}
```

[@tbl-c0-perdoc shuld also be a plot. In general, we prefer showing plots than tables.]{.aside}

### Per-language: act inventory marginals

Pooling all documents per language, how do the two full-pipeline runs compare on act count and the exogenous/endogenous split? Identical marginals are necessary-but-not-sufficient for per-act consistency (the timeline below addresses the rest), but a large marginal gap is itself a finding.

```{r}
#| label: tbl-c0-perlang-counts
#| tbl-cap: "Per-language act inventory: counts and exogeneity split"

if (inv_pending) {
  tibble(Note = "C2 inventory not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_language$counts |>
    transmute(
      Language = toupper(side),
      `C0 acts` = n_acts,
      Exogenous = n_exogenous,
      Endogenous = n_endogenous
    ) |>
    tt() |>
    tt_theme_report()
}
```

```{r}
#| label: tbl-c0-perlang-labels
#| tbl-cap: "Motivation-label marginal distribution by language. A BM-side over-representation of a structural label (e.g. LONG_RUN) surfaces here as a marginal shift — no act-by-act pairing required."

if (inv_pending) {
  tibble(Note = "C2 inventory not yet built") |> tt() |> tt_theme_report()
} else {
  m$per_language$labels |>
    mutate(side = toupper(side)) |>
    pivot_wider(names_from = side, values_from = n, values_fill = 0L) |>
    rename(`Motivation label` = pred_label) |>
    tt() |>
    tt_theme_report()
}
```

[@tbl-c0-perlang-labels should show neat motivation labels.]{.aside}

### Joint EN+BM: cross-language merge rate

Run on the combined EN+BM pool, how often does C0 place an EN surface form and a BM surface form of the same act into one cluster? This is the direct, no-API probe of whether the current `c0_canonicalize` prompt bridges languages (e.g. `Subsidi Bahan Api` ↔ `fuel subsidy`). A low merge rate means C0 treats the languages as disjoint and cross-language aggregation needs a translation step.

```{r}
#| label: tbl-c0-joint
#| tbl-cap: "Joint-pool clustering: mixed (EN+BM) vs single-language clusters"

if (c0_pending || nrow(m$joint) == 0L || m$joint$n_clusters == 0L) {
  tibble(Note = "C0 joint target not yet built") |> tt() |> tt_theme_report()
} else {
  m$joint |>
    transmute(
      `Total clusters` = n_clusters,
      `Mixed (EN+BM)` = n_mixed,
      `EN-only` = n_en_only,
      `BM-only` = n_bm_only,
      `Merge rate` = sprintf("%.1f%%", 100 * merge_rate)
    ) |>
    tt() |>
    tt_theme_report()
}
```

When the merge rate is non-trivial, audit a few mixed clusters by eye (do the EN and BM members really name the same act?). This eyeball check is the human-verified replacement for the dropped Sonnet matcher.

## Headline: final-output consistency

The pipeline runs end-to-end on each language and produces a set of acts, each carrying a C2 motivation label, sign, and exogeneity. The question is whether the two runs land on the same fiscal reasoning. Pure marginals cannot answer this (identical totals can hide compensating per-act errors), so the headline is **two timeline figures**: each act is a point placed on the year axis, coloured by motivation, shaped by sign, drawn solid if exogenous and hollow if endogenous, faceted by language. The reader compares the EN and BM rows directly.

The two figures differ only in the timing source: the first dates each act by the year **extracted from its name** (language-invariant numerals; partial coverage), the second by the **source-document year** (always present).

### Act-name year multiset

A quick deterministic check before the plots: do the two sides surface acts carrying the same set of name-embedded years?

```{r}
#| label: tbl-act-years
#| tbl-cap: "Act-name-year multiset by language (years parsed from act names)"

if (inv_pending || nrow(m$per_language$act_years) == 0L) {
  tibble(Note = "No act-name years available yet") |> tt() |> tt_theme_report()
} else {
  m$per_language$act_years |>
    mutate(side = toupper(side)) |>
    pivot_wider(names_from = side, values_from = n, values_fill = 0L) |>
    rename(`Act-name year` = act_name_year) |>
    arrange(`Act-name year`) |>
    tt() |>
    tt_theme_report()
}
```

[@tbl-act-years should also be a plot.]{.aside}

### Timeline by act-name year

```{r}
#| label: fig-timeline-actname
#| fig-cap: "EN- vs BM-side acts placed by the year extracted from the act name. Colour = motivation; shape = sign (▲ increase, ▼ decrease, ● no change); solid = exogenous, faded = endogenous."
#| fig-width: 9
#| fig-height: 7
#| warning: false

p1 <- plot_malay_act_timeline(inv, timing = "act_name")
if (is.null(p1)) cat("No acts carry a name-embedded year yet.\n") else print(p1)
```



### Timeline by source-document year

```{r}
#| label: fig-timeline-docyear
#| fig-cap: "EN- vs BM-side acts placed by the modal source-document year. Same encodings as above."
#| fig-width: 9
#| fig-height: 7
#| warning: false

p2 <- plot_malay_act_timeline(inv, timing = "doc_year")
if (is.null(p2)) cat("Inventory not yet built.\n") else print(p2)
```


[@fig-timeline-actname and @fig-timeline are quite inteligible. Instead of showing the act name, only count the number of acts in each year-language-motivation-sign cell. ]

## Interpretation

Read the evidence in pipeline order:

- **C1 step.** Any large pre-aggregation drift, foreign-comparator contamination, or the 2018 BM recall gap is a C1 issue and bounds everything downstream — it is not fixable by the aggregator.
- **C0 step.** The per-document compression table closes the loop opened last iteration: does C0 produce **symmetric** act-level aggregation across languages where the JW clusterer fragmented BM more than EN? The joint merge rate separately tells us whether C0 bridges languages at all or treats them as disjoint vocabularies.
- **C2 / headline.** The two timelines are where cross-language consistency of the *full* pipeline is read. Agreement looks like the same years lighting up with the same colours and shapes on both facets; systematic colour shifts (e.g. BM reaching for a structural label) or sign flips on the same act are the disagreements worth chasing. The act-name-year multiset and label-marginal tables are the statistical companions to the visual read.

::: {.callout-note title="Methodological caveat: translation, not independent source"}
The BM Economic Reports are MoF translations of the EN versions, not independent fiscal-policy statements. Two consequences: **correlated errors are invisible** (if Haiku misreads a passage the same way in both languages because the translation preserved the misleading phrasing, this test records agreement); and **some disagreements are translation artifacts** (the translator's word choice, not the model's reasoning). The motivation language R&R rely on — "the actual words used by policymakers" — only fully applies to the EN side. Catching correlated errors requires expert validation on the EN side.
:::

## Implications for deployment scope

This test, conditional on its results, speaks to a narrow population of BM documents:

- **Speaks to:** professional MoF translations of modern Economic Reports (2014+), commissioned alongside the EN versions and expected to be high-quality.
- **Does NOT speak to:** older Economic Report BM (1995-2013, 19 documents — MoF Bahasa originals, not parallel translations), all Budget Speech BM (9 documents, 2000-2008 + 2012 + 2015 — no EN parallels, different register), or crisis-booklet BM (4 documents — branded marketing register).

The pilot window (`pilot_year_max = 2025` in `build_country_configs()`) now spans the full parallel EN+BM Economic Report record (2014–2025). The consistency-test sub-pipeline auto-includes any pair year flagged `has_en & has_bm` in `MANIFEST.csv`, so adding a future pair requires only dropping the two PDFs into `economic_report/` and flipping the manifest row.
