---
title: "The Malaysia Statutory Tax-Change Dataset"
subtitle: "An LLM-assisted, expert-validated dataset of corporate, personal, and consumption tax changes (1990–2023)"
number-sections: true
---

```{r}
#| label: setup
#| cache: false
#| message: false

pacman::p_load(targets, tidyverse, tinytable, here, scales)

here::i_am("notebooks/malaysia_dataset.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))   # pretty_motivation() + motivation palette
source(here("R/tax_shock_report.R"))
set_theme(theme_minimal())

tax_shocks <- tryCatch(tar_read(tax_shocks), error = function(e) tibble::tibble())
have_data  <- nrow(tax_shocks) > 0
```

This page accompanies the dataset of **statutory tax changes for Malaysia** built
for *"Scaling Narrative Fiscal Shock Identification with LLMs"* (Degetau & Samano,
World Bank). It is written for the reviewer: it shows the headline picture, lets
you download the data, walks two acts in full, and explains how the dataset was
produced.

Each row is **one narratively-announced act × tax type** — a corporate income tax
(CIT), personal income tax (PIT), or broad consumption tax (CONSUMPTION =
VAT/GST/SST) change — carrying the statutory rate change, a motivation label, and
an exogenous/endogenous read produced with the validated C2b codebook. The unit of
observation and the motivation taxonomy follow the narrative approach of
@romer2010; the LLM pipeline that produces them is validated with the framework of
@halterman2025.

::: {.callout-important}
**Exogeneity here is a preliminary input, not a finding.** The exogenous/endogenous
classification is the LLM's read of the narrative record and is provided to *assist*
an expert, not to replace one. It is pending expert adjudication. Throughout the
project we keep the human in the loop: the model proposes, the expert disposes. The
two worked examples below were both chosen because the model's read and the
preliminary narrative read disagree — exactly the cases an expert must settle.
:::

```{r}
#| label: pending
#| eval: !expr "!have_data"
#| echo: false
#| output: asis

cat("::: {.callout-warning}\nThe `tax_shocks` target is empty or unbuilt, so the",
    "figures, download, and exhibits below cannot render. Build the deliverable",
    "(see the pipeline notebooks) and re-render this page.\n:::\n")
```

## The dataset at a glance {#sec-inventory}

```{r}
#| label: tbl-inventory
#| tbl-cap: "**Statutory tax-change inventory.** One row per announced act × tax type. Δpp is the change in the headline statutory rate; motivation and the exogenous/endogenous read are from the validated C2b codebook (v0.9.1)."
#| eval: !expr have_data

tax_inventory_table(tax_shocks)
```

```{r}
#| label: fig-rate-paths
#| fig-cap: "**Statutory rate paths by tax type.** Step path of the headline statutory rate implied by the rate-bearing acts. Non-rate acts (one-off levies, regime introductions without a standing rate) are omitted from the path."
#| eval: !expr have_data

plot_tax_rate_paths(tax_shocks)
```

```{r}
#| label: fig-timeline
#| fig-cap: "**Tax-change timeline.** Acts per year × tax type as a diverging stacked bar: hikes grow the bar upward, cuts downward. Colour denotes the C2b motivation. Faceted by tax type."
#| eval: !expr have_data

plot_tax_timeline(tax_shocks)
```

## Download the dataset {#sec-download}

A clean, flat version of the dataset (one row per act × tax type, no nested
provenance columns) is available in three formats. The Stata file carries variable
labels; the Excel workbook includes a second sheet with the data dictionary
reproduced below.

```{r}
#| label: download-check
#| echo: false
#| eval: !expr have_data
#| output: asis

# The files are written reproducibly by the `tax_shocks_clean_files` target
# (R/tax_shock_report.R). We only assert they exist; the links below are
# relative paths the published site resolves.
ok <- tryCatch(all(file.exists(tar_read(tax_shocks_clean_files))),
               error = function(e) FALSE)
if (!ok) cat("::: {.callout-warning}\nDownload files not found —",
             "run `tar_make(tax_shocks_clean_files)`.\n:::\n")
```

- [Download CSV](../data/validated/MY_tax_shocks_clean.csv)
- [Download Excel (.xlsx, with data dictionary)](../data/validated/MY_tax_shocks_clean.xlsx)
- [Download Stata (.dta, labelled)](../data/validated/MY_tax_shocks_clean.dta)

```{r}
#| label: tbl-dictionary
#| tbl-cap: "**Data dictionary.** Columns of the downloadable dataset."

tax_export_dictionary_table()
```

## Two acts in full {#sec-exhibits}

To show how the dataset is built act-by-act, we reproduce two narrative summaries
in the style of @romer2010 (their Exhibits 1 and 2, p. 772): a header with the
classification, the identification and motivation reasoning, the most diagnostic
source quotation, and links to the source documents. We deliberately picked two
acts where the model's classification and the preliminary narrative read **diverge**
— the methodologically consequential cases.

### An endogenous act: the 1998 corporate tax cut {#sec-exhibit-cit}

The Budget 1998 corporate tax cut (30% → 28%, announced October 1997) is an
**omnibus** budget act: alongside the headline CIT cut it bundled import-duty
changes and productivity incentives. Asked for the sign of the effect on tax
liabilities, the model returned `"mixed"` — a value outside the permitted
`{increase, decrease, no_change}` set, because the package has no single sign. The
pipeline degrades gracefully: it keeps the otherwise-valid classification
(countercyclical, endogenous) and records the signed effect as `NA`, preserving the
raw `"mixed"` for the audit trail. The act also illustrates the preliminary-vs-model
tension: read narrowly, the rate cut reads as a long-run competitiveness move
(exogenous); read in its Asian-crisis context, the budget reads as cyclical
stabilization (endogenous). This is the kind of call we hand to an expert.

```{r}
#| label: exhibit-cit
#| echo: false
#| eval: !expr have_data
#| output: asis

tax_exhibit(tax_shocks, "MY-CIT-03", exhibit_label = "Exhibit 1")
```

### An exogenous act: the 2000 personal income tax cut {#sec-exhibit-pit}

The Budget 2000 across-the-board one-percentage-point cut to personal income tax
illustrates the opposite tension. Coming shortly after the 1997–98 crisis, the
preliminary narrative read flags it as part of a demand-recovery package
(endogenous). The model, weighing the budget's stated rationale — global
competition (WTO, AFTA), R&D, human-capital and competitiveness — classifies it as
a long-run, exogenous change. Again, the two reads are kept side by side for the
expert.

```{r}
#| label: exhibit-pit
#| echo: false
#| eval: !expr have_data
#| output: asis

tax_exhibit(tax_shocks, "MY-PIT-01", exhibit_label = "Exhibit 2")
```

## How the dataset was built {#sec-process}

The dataset is the tail of a validated, country-agnostic pipeline. Each stage has
its own notebook documenting its construction and validation:

1. **Corpus.** Primary-source fiscal documents (Economic Reports, Budget Speeches,
   Bank Negara annual reports, 1990–2023) are extracted and quality-checked
   ([corpus verification](verify_country_body.qmd)).
2. **Measure identification (C1).** An LLM codebook surfaces candidate fiscal
   measures from each document chunk, validated against US ground truth with the
   @halterman2025 framework ([C1 evaluation](c1_measure_id.qmd)).
3. **Act aggregation (C0).** Surface forms of the same act, scattered across
   documents and languages, are collapsed into canonical acts
   ([C0 method comparison](c0_aggregator.qmd)).
4. **Motivation & exogeneity (C2).** A two-stage codebook extracts motivation
   evidence (C2a) and classifies each act's motivation and exogeneity (C2b), using
   the @romer2010 taxonomy (spending-driven, countercyclical, deficit-driven,
   long-run) ([cross-country deployment](deployment.qmd)).
5. **Statutory tax-shock identification.** For each instrument, an AI-assisted
   narrative pass reads the surfaced evidence and the source documents directly,
   recovers misses, and consolidates events at the announced-act grain, with a
   recall scorecard and a human stamp
   ([CIT](cit_identification.qmd) · [PIT](pit_identification.qmd) ·
   [consumption](vat_identification.qmd)).
6. **Binding & enrichment.** The frozen per-instrument datasets are bound and each
   act is enriched with the C2b motivation and exogeneity classification
   ([deliverable assembly](tax_shocks.qmd)).

The result is the dataset on this page: a transparent, reproducible starting point
for expert validation, not a finished verdict. Reviewer disagreements — especially
on the exogeneity column — are the intended next input, in keeping with the
narrative tradition of @romer2010, where every classification is documented so
others can check it.
