---
title: "Manual Identification of Malaysia's Government Spending Shocks"
subtitle: "AI-assisted narrative pass over the deployment corpus (Das 2026-style spending side)"
number-sections: true
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, qs2)

here::i_am("notebooks/spending_identification.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
set_theme(theme_minimal())

# Frozen, human-stamped spending shock dataset (produced by the
# /identify-spending skill). Loaded here so both the recall scorecard
# (@sec-recall) and the structured dataset (@sec-frozen) render from the single
# source of truth when it exists.
frozen_path <- here("data", "validated", "MY_SPENDING_shocks.qs")
have_frozen <- file.exists(frozen_path)
if (have_frozen) spend_frozen <- qs2::qs_read(frozen_path)
```

## Purpose

This note documents a focused, **AI-assisted manual identification** of Malaysia's
*discretionary government spending changes* — named stimulus/relief packages,
subsidy-policy changes, large allocations, and Five-Year Plan launches — run
directly over the deployment corpus. It is the spending-side analogue of
`notebooks/cit_identification.qmd`, and it deliberately follows the spirit of the
narrative-identification literature [@romer2010; @das_ai_2026]: close reading of source
documents, with the LLM pipeline as a retrieval-and-extraction aide, not as the
adjudicator.

The method mirrors our `/manual-analysis` protocol. Unlike the tax side, **C1 does
not pre-screen spending** (C1 is scoped to tax liabilities only), so identification
is driven by **direct reading of the corpus** (`country_body` / `country_chunks`),
not the C1 measure pool. A human reads the surfaced evidence and makes the final
call on *what counts as a spending act*, *how scattered surface forms consolidate*,
and *whether each change is exogenous*. Claude pattern-matches against the source
text; the researcher confers meaning.

::: {.callout-important}
The exogenous/endogenous column in @tbl-exogeneity is a **preliminary suggestion
with its supporting quote, pending expert adjudication**, produced by the Das
(2026) two-condition screen. It is not a research finding, and it is carried
*alongside* — never replacing — C2b's downstream label.
:::

## What the corpus surfaced, and what C1 missed {#sec-recall}

Because C1 is tax-scoped, the keyword scan of the C1 pool is expected to surface
almost no spending events — that is itself a finding. The events below are
recovered by **reading the anchor documents directly** (Budget Speeches'
Development Expenditure sections, Five-Year Plans, Economic Reports, BNM Annual
Reports). The mandatory recall scorecard records what was searched and missed.

```{r}
#| label: tbl-recall
#| tbl-cap: "**Recall scorecard**. How each headline spending event fared through the identification pass. The near-empty C1 floor is logged as a finding; 'Recovered manually' marks events recovered by direct reading."

recall <- if (have_frozen) {
  spend_frozen$recall_scorecard[[1]] |>
    dplyr::rename(Stage = stage, Outcome = outcome)
} else {
  tibble::tribble(
    ~Stage,                       ~Outcome,
    "C1 keyword scan",            "Expected near-empty: C1 is scoped to tax liabilities, surfaces ~0 spending measures",
    "Direct corpus sweep",        "Primary method — read Development Expenditure / Five-Year Plan / Economic Report sections directly",
    "Recall checkpoint — 1998",   "National Economic Recovery Plan (Asian-crisis fiscal response) — recovered from source",
    "Recall checkpoint — 2009",   "GFC stimulus packages — recovered from source",
    "Recall checkpoint — 2020",   "COVID-19 packages (PRIHATIN / PENJANA / PEMERKASA, RM250bn+) — recovered from source"
  )
}

recall |>
  tt() |>
  tt_theme_report()
```

## Headline spending acts {#sec-headline}

```{r}
#| label: tbl-headline
#| tbl-cap: "**Headline spending events**. Identified discretionary government spending acts, with category, direction, magnitude, and the budget/plan that announced each. Renders from the frozen dataset once the /identify-spending skill has stamped it."
#| eval: !expr have_frozen

spend_frozen |>
  dplyr::select(shock_id, act_label,
                Category = spending_category, Direction = direction,
                Magnitude = magnitude_note,
                Announced = announced_year, Effective = effective_year) |>
  tt() |>
  tt_theme_report()
```

::: {.content-visible when-format="html"}
```{r}
#| label: headline-pending
#| eval: !expr "!have_frozen"
#| echo: false
#| output: asis

cat("::: {.callout-note}\nThe frozen spending dataset has not been produced yet.",
    "Run the `/identify-spending` skill to identify, consolidate, and stamp the",
    "spending acts; this section will then render the headline table.\n:::\n")
```
:::

## Motivation evidence and preliminary exogeneity {#sec-exogeneity}

For each spending act the corpus carries motivation language. @tbl-exogeneity pairs
each act with the most diagnostic quote and a **preliminary** exogenous/endogenous
read from the Das et al. [-@das_ai_2026] two-condition screen (exogenous iff the motive is
non-cyclical *and* the narrative does not cite contemporaneous macro conditions as
the rationale; "acknowledgment ≠ endogeneity").

```{r}
#| label: tbl-exogeneity
#| tbl-cap: "**Motivation evidence, pending adjudication**. The stated rationale for each spending act and a preliminary exogenous/endogenous read. Suggestions anchored in quoted text, not settled classifications; the exogeneity calls are the object of expert adjudication, and C2b assigns the downstream label."
#| eval: !expr have_frozen

spend_frozen |>
  dplyr::select(shock_id,
                `Preliminary read` = exogenous_preliminary,
                `Motivation quote (source text)` = exogeneity_quote) |>
  tt() |>
  tt_theme_report()
```

## Structured frozen dataset {#sec-frozen}

The narrative identification above is the human-curated content; the
machine-readable form is frozen by the `/identify-spending` skill to
`data/validated/MY_SPENDING_shocks.qs`, following the contract in
`docs/phase_1/spending_shock_schema.md` (one row per announced spending act, with a
`member_chunks` manifest linking each shock to its corpus chunks and a mandatory
recall scorecard). That frozen file — not this notebook's tribbles — is the
reference input the spending pipeline binds and feeds to the frozen C2a/C2b for
motivation and exogeneity (`tar_read(spending_shocks)`).

```{r}
#| label: tbl-frozen-manifest
#| tbl-cap: "**Member-chunk manifest**. Number of corpus chunks attributed to each spending act, and how many were recovered (present in the corpus but not surfaced by C1 — for spending, expected to be most). The recovered chunks drive the downstream C2a re-run."
#| eval: !expr have_frozen

spend_frozen |>
  dplyr::mutate(
    `Member chunks`    = purrr::map_int(member_chunks, nrow),
    `Recovered chunks` = purrr::map_int(recovered_chunks, nrow)
  ) |>
  dplyr::select(shock_id, act_label, `Member chunks`, `Recovered chunks`) |>
  tt() |>
  tt_theme_report()
```

## References {.unnumbered}

::: {#refs}
:::
