---
title: "Malaysia Statutory Tax-Change Dataset"
subtitle: "Identified tax shocks (CIT / PIT / consumption) with motivation and exogeneity"
number-sections: true
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, scales)

here::i_am("notebooks/tax_shocks.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
source(here("R/malay_consistency.R"))  # pretty_motivation() + motivation palette
set_theme(theme_minimal())

tax_shocks <- tryCatch(tar_read(tax_shocks), error = function(e) tibble::tibble())
have_data <- nrow(tax_shocks) > 0
```

## Purpose

This is the deployment deliverable: the bound, classified dataset of **statutory
tax changes** for Malaysia, produced by the per-instrument identification skills
(`/identify-cit`, `/identify-pit`, `/identify-vat`) and the `/identify-tax-shocks`
orchestrator. Each row is one narratively-announced act × tax type, carrying the
identified rate change (Vegh & Vuletin 2015-style), plus a motivation label and an
exogenous/endogenous read from the validated C2b codebook.

The dataset is built to be sent to a human expert for validation. Following the
project's narrative-identification discipline, the exogeneity columns are
**preliminary inputs requiring adjudication, not findings**: the deliverable keeps
both Claude's narrative read (`exogenous_preliminary`) and C2b's classification
(`c2b_exogenous`) side by side, and @sec-disagree flags where they differ.

::: {.content-visible when-format="html"}
```{r}
#| label: pending
#| eval: !expr "!have_data"
#| echo: false
#| output: asis

cat("::: {.callout-important}\nThe `tax_shocks` target is empty or unbuilt. Run the",
    "`/identify-tax-shocks` orchestrator (freeze the per-instrument datasets, then",
    "`tar_make(tax_shocks_evidence)` → `tar_make(tax_shocks_c2b)` → `tar_make(tax_shocks)`,",
    "each gated). This notebook will then render the inventory, rate paths, and timeline.\n:::\n")
```
:::

## Statutory tax-change inventory {#sec-inventory}

```{r}
#| label: tbl-inventory
#| tbl-cap: "**Statutory tax-change inventory**. One row per announced act × tax type. `Δpp` is the change in the headline statutory rate; motivation and exogeneity are from the validated C2b codebook (v0.9.1). The preliminary narrative exogeneity read is shown for comparison and is pending expert adjudication."
#| eval: !expr have_data

tax_shocks |>
  dplyr::arrange(tax_type, effective_year) |>
  dplyr::transmute(
    `Shock`        = shock_id,
    `Act`          = act_label,
    `Tax`          = tax_type,
    `Direction`    = direction,
    `Δpp`          = delta_pp,
    `Effective`    = effective_year,
    `Motivation`   = pretty_motivation(c2b_label),
    `Exo. (C2b)`   = c2b_exogenous,
    `Exo. (prelim)` = exogenous_preliminary
  ) |>
  tt() |>
  tt_theme_report()
```

## Rate paths {#sec-rate-paths}

```{r}
#| label: fig-rate-paths
#| fig-cap: "**Statutory rate paths by tax type**. Step path of the headline statutory rate implied by the identified rate-bearing shocks, faceted by tax type. Non-rate shocks (e.g. one-off levies, regime introductions without a standing rate) are omitted from the path."
#| eval: !expr have_data

rate_path <- tax_shocks |>
  dplyr::filter(!is.na(rate_to), !is.na(effective_year)) |>
  dplyr::arrange(tax_type, effective_year)

if (nrow(rate_path) > 0) {
  ggplot(rate_path, aes(effective_year, rate_to)) +
    geom_step(direction = "hv", linewidth = 0.8, colour = "grey20") +
    geom_point(size = 2) +
    facet_wrap(~ tax_type, ncol = 1, scales = "free_y") +
    scale_y_continuous(labels = scales::label_percent(scale = 1)) +
    scale_x_continuous(breaks = scales::breaks_width(5)) +
    labs(x = NULL, y = "Statutory rate")
} else {
  knitr::asis_output("_No rate-bearing shocks to plot._")
}
```

## Timeline {#sec-timeline}

```{r}
#| label: fig-timeline
#| fig-cap: "**Tax-change timeline**. Acts counted per year × tax type × motivation × direction as a diverging stacked bar: hikes grow the bar upward, cuts downward, neutral acts sit at the baseline. Colour = C2b motivation. Faceted by tax type."
#| eval: !expr have_data

tl <- tax_shocks |>
  dplyr::filter(!is.na(effective_year)) |>
  dplyr::mutate(
    sign = dplyr::case_when(direction == "Hike" ~ "+",
                            direction == "Cut"  ~ "-",
                            TRUE                 ~ "0"),
    motivation = dplyr::if_else(sign == "0", "No change",
                                as.character(pretty_motivation(c2b_label)))
  ) |>
  dplyr::count(tax_type, effective_year, sign, motivation, name = "n") |>
  dplyr::mutate(
    signed_n   = dplyr::if_else(sign == "-", -n, n),
    motivation = factor(motivation,
                        levels = c(unname(.malay_motivation_labels), "No change"))
  )

if (nrow(tl) > 0) {
  ggplot(tl, aes(effective_year, signed_n, fill = motivation)) +
    geom_col(width = 0.7) +
    geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey40") +
    scale_x_continuous(breaks = scales::breaks_width(5)) +
    scale_fill_manual(values = .malay_motivation_palette, name = "Motivation",
                      drop = FALSE) +
    facet_wrap(~ tax_type, ncol = 1) +
    labs(x = NULL, y = "Acts (hike ↑ / cut ↓)")
} else {
  knitr::asis_output("_No dated shocks to plot._")
}
```

## Where the preliminary read and C2b disagree {#sec-disagree}

The cases where Claude's preliminary narrative read and C2b's classification
diverge are the priority for expert adjudication — they are the methodologically
consequential calls, not noise.

```{r}
#| label: tbl-disagree
#| tbl-cap: "**Exogeneity disagreements**. Shocks where the preliminary narrative read and the C2b exogenous/endogenous classification differ. These are the priority cases for expert adjudication."
#| eval: !expr have_data

disagree <- tax_shocks |>
  dplyr::mutate(prelim_exo = dplyr::case_when(
    exogenous_preliminary %in% c("TRUE", "true")   ~ TRUE,
    exogenous_preliminary %in% c("FALSE", "false") ~ FALSE,
    TRUE                                           ~ NA
  )) |>
  dplyr::filter(!is.na(prelim_exo), !is.na(c2b_exogenous), prelim_exo != c2b_exogenous)

if (nrow(disagree) > 0) {
  disagree |>
    dplyr::transmute(
      `Shock` = shock_id, `Act` = act_label, `Tax` = tax_type,
      `Exo. (prelim)` = prelim_exo, `Exo. (C2b)` = c2b_exogenous,
      `Motivation (C2b)` = pretty_motivation(c2b_label)
    ) |>
    tt() |>
    tt_theme_report()
} else {
  knitr::asis_output("_No disagreements: the preliminary read and C2b agree on every shock._")
}
```
