---
title: "Manual Identification of Malaysia's Statutory Consumption-Tax Shocks"
subtitle: "AI-assisted narrative pass over the C1 / C0 deployment evidence"
number-sections: true
---

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, tinytable, here, qs2)

here::i_am("notebooks/vat_identification.qmd")
tar_config_set(store = here("_targets"))
source(here("R/tt_theme.R"))
set_theme(theme_minimal())

# Full Malaysia C1 measure pool (pre-relevance-cut deployment output)
c1 <- tar_read(country_c1_measures)[[1]]

vat_kw <- stringr::regex(
  paste0("goods and services tax|GST|cukai barang|cukai barangan|sales and service|",
         "sales tax|service tax|\\bSST\\b|cukai jualan|cukai perkhidmatan|",
         "value.added|\\bVAT\\b|consumption tax"),
  ignore_case = TRUE
)
n_pool <- nrow(c1)
n_vat  <- c1 |> dplyr::filter(stringr::str_detect(measure_name, vat_kw)) |> nrow()

# Frozen, human-stamped consumption-tax shock dataset (produced by /identify-vat).
# Loaded here so both the recall scorecard (@sec-recall) and the structured
# dataset (@sec-frozen) render from the single source of truth when it exists.
frozen_path <- here("data", "validated", "MY_CONSUMPTION_shocks.qs")
have_frozen <- file.exists(frozen_path)
if (have_frozen) vat_frozen <- qs2::qs_read(frozen_path)
```

## Purpose

This note documents a focused, **AI-assisted manual identification** of Malaysia's *statutory broad consumption-tax changes* — the value-added / sales-and-service tax in its successive Malaysian forms — run directly over the evidence our deployment pipeline already surfaced. It is the consumption-tax counterpart to the CIT and PIT seeds, and it deliberately follows the spirit of the narrative-identification literature [@romer2010]: close reading of source documents, with the LLM pipeline as a retrieval-and-extraction aide, not as the adjudicator.

The method mirrors our `/manual-analysis` protocol. The pipeline (C1 measure identification, C0 act aggregation) pre-screens candidate measures; a human reads the surfaced evidence and makes the final call on *what counts as an event*, *how scattered surface forms consolidate*, and *whether each change is exogenous*. Claude pattern-matches against the source text; the researcher confers meaning.

We identify by the **concept** — a broad statutory tax on consumption / value added / sales of goods and services — using whichever surface terms Malaysia employs (GST, SST, sales tax, service tax). Unlike a Western-style standing VAT, Malaysia's broad consumption tax changed *regime* twice in the corpus window: the Sales Tax (1972) and Service Tax (1975) baseline was replaced by the Goods and Services Tax (GST) in 2015, which was then itself abolished and replaced by a reinstated Sales and Service Tax (SST) in 2018. **Regime introductions and abolitions count as shocks**, which is why this instrument carries more structural events than a simple rate path.

We do **not** drive identification from C0. As @sec-recall notes, C0 fragments the GST measures across many clusters and never bridges the English/Malay surface forms, so it cannot serve as the event layer. We retrieve from the full C1 pool (`r n_pool` measures, pre-relevance-cut) and consolidate by hand.

::: {.callout-important}
The exogenous/endogenous column in @tbl-exogeneity is a **preliminary suggestion with its supporting quote, pending expert adjudication**. It is not a research finding. The whole point of this pass is that a human adjudicates these calls; see the open questions in @sec-open.
:::

## What C1 surfaced, and what it missed {#sec-recall}

A keyword scan of the full C1 pool returns **`r n_vat` measures** touching the consumption-tax regex, the bulk of them GST surface forms (English and Malay), spanning 2001–2022. A broader semantic sweep over all `r n_pool` measures — reading every consumption-tax surface form and then the anchor source documents directly — confirms **four** headline statutory events and rules out any further broad-rate change across 1980–2023.

Recall here is **better than for CIT**: C1 surfaced all four events as measures. What needed manual recovery was finer detail — the reinstated-SST rate schedule (Sales Tax 5% / 10%, Service Tax 6%) and the SST commencement passages, plus a Bank Negara corroboration of the 2011 service-tax rise whose source document is not in the chunk store.

```{r}
#| label: tbl-recall
#| tbl-cap: "**Recall scorecard**. How each headline consumption-tax event fared through C1 measure identification and C0 act aggregation, and which supporting detail had to be recovered by reading the source documents directly."

recall <- if (have_frozen) {
  vat_frozen$recall_scorecard[[1]] |>
    dplyr::rename(Stage = stage, Outcome = outcome)
} else {
  tibble::tribble(
    ~Stage,                       ~Outcome,
    "C1 measure identification",  "Surfaced all four headline events (service tax 5->6, GST introduction, GST abolition, SST reintroduction)",
    "Recall miss - SST detail",   "SST rate schedule and commencement passages recovered from Economic Report 2019 and BNM AR 2018",
    "C0 act aggregation",         "Fragmented the GST measures across many clusters; EN/BM never bridged"
  )
}

recall |>
  tt() |>
  tt_theme_report()
```

The retrieval lesson echoes the CIT and PIT passes: a consumption-tax change is maximally relevant *because it is a rate or regime change*, however briefly the documents discuss it. An instrument-targeted recall pass plus human reading is needed rather than trust in discussion volume or in C0's clustering.

## The headline statutory consumption tax {#sec-headline}

Reading the consolidated evidence yields a clean structural narrative: a **service-tax rise from 5% to 6%** (2011), the **introduction of GST at 6%** (2015, replacing the SST regime), the **abolition / zero-rating of GST** (2018), and the **reinstatement of SST** (2018). @tbl-headline gives the identified events; @fig-headline-path traces the standard consumption-tax rate and marks the two regime transitions.

```{r}
#| label: tbl-headline
#| tbl-cap: "**Headline consumption-tax events**. Identified statutory broad consumption-tax changes, with the budget that announced each and the year it took effect. SST is a multi-rate regime (Sales Tax 5%/10%, Service Tax 6%), so its reinstatement carries no single 'to' rate."

headline <- tibble::tribble(
  ~Event, ~Change,                                   ~`From -> To`, ~`Effective`,  ~`Announced`,             ~`Source surface form / doc`,
  "V1",   "Service tax rate increase",               "5% -> 6%",    "2011 (1 Jan)", "Budget 2011 (Oct 2010)", "Budget Speech 2011; BNM AR 2010",
  "V2",   "GST introduction (replacing SST)",        "0% -> 6%",    "2015 (1 Apr)", "Budget 2014 (Oct 2013)", "Budget Speech 2014; ECON Report 2015",
  "V3",   "GST abolition / zero-rating",             "6% -> 0%",    "2018 (1 Jun)", "PH reform (2018)",       "Budget Speech 2019; ECON Report 2019",
  "V4",   "SST reinstatement (Sales 5/10%, Svc 6%)", "0% -> SST",   "2018 (1 Sep)", "PH reform (2018)",       "BNM AR 2018; ECON Report 2019"
)

headline |>
  tt() |>
  tt_theme_report()
```

```{r}
#| label: fig-headline-path
#| fig-cap: "**Standard statutory consumption-tax rate, Malaysia 2009-2023**. The general service-tax rate (pre-2015), the GST standard rate (2015-2018), and the reinstated Service Tax rate (post-2018). The two regime transitions - SST -> GST (2015) and GST -> SST (2018) - are marked; the SST sales-tax component (5% / 10%) is not a single standard rate and is omitted from the line."

rate_path <- tibble::tribble(
  ~year, ~rate,
  2009,  5,
  2011,  6,
  2015,  6,   # GST introduced at 6% (replaces SST)
  2018,  0,   # GST zero-rated / abolished
  2018.7, 6,  # SST service-tax rate on reinstatement
  2023,  6
)

regimes <- tibble::tribble(
  ~year, ~rate, ~label,
  2015,  6,     "  SST -> GST\n  (1 Apr 2015)",
  2018,  0,     "  GST -> SST\n  (Jun-Sep 2018)"
)

ggplot(rate_path, aes(year, rate)) +
  geom_step(direction = "hv", linewidth = 0.8, colour = "grey20") +
  geom_point(data = regimes, shape = 17, size = 3, colour = "firebrick") +
  geom_text(data = regimes, aes(label = label), hjust = 0, vjust = -0.2,
            size = 3, colour = "firebrick") +
  scale_y_continuous(labels = scales::label_percent(scale = 1), limits = c(-1, 8)) +
  scale_x_continuous(breaks = seq(2009, 2023, 2)) +
  labs(x = NULL, y = "Standard consumption-tax rate")
```

## Motivation evidence and preliminary exogeneity {#sec-exogeneity}

Each headline event carries motivation language in the corpus. @tbl-exogeneity pairs each event with the most diagnostic quote and a **preliminary** exogenous/endogenous read. The methodologically interesting separation is the contrast between the **2015 GST introduction** — framed as a long-run structural reform to broaden the base and reduce oil-revenue dependence — and the **2018 abolition** — a campaign-promise, cost-of-living measure whose classification is genuinely contestable.

```{r}
#| label: tbl-exogeneity
#| tbl-cap: "**Motivation evidence, pending adjudication**. The stated rationale for each headline event and a preliminary exogenous/endogenous read. These are suggestions anchored in the quoted text, not settled classifications; the exogeneity calls are the object of expert adjudication (see @sec-open)."

exo <- if (have_frozen) {
  vat_frozen |>
    dplyr::transmute(
      Event = sub("MY-CONSUMPTION-0", "V", shock_id),
      `Motivation quote (source text)` = exogeneity_quote,
      `Preliminary read` = exogenous_preliminary,
      Adjudication = "PENDING"
    )
} else {
  tibble::tribble(
    ~Event, ~`Motivation quote (source text)`, ~`Preliminary read`, ~Adjudication,
    "V1", "service-tax rise to increase revenue collection",                 "FALSE (revenue)",            "PENDING",
    "V2", "broaden the tax base and reduce dependence on oil revenue",       "TRUE (structural)",          "PENDING",
    "V3", "Pakatan Harapan manifesto; lessen tax burden, reduce prices",     "ambiguous (cost-of-living)", "PENDING",
    "V4", "less burdensome SST; part of the 2018 tax-regime change",         "ambiguous (PH package)",     "PENDING"
  )
}

exo |>
  tt() |>
  tt_theme_report()
```

## Other consumption-tax tracks {#sec-tracks}

For completeness, @tbl-tracks lists consumption-tax-adjacent changes that were *excluded* from the headline series as narrow carve-outs rather than broad standard-rate shocks. They are tracked so the exclusion is auditable.

```{r}
#| label: tbl-tracks
#| tbl-cap: "**Excluded consumption-tax tracks**. Product-specific sales-tax changes and GST base/exemption tweaks that are not changes to the broad standard rate, plus the pre-corpus baseline."

tracks <- tibble::tribble(
  ~Track,                       ~`Identified items`,                                                                 ~Note,
  "Product-specific sales tax", "Liquor 15%->20% and cigarettes 15%->25% (Budget 2001); 1997/98 crisis exemptions on food, medication, paper", "Excise-like carve-outs on selected goods, not the general rate",
  "GST base / exemption tweaks", "Zero-rating expansions and exemption streamlining (Budgets 2016, 2017)",            "Base changes within the GST regime, not a standard-rate shock",
  "Pre-corpus baseline",        "Sales Tax Act 1972 (5%/10%); Service Tax Act 1975 (6%)",                            "Predates the 1980 corpus start; the standing baseline before V1"
)

tracks |>
  tt() |>
  tt_theme_report()
```

## Open adjudication questions {#sec-open}

Three calls are genuinely the researcher's to make; they were flagged at the `/identify-vat` PAUSE point and adjudicated at the freeze.

::: {.callout-tip}
## Resolved at freeze (2026-06-26)

1. **Regime-replacement grain.** The two regime swaps are treated **asymmetrically**, following the corpus's own framing. The **2015 SST -> GST** swap is **one row** (`MY-CONSUMPTION-02`, GST introduction), folding the simultaneous SST abolition into the launch (both dated 1 April 2015). The **2018 GST -> SST** swap is **two rows** (`MY-CONSUMPTION-03` abolition + `MY-CONSUMPTION-04` reinstatement), because the corpus dates them three months apart (GST zero-rated 1 June; SST effective 1 September) with an interim tax holiday.
2. **2015 GST exogeneity.** Frozen as **TRUE (exogenous)** on the long-run structural-reform framing ("broaden the tax base and reduce dependence on oil revenue"), carried alongside — never replacing — C2b's downstream label.
3. **2018 abolition / SST exogeneity.** Frozen as **ambiguous**, preserving the contestable campaign-promise / cost-of-living framing as the object of expert adjudication.

The narrative above is retained as the reasoning record behind these calls.
:::

::: {.callout-note}
## 1. Grain of the two regime replacements
The 2015 swap is one act here and the 2018 swap is two. The alternative is full symmetry — either split 2015 into an SST-abolition row plus a GST-introduction row (five rows total), or collapse the 2018 swap into a single "GST -> SST regime change" row (three rows total). The asymmetric default tracks how the documents themselves date and emphasise each transition.
:::

::: {.callout-note}
## 2. Exogeneity of the 2015 GST introduction
The corpus frames GST as a long-run structural reform to broaden the base and reduce oil-revenue dependence — pulling toward exogenous. The countervailing reading is fiscal-consolidation-driven (still exogenous, different reason) given the deficit context. The preliminary read is TRUE; the reason label is the live question.
:::

::: {.callout-note}
## 3. Exogeneity of the 2018 abolition and SST return
The identical 2018 package is framed as an ideological campaign promise *and* as cost-of-living relief that raises household purchasing power. The first reading leans exogenous (long-run political), the second endogenous (countercyclical). Adjudicating this is the methodologically interesting contribution, not overhead.
:::

## Structured frozen dataset {#sec-frozen}

The narrative identification above is the human-curated content; the machine-readable
form is frozen by the `/identify-vat` skill to `data/validated/MY_CONSUMPTION_shocks.qs`,
following the shared contract in `docs/phase_1/tax_shock_schema.md` (one row per
announced act x tax type, with a `member_chunks` manifest linking each shock to its
corpus chunks and a mandatory recall scorecard). That frozen file — not this
notebook's tribbles — is the reference input the deployment pipeline binds and feeds
to C2a/C2b for motivation and exogeneity (`tar_read(tax_shocks)`).

@sec-recall above is the recall scorecard for this pass. The frozen dataset and its
per-shock member-chunk manifest render below once the skill has produced it.

::: {.content-visible when-format="html"}
```{r}
#| label: frozen-pending
#| eval: !expr "!have_frozen"
#| echo: false
#| output: asis

cat("::: {.callout-note}\nThe frozen consumption-tax dataset has not been produced yet. Run the",
    "`/identify-vat` skill to identify, consolidate, and stamp the shocks; this",
    "section will then render the structured dataset and its member-chunk manifest.\n:::\n")
```
:::

```{r}
#| label: tbl-frozen
#| tbl-cap: "**Frozen consumption-tax shock dataset**. Structured statutory events as stamped by the `/identify-vat` skill (selected columns). The preliminary exogeneity read is pending expert adjudication and is carried alongside - never replaced by - C2b's downstream label."
#| eval: !expr have_frozen

vat_frozen |>
  dplyr::select(shock_id, act_label, direction, rate_from, rate_to, `Delta pp` = delta_pp,
                `Effective` = effective_year, `Preliminary exo.` = exogenous_preliminary) |>
  tt() |>
  tt_theme_report()
```

```{r}
#| label: tbl-frozen-manifest
#| tbl-cap: "**Member-chunk manifest**. Number of corpus chunks attributed to each shock, and how many were recovered (present in the corpus but not surfaced by C1). The recovered chunks drive the downstream C2a re-run."
#| eval: !expr have_frozen

vat_frozen |>
  dplyr::mutate(
    `Member chunks`    = purrr::map_int(member_chunks, nrow),
    `Recovered chunks` = purrr::map_int(recovered_chunks, nrow)
  ) |>
  dplyr::select(shock_id, act_label, `Member chunks`, `Recovered chunks`) |>
  tt() |>
  tt_theme_report()
```

## References {.unnumbered}

::: {#refs}
:::
