---
title: "C2 Input Verification"
subtitle: "Validate C1→C2 handoff: classified chunks and motivation-filtered input"
execute:
  cache: false
---

## Overview

Before building the C2 codebook or running evaluation, this notebook verifies the C1→C2 data handoff. Two targets are checked:

- `c1_classified_chunks`: C1 S2 classifications merged with chunk text and ground truth labels from `aligned_data`
- `c2_input_data`: Filtered to `pred_label == "FISCAL_MEASURE"` and `discusses_motivation == TRUE`

Eight tests check coverage, label completeness, text joins, C1 output fields, filter correctness, post-filter act coverage, class balance, and token distribution.

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, gt, here, scales)

here::i_am("notebooks/verify_c2_inputs.qmd")
tar_config_set(store = here("_targets"))
source(here("R/gt_theme.R"))
```

```{r}
#| label: load-data

c1_classified <- tar_read(c1_classified_chunks)
c2_input      <- tar_read(c2_input_data)
aligned_data  <- tar_read(aligned_data)
```

```{r}
#| label: init-results

test_results <- tibble(
  test_id = integer(),
  test_name = character(),
  result = character(),
  status = character()
)

add_result <- function(id, name, result, status) {
  test_results <<- bind_rows(
    test_results,
    tibble(test_id = id, test_name = name, result = result, status = status)
  )
}
```

## Part A: `c1_classified_chunks` (Precursor)

### Test 1: Act Coverage

Every act in `aligned_data` should appear in `c1_classified_chunks`.

```{r}
#| label: test-1-coverage

aligned_acts <- sort(unique(aligned_data$act_name))
classified_acts <- sort(unique(c1_classified$act_name))

missing <- setdiff(aligned_acts, classified_acts)
extra <- setdiff(classified_acts, aligned_acts)
coverage_ok <- length(missing) == 0 && length(extra) == 0

add_result(
  1, "Act coverage",
  sprintf("%d/%d acts present", length(classified_acts), length(aligned_acts)),
  if (coverage_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: test-1-details
#| results: asis

if (length(missing) > 0) {
  cat("\n\n**Acts in aligned_data but missing from c1_classified_chunks:**\n\n")
  for (act in missing) cat(sprintf("- %s\n", act))
}
if (length(extra) > 0) {
  cat("\n\n**Acts in c1_classified_chunks but not in aligned_data:**\n\n")
  for (act in extra) cat(sprintf("- %s\n", act))
}
```

**Test 1**: `r if(coverage_ok) "PASS" else paste0("FAIL — ", length(missing), " missing, ", length(extra), " extra")`

### Test 2: Label Completeness

Every row must have non-NA `motivation_category` and `exogenous_flag`.

```{r}
#| label: test-2-labels

n_missing_motivation <- sum(is.na(c1_classified$motivation_category))
n_missing_exogenous <- sum(is.na(c1_classified$exogenous_flag))
labels_ok <- n_missing_motivation == 0 && n_missing_exogenous == 0

add_result(
  2, "Label completeness",
  sprintf("motivation: %d NA, exogenous: %d NA",
          n_missing_motivation, n_missing_exogenous),
  if (labels_ok) "PASS" else "FAIL"
)
```

**Test 2**: `r if(labels_ok) "PASS" else paste0("FAIL — ", n_missing_motivation + n_missing_exogenous, " missing labels")`

### Test 3: Text Join Completeness

All rows should have non-NA, non-empty `text` after joining from `chunks`.

```{r}
#| label: test-3-text

n_na_text <- sum(is.na(c1_classified$text))
n_empty_text <- sum(!is.na(c1_classified$text) & nchar(c1_classified$text) == 0)
text_ok <- n_na_text == 0 && n_empty_text == 0

add_result(
  3, "Text join completeness",
  sprintf("%d NA, %d empty (of %d rows)", n_na_text, n_empty_text,
          nrow(c1_classified)),
  if (text_ok) "PASS" else "FAIL"
)
```

**Test 3**: `r if(text_ok) "PASS" else paste0("FAIL — ", n_na_text, " NA + ", n_empty_text, " empty")`

### Test 4: C1 Output Field Completeness

`pred_label` should be non-NA for all rows. Report the distribution of `discusses_motivation`, `discusses_timing`, and `discusses_magnitude`.

```{r}
#| label: test-4-c1-fields

n_na_pred <- sum(is.na(c1_classified$pred_label))
pred_ok <- n_na_pred == 0

add_result(
  4, "C1 output fields",
  sprintf("pred_label: %d NA; discusses_motivation: %d TRUE / %d FALSE / %d NA",
          n_na_pred,
          sum(c1_classified$discusses_motivation == TRUE, na.rm = TRUE),
          sum(c1_classified$discusses_motivation == FALSE, na.rm = TRUE),
          sum(is.na(c1_classified$discusses_motivation))),
  if (pred_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: tbl-c1-pred-dist
#| tbl-cap: "C1 prediction distribution"

c1_classified |>
  count(pred_label, name = "n") |>
  mutate(pct = n / sum(n)) |>
  gt() |>
  cols_label(pred_label = "Predicted Label", n = "Chunks", pct = "Share") |>
  fmt_percent(columns = pct, decimals = 1) |>
  gt_theme_report()
```

```{r}
#| label: tbl-discusses-flags
#| tbl-cap: "Extra output field distribution (FISCAL_MEASURE chunks only)"

c1_classified |>
  filter(pred_label == "FISCAL_MEASURE") |>
  summarise(
    motivation_TRUE = sum(discusses_motivation == TRUE, na.rm = TRUE),
    motivation_FALSE = sum(discusses_motivation == FALSE, na.rm = TRUE),
    motivation_NA = sum(is.na(discusses_motivation)),
    timing_TRUE = sum(discusses_timing == TRUE, na.rm = TRUE),
    timing_FALSE = sum(discusses_timing == FALSE, na.rm = TRUE),
    timing_NA = sum(is.na(discusses_timing)),
    magnitude_TRUE = sum(discusses_magnitude == TRUE, na.rm = TRUE),
    magnitude_FALSE = sum(discusses_magnitude == FALSE, na.rm = TRUE),
    magnitude_NA = sum(is.na(discusses_magnitude))
  ) |>
  pivot_longer(everything(),
               names_to = c("field", "value"),
               names_sep = "_",
               values_to = "n") |>
  pivot_wider(names_from = value, values_from = n) |>
  gt() |>
  cols_label(field = "Field") |>
  gt_theme_report()
```

**Test 4**: `r if(pred_ok) "PASS" else paste0("FAIL — ", n_na_pred, " NA predictions")`

## Part B: `c2_input_data` (Filtered)

### Test 5: Filter Correctness

All rows must have `pred_label == "FISCAL_MEASURE"` and `discusses_motivation == TRUE`.

```{r}
#| label: test-5-filter

n_wrong_label <- sum(c2_input$pred_label != "FISCAL_MEASURE", na.rm = TRUE)
n_wrong_motiv <- sum(c2_input$discusses_motivation != TRUE, na.rm = TRUE)
filter_ok <- n_wrong_label == 0 && n_wrong_motiv == 0

add_result(
  5, "Filter correctness",
  sprintf("Non-FISCAL_MEASURE: %d, discusses_motivation != TRUE: %d",
          n_wrong_label, n_wrong_motiv),
  if (filter_ok) "PASS" else "FAIL"
)
```

**Test 5**: `r if(filter_ok) "PASS" else "FAIL — filter leakage detected"`

### Test 6: Act Coverage Post-Filter

How many aligned acts survive the motivation filter? Acts that lose all chunks cannot be evaluated by C2.

```{r}
#| label: test-6-post-filter

c2_acts <- unique(c2_input$act_name)
lost_acts <- setdiff(aligned_acts, c2_acts)
n_lost <- length(lost_acts)
post_filter_ok <- n_lost == 0

add_result(
  6, "Act coverage post-filter",
  sprintf("%d/%d acts retained (%d lost)",
          length(c2_acts), length(aligned_acts), n_lost),
  if (post_filter_ok) "PASS" else "WARN"
)
```

```{r}
#| label: test-6-details
#| results: asis

if (n_lost > 0) {
  cat("\n\n**Acts with no FISCAL_MEASURE + discusses_motivation chunks:**\n\n")
  lost_info <- aligned_data |>
    filter(act_name %in% lost_acts) |>
    select(act_name, year, motivation_category) |>
    distinct()
  for (i in seq_len(nrow(lost_info))) {
    cat(sprintf("- %s (%d) — %s\n",
                lost_info$act_name[i], lost_info$year[i],
                lost_info$motivation_category[i]))
  }
}
```

**Test 6**: `r if(post_filter_ok) "PASS" else paste0("WARN — ", n_lost, " act(s) lost all chunks")`

### Test 7: Motivation Class Balance

Distribution of `motivation_category` across C2 input chunks, compared to `aligned_data`.

```{r}
#| label: test-7-balance

c2_act_dist <- c2_input |>
  distinct(act_name, motivation_category) |>
  count(motivation_category, name = "n_c2") |>
  arrange(desc(n_c2))

aligned_dist <- aligned_data |>
  count(motivation_category, name = "n_aligned") |>
  arrange(desc(n_aligned))

balance_check <- c2_act_dist |>
  full_join(aligned_dist, by = "motivation_category") |>
  mutate(
    n_c2 = replace_na(n_c2, 0L),
    n_aligned = replace_na(n_aligned, 0L),
    match = n_c2 == n_aligned
  )

balance_ok <- all(balance_check$match)

add_result(
  7, "Motivation class balance",
  sprintf("%d/%d categories match aligned_data",
          sum(balance_check$match), nrow(balance_check)),
  if (balance_ok) "PASS" else "WARN"
)
```

```{r}
#| label: tbl-balance
#| tbl-cap: "Motivation category: acts in c2_input vs aligned_data"

exog_map <- aligned_data |>
  distinct(motivation_category, exogenous_flag)

balance_check |>
  left_join(exog_map, by = "motivation_category") |>
  select(motivation_category, exogenous_flag, n_c2, n_aligned) |>
  gt() |>
  cols_label(
    motivation_category = "Category",
    exogenous_flag = "Exogenous?",
    n_c2 = "c2_input",
    n_aligned = "aligned_data"
  ) |>
  gt_theme_report()
```

```{r}
#| label: tbl-chunks-per-category
#| tbl-cap: "Chunk count per motivation category in c2_input"

c2_input |>
  count(motivation_category, name = "n_chunks") |>
  arrange(desc(n_chunks)) |>
  gt() |>
  cols_label(motivation_category = "Category", n_chunks = "Chunks") |>
  gt_theme_report()
```

**Test 7**: `r if(balance_ok) "PASS" else "WARN — some categories lost acts after filtering"`

### Test 8: Token Distribution

Per-chunk token distribution. Chunks were constructed to fit within the context window, so all should be well under 200K.

```{r}
#| label: test-8-tokens

token_summary <- c2_input |>
  summarise(
    min = min(approx_tokens),
    p25 = quantile(approx_tokens, 0.25),
    median = median(approx_tokens),
    mean = round(mean(approx_tokens)),
    p75 = quantile(approx_tokens, 0.75),
    max = max(approx_tokens)
  )

overhead_tokens <- 10000
context_limit <- 200000
effective_limit <- context_limit - overhead_tokens
n_over <- sum(c2_input$approx_tokens > effective_limit)
tokens_ok <- n_over == 0

add_result(
  8, "Token budget fit",
  sprintf("Max: %s tokens (limit: %s)",
          comma(token_summary$max), comma(effective_limit)),
  if (tokens_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: tbl-token-summary
#| tbl-cap: "Token distribution across C2 input chunks"

token_summary |>
  pivot_longer(everything(), names_to = "Statistic", values_to = "Tokens") |>
  gt() |>
  fmt_number(columns = Tokens, decimals = 0) |>
  gt_theme_report()
```

```{r}
#| label: fig-token-hist
#| fig-cap: "Approximate token count per chunk (C2 input)"
#| fig-width: 8
#| fig-height: 4

ggplot(c2_input, aes(x = approx_tokens)) +
  geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = effective_limit, color = "red",
             linetype = "dashed", linewidth = 0.8) +
  annotate("text", x = effective_limit, y = Inf,
           label = "Context limit", hjust = 1.1, vjust = 2,
           color = "red", size = 3.5) +
  scale_x_continuous(labels = comma) +
  labs(x = "Approximate Tokens", y = "Chunks") +
  theme_minimal()
```

**Test 8**: `r if(tokens_ok) "PASS" else paste0("FAIL — ", n_over, " chunks exceed context limit")`

## Per-Act Summary

```{r}
#| label: tbl-per-act
#| tbl-cap: "Per-act chunk summary: classified vs C2 input"

classified_summary <- c1_classified |>
  group_by(act_name) |>
  summarise(
    n_total = n(),
    n_fiscal = sum(pred_label == "FISCAL_MEASURE", na.rm = TRUE),
    .groups = "drop"
  )

c2_summary <- c2_input |>
  count(act_name, name = "n_c2")

per_act <- aligned_data |>
  distinct(act_name, year, motivation_category) |>
  left_join(classified_summary, by = "act_name") |>
  left_join(c2_summary, by = "act_name") |>
  mutate(
    n_total = replace_na(n_total, 0L),
    n_fiscal = replace_na(n_fiscal, 0L),
    n_c2 = replace_na(n_c2, 0L)
  ) |>
  arrange(desc(n_c2))

per_act |>
  gt() |>
  cols_label(
    act_name = "Act",
    year = "Year",
    motivation_category = "Category",
    n_total = "Classified",
    n_fiscal = "FISCAL_MEASURE",
    n_c2 = "C2 Input"
  ) |>
  gt_theme_report()
```

## Summary Dashboard

```{r}
#| label: tbl-dashboard
#| tbl-cap: "C2 input verification dashboard"

dashboard <- test_results |>
  mutate(test_label = sprintf("T%d: %s", test_id, test_name)) |>
  select(test_label, result, status)

n_pass <- sum(dashboard$status == "PASS")
n_warn <- sum(dashboard$status == "WARN")
n_fail <- sum(dashboard$status == "FAIL")

dashboard |>
  gt() |>
  cols_label(
    test_label = "Test",
    result = "Result",
    status = "Status"
  ) |>
  tab_style(
    style = cell_fill(color = "#ccffcc"),
    locations = cells_body(columns = status, rows = status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3cd"),
    locations = cells_body(columns = status, rows = status == "WARN")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcccc"),
    locations = cells_body(columns = status, rows = status == "FAIL")
  ) |>
  gt_theme_report()
```

**Overall**: `r n_pass` PASS, `r n_warn` WARN, `r n_fail` FAIL out of `r nrow(dashboard)` tests.

`r if(n_fail > 0) "**ACTION REQUIRED:** Fix all FAIL tests before proceeding to C2 codebook development." else if(n_warn > 0) "**Review warnings.** WARN tests flag potential issues (e.g., acts lost after filtering) that may affect C2 evaluation coverage." else "**All clear:** C1→C2 handoff validated. Data is ready for C2 codebook development."`
