---
title: "Pre-Flight API Input Verification"
subtitle: "Systematic validation of data integrity before S2 LOOCV"
execute:
  cache: false
---

## Overview

Before running expensive S2 LOOCV API calls, this notebook runs 12 systematic checks on the data pipeline outputs. Each test produces a PASS/WARN/FAIL result. All tests must pass (or have documented exceptions) before proceeding.

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, gt, here, jsonlite)

here::i_am("notebooks/verify_api_inputs.qmd")
tar_config_set(store = here("_targets"))
source(here("R/gt_theme.R"))
source(here("R/identify_chunk_tiers.R"))
source(here("R/codebook_stage_0.R"))
source(here("R/generate_c1_examples.R"))
set_theme(theme_minimal())
```

```{r}
#| label: load-data

us_body      <- tar_read(us_body)
chunks       <- tar_read(chunks)
aligned_data <- tar_read(aligned_data)
c1_chunk_data <- tar_read(c1_chunk_data)
c1_codebook  <- tar_read(c1_codebook)
```

```{r}
#| label: init-results

# Accumulator for test results
test_results <- tibble(
  test_id = integer(),
  test_name = character(),
  result = character(),
  status = character()
)

add_result <- function(id, name, result, status) {
  test_results <<- bind_rows(
    test_results,
    tibble(test_id = id, test_name = name, result = result, status = status)
  )
}
```

## Test 1: Chunking Completeness

Every document in the corpus should be fully chunked: the first chunk starts at page 1 and the last chunk ends at or beyond the document's page count.

```{r}
#| label: test-1-chunking-completeness

# Get total pages per document from us_body (aggregated by package_id)
id_col <- if ("package_id" %in% names(us_body)) "package_id" else "doc_id"

doc_pages <- us_body |>
  group_by(across(all_of(id_col))) |>
  summarize(
    n_pages_doc = sum(map_int(text, length)),
    .groups = "drop"
  ) |>
  rename(doc_id = all_of(id_col))

# Get chunk coverage per document
chunk_coverage <- chunks |>
  group_by(doc_id) |>
  summarize(
    min_start = min(start_page),
    max_end = max(end_page),
    n_chunks = n(),
    .groups = "drop"
  )

# Join and check
coverage_check <- doc_pages |>
  filter(n_pages_doc > 0) |>
  inner_join(chunk_coverage, by = "doc_id") |>
  mutate(
    starts_at_1 = min_start == 1,
    covers_end = max_end >= n_pages_doc,
    ok = starts_at_1 & covers_end
  )

n_fail_start <- sum(!coverage_check$starts_at_1)
n_fail_end <- sum(!coverage_check$covers_end)
n_docs_checked <- nrow(coverage_check)
all_ok <- all(coverage_check$ok)
```

```{r}
#| label: test-1-details

if (!all_ok) {
  coverage_check |>
    filter(!ok) |>
    select(doc_id, n_pages_doc, min_start, max_end) |>
    slice_head(n = 10) |>
    gt() |>
    cols_label(
      doc_id = "Document",
      n_pages_doc = "Total Pages",
      min_start = "First Chunk Start",
      max_end = "Last Chunk End"
    ) |>
    gt_theme_report()
}

add_result(
  1, "Chunking completeness",
  sprintf("%d/%d docs fully chunked", sum(coverage_check$ok), n_docs_checked),
  if (all_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: free-us-body

# Free us_body to reduce memory pressure (only needed for Test 1)
rm(us_body)
gc()
```

**Test 1**: `r if(all_ok) "PASS" else paste0("FAIL - ", n_fail_start, " start gaps, ", n_fail_end, " end gaps")`

## Test 2: No Short Chunks (min_chars Filter)

`make_chunks()` drops chunks with 100 or fewer characters as extraction artifacts (`min_chars = 100L`). This test verifies the filter is active.

```{r}
#| label: test-2-short-chunks

chunk_lengths <- nchar(chunks$text)
n_short <- sum(chunk_lengths <= 100)
shortest <- if (n_short > 0) {
  chunks |>
    filter(nchar(text) <= 100) |>
    select(doc_id, chunk_id, n_pages, approx_tokens) |>
    slice_head(n = 5)
} else {
  NULL
}

add_result(
  2, "No short chunks (min_chars filter)",
  sprintf("%d chunks with <=100 chars (of %d total)", n_short, nrow(chunks)),
  if (n_short == 0) "PASS" else "FAIL"
)
```

```{r}
#| label: test-2-details

if (!is.null(shortest)) {
  shortest |>
    gt() |>
    cols_label(doc_id = "Doc", chunk_id = "Chunk", n_pages = "Pages",
               approx_tokens = "Tokens") |>
    gt_theme_report()
}
```

**Test 2**: `r if(n_short == 0) "PASS (min_chars filter active)" else paste0("FAIL - ", n_short, " short chunks leaked through filter")`

## Test 3: No Text-Less Tier Rows

Every Tier 1 and Tier 2 row must have non-empty text after the join from `assemble_c1_chunk_data()`.

```{r}
#| label: test-3-textless

tier1_missing <- c1_chunk_data$tier1 |>
  filter(is.na(text) | nchar(text) == 0) |>
  nrow()

tier2_missing <- c1_chunk_data$tier2 |>
  filter(is.na(text) | nchar(text) == 0) |>
  nrow()

n_textless <- tier1_missing + tier2_missing

add_result(
  3, "No text-less tier rows",
  sprintf("Tier 1: %d missing, Tier 2: %d missing", tier1_missing, tier2_missing),
  if (n_textless == 0) "PASS" else "FAIL"
)
```

**Test 3**: `r if(n_textless == 0) "PASS" else paste0("FAIL - ", n_textless, " rows without text")`

## Test 4: Act Name Alignment

The act names in `aligned_data` should match exactly with the act names in the tier data (both directions).

```{r}
#| label: test-4-act-alignment

aligned_acts <- sort(unique(aligned_data$act_name))
tier_acts <- sort(unique(c(
  c1_chunk_data$tier1$act_name,
  c1_chunk_data$tier2$act_name
)))

in_aligned_not_tiers <- setdiff(aligned_acts, tier_acts)
in_tiers_not_aligned <- setdiff(tier_acts, aligned_acts)

alignment_ok <- length(in_aligned_not_tiers) == 0 && length(in_tiers_not_aligned) == 0

add_result(
  4, "Act name alignment",
  sprintf("Aligned→Tiers: %d missing, Tiers→Aligned: %d extra",
          length(in_aligned_not_tiers), length(in_tiers_not_aligned)),
  if (alignment_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: test-4-details
#| results: asis

if (length(in_aligned_not_tiers) > 0) {
  cat("\n\n**Acts in aligned_data but not in tiers:**\n\n")
  for (act in in_aligned_not_tiers) cat(sprintf("- %s\n", act))
}
if (length(in_tiers_not_aligned) > 0) {
  cat("\n\n**Acts in tiers but not in aligned_data:**\n\n")
  for (act in in_tiers_not_aligned) cat(sprintf("- %s\n", act))
}
```

**Test 4**: `r if(alignment_ok) "PASS" else "FAIL"`

## Test 5: Passage Delimiter Integrity

For acts with multiple passages (`n_passages > 1`), splitting on `\n\n` should produce multiple non-empty elements.

```{r}
#| label: test-5-delimiter

multi_passage <- aligned_data |>
  filter(n_passages > 1)

if (nrow(multi_passage) > 0) {
  split_check <- multi_passage |>
    mutate(
      split_parts = map(passages_text, ~ str_split(.x, "\n\n")[[1]]),
      n_nonempty = map_int(split_parts, ~ sum(nchar(str_trim(.x)) > 0))
    )

  delimiter_failures <- split_check |>
    filter(n_nonempty <= 1)

  n_delim_fail <- nrow(delimiter_failures)
} else {
  n_delim_fail <- 0
}

add_result(
  5, "Passage delimiter integrity",
  sprintf("%d/%d multi-passage acts split correctly",
          nrow(multi_passage) - n_delim_fail, nrow(multi_passage)),
  if (n_delim_fail == 0) "PASS" else "FAIL"
)
```

**Test 5**: `r if(n_delim_fail == 0) "PASS" else paste0("FAIL - ", n_delim_fail, " acts with split failure")`

## Test 6: Usable Passages Per Act

After splitting on `\n\n` and filtering to passages with more than 50 characters, every act should have at least one usable passage.

```{r}
#| label: test-6-usable-passages

passage_counts <- aligned_data |>
  mutate(
    passages = map(passages_text, ~ {
      parts <- str_split(.x, "\n\n")[[1]]
      parts <- str_trim(parts)
      parts[nchar(parts) > 50]
    }),
    n_usable = map_int(passages, length)
  )

n_no_passages <- sum(passage_counts$n_usable == 0)

add_result(
  6, "Usable passages per act",
  sprintf("%d/%d acts have >=1 usable passage",
          sum(passage_counts$n_usable > 0), nrow(aligned_data)),
  if (n_no_passages == 0) "PASS" else "FAIL"
)
```

```{r}
#| label: test-6-details

if (n_no_passages > 0) {
  passage_counts |>
    filter(n_usable == 0) |>
    select(act_name, n_passages, n_usable) |>
    gt() |>
    cols_label(
      act_name = "Act", n_passages = "Declared Passages", n_usable = "Usable"
    ) |>
    gt_theme_report()
}
```

**Test 6**: `r if(n_no_passages == 0) "PASS" else paste0("FAIL - ", n_no_passages, " acts without usable passages")`

## Test 7: No Within-Tier Duplicates

Tier 1 should have no duplicate `(doc_id, chunk_id, act_name, passage_idx)` combinations. Tier 2 should have no duplicate `(doc_id, chunk_id, act_name)` combinations. Multiple passages matching the same chunk are expected for Tier 1 and flagged as WARN.

```{r}
#| label: test-7-duplicates

# Tier 1: true duplicates (same chunk, act, passage)
tier1_n_before <- nrow(c1_chunk_data$tier1)
tier1_n_distinct <- c1_chunk_data$tier1 |>
  distinct(doc_id, chunk_id, act_name, passage_idx) |>
  nrow()
tier1_dupes <- tier1_n_before - tier1_n_distinct

# Tier 1: multi-passage matches (same chunk, different passages)
tier1_multi <- c1_chunk_data$tier1 |>
  count(doc_id, chunk_id, act_name) |>
  filter(n > 1) |>
  nrow()

# Tier 2: true duplicates
tier2_n_before <- nrow(c1_chunk_data$tier2)
tier2_n_distinct <- c1_chunk_data$tier2 |>
  distinct(doc_id, chunk_id, act_name) |>
  nrow()
tier2_dupes <- tier2_n_before - tier2_n_distinct

total_dupes <- tier1_dupes + tier2_dupes

add_result(
  7, "No within-tier duplicates",
  sprintf("T1: %d true dupes, %d multi-passage chunks; T2: %d dupes",
          tier1_dupes, tier1_multi, tier2_dupes),
  if (total_dupes == 0 && tier1_multi == 0) "PASS"
  else if (total_dupes == 0 && tier1_multi > 0) "WARN"
  else "FAIL"
)
```

**Test 7**: `r if(total_dupes == 0) "PASS" else paste0("FAIL - ", total_dupes, " true duplicates")``r if(tier1_multi > 0) paste0(" (WARN: ", tier1_multi, " multi-passage chunks in Tier 1)") else ""`

## Test 8: Token Budget Fit

All chunks plus overhead (codebook prompt + few-shot examples + system prompt + output buffer) must fit within the 200K context window.

```{r}
#| label: test-8-token-budget

# Compute actual codebook prompt tokens
codebook_prompt <- construct_codebook_prompt(c1_codebook)
codebook_tokens <- nchar(codebook_prompt) / 4

# Overhead components
few_shot_tokens <- 3100   # estimated from generate_c1_loocv_fold_examples()
system_tokens <- 125      # system prompt
output_tokens <- 2000     # output buffer
overhead <- codebook_tokens + few_shot_tokens + system_tokens + output_tokens

context_limit <- 200000

# Get max tokens across all tiers
max_tier1_tokens <- if (nrow(c1_chunk_data$tier1) > 0) max(c1_chunk_data$tier1$approx_tokens, na.rm = TRUE) else 0
max_tier2_tokens <- if (nrow(c1_chunk_data$tier2) > 0) max(c1_chunk_data$tier2$approx_tokens, na.rm = TRUE) else 0
max_neg_tokens <- if (nrow(c1_chunk_data$negatives) > 0) max(c1_chunk_data$negatives$approx_tokens, na.rm = TRUE) else 0
max_chunk_tokens <- max(max_tier1_tokens, max_tier2_tokens, max_neg_tokens)

total_max <- max_chunk_tokens + overhead
fits <- total_max < context_limit

budget_tbl <- tibble(
  component = c("Codebook prompt", "Few-shot examples", "System prompt",
                "Output buffer", "Max chunk", "Total"),
  tokens = c(round(codebook_tokens), few_shot_tokens, system_tokens,
             output_tokens, max_chunk_tokens, round(total_max))
)

budget_tbl |>
  gt() |>
  cols_label(component = "Component", tokens = "Tokens (est.)") |>
  fmt_number(columns = tokens, decimals = 0) |>
  tab_style(
    style = cell_fill(color = if (fits) "#ccffcc" else "#ffcccc"),
    locations = cells_body(rows = component == "Total")
  ) |>
  gt_theme_report()

add_result(
  8, "Token budget fit",
  sprintf("Max total: %s / %s", scales::comma(round(total_max)), scales::comma(context_limit)),
  if (fits) "PASS" else "FAIL"
)
```

**Test 8**: `r if(fits) "PASS" else "FAIL"` (max `r scales::comma(round(total_max))` tokens vs. `r scales::comma(context_limit)` limit)

## Test 9: Text Encoding Safety

Sample chunks from each tier and check for null bytes and control characters that could corrupt API calls.

```{r}
#| label: test-9-encoding

# Control character pattern (excludes \x00 which R can't store, and \t \n \r which are safe)
control_pattern <- "[\x01-\x08\x0b\x0c\x0e-\x1f]"

check_encoding <- function(texts, label) {
  n <- min(100, length(texts))
  if (n == 0) return(tibble(tier = label, n_checked = 0, n_unsafe = 0))
  sample_texts <- texts[sample(length(texts), n)]
  n_unsafe <- sum(str_detect(sample_texts, control_pattern))
  tibble(tier = label, n_checked = n, n_unsafe = n_unsafe)
}

set.seed(44)
encoding_results <- bind_rows(
  check_encoding(c1_chunk_data$tier1$text, "Tier 1"),
  check_encoding(c1_chunk_data$tier2$text, "Tier 2"),
  check_encoding(c1_chunk_data$negatives$text, "Negative")
)

total_unsafe <- sum(encoding_results$n_unsafe)

encoding_results |>
  gt() |>
  cols_label(tier = "Tier", n_checked = "Sampled", n_unsafe = "Unsafe") |>
  gt_theme_report()

add_result(
  9, "Text encoding safety",
  sprintf("%d unsafe chars in %d sampled chunks",
          total_unsafe, sum(encoding_results$n_checked)),
  if (total_unsafe == 0) "PASS" else "WARN"
)
```

**Test 9**: `r if(total_unsafe == 0) "PASS" else paste0("WARN - ", total_unsafe, " chunks with control characters")`

## Test 10: Year Filter Consistency

The tier data should respect the `max_doc_year = 2007` parameter. No chunk in Tier 1 or Tier 2 should come from a document published after 2007.

```{r}
#| label: test-10-year-filter

max_doc_year <- 2007L

tier1_max_year <- if (nrow(c1_chunk_data$tier1) > 0) max(c1_chunk_data$tier1$year, na.rm = TRUE) else NA
tier2_max_year <- if (nrow(c1_chunk_data$tier2) > 0) max(c1_chunk_data$tier2$year, na.rm = TRUE) else NA

year_ok <- (is.na(tier1_max_year) || tier1_max_year <= max_doc_year) &&
           (is.na(tier2_max_year) || tier2_max_year <= max_doc_year)

add_result(
  10, "Year filter consistency",
  sprintf("Tier 1 max year: %s, Tier 2 max year: %s (limit: %d)",
          as.character(tier1_max_year), as.character(tier2_max_year), max_doc_year),
  if (year_ok) "PASS" else "FAIL"
)
```

**Test 10**: `r if(year_ok) "PASS" else "FAIL"`

## Test 11: Few-Shot Example Preview

Generate examples for one LOOCV fold (first act held out) and verify they render correctly. This confirms that `generate_c1_loocv_fold_examples()` produces usable training data.

```{r}
#| label: test-11-fewshot-setup

# Hold out the first act
held_out_act <- aligned_data$act_name[1]
train_data <- aligned_data |> filter(act_name != held_out_act)

examples <- generate_c1_loocv_fold_examples(
  train_data,
  c1_chunk_data$negatives,
  n_per_class = 5,
  codebook = c1_codebook,
  seed = 20251206
)

n_examples <- length(examples)
n_positive <- sum(map_chr(examples, ~ .x$output$label) == "FISCAL_MEASURE")
n_negative <- n_examples - n_positive

add_result(
  11, "Few-shot example preview",
  sprintf("%d examples (%d pos, %d neg) for fold '%s'",
          n_examples, n_positive, n_negative,
          str_trunc(held_out_act, 30)),
  if (n_examples >= 2) "PASS" else "FAIL"
)
```

```{r}
#| label: test-11-render
#| results: asis

cat(sprintf("\n\n**Held-out act:** %s\n\n", held_out_act))
cat(sprintf("**Examples generated:** %d (%d positive, %d negative)\n\n",
            n_examples, n_positive, n_negative))

for (i in seq_along(examples)) {
  ex <- examples[[i]]
  label <- ex$output$label
  act_name <- ex$output$measure_name
  reasoning <- ex$output$reasoning
  input_trunc <- str_trunc(ex$input, 400)

  cat(sprintf("\n\n#### Example %d: %s\n\n", i, label))
  if (!is.null(act_name)) cat(sprintf("**Act:** %s\n\n", act_name))
  cat(sprintf("**Reasoning:** %s\n\n", reasoning))
  cat(sprintf("**Input (truncated):**\n\n> %s\n\n",
              str_replace_all(input_trunc, "\n", "\n> ")))
}
```

## Test 12: Summary Dashboard

```{r}
#| label: test-12-dashboard
#| tbl-cap: "Pre-flight verification dashboard"

add_result(
  12, "Summary dashboard",
  sprintf("%d tests completed", nrow(test_results)),
  "INFO"
)

# Build final dashboard (exclude self-referential Test 12)
dashboard <- test_results |>
  filter(test_id < 12) |>
  mutate(
    test_label = sprintf("T%d: %s", test_id, test_name)
  ) |>
  select(test_label, result, status)

n_pass <- sum(dashboard$status == "PASS")
n_warn <- sum(dashboard$status == "WARN")
n_fail <- sum(dashboard$status == "FAIL")

dashboard |>
  gt() |>
  cols_label(
    test_label = "Test",
    result = "Result",
    status = "Status"
  ) |>
  tab_style(
    style = cell_fill(color = "#ccffcc"),
    locations = cells_body(columns = status, rows = status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffffcc"),
    locations = cells_body(columns = status, rows = status == "WARN")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcccc"),
    locations = cells_body(columns = status, rows = status == "FAIL")
  ) |>
  tab_style(
    style = cell_fill(color = "#e0e0e0"),
    locations = cells_body(columns = status, rows = status == "INFO")
  ) |>
  gt_theme_report()
```

**Overall**: `r n_pass` PASS, `r n_warn` WARN, `r n_fail` FAIL out of `r nrow(dashboard)` tests.

`r if(n_fail > 0) "**ACTION REQUIRED:** Fix all FAIL tests before running S2 LOOCV." else if(n_warn > 0) "**Proceed with caution:** All critical tests pass but warnings should be reviewed." else "**All clear:** Data pipeline is validated for S2 LOOCV."`
