---
title: "Malaysia PDF URL Manifest: Human Verification Dashboard"
subtitle: "Phase 2 deployment — clickable checklist for human download"
date: today
execute:
  cache: false
  warning: false
  message: false
---

## Setup

```{r setup}
library(tidyverse)
library(targets)
library(here)
library(gt)

here::i_am("notebooks/verify_malaysia_urls.qmd")
tar_config_set(store = here("_targets"))

source(here("R/gt_theme.R"))

country_urls_list <- tar_read(country_urls)
malaysia_urls <- country_urls_list |>
  purrr::keep(~ "country" %in% names(.x) &&
                any(.x$country == "malaysia", na.rm = TRUE)) |>
  dplyr::bind_rows()
```

This dashboard renders the Malaysia URL manifest produced by `R/pull_malaysia.R`. Each row is one PDF target. Rows are grouped by `access_status`:

- **`auto`** — direct PDF URL verified; the deployment pipeline fetches automatically.
- **`manual_pending`** — human action required. Click the URL, navigate to the PDF, save it to the displayed local path. After re-running `tar_make(country_urls)`, the row flips to `manual_ready`.
- **`manual_ready`** — file present at `local_path`; pipeline ingests it on the next text-extraction step.

## Overview

```{r overview}
overview <- malaysia_urls |>
  summarise(
    total_targets = n(),
    distinct_series = n_distinct(body),
    distinct_years = n_distinct(year),
    year_min = min(year, na.rm = TRUE),
    year_max = max(year, na.rm = TRUE),
    auto = sum(access_status == "auto"),
    manual_ready = sum(access_status == "manual_ready"),
    manual_pending = sum(access_status == "manual_pending")
  ) |>
  mutate(across(everything(), as.character)) |>
  pivot_longer(everything(), names_to = "metric", values_to = "value") |>
  mutate(metric = str_replace_all(metric, "_", " ") |> str_to_title())

overview |>
  gt() |>
  tab_header(title = "Malaysia URL manifest — overview") |>
  gt_theme_report()
```

```{r status-by-series}
malaysia_urls |>
  count(body, access_status) |>
  pivot_wider(names_from = access_status, values_from = n, values_fill = 0L) |>
  gt() |>
  tab_header(title = "Status counts by document series") |>
  gt_theme_report()
```

## Auto — verified direct PDF URLs

These will be fetched automatically by the deployment pipeline. No human action required.

```{r auto-rows}
auto_rows <- malaysia_urls |>
  filter(access_status == "auto") |>
  arrange(year, package_id)

if (nrow(auto_rows) == 0L) {
  cat("(none — no verified direct PDF URLs in the current manifest)\n")
} else {
  auto_rows |>
    mutate(
      pdf_link = sprintf('<a href="%s" target="_blank">%s</a>',
                         pdf_url, package_id)
    ) |>
    select(year, body, pdf_link, source, notes) |>
    gt() |>
    fmt_markdown(columns = pdf_link) |>
    cols_label(
      year = "Year", body = "Series", pdf_link = "Package (link)",
      source = "Source", notes = "Notes"
    ) |>
    tab_header(title = "Auto-fetch URLs (verified direct PDFs)") |>
    gt_theme_report()
}
```

## Manual ready — files dropped, awaiting extraction

Rows where the human has already placed a PDF at `local_path`. The pipeline will extract these on the next text-extraction step.

```{r manual-ready}
ready_rows <- malaysia_urls |>
  filter(access_status == "manual_ready") |>
  arrange(year, package_id)

if (nrow(ready_rows) == 0L) {
  cat("(none — no files dropped yet)\n")
} else {
  ready_rows |>
    select(year, body, package_id, local_path, notes) |>
    gt() |>
    cols_label(
      year = "Year", body = "Series", package_id = "Package",
      local_path = "Local path", notes = "Notes"
    ) |>
    tab_header(title = "Manual-ready (file present, ready to extract)") |>
    gt_theme_report()
}
```

## Manual pending — human download required

This is the work queue. For each row:

1. Click the **URL** column to open the page in a browser.
2. Navigate to the PDF if the URL is a landing page (some archives require year selection or anti-bot click-through).
3. Save the PDF to the displayed **Local path**. Create subdirectories as needed.
4. Re-render this notebook (or re-run `tar_make(country_urls)`) to confirm the row flipped to `manual_ready`.

```{r manual-pending-by-series}
#| results: asis
pending_rows <- malaysia_urls |>
  filter(access_status == "manual_pending") |>
  arrange(body, year, package_id)

if (nrow(pending_rows) == 0L) {
  cat("(none — all manual rows already resolved)\n")
} else {
  for (series in unique(pending_rows$body)) {
    cat("\n### ", series, "\n\n", sep = "")
    series_rows <- pending_rows |> filter(body == series)
    tbl <- series_rows |>
      mutate(
        pdf_link = ifelse(
          is.na(pdf_url) | !nzchar(pdf_url),
          "(no URL — physical archive)",
          sprintf('<a href="%s" target="_blank">open</a>', pdf_url)
        )
      ) |>
      select(year, package_id, pdf_link, local_path, notes) |>
      gt() |>
      fmt_markdown(columns = pdf_link) |>
      cols_label(
        year = "Year", package_id = "Package", pdf_link = "URL",
        local_path = "Drop file here", notes = "Notes"
      ) |>
      gt_theme_report()
    cat(gt::as_raw_html(tbl), "\n\n")
  }
}
```

## Drop convention

Per `data/manual/malaysia/README.md`, downloaded PDFs go in:

```
data/manual/malaysia/
├── economic_report/   <year>.pdf
├── budget_speech/     <year>.pdf
├── bnm_annual_report/ <year>.pdf
├── rmk/               rmk<plan>_{plan,mtr}_<year>.pdf
└── stimulus/          named per package_id
```

After dropping files, re-run `tar_make(country_urls)` (cheap; no API calls) to refresh the manifest.
