---
title: "C2 Input Verification"
subtitle: "Validate act-level assembled data for motivation classification"
execute:
  cache: false
---

## Overview

Before building the C2 codebook or running evaluation, this notebook verifies that `c2_act_data` correctly assembles chunk text by act and attaches motivation labels. Five tests check coverage, label completeness, token budgets, class balance, and concatenation structure.

```{r}
#| label: setup
#| cache: false

pacman::p_load(targets, tidyverse, gt, here, scales)

here::i_am("notebooks/verify_c2_inputs.qmd")
tar_config_set(store = here("_targets"))
source(here("R/gt_theme.R"))
```

```{r}
#| label: load-data

c2_act_data  <- tar_read(c2_act_data)
aligned_data <- tar_read(aligned_data)
```

```{r}
#| label: init-results

test_results <- tibble(
  test_id = integer(),
  test_name = character(),
  result = character(),
  status = character()
)

add_result <- function(id, name, result, status) {
  test_results <<- bind_rows(
    test_results,
    tibble(test_id = id, test_name = name, result = result, status = status)
  )
}
```

## Test 1: Act Coverage

Every act in `aligned_data` should appear in `c2_act_data`.

```{r}
#| label: test-1-coverage

aligned_acts <- sort(unique(aligned_data$act_name))
c2_acts <- sort(unique(c2_act_data$act_name))

missing_from_c2 <- setdiff(aligned_acts, c2_acts)
extra_in_c2 <- setdiff(c2_acts, aligned_acts)
coverage_ok <- length(missing_from_c2) == 0 && length(extra_in_c2) == 0

add_result(
  1, "Act coverage",
  sprintf("%d/%d acts present", length(c2_acts), length(aligned_acts)),
  if (coverage_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: test-1-details
#| results: asis

if (length(missing_from_c2) > 0) {
  cat("\n\n**Acts in aligned_data but missing from c2_act_data:**\n\n")
  for (act in missing_from_c2) cat(sprintf("- %s\n", act))
}
if (length(extra_in_c2) > 0) {
  cat("\n\n**Acts in c2_act_data but not in aligned_data:**\n\n")
  for (act in extra_in_c2) cat(sprintf("- %s\n", act))
}
```

**Test 1**: `r if(coverage_ok) "PASS" else paste0("FAIL - ", length(missing_from_c2), " missing, ", length(extra_in_c2), " extra")`

## Test 2: Label Completeness

Every row must have non-NA `motivation_category` and `exogenous_flag`.

```{r}
#| label: test-2-labels

n_missing_motivation <- sum(is.na(c2_act_data$motivation_category))
n_missing_exogenous <- sum(is.na(c2_act_data$exogenous_flag))
labels_ok <- n_missing_motivation == 0 && n_missing_exogenous == 0

add_result(
  2, "Label completeness",
  sprintf("motivation: %d NA, exogenous: %d NA", n_missing_motivation, n_missing_exogenous),
  if (labels_ok) "PASS" else "FAIL"
)
```

**Test 2**: `r if(labels_ok) "PASS" else paste0("FAIL - ", n_missing_motivation + n_missing_exogenous, " missing labels")`

## Test 3: Token Distribution

All assembled texts must fit within the 200K context window after accounting for overhead (codebook prompt + output buffer).

```{r}
#| label: test-3-tokens

overhead_tokens <- 10000  # conservative estimate for codebook + output
context_limit <- 200000
effective_limit <- context_limit - overhead_tokens

token_summary <- c2_act_data |>
  summarise(
    min = min(approx_tokens),
    p25 = quantile(approx_tokens, 0.25),
    median = median(approx_tokens),
    mean = round(mean(approx_tokens)),
    p75 = quantile(approx_tokens, 0.75),
    p95 = quantile(approx_tokens, 0.95),
    max = max(approx_tokens)
  )

n_over_limit <- sum(c2_act_data$approx_tokens > effective_limit)
tokens_ok <- n_over_limit == 0

add_result(
  3, "Token budget fit",
  sprintf("Max: %s tokens (limit: %s)", comma(token_summary$max), comma(effective_limit)),
  if (tokens_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: test-3-histogram
#| fig-cap: "Approximate token count per act (assembled text)"
#| fig-width: 8
#| fig-height: 4

ggplot(c2_act_data, aes(x = approx_tokens)) +
  geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = effective_limit, color = "red", linetype = "dashed",
             linewidth = 0.8) +
  annotate("text", x = effective_limit, y = Inf, label = "Context limit",
           hjust = 1.1, vjust = 2, color = "red", size = 3.5) +
  scale_x_continuous(labels = comma) +
  labs(x = "Approximate Tokens", y = "Acts") +
  theme_minimal()
```

```{r}
#| label: tbl-token-summary
#| tbl-cap: "Token distribution summary across acts"

token_summary |>
  pivot_longer(everything(), names_to = "Statistic", values_to = "Tokens") |>
  gt() |>
  fmt_number(columns = Tokens, decimals = 0) |>
  gt_theme_report()
```

**Test 3**: `r if(tokens_ok) "PASS" else paste0("FAIL - ", n_over_limit, " acts exceed context limit")`

## Test 4: Motivation Class Balance

Distribution should match `aligned_data`: Long-run (15), Spending-driven (10), Deficit-driven (8), Countercyclical (6).

```{r}
#| label: test-4-balance

c2_dist <- c2_act_data |>
  count(motivation_category, name = "n_c2") |>
  arrange(desc(n_c2))

aligned_dist <- aligned_data |>
  count(motivation_category, name = "n_aligned") |>
  arrange(desc(n_aligned))

balance_check <- c2_dist |>
  left_join(aligned_dist, by = "motivation_category") |>
  mutate(match = n_c2 == n_aligned)

balance_ok <- all(balance_check$match)

add_result(
  4, "Motivation class balance",
  sprintf("%d/%d categories match aligned_data", sum(balance_check$match), nrow(balance_check)),
  if (balance_ok) "PASS" else "FAIL"
)
```

```{r}
#| label: tbl-balance
#| tbl-cap: "Motivation category distribution"

exog_map <- c2_act_data |>
  distinct(motivation_category, exogenous_flag)

balance_check |>
  left_join(exog_map, by = "motivation_category") |>
  select(motivation_category, exogenous_flag, n_c2, n_aligned) |>
  gt() |>
  cols_label(
    motivation_category = "Category",
    exogenous_flag = "Exogenous?",
    n_c2 = "c2_act_data",
    n_aligned = "aligned_data"
  ) |>
  gt_theme_report()
```

**Test 4**: `r if(balance_ok) "PASS" else "FAIL — distribution mismatch"`

## Test 5: Rendered Example

Visual inspection of the assembled text for a mid-length act. The `---` delimiters should clearly separate individual chunks.

```{r}
#| label: test-5-example
#| results: asis

# Pick the act closest to median token count
median_tokens <- median(c2_act_data$approx_tokens)
example_act <- c2_act_data |>
  mutate(dist = abs(approx_tokens - median_tokens)) |>
  slice_min(dist, n = 1, with_ties = FALSE)

cat(sprintf("\n\n### %s\n\n", example_act$act_name))
cat(sprintf("**Year:** %d | **Category:** %s | **Exogenous:** %s\n\n",
            example_act$year, example_act$motivation_category, example_act$exogenous_flag))
cat(sprintf("**Chunks used:** %d (%d Tier 1, %d Tier 2) | **Tokens:** %s\n\n",
            example_act$n_chunks_used, example_act$n_tier1, example_act$n_tier2,
            comma(example_act$approx_tokens)))

# Show first 3000 chars to keep notebook readable
text_preview <- substr(example_act$assembled_text, 1, 3000)
if (nchar(example_act$assembled_text) > 3000) {
  text_preview <- paste0(text_preview, "\n\n[... truncated at 3000 chars ...]")
}

cat("\n\n::: {.callout-note collapse=\"true\"}\n## Full assembled text (click to expand)\n\n")
cat("```\n")
cat(text_preview)
cat("\n```\n\n:::\n")
```

```{r}
#| label: tbl-per-act
#| tbl-cap: "Per-act chunk and token summary"

c2_act_data |>
  select(act_name, year, motivation_category, n_chunks_used, n_tier1, n_tier2, approx_tokens) |>
  arrange(desc(approx_tokens)) |>
  gt() |>
  cols_label(
    act_name = "Act",
    year = "Year",
    motivation_category = "Category",
    n_chunks_used = "Chunks",
    n_tier1 = "T1",
    n_tier2 = "T2",
    approx_tokens = "Tokens"
  ) |>
  fmt_number(columns = approx_tokens, decimals = 0) |>
  gt_theme_report()
```

## Summary Dashboard

```{r}
#| label: tbl-dashboard
#| tbl-cap: "C2 input verification dashboard"

dashboard <- test_results |>
  mutate(test_label = sprintf("T%d: %s", test_id, test_name)) |>
  select(test_label, result, status)

n_pass <- sum(dashboard$status == "PASS")
n_fail <- sum(dashboard$status == "FAIL")

dashboard |>
  gt() |>
  cols_label(
    test_label = "Test",
    result = "Result",
    status = "Status"
  ) |>
  tab_style(
    style = cell_fill(color = "#ccffcc"),
    locations = cells_body(columns = status, rows = status == "PASS")
  ) |>
  tab_style(
    style = cell_fill(color = "#ffcccc"),
    locations = cells_body(columns = status, rows = status == "FAIL")
  ) |>
  gt_theme_report()
```

**Overall**: `r n_pass` PASS, `r n_fail` FAIL out of `r nrow(dashboard)` tests.

`r if(n_fail > 0) "**ACTION REQUIRED:** Fix all FAIL tests before proceeding to C2 codebook development." else "**All clear:** Act-level data is validated for C2 motivation classification."`
