---
title: "Literature base"
date: 2026-06-24
date-format: long
toc: true
toc-depth: 1
---

```{r}
#| label: setup
#| include: false
pacman::p_load(tinytable, dplyr, here)

here::i_am("docs/lit_base.qmd")
```

*Role: the annotated bibliography for the methodological paper — one entry per source, recording (i) its status in the library, (ii) a summary of its contents, and (iii) its relation to our project and the rest of the literature. This is the per-source knowledge base that feeds `docs/lit_review.qmd` (synthesis) and is distinct from `docs/lit_log.md` (process decisions). Maintained by the `/lit-intake` (status + stubs), `/lit-digest` (summary + relation), and `/lit-synthesize` (reads from here) skills.*

The `consultation` field is the shared state: `pending` (stub, not yet read), `abstract` (read at abstract depth), or `full` (full-text read and distilled). `/lit-digest` advances it. Citekeys come from `references.bib`; sources in Zotero but not yet exported show "pending re-export" plus their Zotero item key.

```{r}
#| label: tbl-status
#| tbl-cap: "Library status by source. File = PDF in Zotero; Tag = section/role tags applied; Consult = consultation depth."
status <- tibble::tribble(
  ~Source, ~Sec, ~Role, ~File, ~Tag, ~Consult,
  # --- s1 motivation ---
  "vegh_how_2015",            "§1", "cite-only",   "yes", "yes", "pending",
  "jha_effectiveness_2014",   "§1", "cite-only",   "yes", "yes", "pending",
  "frankel_graduation_2013",  "§1", "cite-only",   "yes", "yes", "pending",
  "ilzetzki_how_2013",        "§1", "cite-only",   "yes", "yes", "pending",
  "acosta_ormaechea_tax_2012","§1", "cite-only",   "yes", "yes", "pending",
  "crispolti_cross-country_2022","§1","cite-only", "yes", "yes", "pending",
  "kaminsky_when_2004",       "§1", "cite-only",   "yes", "yes", "pending",
  "romer_fiscal_2019",        "§1", "cite-only",   "yes", "yes", "pending",
  "romer_new_2017",           "§1", "cite-only",   "yes", "yes", "pending",
  "angrist_credibility_2010", "§1", "cite-only",   "yes", "yes", "pending",
  "romer_trouble_nodate",     "§1", "cite-only",   "yes", "yes", "pending",
  "burke_global_2015",        "§1", "exemplar",    "yes", "yes", "abstract",
  "hsiang_quantifying_2013",  "§1", "exemplar",    "yes", "yes", "abstract",
  # --- s2 narrative ---
  "romer2010",                "§2", "read-deep",   "yes", "yes", "full",
  "romer",                    "§2", "read-deep",   "yes", "yes", "full",
  "romer_presidential_2023",  "§2", "cite-only",   "yes", "yes", "pending",
  "mertens2013",              "§2", "read-deep",   "yes", "yes", "abstract",
  "mertens_reconciliation_2014","§2","read-deep",  "yes", "yes", "abstract",
  "cloyne_discretionary_2013","§2", "read-deep",   "yes", "yes", "abstract",
  "tillmann_disagreement_2024","§2","cite-only",   "yes", "yes", "pending",
  "latifi_fiscal_nodate",     "§2", "cite-only",   "yes", "yes", "pending",
  "kim_understanding_2015",   "§2", "cite-only",   "yes", "yes", "pending",
  "hebous_revisiting_2018",   "§2", "cite-only",   "yes", "yes", "pending",
  # --- s3 effects ---
  "blanchard_empirical_nodate","§3","cite-only",   "yes", "yes", "pending",
  "auerbach_measuring_2012",  "§3", "cite-only",   "yes", "yes", "pending",
  "gechert_what_2015",        "§3", "cite-only",   "yes", "yes", "pending",
  "riera-crichton_tax_2016",  "§3", "cite-only",   "yes", "yes", "pending",
  "ramey_ten_2019",           "§3", "cite-only",   "yes", "yes", "pending",
  "ramey_macroeconomic_2016", "§3", "cite-only",   "yes", "yes", "pending",
  "ramey_identifying_2011",   "§3", "cite-only",   "yes", "yes", "pending",
  "jorda_estimation_2005",    "§3", "cite-only",   "yes", "yes", "pending",
  "jorda_local_nodate",       "§3", "cite-only",   "yes", "yes", "pending",
  "dube",                     "§3", "cite-only",   "yes", "yes", "pending",
  "ramey_government_2018",    "§3", "cite-only",   "yes", "yes", "pending",
  "dabla-norris_macroeconomic_2023","§3","cite-only","yes","yes","pending",
  # --- s4 llm text-as-data ---
  "das_mapping_nodate",       "§4", "competitor",  "yes", "yes", "full",
  "bhasin_identifying_2026",  "§4", "competitor",  "yes", "yes", "full",
  "fritsch_high-frequency_nodate","§4","competitor","yes","yes", "full",
  "aruoba_identifying_nodate","§4", "exemplar",    "yes", "yes", "abstract",
  "fernandez-fuertes_monetary_2025","§4","exemplar","yes","yes","abstract",
  "torreblanca_credibility_2025","§4","exemplar",  "yes", "yes", "abstract",
  "gentzkow_text_2019",       "§4", "cite-only",   "yes", "yes", "pending",
  "korinek_ai_2025",          "§4", "cite-only",   "yes", "yes", "pending",
  "feuerriegel_using_2025",   "§4", "cite-only",   "yes", "yes", "pending",
  "ash_text_2023",            "§4", "cite-only",   "yes", "yes", "pending",
  "grimmer_text_2022",        "§4", "cite-only",   "yes", "yes", "pending",
  "gilardi_chatgpt_2023",     "§4", "cite-only",   "yes", "yes", "pending",
  "dell_deep_2025",           "§4", "cite-only",   "yes", "yes", "pending",
  "ziems_can_2024",           "§4", "cite-only",   "yes", "yes", "pending",
  "tornberg_large_2025",      "§4", "cite-only",   "yes", "yes", "pending",
  "gambacorta_cb-lms_2024",   "§4", "cite-only",   "yes", "yes", "pending",
  "araujo_artificial_2024",   "§4", "cite-only",   "yes", "yes", "pending",
  # --- s5 validation ---
  "halterman2025",            "§5", "read-deep",   "yes", "yes", "full",
  "asirvatham_gpt_2026",      "§5", "read-deep",   "yes", "yes", "full",
  "rathje_gpt_2024",          "§5", "exemplar",    "yes", "yes", "abstract",
  "singh_deadly_2026",        "§5", "exemplar",    "yes", "yes", "abstract",
  "zhao_agentic_2026",        "§5", "exemplar",    "yes", "yes", "abstract",
  "fellegi_theory_1969",      "§5", "cite-only",   "yes", "yes", "pending",
  "abramitzky_automated_2021","§5", "cite-only",   "yes", "yes", "pending",
  "binette_almost_2022",      "§5", "cite-only",   "yes", "yes", "pending",
  "carlini_quantifying_2022", "§5", "read-deep",   "yes", "yes", "full",
  "golchin_time_2024",        "§5", "cite-only",   "yes", "yes", "pending",
  "deng_unveiling_2024",      "§5", "cite-only",   "yes", "yes", "pending",
  "ahuja_mega_2023",          "§5", "cite-only",   "yes", "yes", "pending",
  "singh_indicgenbench_2024", "§5", "cite-only",   "yes", "yes", "pending",
  "xuan_mmlu-prox_2025",      "§5", "cite-only",   "yes", "yes", "pending",
  # --- s6 generated-regressor ---
  "ludwig_large_2025",        "§6", "read-deep",   "yes", "yes", "full",
  "battaglia_inference_2024", "§6", "read-deep",   "yes", "yes", "full",
  "angelopoulos_prediction-powered_2023","§6","read-deep","yes","yes","full",
  "egami_using_2023",         "§6", "read-deep",   "yes", "yes", "full",
  # --- s7 datasets ---
  "pescatori_new_2011",       "§7", "benchmark",   "yes", "yes", "abstract",
  "adler_updated_2024",       "§7", "benchmark",   "yes", "yes", "abstract",
  "guajardo_expansionary_2014","§7","benchmark",   "yes", "yes", "abstract",
  "romer_replication_2019",   "§7", "benchmark",   "yes", "yes", "abstract"
)

tt(status)
```

# §1 Motivation

## Vegh & Vuletin 2015 — How Is Tax Policy Conducted over the Business Cycle? {#sec-vegh_how_2015}

**Citekey** `@vegh_how_2015` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Documents that tax *policy* (statutory rates), not just tax revenue, is procyclical in developing countries and acyclical in industrial ones, using a novel 62-country rate dataset.

**Relation.** *Pending `/lit-digest`.* Seed: names the `@romer2010` narrative classification as the missing input and concedes the procyclical correlation "could just reflect the effect of tax multipliers" — the gap our pipeline fills. Links: [[#sec-romer2010]].

## Jha et al. 2014 — Effectiveness of Countercyclical Fiscal Policy: Developing Asia {#sec-jha_effectiveness_2014}

**Citekey** `@jha_effectiveness_2014` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Finds tax cuts have larger countercyclical output effects than spending in developing Asia.

**Relation.** *Pending `/lit-digest`.* Seed: concedes ~70% of EM revenue variation is non-discretionary — the endogeneity narrative motivation removes.

## Frankel, Végh & Vuletin 2013 — On Graduation from Fiscal Procyclicality {#sec-frankel_graduation_2013}

**Citekey** `@frankel_graduation_2013` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** About a third of developing countries have "graduated" to countercyclical fiscal policy; institutional quality is the key determinant.

**Relation.** *Pending `/lit-digest`.* Seed: supplies the "graduation" angle (does the *effectiveness* of tax policy change as institutions improve?). Now filed in `_Fiscal Shocks` (Zotero `R64M3ZMY`).

## Ilzetzki, Mendoza & Végh 2013 — How Big (Small?) Are Fiscal Multipliers? {#sec-ilzetzki_how_2013}

**Citekey** `@ilzetzki_how_2013` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Government-spending multipliers depend on country characteristics: near zero under flexible rates, large under fixed, negative at high debt.

**Relation.** *Pending `/lit-digest`.* Seed: the spending result whose tax-side analog is our Candidate-2 flagship. Now filed in `_Fiscal Shocks` (Zotero `XWMP6F75`).

## Kaminsky, Reinhart & Végh 2004 — When It Rains, It Pours {#sec-kaminsky_when_2004}

**Citekey** `@kaminsky_when_2004` · Zotero `C375EN8P` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Origin of the "when it rains, it pours" procyclicality stylized fact for emerging-market capital flows and macro policy.

**Relation.** *Pending `/lit-digest`.* Seed: the named puzzle our flagship decomposes. Note: the former duplicate `QJWG8RSN` no longer appears in the collection (likely merged at source); citekey now resolves in the re-exported bib.

## Romer & Romer 2019 — Fiscal Space and the Aftermath of Financial Crises {#sec-rr_fiscal_space}

**Citekey** `@romer_fiscal_2019` · Zotero `5JBXFRX9` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Fiscal-space framing for the severity of post-crisis downturns.

**Relation.** *Pending `/lit-digest`.* Seed: motivation-side framing; R&R authors, but not the tax-narrative method.

## Romer & Romer 2017 — New Evidence on the Aftermath of Financial Crises {#sec-rr_new_evidence}

**Citekey** `@romer_new_2017` · Zotero `HYNHB8X3` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Empirical evidence on output costs of financial crises in advanced economies.

**Relation.** *Pending `/lit-digest`.* Seed: motivation-side framing companion to [[#sec-rr_fiscal_space]].

## Angrist & Pischke 2010 — The Credibility Revolution in Empirical Economics {#sec-angrist_credibility_2010}

**Citekey** `@angrist_credibility_2010` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** The design-based "credibility revolution" in empirical micro.

**Relation.** *Pending `/lit-digest`.* Seed: framing for why credible identification matters; sets up the narrative-identification stakes.

## Romer (Paul) — The Trouble With Macroeconomics {#sec-romer_trouble_nodate}

**Citekey** `@romer_trouble_nodate` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** Polemic that macroeconomic identification remains opaque and under-disciplined by fact.

**Relation.** *Pending `/lit-digest`.* Seed: the identification critique our narrative + validation approach answers.

## Burke, Hsiang & Miguel 2015 — Global Non-linear Effect of Temperature on Economic Production {#sec-burke_global_2015}

**Citekey** `@burke_global_2015` · Zotero `PU2P7CVS` · **Status** file ✓ · tag ✓ · §1 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Burke, Hsiang & Miguel (2015, *Nature*) estimate a single, globally common non-linear relationship between annual temperature and aggregate economic production from a large subnational country panel, finding output peaks near 13 °C and falls sharply at higher temperatures — the same dose–response in rich and poor countries — then project large global GDP losses under continued warming.

**Relation.** The paper-architecture template: invent one new, comparable cross-country unit and estimate a single transferable function over it. That is the structural move our project makes by producing comparable *exogenous fiscal-shock* series across countries from heterogeneous national records. Imperfect §-fit (climate, not fiscal) — filed in §1 as a framing exemplar. Links: [[#sec-hsiang_quantifying_2013]].

## Hsiang, Burke & Miguel 2013 — Quantifying the Influence of Climate on Human Conflict {#sec-hsiang_quantifying_2013}

**Citekey** `@hsiang_quantifying_2013` · Zotero `K4C3I9HW` · **Status** file ✓ · tag ✓ · §1 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Hsiang, Burke & Miguel (2013, *Science*) meta-analyze dozens of studies across archaeology, criminology, economics, geography, history and political science onto a common standardized climate–conflict effect size, finding that warmer temperatures and rainfall extremes robustly raise both interpersonal and intergroup conflict.

**Relation.** The second Burke–Hsiang architecture template — converging heterogeneous studies onto one comparable unit — the rhetorical model for presenting a single cross-country fiscal-shock measure built from disparate national sources. Note: two Zotero copies exist (`K4C3I9HW`, `7C6BBH7P`); merge at source (citekey `hsiang_quantifying_2013` unaffected). Links: [[#sec-burke_global_2015]].

## Acosta Ormaechea & Yoo 2012 — Tax Composition and Growth: A Broad Cross-Country Perspective {#sec-acosta_ormaechea_tax_2012}

**Citekey** `@acosta_ormaechea_tax_2012` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** IMF cross-country panel linking tax composition (income vs. consumption/property) to long-run growth.

**Relation.** *Pending `/lit-digest`.* Seed: the canonical "tax composition and growth" anchor the motivation needs — comparable cross-country shock series are both scarce and valuable. Fills the §1 PENDING ACQUISITION slot in `lit_review.qmd`. Links: [[#sec-vegh_how_2015]].

## Crispolti, Amaglobeli & Sheng 2022 — Cross-Country Evidence on the Revenue Impact of Tax Reforms {#sec-crispolti_cross-country_2022}

**Citekey** `@crispolti_cross-country_2022` · **Status** file ✓ · tag ✓ · §1 · consultation: pending
**Role** cite-only

**Summary.** IMF cross-country study estimating the revenue impact of discretionary tax-policy reforms, built on a tax-reform action dataset.

**Relation.** *Pending `/lit-digest`.* Seed: cross-country tax-reform/revenue evidence motivating why comparable exogenous tax measures matter. **Dual-fit:** filed §1 (motivation) but defensibly §7 (adjacent to the IMF Tax Policy Reform Database / statutory-rate panels) — revisit at synthesis. Links: [[#sec-acosta_ormaechea_tax_2012]], [[#sec-dabla-norris_macroeconomic_2023]].

# §2 Narrative identification (pillar)

## Romer & Romer 2010 — Macroeconomic Effects of Tax Changes {#sec-romer2010}

**Citekey** `@romer2010` · **Status** file ✓ · tag ✓ · §2 · consultation: full
**Role** read-deep

**Summary.** The foundational narrative method: read historical documents to identify *why* taxes changed and separate exogenous changes (long-run growth / deficit-reduction motivated) from endogenous ones (responding to current or prospective output). Full implementation detail lives in `docs/codebook_sources.md` §1.

**Relation.** The method we automate — the spine of the whole project. Every codebook (C1–C4) operationalizes a step of it. Links: [[#sec-romer]], [[#sec-cloyne_discretionary_2013]], [[#sec-romer_presidential_2023]].

## Romer & Romer — A Narrative Analysis of Postwar Tax Changes {#sec-romer}

**Citekey** `@romer` · **Status** file ✓ · tag ✓ · §2 · consultation: full
**Role** read-deep

**Summary.** The companion act-by-act narrative classification underlying `@romer2010`; the source of the 50-act reference index distilled in `docs/codebook_sources.md` §1.6.

**Relation.** Our gold reference for US validation (`us_shocks.csv` lineage). Links: [[#sec-romer2010]].

## Romer & Romer 2023 — Presidential Address: Does Monetary Policy Matter? {#sec-romer_presidential_2023}

**Citekey** `@romer_presidential_2023` · **Status** file ✓ · tag ✓ · §2 · consultation: pending
**Role** cite-only

**Summary.** States the requirements for rigorous narrative identification (the four-requirements framing), applied to monetary policy.

**Relation.** *Pending `/lit-digest`.* Seed: the narrative-method doctrine; Das et al. adopt its four requirements (see `codebook_sources.md` §2.1).

## Mertens & Ravn 2013 — Dynamic Effects of Personal and Corporate Income Tax Changes {#sec-mertens2013}

**Citekey** `@mertens2013` · **Status** file ✓ · tag ✓ · §2 · consultation: abstract
**Role** read-deep

**Summary.** Personal income tax changes are far more potent for output than corporate ones, a composition decomposition only the narrative record enables.

**Relation.** Motivates carrying PIT/CIT/VAT composition through our schema. Links: [[#sec-mertens_reconciliation_2014]].

## Mertens & Ravn 2014 — A Reconciliation of SVAR and Narrative Tax-Multiplier Estimates {#sec-mertens_reconciliation_2014}

**Citekey** `@mertens_reconciliation_2014` · **Status** file ✓ · tag ✓ · §2 · consultation: abstract
**Role** read-deep

**Summary.** Pure SVARs cannot recover the tax multiplier without assuming the output-elasticity of revenue; narrative proxies as instruments yield larger multipliers (~2 on impact, up to 3), and narrative measurement error is itself consequential.

**Relation.** Method-scrutiny anchor: why the narrative record is the necessary input, and a direct bridge to the §6 generated-regressor/measurement-error problem. Links: [[#sec-ludwig_large_2025]].

## Cloyne 2013 — Discretionary Tax Changes and the Macroeconomy (UK) {#sec-cloyne_discretionary_2013}

**Citekey** `@cloyne_discretionary_2013` · **Status** file ✓ · tag ✓ · §2 · consultation: abstract
**Role** read-deep

**Summary.** Applies the R&R narrative strategy to the UK and finds "remarkably similar" magnitudes (1% tax cut → +0.6% GDP on impact, +2.5% over three years).

**Relation.** The single most important precedent for our cross-country transfer claim — the only non-US tax-narrative series. Links: [[#sec-romer2010]].

## Tillmann et al. 2024 — Disagreement about Fiscal Policy {#sec-tillmann_disagreement_2024}

**Citekey** `@tillmann_disagreement_2024` · **Status** file ✓ · tag ✓ · §2 · consultation: pending
**Role** cite-only

**Summary.** Uses NLP on Bundestag speeches (1960–) to construct fiscal-disagreement series and trace macro effects.

**Relation.** *Pending `/lit-digest`.* Seed: the German textual-analysis tradition — reads parliamentary speech rather than budget documents. Links: [[#sec-latifi_fiscal_nodate]].

## Latifi et al. — Fiscal Policy in the Bundestag {#sec-latifi_fiscal_nodate}

**Citekey** `@latifi_fiscal_nodate` · **Status** file ✓ · tag ✓ · §2 · consultation: pending
**Role** cite-only

**Summary.** Embedding-based measurement of fiscal-policy sentiment in Bundestag debates, fed into VARs to identify spending shocks.

**Relation.** *Pending `/lit-digest`.* Seed: a contrast to the document-based R&R lineage we follow. Links: [[#sec-tillmann_disagreement_2024]].

## Kim 2015 — Understanding Narrative Inquiry {#sec-kim_understanding_2015}

**Citekey** `@kim_understanding_2015` · **Status** file ✓ · tag ✓ · §2 · consultation: pending
**Role** cite-only

**Summary.** A social-science methodology text on narrative inquiry.

**Relation.** *Pending `/lit-digest`.* Seed: narrative-methodology background; weak fit, retained as a framing cite.

## Hebous & Zimmermann 2018 — Revisiting the Narrative Approach of Estimating Tax Multipliers {#sec-hebous_revisiting_2018}

**Citekey** `@hebous_revisiting_2018` · **Status** file ✓ · tag ✓ · §2 · consultation: pending
**Role** cite-only

**Summary.** Re-examines the narrative tax-multiplier estimates, probing sensitivity of the R&R/narrative series and the multiplier magnitudes they imply.

**Relation.** *Pending `/lit-digest`.* Seed: the "method scrutiny" cite that rounds out §2 alongside [[#sec-mertens_reconciliation_2014]] — fills the §2 PENDING ACQUISITION slot in `lit_review.qmd`. Links: [[#sec-romer2010]].

# §3 Effects of fiscal policy

## Blanchard & Perotti 2002 — Dynamic Effects of Spending and Taxes on Output {#sec-blanchard_perotti}

**Citekey** `@blanchard_empirical_nodate` · Zotero `XDDZP426` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** The SVAR baseline for identifying fiscal shocks via institutional-timing assumptions.

**Relation.** *Pending `/lit-digest`.* Seed: the SVAR tradition our narrative approach is an alternative to. Citekey now resolves in the re-exported bib.

## Auerbach & Gorodnichenko 2012 — Measuring the Output Responses to Fiscal Policy {#sec-auerbach_gorodnichenko}

**Citekey** `@auerbach_measuring_2012` · Zotero `RKVTWKCU` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** State-dependent fiscal multipliers — larger in recessions than expansions.

**Relation.** *Pending `/lit-digest`.* Seed: motivates state-dependence as a (de-prioritized) downstream question. Citekey now resolves in the re-exported bib.

## Gechert 2015 — What Fiscal Policy Is Most Effective? A Meta-Regression {#sec-gechert}

**Citekey** `@gechert_what_2015` · Zotero `PN2C6PIY` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Meta-regression synthesizing multiplier estimates across the literature.

**Relation.** *Pending `/lit-digest`.* Seed: reference point for plausible multiplier magnitudes. Citekey now resolves in the re-exported bib.

## Riera-Crichton, Végh & Vuletin 2016 — Tax Multipliers: Pitfalls in Measurement {#sec-riera_crichton}

**Citekey** `@riera-crichton_tax_2016` · Zotero `TNSX9KDL` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Catalogues measurement and identification pitfalls specific to tax multipliers.

**Relation.** *Pending `/lit-digest`.* Seed: warns of the exact errors our magnitude/exogeneity outputs must avoid. Citekey now resolves in the re-exported bib.

## Ramey 2019 — Ten Years after the Financial Crisis: Renaissance in Fiscal Research {#sec-ramey_2019}

**Citekey** `@ramey_ten_2019` · Zotero `SYE7PVL2` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Survey of the post-crisis fiscal-multiplier literature.

**Relation.** *Pending `/lit-digest`.* Seed: situates demand for credible shock series. Citekey now resolves in the re-exported bib.

## Ramey 2016 — Macroeconomic Shocks and Their Propagation {#sec-ramey_macroeconomic_2016}

**Citekey** `@ramey_macroeconomic_2016` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Handbook chapter synthesizing identification of monetary, fiscal, and technology shocks.

**Relation.** *Pending `/lit-digest`.* Seed: the authoritative gap citation — the handbook's fiscal section is US-only; "emerging"/"developing" never appear.

## Ramey 2011 — Identifying Government Spending Shocks: It's All in the Timing {#sec-ramey_identifying_2011}

**Citekey** `@ramey_identifying_2011` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Anticipation of fiscal changes biases shock identification; timing is decisive.

**Relation.** *Pending `/lit-digest`.* Seed: the anticipation/timing dimension any shock series (and C3) must respect.

## Jordà 2005 — Estimation and Inference of Impulse Responses by Local Projections {#sec-jorda_estimation_2005}

**Citekey** `@jorda_estimation_2005` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Introduces local projections for impulse-response estimation.

**Relation.** *Pending `/lit-digest`.* Seed: the estimation toolkit our downstream payoff would use. Links: [[#sec-jorda_local_nodate]], [[#sec-dube]].

## Jordà — Local Projections for Applied Economics {#sec-jorda_local_nodate}

**Citekey** `@jorda_local_nodate` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Review connecting local projections to the potential-outcomes/treatment-effects framework.

**Relation.** *Pending `/lit-digest`.* Seed: bridges the macro IRF and applied-micro languages. Links: [[#sec-jorda_estimation_2005]].

## Dube et al. — A Local Projections Approach to DiD Event Studies {#sec-dube}

**Citekey** `@dube` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** LP-based difference-in-differences event-study estimator.

**Relation.** *Pending `/lit-digest`.* Seed: a candidate downstream estimator for the shock series.

## Ramey & Zubairy 2018 — Government Spending Multipliers in Good Times and in Bad {#sec-ramey_government_2018}

**Citekey** `@ramey_government_2018` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Estimates state-dependent government-spending multipliers from US historical data using local projections, finding little evidence multipliers are larger in slack.

**Relation.** *Pending `/lit-digest`.* Seed: strengthens the state-dependence point in §3 and pairs the LP toolkit with a fiscal application — fills the §3 PENDING ACQUISITION slot in `lit_review.qmd`. Note: two Zotero copies exist — the PDF copy `WRUHDRNP` (tagged, `duplicate-review`) and a no-PDF copy `46AXGD36`; merge at source keeping the PDF copy (citekey `ramey_government_2018` unaffected). Links: [[#sec-jorda_estimation_2005]].

## Dabla-Norris & Lima 2023 — Macroeconomic Effects of Tax Rate and Base Changes {#sec-dabla-norris_macroeconomic_2023}

**Citekey** `@dabla-norris_macroeconomic_2023` · **Status** file ✓ · tag ✓ · §3 · consultation: pending
**Role** cite-only

**Summary.** Estimates the macroeconomic effects of discretionary tax changes, separating tax-rate from tax-base reforms, using evidence drawn from action-based fiscal-consolidation episodes.

**Relation.** *Pending `/lit-digest`.* Seed: an effects-of-tax-changes study whose rate-vs-base decomposition speaks to our magnitude/composition schema. **Dual-fit:** filed §3 (effects) but defensibly §7 (built on the action-based consolidation dataset lineage) — revisit at synthesis. Links: [[#sec-mertens2013]], [[#sec-crispolti_cross-country_2022]], [[#sec-pescatori_new_2011]].

# §4 LLMs as measurement tools

## Das et al. 2026 — Mapping Government Spending Actions Across 64 Countries {#sec-das_mapping_nodate}

**Citekey** `@das_mapping_nodate` · Zotero `WBYNNC32` · **Status** file ✓ · tag ✓ · §4 · consultation: full
**Role** competitor

**Summary.** Das, Furceri, Patel & Peralta-Alva (2026, IMF WP/26/43) build the first global *quarterly* narrative database of discretionary government-spending actions by applying a fixed, zero-shot GPT-4.1 prompt to Economist Intelligence Unit Country Reports (64 countries, 1952–2023; 8,636 episodes). The prompt screens R&R-style for *exogenous* spending actions, but the production output is a discrete direction proxy z ∈ {−1, 0, +1} — spending-only, no tax side, no %GDP magnitude (an optional intensity field is excluded from the econometrics). Validation is fourfold: replicating R&R (2019) expert coding (>93%), a 51-run replicability test, local-projection prediction of realized spending, and ≥96% alignment with Adler et al. (2024) %GDP consolidations. Headline: a median cumulative spending multiplier ~0.7, larger in less-open economies, under fixed exchange rates, in downturns, and at the ZLB, identified via country-level Bayesian proxy-SVAR.

**Relation.** The flagship competitor — it shares our exact thesis that a fixed, country-agnostic, off-the-shelf LLM prompt can scale R&R narrative identification cross-country *without retraining*, and even shares design instincts we use (conservative net-aggregation of multi-action passages, verbatim-evidence retention, separate validation vs. production prompts). We differ on all four differentiators: (1) spending-only vs. our tax-side depth; (2) **direction only** — the authors explicitly "do not observe dollar-equivalent sizes … only their qualitative direction," and their 51-run test shows the granular intensity field is the *least* replicable while coarse direction is stable, corroborating our position that quantitative %GDP sizing is the hard part and our contribution; (3) no Halterman & Keith behavioral/zero-shot/error-analysis framework vs. our S0–S3; (4) English-only EIU reports vs. our EN/BM bilingual transfer. Their three-tier no-ground-truth validation (expert-replication % + aggregate prediction + independent-dataset alignment) is a reusable template for our Malaysia/SEA setting. Note: Das positions only against Aruoba & Drechsel (monetary) and does *not* cite Bhasin or Fritsch — the competitor cluster is one we draw, not one they assert. Links: [[#sec-bhasin_identifying_2026]], [[#sec-fritsch_high-frequency_nodate]], [[#sec-adler_2024]], [[#sec-aruoba_identifying_nodate]].

## Bhasin & Loungani 2026 — Identifying Episodes of Fiscal Austerity {#sec-bhasin_identifying_2026}

**Citekey** `@bhasin_identifying_2026` · Zotero `BLQ3YEEH` · **Status** file ✓ · tag ✓ · §4 · consultation: full
**Role** competitor

**Summary.** Bhasin & Loungani (2026, *J. Economic Dynamics & Control*) build a hierarchical five-prompt LLM cascade to identify fiscal-austerity (consolidation) episodes from IMF Article IV staff reports (2004–2020, 17 OECD countries): country filter → fiscal-policy gate → four-pillar austerity definition → R&R-style exogeneity filter → %GDP magnitude. Core model Llama-3.3-70B (GPT-4.1 / DeepSeek-R1 robustness checks converge). Validated against the Guajardo et al. (2014) / Adler et al. (2024) consolidation dataset over 289 country-years: precision 0.93, recall 0.90, F1 0.91. Headline: LLM-identified shocks yield *smaller* multipliers than the narrative benchmark, and a swap-decomposition attributes the divergence to shock *sizing*, not event selection.

**Relation.** The closest direct competitor on method — LLM-automated R&R identification with an explicit exogeneity gate *and* a %GDP magnitude step — sharing our cross-country ambition. It diverges on our differentiators: consolidation episodes broadly (revenue *or* expenditure lumped into a binary austerity label) vs. our tax decomposition; magnitude secondary to a binary label (size correlation only 0.59, systematically smaller than benchmark) vs. our first-class per-act %GDP; a single harmonized *secondary* source (Article IV, written *about* countries by IMF staff), validated against existing ground truth, vs. our *primary* budget/treasury documents with no Malaysia/SEA ground truth and H&K/expert-agreement gates they do not use. **Correction to the prior seed:** the paper does say "shock sizing, not event selection," but frames it *oppositely* — it argues the *narrative benchmark's* larger multipliers are inflated by persistence/residual endogeneity (ρ≈0.73, predictable from lagged GDP) and presents its own smaller, less-persistent sizes as *better-identified*. So it concedes only that Article IV reports yield conservative ("headline-package") magnitudes, not that magnitude is a binding accuracy constraint — our positioning should not overstate this. Their Prompt-4 exogeneity exemplars (Maastricht/SGP compliance, constitutional debt-brake, pre-announced medium-term programs) are concrete worked examples that could sharpen our C2 exogenous-side examples. Links: [[#sec-das_mapping_nodate]], [[#sec-fritsch_high-frequency_nodate]], [[#sec-guajardo_leigh_pescatori]], [[#sec-adler_2024]].

## Fritsch & Mazlish — High-Frequency Fiscal Shocks {#sec-fritsch_high-frequency_nodate}

**Citekey** `@fritsch_high-frequency_nodate` · Zotero `FWILZBHC` · **Status** file ✓ · tag ✓ · §4 · consultation: full
**Role** competitor

**Summary.** Fritsch & Mazlish ("High-Frequency Fiscal Shocks," Oxford, preliminary) build a two-agent LLM pipeline producing daily-frequency US fiscal shocks (1947–2025): a *forecaster* agent reads each day's fiscal/macro news (full Reuters/WSJ articles for 2005–25; summaries pre-2005) and updates a present-value-of-deficits forecast, while a *classifier* agent labels each change endogenous / fiscal-event / both, apportions it into purchases (G), taxes/transfers (T) and interest (I), and tags permanence and stabilization. It thus recovers signed, %GDP-scaled magnitudes *with* a tax-vs-spending split at daily/event frequency, with a prompt-imposed knowledge-cutoff guarding against look-ahead bias. Validation: correlation 0.71 with Ramey (2011) defense shocks and **0.89 with R&R (2010) tax shocks**, plus high-frequency asset-price reactions (10Y yields, dollar, term spread) and multipliers.

**Relation.** The most direct competitor on three of our differentiators at once — signed %GDP magnitude, R&R-style exogeneity, and a tax-side decomposition — and its 0.89 correlation with R&R tax shocks is a strong validation benchmark. Our defensible distinctions: (i) it is **US-only**, with cross-country deployment merely *promised* as future work, whereas our contribution is the *executed* country-agnostic transfer (US→Malaysia→SEA) with frozen codebooks and expert-agreement gates; (ii) its inputs are **news wires** building a *forecast/expectations* series, fundamentally different from our reading of *primary* budget/treasury documents to classify *enacted* measures; (iii) it lacks the formal H&K 5-stage scaffolding we use, relying on correlation-recovery and asset-price/multiplier validation. Its high-frequency asset-price identification is *orthogonal* to our document-based approach (our unit is the enacted act at quarterly/annual frequency, not a datestamped market event), so it is a complement to cite, not a method to adopt. Links: [[#sec-das_mapping_nodate]], [[#sec-bhasin_identifying_2026]].

## Aruoba & Drechsel — Identifying Monetary Policy Shocks: A Natural-Language Approach {#sec-aruoba_identifying_nodate}

**Citekey** `@aruoba_identifying_nodate` · Zotero `NP8WKD26` · **Status** file ✓ · tag ✓ · §4 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Aruoba & Drechsel identify monetary-policy shocks by applying natural-language-processing techniques to the documents Federal Reserve staff prepare in advance of policy decisions, extracting the information set behind each decision and purging the systematic component; the residual surprises are validated by the absence of forecast-error predictability and by theory-consistent impulse responses.

**Relation.** The validate-without-ground-truth precedent we follow: measurement validity is established through forecast-error orthogonality and theory-consistent IRFs rather than concordance against an existing series — the template for our no-ground-truth Phase 2 gates. Das et al. (2026) likewise position against this as their main narrative-identification precedent. Note: two Zotero copies (`NP8WKD26` journalArticle, no date; `I43BYWTD` 2024 preprint); merge at source. Links: [[#sec-asirvatham_gpt_2026]], [[#sec-das_mapping_nodate]].

## Fernandez-Fuertes 2025 — Monetary Policy Shocks: LLMs and Central-Bank Communication {#sec-fernandez_fuertes}

**Citekey** `@fernandez-fuertes_monetary_2025` · Zotero `QBBU9ZQ2` · **Status** file ✓ · tag ✓ · §4 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Fernandez-Fuertes (2025, SSRN preprint) develops a multi-agent LLM framework that reads Fed communications (Beige Books, Minutes) released before each FOMC meeting to generate conditional expectations and construct narrative monetary-policy surprises that are *less noisy* than market-based measures, yielding theoretically consistent impulse responses and profitable yield-curve strategies. By extracting expectations directly rather than cleaning surprises ex post, it implements narrative identification at scale without high-frequency contamination.

**Relation.** A multi-agent-LLM-as-measurement deployment outside fiscal policy, and the closest monetary-side analog to our pipeline: it shares the move of extracting measurements/expectations directly from policy documents, and its explicit contamination-avoidance concern echoes the threat our S1 memorization test guards against. A companion to Aruoba & Drechsel on the monetary side of the LLM-narrative-identification frontier. Links: [[#sec-aruoba_identifying_nodate]], [[#sec-asirvatham_gpt_2026]].

## Torreblanca et al. 2025 — The Credibility Revolution in Political Science {#sec-torreblanca_credibility_2025}

**Citekey** `@torreblanca_credibility_2025` · Zotero `NUBCCS2W` · **Status** file ✓ · tag ✓ · §4 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Torreblanca, Dinneen, Grossman & Xu (2025, preprint) use an LLM to classify 91,632 political-science articles published 2003–2023 across 174 journals by causal research design, charting how the credibility revolution reshaped the discipline.

**Relation.** A large-scale LLM-as-classifier exemplar in a social science — demonstrating reliable LLM annotation at the corpus scale our deployment operates in, and pairing the LLM measurement tool with the credibility-revolution framing (cf. Angrist & Pischke) that motivates our identification stakes. Links: [[#sec-angrist_credibility_2010]].

## Gentzkow, Kelly & Taddy 2019 — Text as Data {#sec-gentzkow_text_2019}

**Citekey** `@gentzkow_text_2019` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Canonical survey of text-as-data methods for economics.

**Relation.** *Pending `/lit-digest`.* Seed: the canon — and the adversarial flag that fixed text measures notoriously fail to transfer across domains (which our frozen codebooks must survive). Links: [[#sec-halterman2025]].

## Korinek 2025 — AI Agents for Economic Research {#sec-korinek_ai_2025}

**Citekey** `@korinek_ai_2025` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Practical guide to building LLM agents for economic research workflows.

**Relation.** *Pending `/lit-digest`.* Seed: the LLM-for-economists meta layer; catalogues hallucination/error-cascade/prompt-brittleness caveats and declines to vouch for LLMs in causal inference.

## Feuerriegel et al. 2025 — Using NLP to Analyse Text Data in Behavioural Science {#sec-feuerriegel_using_2025}

**Citekey** `@feuerriegel_using_2025` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Review of the NLP pipeline and modelling choices for behavioural-science text analysis.

**Relation.** *Pending `/lit-digest`.* Seed: situates our tooling in the broader NLP-for-social-science methods literature.

## Ash & Hansen 2023 — Text Algorithms in Economics {#sec-ash_text_2023}

**Citekey** `@ash_text_2023` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Annual Review survey of text-as-data algorithms (embeddings, topic models, supervised learning) and their economics applications.

**Relation.** *Pending `/lit-digest`.* Seed: the econ text-as-data canon alongside [[#sec-gentzkow_text_2019]]; situates our LLM tooling in the methods lineage. Fills part of the §4 PENDING ACQUISITION slot.

## Grimmer, Roberts & Stewart 2022 — Text as Data: A New Framework for Machine Learning and the Social Sciences {#sec-grimmer_text_2022}

**Citekey** `@grimmer_text_2022` · **Status** file ✓ (book; no PDF) · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Textbook framework for text-as-data in the social sciences: the discovery → measurement → inference workflow, and the validation discipline measurement requires.

**Relation.** *Pending `/lit-digest`.* Seed: the measurement-validity framing our H&K pipeline instantiates; canonical §4 cite. Book (cannot be PDF-filed) — citekey resolves in `references.bib`. Links: [[#sec-halterman2025]].

## Gilardi, Alizadeh & Kubli 2023 — ChatGPT Outperforms Crowd Workers for Text-Annotation Tasks {#sec-gilardi_chatgpt_2023}

**Citekey** `@gilardi_chatgpt_2023` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Shows zero-shot ChatGPT exceeds crowd workers (and approaches experts) on several text-annotation tasks in political science.

**Relation.** *Pending `/lit-digest`.* Seed: the LLM-annotation-validity anchor for §4 — the social-science precedent that LLMs can stand in for human coders, which our expert-agreement gates both rely on and stress-test. Links: [[#sec-halterman2025]].

## Ziems et al. 2024 — Can Large Language Models Transform Computational Social Science? {#sec-ziems_can_2024}

**Citekey** `@ziems_can_2024` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Benchmarks zero-shot LLMs across a suite of computational-social-science annotation tasks, mapping where they can and cannot substitute for human coders.

**Relation.** *Pending `/lit-digest`.* Seed: the capabilities-and-limits survey that bounds the LLM-as-annotator claim our pipeline rests on. Links: [[#sec-gilardi_chatgpt_2023]], [[#sec-tornberg_large_2025]].

## Törnberg 2025 — LLMs Outperform Expert Coders and Supervised Classifiers at Annotating Political Social Media {#sec-tornberg_large_2025}

**Citekey** `@tornberg_large_2025` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Finds zero-shot LLMs exceed both trained expert coders and supervised classifiers at annotating political social-media messages.

**Relation.** *Pending `/lit-digest`.* Seed: a strong-form LLM-annotation-validity cite reinforcing [[#sec-gilardi_chatgpt_2023]]; the precedent our expert-agreement gates lean on. Links: [[#sec-ziems_can_2024]].

## Dell 2025 — Deep Learning for Economists {#sec-dell_deep_2025}

**Citekey** `@dell_deep_2025` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Survey of deep-learning and LLM methods for economists, covering text-as-data measurement, embeddings, and structured extraction.

**Relation.** *Pending `/lit-digest`.* Seed: the LLM-for-economists meta layer situating our tooling in the discipline's methods canon. Links: [[#sec-korinek_ai_2025]], [[#sec-gentzkow_text_2019]].

## Gambacorta et al. 2024 — CB-LMs: Language Models for Central Banking {#sec-gambacorta_cb-lms_2024}

**Citekey** `@gambacorta_cb-lms_2024` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** Introduces encoder-only language models retrained on a central-bank corpus (speeches, policy documents, research), outperforming foundational models on central-bank idiom tasks.

**Relation.** *Pending `/lit-digest`.* Seed: a domain-specialized-LM exemplar for applying LLMs to government/policy economic text. Links: [[#sec-araujo_artificial_2024]].

## Araujo et al. 2024 — Artificial Intelligence in Central Banking {#sec-araujo_artificial_2024}

**Citekey** `@araujo_artificial_2024` · **Status** file ✓ · tag ✓ · §4 · consultation: pending
**Role** cite-only

**Summary.** BIS overview of AI/LLM adoption across central-bank functions, including text analysis of policy and economic documents.

**Relation.** *Pending `/lit-digest`.* Seed: the BIS LLM-primer the plan names for §4; situates LLM economic-text analysis institutionally. Links: [[#sec-gambacorta_cb-lms_2024]].

# §5 Validation (pillar)

## Halterman & Keith 2025 — Codebook LLMs {#sec-halterman2025}

**Citekey** `@halterman2025` · **Status** file ✓ · tag ✓ · §5 · consultation: full
**Role** read-deep

**Summary.** The five-stage codebook-LLM measurement framework (prepare → behavioral tests → zero-shot eval → error analysis → fine-tune), built on the premise that codebook conceptualization is a first-order concern. Distilled in `docs/codebook_sources.md` §3.

**Relation.** The validation spine of the project — our S0–S3 pipeline is its instantiation. Links: [[#sec-asirvatham_gpt_2026]], [[#sec-gentzkow_text_2019]].

## Asirvatham, Mokski & Shleifer 2026 — GPT as a Measurement Tool {#sec-asirvatham_gpt_2026}

**Citekey** `@asirvatham_gpt_2026` · **Status** file ✓ · tag ✓ · §5 · consultation: full
**Role** read-deep

**Summary.** The GABRIEL package; separates **accuracy** (benchmark agreement) from **directness** (the measurement comes from the text, via a signal-stripping test); contamination-via-cutoff and prompt-robustness evidence; human-parity framing. Distilled in `docs/codebook_sources.md` §5.2.

**Relation.** Closest external template for "is the LLM output a valid measurement"; seeds a proposed C2 motivation-masking robustness test and the human-parity framing for our no-ground-truth Phase 2 gates. Links: [[#sec-halterman2025]], [[#sec-ludwig_large_2025]].

## Rathje et al. 2024 — GPT for Multilingual Psychological Text Analysis {#sec-rathje_gpt_2024}

**Citekey** `@rathje_gpt_2024` · Zotero `8LABL4VM` · **Status** file ✓ · tag ✓ · §5 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Rathje et al. (2024, *PNAS Nexus*) show that GPT detects psychological constructs (sentiment, discrete emotions, offensiveness) across 12 languages with accessible zero-shot prompts, outperforming established dictionary methods and approaching fine-tuned models without per-language retraining.

**Relation.** Directly relevant to our EN/BM cross-language stress test: evidence that a single LLM can transfer a measurement construct across languages with no per-language retraining — the capability our Malaysia bilingual leg depends on — while the cross-lingual benchmark literature warns that this transfer degrades for lower-resource languages. Links: [[#sec-ahuja_mega_2023]], [[#sec-singh_indicgenbench_2024]], [[#sec-xuan_mmlu-prox_2025]], [[#sec-halterman2025]].

## Singh 2026 — Deadly Stigma {#sec-singh_deadly_2026}

**Citekey** `@singh_deadly_2026` · Zotero `C3AVEYPB` · **Status** file ✓ · tag ✓ · §5 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Singh (2026, preprint "Deadly Stigma") uses a physician-validated LLM to detect stigmatizing language in clinical notes, then exploits exogenous variation in stigma to estimate its real-world harm — solving the measurement problem (stigma is hard to observe) so that a credible research design becomes feasible.

**Relation.** An LLM-as-measurement validated against expert (physician) judgment and then fed into a causal design — the same two-step pattern as our pipeline (LLM measurement → downstream inference), with expert validation as the credibility anchor. A close analog to our expert-agreement gates and to the §6 generated-regressor concern of carrying LLM measurement error into downstream estimates. Links: [[#sec-zhao_agentic_2026]], [[#sec-halterman2025]].

## Zhao et al. 2026 — An Agentic System for Rare-Disease Diagnosis {#sec-zhao_agentic_2026}

**Citekey** `@zhao_agentic_2026` · Zotero `ATXC3JZM` · **Status** file ✓ · tag ✓ · §5 · consultation: abstract
**Role** exemplar-other-domain

**Summary.** Zhao et al. (2026, *Nature*) present a multi-agent LLM system for rare-disease diagnosis with *traceable reasoning* chains, reporting ~95% expert agreement on its reasoning steps (not just final labels) and strong diagnostic accuracy across a large patient cohort.

**Relation.** The traceable-reasoning + expert-agreement validation pattern: experts verify the reasoning that produced a label, not only the label itself. A direct model for how our expert-validation gates could assess C2 motivation *rationales* (the evidence-and-reasoning behind an exogenous/endogenous call), strengthening the credibility argument beyond marginal-label agreement. Links: [[#sec-singh_deadly_2026]].

## Fellegi & Sunter 1969 — A Theory for Record Linkage {#sec-fellegi_theory_1969}

**Citekey** `@fellegi_theory_1969` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** The foundational probabilistic record-linkage theory: decision rule for matching/non-matching record pairs under controlled error rates.

**Relation.** *Pending `/lit-digest`.* Seed: the classical anchor for the C0 aggregation / entity-resolution sub-task — the "is this the same act" problem our canonical clustering solves with an LLM. Fills the §5-core record-linkage PENDING ACQUISITION slot. Links: [[#sec-abramitzky_automated_2021]], [[#sec-binette_almost_2022]].

## Abramitzky, Boustan & Eriksson et al. 2021 — Automated Linking of Historical Data {#sec-abramitzky_automated_2021}

**Citekey** `@abramitzky_automated_2021` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** JEL survey of automated record-linkage methods in economic history, benchmarking linking algorithms and their error/bias trade-offs.

**Relation.** *Pending `/lit-digest`.* Seed: the econ-history precedent for automated linking that legitimizes a machine-assisted C0 aggregation step. Links: [[#sec-fellegi_theory_1969]], [[#sec-binette_almost_2022]].

## Binette & Steorts 2022 — (Almost) All of Entity Resolution {#sec-binette_almost_2022}

**Citekey** `@binette_almost_2022` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** Modern survey of entity resolution: blocking, pairwise matching, clustering, and evaluation against ground truth.

**Relation.** *Pending `/lit-digest`.* Seed: the state-of-the-art ER framing for C0 — blocking + clustering vocabulary maps onto our year-blocked and canonical-clustering methods. Links: [[#sec-fellegi_theory_1969]], [[#sec-abramitzky_automated_2021]].

## Carlini et al. 2022 — Quantifying Memorization Across Neural Language Models {#sec-carlini_quantifying_2022}

**Citekey** `@carlini_quantifying_2022` · **Status** file ✓ · tag ✓ · §5 · consultation: full
**Role** read-deep

**Summary.** Measures verbatim memorization by prompting models with true training-set prefixes and checking exact-suffix reproduction under greedy decoding. Establishes three log-linear laws: extractable memorization rises with model size (≈+19 pp per 10× parameters; 2–5× across a family), with training-data duplication, and with prompt-context length — the "discoverability" effect, 33%→65% extractable from 50→450 tokens of context. A GPT-2-vs-GPT-Neo control (prompt a model on data it could *not* have seen and compare) isolates genuine memorization from generic predictive competence, and estimates ≥1% of The Pile is verbatim-extractable from a 6B model.

**Relation.** The mechanistic backing for our H&K **S1 memorization test**. The threat is concrete: the published 44-act R&R list and the source ERP/Treasury documents are exactly the high-value, plausibly-duplicated public text most at risk of verbatim recall, so a model could "recognize" an act by name/year rather than reason from the supplied chunk. Carlini supplies both the risk levers (larger models, duplication, long name-and-year-rich prefixes) and the control logic our S1 probe borrows — withhold or perturb the identifying surface forms (act name, P.L. number, year) and check whether the classification survives. Because memorization scales with model size, S1 must be re-run when the deployment model changes, not treated as a one-time check. Pillar deep-read distilled into `docs/codebook_sources.md` §3. Links: [[#sec-halterman2025]], [[#sec-golchin_time_2024]], [[#sec-deng_unveiling_2024]].

## Golchin & Surdeanu 2024 — Time Travel in LLMs: Tracing Data Contamination {#sec-golchin_time_2024}

**Citekey** `@golchin_time_2024` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** Proposes prompt-based methods to detect whether specific datasets leaked into an LLM's pretraining data.

**Relation.** *Pending `/lit-digest`.* Seed: a contamination-detection technique relevant to certifying our US historical corpus is read, not recalled. Links: [[#sec-carlini_quantifying_2022]], [[#sec-deng_unveiling_2024]].

## Deng et al. 2024 — Unveiling the Spectrum of Data Contamination: A Survey {#sec-deng_unveiling_2024}

**Citekey** `@deng_unveiling_2024` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** Survey of data-contamination detection and remediation methods across the LLM evaluation literature.

**Relation.** *Pending `/lit-digest`.* Seed: the survey anchor for the contamination/memorization risk our S1 test guards against. Links: [[#sec-carlini_quantifying_2022]], [[#sec-golchin_time_2024]].

## Ahuja et al. 2023 — MEGA: Multilingual Evaluation of Generative AI {#sec-ahuja_mega_2023}

**Citekey** `@ahuja_mega_2023` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** Benchmarks generative LLMs across many languages and tasks, documenting systematic degradation on lower-resource languages relative to English.

**Relation.** *Pending `/lit-digest`.* Seed: evidence of the cross-lingual performance gap our EN/BM (and SEA-extension) stress test must measure. Links: [[#sec-singh_indicgenbench_2024]], [[#sec-xuan_mmlu-prox_2025]], [[#sec-rathje_gpt_2024]].

## Singh et al. 2024 — IndicGenBench: Multilingual Benchmark for LLM Generation {#sec-singh_indicgenbench_2024}

**Citekey** `@singh_indicgenbench_2024` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** A large multi-way parallel benchmark across 29 Indic languages, finding persistent gaps versus English even for the strongest models.

**Relation.** *Pending `/lit-digest`.* Seed: a low-resource-language evaluation precedent directly analogous to the Bahasa Malaysia leg of our cross-language test. Links: [[#sec-ahuja_mega_2023]], [[#sec-xuan_mmlu-prox_2025]].

## Xuan et al. 2025 — MMLU-ProX: A Multilingual Benchmark for Advanced LLM Evaluation {#sec-xuan_mmlu-prox_2025}

**Citekey** `@xuan_mmlu-prox_2025` · **Status** file ✓ · tag ✓ · §5 · consultation: pending
**Role** cite-only

**Summary.** A 29-language parallel benchmark (identical questions per language) showing marked accuracy declines in low-resource languages.

**Relation.** *Pending `/lit-digest`.* Seed: the parallel-question design is the methodological template for a like-for-like EN/BM comparison; reinforces the cross-lingual-risk evidence base. Links: [[#sec-ahuja_mega_2023]], [[#sec-singh_indicgenbench_2024]].

# §6 Generated-regressor inference (pillar)

## Ludwig, Mullainathan & Rambachan 2025 — LLMs: An Applied Econometric Framework {#sec-ludwig_large_2025}

**Citekey** `@ludwig_large_2025` · **Status** file ✓ · tag ✓ · §6 · consultation: full
**Role** read-deep

**Summary.** Splits LLM uses into prediction (needs "no training leakage") and estimation/measurement (needs a validation sample for valid downstream inference). Accuracy ≠ validity (bias = correlation of LLM error with regressors); debias against a small random validation sample (125–250 items help); model/prompt choice can flip downstream estimates absent debiasing. Distilled in `docs/codebook_sources.md` §5.1.

**Relation.** A §6 pillar anchor — our LLM fiscal shocks are exactly their "estimation problem" feeding a VAR/LP. The de-risking trio (Battaglia, Angelopoulos, Egami) is now filed, ending the single-source risk. Links: [[#sec-mertens_reconciliation_2014]], [[#sec-asirvatham_gpt_2026]], [[#sec-battaglia_inference_2024]], [[#sec-angelopoulos_prediction-powered_2023]], [[#sec-egami_using_2023]].

## Battaglia, Christensen, Hansen & Sacher 2024 — Inference for Regression with Variables Generated from Unstructured Data {#sec-battaglia_inference_2024}

**Citekey** `@battaglia_inference_2024` · **Status** file ✓ · tag ✓ · §6 · consultation: full
**Role** read-deep

**Summary.** Formalizes inference for the two-step strategy in which an upstream model estimates latent variables from text (e.g. topic shares) and a downstream regression treats them as observed. Proves this is a measurement-error problem whose bias scales with the average inverse information per observation (κ = lim √n · E[1/Cᵢ]): when κ > 0 the OLS estimator is consistent but its asymptotic distribution is *re-centered*, so ordinary (Eicker–Huber–White) standard errors have the right *width* but the wrong *center* and under-cover; only κ = 0 gives valid two-step inference. Their baseline correction is a "one-step" joint upstream+downstream likelihood estimated by HMC — which presumes a generative IR likelihood an LLM judge does not provide. Extends explicitly to VARs/IRFs and similarity-measure regressions.

**Relation.** The most directly on-point external source for §6: our LLM-measured shocks are exactly a "variable generated from unstructured data" feeding a VAR/LP, and their VAR/IRF extension is our downstream object. Two transfer caveats: their κ formula keys the bias to document length under a topic-model likelihood (our LLM-classification error is not literally a 1/Cᵢ sampling process — the qualitative lesson transfers, the exact formula does not), and their joint-MLE fix needs a likelihood the LLM lacks (the authors flag this as the open frontier). Because we *have* a validation sample (44 US labeled acts; expert-labeled Malaysia), the operative correction for us is the labeled-subset strand they contrast against, with Battaglia supplying the bias *diagnosis*: the error mode is mis-centering, not mis-sizing, so it cannot be detected by inspecting reported standard errors. Pillar deep-read distilled into `docs/codebook_sources.md` §5.4. Links: [[#sec-ludwig_large_2025]], [[#sec-egami_using_2023]], [[#sec-angelopoulos_prediction-powered_2023]], [[#sec-mertens_reconciliation_2014]].

## Angelopoulos, Bates, Fannjiang, Jordan & Zrnic 2023 — Prediction-Powered Inference {#sec-angelopoulos_prediction-powered_2023}

**Citekey** `@angelopoulos_prediction-powered_2023` · **Status** file ✓ · tag ✓ · §6 · consultation: full
**Role** read-deep

**Summary.** Prediction-powered inference (PPI) builds provably valid confidence intervals and P-values for a population estimand from a small gold-labeled sample (n) plus a large unlabeled sample (N ≫ n) carrying ML predictions. It combines a *measure of fit* on the large imputed set with a *rectifier* — the prediction error estimated only on the gold sample — to debias the imputation; coverage holds for *any* ML algorithm and *any* data distribution, with no assumption on model quality. Intervals are strictly narrower than gold-only when predictions are decent and N is large; the framework covers means, quantiles, OLS/logistic coefficients (any convex-objective estimand), plus covariate- and label-shift variants. It offers no gain when predictions are too weak or N/n too small.

**Relation.** The general, assumption-light machinery underwriting our Phase 2 design: a small expert-labeled subset (PPI's gold sample) plus a large LLM-labeled corpus (the prediction sample) is a textbook PPI instance. It converts the expert subset from a one-off agreement check into the rectifier that yields valid downstream confidence intervals over the LLM-measured shock series — even when the LLM is biased, which is exactly our cross-country, no-retraining situation. The naive "treat LLM labels as truth" path is the imputation estimator the paper shows fails to cover; expert-only is valid but underpowered; PPI is the principled middle. Its stated limitation (no gain when predictions are weak or N not large relative to n) is a live risk at SEA corpus sizes, and a `ppi-py` reference implementation exists. Links: [[#sec-egami_using_2023]], [[#sec-battaglia_inference_2024]], [[#sec-ludwig_large_2025]], [[#sec-asirvatham_gpt_2026]].

## Egami, Hinck, Stewart & Wei 2023 — Using Imperfect Surrogates for Downstream Inference {#sec-egami_using_2023}

**Citekey** `@egami_using_2023` · **Status** file ✓ · tag ✓ · §6 · consultation: full
**Role** read-deep

**Summary.** Design-based supervised learning (DSL) uses imperfect, possibly-biased LLM/ML "surrogate" labels as outcomes in downstream regressions while preserving valid inference. Surrogates are observed for all n documents, but expert gold labels only for a *probability-sampled* subset drawn with a known sampling probability π bounded away from zero; DSL forms a doubly-robust pseudo-outcome (cross-fitted prediction + inverse-propensity residual correction on the gold rows) and substitutes it into the downstream moment equations. Because π is known *by design*, estimates are consistent with valid confidence intervals *regardless* of how biased the surrogate is — surrogate quality affects only efficiency. Naive surrogate-only use gave ~40% coverage of nominal-95% CIs at 90% accuracy in their simulations; DSL matched gold-only bias/coverage at materially lower RMSE across 18 datasets.

**Relation.** The social-science-facing formalization closest to our expert-validation workflow: LLM labels over the full act corpus, expert labels on a validated subset, feeding downstream macro inference. Its load-bearing requirement reshapes our protocol — the validated subset must be *probability-sampled* with a recorded π (simple or stratified), not a convenience or borderline-acts sample — and in return the expert validation does double duty: not just agreement measurement (our ≥80%/≥70% targets) but active debiasing, with SEs that survive arbitrarily biased LLM labels. This reframes the agreement target as an *efficiency* lever rather than a validity gate. Distinct from PPI in exploiting researcher-controlled sampling (removing the nuisance-rate requirement), which our fully-controlled corpus naturally satisfies. Scope caveat: DSL targets the known-corpus outcome-regression case, so our US→Malaysia cross-country transfer sits at the edge of its domain-shift scope. Links: [[#sec-angelopoulos_prediction-powered_2023]], [[#sec-battaglia_inference_2024]], [[#sec-ludwig_large_2025]].

# §7 Datasets

## Devries, Guajardo, Leigh & Pescatori 2011 — A New Action-Based Dataset of Fiscal Consolidation {#sec-pescatori_new_2011}

**Citekey** `@pescatori_new_2011` · Zotero `C9YM26FG` · **Status** file ✓ · tag ✓ · §7 · consultation: abstract
**Role** benchmark

**Summary.** Devries, Guajardo, Leigh & Pescatori (2011, IMF WP) assemble the original action-based narrative dataset of fiscal-consolidation measures for 17 OECD economies, reading contemporaneous policy documents to identify changes motivated by deficit reduction rather than by a response to prospective economic conditions.

**Relation.** The founding narrative-comparator dataset that Guajardo et al. (2014) and Adler et al. (2024) extend, and the lineage Das and Bhasin validate against — the benchmark class our output landscape is measured against. Fills the §7 PENDING ACQUISITION slot in `lit_review.qmd`. Links: [[#sec-guajardo_leigh_pescatori]], [[#sec-adler_2024]].

## Adler et al. 2024 — An Updated Action-Based Dataset of Fiscal Consolidation {#sec-adler_2024}

**Citekey** `@adler_updated_2024` · Zotero `7ZQA6BJ3` · **Status** file ✓ · tag ✓ · §7 · consultation: abstract
**Role** benchmark

**Summary.** Adler, Allen, Ganelli & Leigh (2024, IMF WP 2024/210) extend the action-based fiscal-consolidation dataset to a longer, broader OECD panel, supplying %GDP-sized consolidation measures identified from contemporaneous policy documents.

**Relation.** The OECD %GDP benchmark that the LLM competitors lean on — Das et al. report ≥96% alignment against it and Bhasin validates against the same lineage — so it is the quantitative-magnitude comparator our per-act %GDP output is positioned beside. Links: [[#sec-guajardo_leigh_pescatori]], [[#sec-pescatori_new_2011]], [[#sec-das_mapping_nodate]], [[#sec-bhasin_identifying_2026]].

## Guajardo, Leigh & Pescatori 2014 — Expansionary Austerity? International Evidence {#sec-guajardo_leigh_pescatori}

**Citekey** `@guajardo_expansionary_2014` · Zotero `MFGJ6A8H` · **Status** file ✓ · tag ✓ · §7 · consultation: abstract
**Role** benchmark

**Summary.** Guajardo, Leigh & Pescatori (2014, *JEEA*) use the action-based narrative dataset to estimate the short-run effects of fiscal consolidation on OECD output, finding consolidation is contractionary for private demand and GDP — and that conventional cyclically-adjusted measures bias estimates toward the "expansionary austerity" view.

**Relation.** The narrative-comparator study that demonstrates *why* exogenous (action-based) identification overturns conclusions drawn from endogenous fiscal measures — the exact bias our exogeneity classification removes — and the benchmark Bhasin validates against. Links: [[#sec-adler_2024]], [[#sec-pescatori_new_2011]], [[#sec-bhasin_identifying_2026]].

## Romer & Romer 2019 — Replication Data for the 2010 Tax-Shock Measure {#sec-romer_replication_2019}

**Citekey** `@romer_replication_2019` · Zotero `YCM7BMYI` · **Status** file ✓ · tag ✓ · §7 · consultation: abstract
**Role** benchmark

**Summary.** Romer & Romer (2019) ICPSR/openICPSR replication archive for the 2010 tax-shock measure — the act-by-act narrative classification and constructed exogenous tax-change series underlying "The Macroeconomic Effects of Tax Changes."

**Relation.** The US reference series (`us_shocks.csv` lineage) our pipeline replicates and validates against end-to-end in Phase 1, and the source the §4 LLM competitors (Das, Fritsch) benchmark their recovery against. Links: [[#sec-romer2010]].

# References

::: {#refs}
:::
